### 2022-03-17 InfHyp.Rmd
first commit

---
title: "Bayesian Evidence Synthesis Methods"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Matrix)
library(lavaan)
library(BFpack)
library(ggplot2)
library(metafor)
library(bain)
```

## Context
Caspar developed a method to compute an evidence synthesis BF. To do so, he (more or less) creates a datasets with $k$ independent groups, where each group represents a separate study. With this multigroup dataset, we can use `bain` or `BFpack` to calculate 'overall Bayes factors'. For $BF_{i,u}$ this is the same as the product BF that we use in Bayesian evidence synthesis. For $BF_{i,c}$ it is not. Below, I explain the difference and the link between both methods. The remaining challenge is to decide what is conceptually correct.

## Example Data + Results
In this example our evidence synthesis hypothesis with $k=3$ datasets is $H_i: r_{data1}>0.1 \& r_{data2}>0.1 \& r_{data3}>0.1$.

```{r sim}
#hyperparameters simulation
#es = effect size
#tau2 = between-study variance in es
#participants
#k=number of datasets
hyperparams <- list(
    es = 0.3,
    tau2 = 0,
    pp = 120,
    k = 3
)

#hyp_value = threshold for > in informative hypothesis
hyp_value <- .1

sim_cond <- do.call(expand.grid, hyperparams)

#hyperparameters population
#true_es = effect size cor in population
sim_data <- function(true_es, tau2, pp){
    rho <- rnorm(1, true_es, sqrt(tau2)) #reliability
    x = rnorm(pp, 0, 1) # vector of predictor?
    y = rnorm(pp, rho*x, sqrt(1-rho^2)) #outcome
    list(x = x, y = y)
}

colSD <- function(x) apply(x, 2, sd, na.rm = TRUE)

# generate data
set.seed(6164900)
dfs <- lapply(1:hyperparams$k, function(n){
    sim_data(true_es = hyperparams$es, tau2 = hyperparams$tau2, pp = hyperparams$pp)
})
# Compute correlations and se (correlations between x and y)
res <- sapply(dfs, function(x){
    r = do.call(cor.test, x)
    unname(c(r$estimate, sqrt((1 - r$estimate^2)/r$parameter)))
})
# Prepare for BFpack calls
est <- res[1, ]
names(est) <- paste0("r", 1:length(est))
Sig <- diag(res[2, ])
# Get individual BFs for each sample
bf_individual <- lapply(paste0(names(est), ">", hyp_value),
                        bain,
                        x = est,
                        Sigma = Sig,
                        n = hyperparams$pp)

fit.i <- sapply(bf_individual, function(x){x$fit$Fit[1]}) #fits of hypotheses (posterior probability of hypotheses given the data)
#in this case, posterior probability that rho > .1 given the data in each df.
com.i <- sapply(bf_individual, function(x){x$fit$Com[1]}) #fit of complement given the data.
bf_prod_c <- sapply(bf_individual, function(x){x$fit$BF.c[1]}) #BFic this is the complement and thus it is just fit.i/(1-fit.i)
bf_prod_u <- sapply(bf_individual, function(x){x$fit$BF.u[1]}) #BFiu this is just fit.i/com.i, so posterior prop of h1/posterior prop of hu

# Get overall BF
bf_together <- bain(x = est,
                    hypothesis = paste0("(", paste0(names(est), collapse = ", "), ") > ", hyp_value),
                    Sigma = Sig,
                    n = hyperparams$k*hyperparams$pp)

fit.t <- bf_together$fit$Fit[1]
com.t <- bf_together$fit$Com[1]

results = list(fit.i=fit.i,
               prod.fit=prod(fit.i), #product of informative hypotheses fits
               com.i=com.i,
               prod.com=prod(com.i),
               bf.ic=bf_prod_c,
               bf.ic.prod=prod(bf_prod_c),
               fit.t=fit.t,
               com.t=com.t,
               bf.ct=bf_together$fit$BF.c[1],
               bf.ict=prod(fit.i)/prod(com.i)/((1-prod(fit.i))/(1-prod(com.i))),
               bf.iu.prod=prod(bf_prod_u),
               bf.ut=bf_together$fit$BF.u[1])

print(results)

```

## Explanation
The product of the individual $BF_{i,u}$ = `r results$bf.iu.prod`, this equals the combined data product `r results$bf.ut`.

The product of the individual $BF_{i,c}$ = `r results$bf.ic.prod`, whereas the combined data product is `r results$bf.ct`.

The product Bayes factor equals $\frac{\prod f_i / \prod c_i}{(\prod 1-f_i)/(\prod 1-c_i)}$ = $\frac{`r results$prod.fit` / `r results$prod.com`}{`r prod(1-results$fit.i)` / `r prod(1-results$com.i)`}$ = `r results$bf.ic.prod`.

The combined data product can be recomputed based on the elements of the individual Bayes factors by $\frac{\prod f_i / \prod c_i}{(1-\prod f_i)/(1-\prod c_i)}$ = $\frac{`r results$prod.fit` / `r results$prod.com`}{`r 1-results$prod.fit` / `r 1-results$prod.com`}$ = `r results$bf.ict`.

The difference is in the denominator: should we look at the within-study complement or is it (also) correct to take a complement of the synthesized hypothesis?

With the individual product method, our complement is: "r_data_1<0.1 \& r_data_2<0.1 \& r_data_3<0.1", which has the same complexity as our hypothesis "r_data_1>0.1 \& r_data_2>0.1 \& r_data_3>0.1" = $0.5^k$ = `r 0.5^3`. In the overall method, the complement is 'not Hi', which encompasses more options = $1-0.5^k$ = `r 1-0.5^3`.

The question is whether we first test against the complement in each study and take $BF_{i,c} \times BF_{i,c} \times BF_{i,c}$ or if we can compute the synthesized $BF_{i,c}$ for $H_i: r_{data1}>0.1 \& r_{data2}>0.1 \& r_{data3}>0.1$ vs "not $H_i$".

&nbsp;

### 2022-03-17 Sim_Eli/Simulatie_EliV2.R

@@ -1,0 +1,121 @@
+#clear console and env
+rm(list=ls(all.names = T))
+cat("\014")
+
+#load necessary packages
+dependencies <- c('MASS', 'bain')
+lapply(dependencies, function(x){library(x, character.only = T)})
+
+# Check package versions
+versions <- c(
+  compareVersion(as.character(packageVersion("bain")), "0.2.8"),
+  compareVersion(as.character(packageVersion("MASS")), "7.3.55"))
+if(!all(versions == 0)) stop("Using the incorrect version of one or more packages.")
+
+# Load simulation functions from source -----------------------------------
+source('Sim_Eli/functions.R')
+
+# set conditions for simulation
+hyper_parameters<-list(
+  ndataset = 1:2,            # number of replications per condition
+  es = c(0, 0.1),        # true effect size = true correlation with outcome
+  tau2 = c(0, .1),            # reliability of outcome (standard deviation of outcome) (NOTE)
+  n = c(30, 300),             # mean sample size per group
+  k = c(3, 10),               # number of groups
+  hyp_val = c(0, 0.1)    # thresholds for informative hypotheses
+)
+
+# Create hypergrid with simulation parameters and save it as .RData file extension
+summarydata <- expand.grid(hyper_parameters, stringsAsFactors = FALSE)
+saveRDS(summarydata, file = "./Sim_Eli/summarydata.RData")
+# summarydata<-readRDS("./Sim_Eli/summarydata.RData")
+
+
+# Delete folder where previous results are in and create new 'Results' folder
+unlink("./sim_Eli/Results", recursive = T)
+dir.create("./Sim_Eli/Results")
+
+# number of algorithms used
+nalg <- 6
+
+# prepare parallel processing
+library(doSNOW)
+nclust <- parallel::detectCores() 
+cl <- makeCluster(nclust) 
+registerDoSNOW(cl) 
+
+# add progression bar
+pb <- txtProgressBar(min = 0, max = nrow(summarydata), style = 3)
+opts <- list(progress = function(n) setTxtProgressBar(pb, n))
+
+# export nalg and vector of seeds to all clusters
+set.seed(6164900)
+seeds <- sample(1:.Machine$integer.max, nrow(summarydata))
+clusterExport(cl, c('nalg', 'seeds'))
+
+# run simulation
+foreach(rownum = 1:nrow(summarydata), .options.snow = opts, .packages = c("bain", "MASS")) %dopar% {
+  
[…]

### 2022-03-17 Sim_Eli/functions.R

@@ -1,0 +1,108 @@
+#simulate data function
+simdata <- function(es, n, tau2){
+  S <- diag(2) * 1                  # initiate correlation matrix
+  rho <- rnorm(1, es, sqrt(tau2))   # generate correlation coefficient
+  
+  while(rho > 1 | rho < -1){
+    rho <- rnorm(1, es, sqrt(tau2)) # make sure correlation coefficient is a possible value
+  }
+  
+  # evt alternative for while loop
+  # ifelse(rho > 1, 1, rho) 
+  # ifelse(rho < -1, -1, rho)
+  
+  S[row(S) != col(S)] <- rho        # correlation between predictor and outcome (off-diagonals)
+  sds <- c(1,1)                     # reliability of predictor and outcome, vary outcome reliability
+  S <- diag(sds)%*%S%*%diag(sds)    # convert correlation to covariance matrix
+  
+  df <-  mvrnorm(n, mu = rep(0,2), S, tol = 1E-6, empirical = F) #generate data
+  colnames(df) <- c('y', 'x')       # give names
+  return(df)
+}
+
+
+BFs <- function(es, n, hyp_val, k, tau2){
+  
+  # create dataframe for each group
+  dfs <- lapply(1:k, function(i){
+    n_df <- floor(rnorm(1, n, n/3))     # sample a sample size
+    n_df <- ifelse(n_df < 10, 10, n_df) # make sure sample is at least of size 10
+    simdata(es, n_df, tau2)             # return for every group a dataset
+  })
+  
+  # obtain estimates for correlations and their standard errors for every dataset
+  res <- sapply(dfs, function(x){
+    est <- cor(x)[,1][-1]                     #estimate for correlation between x and y
+    se_est <- sqrt((1 - est^2)/(nrow(x) - 2)) #estimate for standard error = sqrt((1-r^2)/df) with df = N - 2
+    cbind(est, se_est, nrow(x))               #return estimate, se and sample size of particular set
+  })
+  
+  # necessary naming for bain and further preparing
+  colnames(res) <- paste0('r', 1:k)
+  sig <- lapply(res[2,], matrix)    # make list of covariance matrices for the datasets, as shown in the Bain vignette
+  ngroup <- sapply(dfs, nrow)       # obtain sample size per group
+  
+  #run bf_individual to extract product bf and geometric product bf
[…]

### 2022-03-17 Sim_Eli/preliminary_analysis.R

@@ -1,0 +1,66 @@
+########################
+# Preliminary Analyses #
+########################
+library(data.table)
+rm(list=ls())
+
+# algorithms we used
+algorithms <- c("gpbf", "prodbf", "tbf")
+hyps <- c("_ic", "_iu")
+alg_names <- paste0(rep(algorithms, each = length(hyps)), hyps)
+
+# some preparation
+summarydata <- readRDS("./Sim_Eli/summarydata.RData")
+conditions <- c("ndataset", "es", "tau2", "n", "k", "hyp_val")
+res <- readRDS(file.path("Sim_Eli", "sim_results_2022-03-14.RData"))
+lc <- length(conditions)
+
+# obtain unique conditions
+unique_conds <- unique(summarydata[,2:lc])
+unique_conds <- t(apply(unique_conds, 1, function(x){
+  paste(paste0(colnames(unique_conds), " == ", x), collapse = " & ")
+}))
+
+# convert res to datatable and make outcome values numeric
+res <- as.data.table(res)
+res[, c(lc:ncol(res)):=lapply(.SD, as.numeric), .SDcols=c(lc:ncol(res))]
+
+# we will sample 16 conditions to plot
+N <- 16
+colors <- c('#c15cf7', '#dda7fa', '#5c8df7', '#acc5fa', '#f5bd3b', '#ffe7b0')
+
+# Sample 16 integers from 1 to length(unique_conds)
+set.seed(6164900)
+sample_cond <- sample(1:length(unique_conds), 16)
+
+
+# function to plot histogram of BFs of a condition which is called in the next function
+# while subsetting the res datatable.
+plotcond <- function(outcome, name, condition, col){
+  condition <- unlist(strsplit(condition, "&"))
+  condition <- gsub("==", "=", condition)
+  condition <- paste(condition, collapse = " ")
+  
+  propBF3 <- sum(outcome > 3) / length(outcome)
+  hist(outcome, main = condition, col = col,
+       xlab = paste0("median: ", round(median(outcome),2),
+                     ";    sd: ", round(sd(outcome),2), 
+                     ";    prop BF > 3: ", round(propBF3,2)),
+       cex.lab = 1.3)
+}
+
+
+# Plot histogram s of the calculated BayesFactors from 16 random conditions in a single grid
+plot16 <- function(name, line = -1.5, colr){
+  par(mfrow = c(4,4))
+  for(i in seq_along(sample_cond)){
[…]

### 2022-03-19 Sim_Eli/Simulatie_EliV2.R
fix sim

@@ -2,6 +2,10 @@
 rm(list=ls(all.names = T))
 cat("\014")
 
+# Delete folder where previous results are in and create new 'Results' folder
+unlink("./sim_Eli/Results", recursive = T)
+dir.create("./Sim_Eli/Results")
+
 #load necessary packages
 dependencies <- c('MASS', 'bain')
 lapply(dependencies, function(x){library(x, character.only = T)})
@@ -18,25 +22,19 @@ source('Sim_Eli/functions.R')
 # set conditions for simulation
 hyper_parameters<-list(
   ndataset = 1:2,            # number of replications per condition
-  es = c(0, 0.1),        # true effect size = true correlation with outcome
-  tau2 = c(0, .1),            # reliability of outcome (standard deviation of outcome) (NOTE)
-  n = c(30, 300),             # mean sample size per group
-  k = c(3, 10),               # number of groups
-  hyp_val = c(0, 0.1)    # thresholds for informative hypotheses
+  es = c(0, 0.1, .2),        # true effect size = true correlation with outcome
+  errorsd = c(0.81, .5, 0),   # corresponds to reliability of .6, .8, and 1
+  n = c(20, 80, 200, 500),             # mean sample size per group
+  k = c(1, 3, 10),               # number of groups
+  hyp_val = c(0.1)    # thresholds for informative hypotheses
 )
 
 # Create hypergrid with simulation parameters and save it as .RData file extension
 summarydata <- expand.grid(hyper_parameters, stringsAsFactors = FALSE)
+set.seed(6164900)
+summarydata$seed <- sample(1:.Machine$integer.max, nrow(summarydata))
 saveRDS(summarydata, file = "./Sim_Eli/summarydata.RData")
 # summarydata<-readRDS("./Sim_Eli/summarydata.RData")
-
-
-# Delete folder where previous results are in and create new 'Results' folder
-unlink("./sim_Eli/Results", recursive = T)
-dir.create("./Sim_Eli/Results")
-
-# number of algorithms used
-nalg <- 6
 
 # prepare parallel processing
 library(doSNOW)
@@ -48,28 +46,67 @@ registerDoSNOW(cl)
 pb <- txtProgressBar(min = 0, max = nrow(summarydata), style = 3)
 opts <- list(progress = function(n) setTxtProgressBar(pb, n))
 
-# export nalg and vector of seeds to all clusters
-set.seed(6164900)
[…]

### 2022-03-19 Sim_Eli/functions.R

@@ -1,53 +1,37 @@
-#simulate data function
-simdata <- function(es, n, tau2){
-  S <- diag(2) * 1                  # initiate correlation matrix
-  rho <- rnorm(1, es, sqrt(tau2))   # generate correlation coefficient
-  
-  while(rho > 1 | rho < -1){
-    rho <- rnorm(1, es, sqrt(tau2)) # make sure correlation coefficient is a possible value
-  }
-  
-  # evt alternative for while loop
-  # ifelse(rho > 1, 1, rho) 
-  # ifelse(rho < -1, -1, rho)
-  
-  S[row(S) != col(S)] <- rho        # correlation between predictor and outcome (off-diagonals)
-  sds <- c(1,1)                     # reliability of predictor and outcome, vary outcome reliability
-  S <- diag(sds)%*%S%*%diag(sds)    # convert correlation to covariance matrix
-  
-  df <-  mvrnorm(n, mu = rep(0,2), S, tol = 1E-6, empirical = F) #generate data
-  colnames(df) <- c('y', 'x')       # give names
-  return(df)
-}
 
 
-BFs <- function(es, n, hyp_val, k, tau2){
+
+BFs <- function(es, n, hyp_val, k, errorsd){
   
   # create dataframe for each group
   dfs <- lapply(1:k, function(i){
-    n_df <- floor(rnorm(1, n, n/3))     # sample a sample size
-    n_df <- ifelse(n_df < 10, 10, n_df) # make sure sample is at least of size 10
-    simdata(es, n_df, tau2)             # return for every group a dataset
+    df <- rmvnorm(n, sigma = matrix(c(1, es, es, 1), nrow = 2))
+    df + rnorm(2*n, sd = errorsd)
   })
   
   # obtain estimates for correlations and their standard errors for every dataset
   res <- sapply(dfs, function(x){
-    est <- cor(x)[,1][-1]                     #estimate for correlation between x and y
-    se_est <- sqrt((1 - est^2)/(nrow(x) - 2)) #estimate for standard error = sqrt((1-r^2)/df) with df = N - 2
-    cbind(est, se_est, nrow(x))               #return estimate, se and sample size of particular set
+    est <- cor(x)[2,1]                        #estimate for correlation between x and y
+    se_est <- sqrt((1 - est^2)/(n - 2)) #estimate for standard error = sqrt((1-r^2)/df) with df = N - 2
[…]

### 2022-03-19 Sim_Eli/Simulatie_EliV2.R
update 2022-03-19 10:25:25

@@ -67,7 +67,7 @@ tab <- foreach(rownum = 1:nrow(summarydata), .options.snow = opts, .packages = c
   
   # Classic approach
   classic_allsig <- all(apply(res, 2, function(x){
-    x[1]-1.96*x[2] > hyp_val
+    (x[1]-hyp_val)/x[2] > 1.644854
   }))
   
   # necessary naming for bain and further preparing

### 2022-03-19 Sim_Eli/functions.R

@@ -18,7 +18,7 @@ BFs <- function(es, n, hyp_val, k, errorsd){
   
   # Classic approach
   classic_allsig <- all(apply(res, 2, function(x){
-    x[1]-1.96*x[2] > hyp_val
+    (x[1]-hyp_val)/x[2] > 1.644854
   }))
   
   # necessary naming for bain and further preparing

### 2022-03-19 Sim_Eli/Simulatie_EliV2.R
update 2022-03-19 10:25:55

@@ -17,7 +17,7 @@ versions <- c(
 if(!all(versions == 0)) stop("Using the incorrect version of one or more packages.")
 
 # Load simulation functions from source -----------------------------------
-source('Sim_Eli/functions.R')
+# source('Sim_Eli/functions.R')
 
 # set conditions for simulation
 hyper_parameters<-list(

### 2022-03-19 Sim_Eli/functions.R

@@ -1,64 +1,64 @@
-
-
-
-BFs <- function(es, n, hyp_val, k, errorsd){
-  
-  # create dataframe for each group
-  dfs <- lapply(1:k, function(i){
-    df <- rmvnorm(n, sigma = matrix(c(1, es, es, 1), nrow = 2))
-    df + rnorm(2*n, sd = errorsd)
-  })
-  
-  # obtain estimates for correlations and their standard errors for every dataset
-  res <- sapply(dfs, function(x){
-    est <- cor(x)[2,1]                        #estimate for correlation between x and y
-    se_est <- sqrt((1 - est^2)/(n - 2)) #estimate for standard error = sqrt((1-r^2)/df) with df = N - 2
-    c(est, se_est)               #return estimate, se and sample size of particular set
-  })
-  
-  # Classic approach
-  classic_allsig <- all(apply(res, 2, function(x){
-    (x[1]-hyp_val)/x[2] > 1.644854
-  }))
-  
-  # necessary naming for bain and further preparing
-  colnames(res) <- paste0('r', 1:k)
-  sig <- lapply(res[2,], matrix)    # make list of covariance matrices for the datasets, as shown in the Bain vignette
-  #ngroup <- rep(n, k)       # obtain sample size per group
-  
-  #run bf_individual to extract product bf and geometric product bf
-  bf_individual <- lapply(paste0(colnames(res), ">", hyp_val), # for every group, we hypothesize that r_k > hyp_val
-                          bain,                 # call bain
-                          x = res[1,],          # estimates
-                          Sigma = sig,          # all rho's are assumed to be independent
-                          n = rep(n, k),           # pass the named vector of sample sizes per group to bain
-                          group_parameters = 1, # every group k has 1 parameter which is rho_xy
-                          joint_parameters = 0) # they do not share parameters 
-
-  # extract BF_ic and BF_iu for the parameter of every group
-  BFs <- t(sapply(bf_individual, function(x){
-    c(x$fit$BF.c[1], x$fit$BF.u[1])
-  }))
-
-  # obtain geometric product and regular product
-  gp_and_prod <- apply(BFs, 2, function(x){
-    prod_bf <- prod(x)         #obtain product bf
-    c(prod_bf^(1/k), prod_bf)  #concatenate geometric product and regular product
-  })
-  
[…]

### 2022-03-19 Sim_Eli/Simulatie_EliV2.R
add results

@@ -21,7 +21,7 @@ if(!all(versions == 0)) stop("Using the incorrect version of one or more package
 
 # set conditions for simulation
 hyper_parameters<-list(
-  ndataset = 1:2,            # number of replications per condition
+  ndataset = 1:100,            # number of replications per condition
   es = c(0, 0.1, .2),        # true effect size = true correlation with outcome
   errorsd = c(0.81, .5, 0),   # corresponds to reliability of .6, .8, and 1
   n = c(20, 80, 200, 500),             # mean sample size per group

### 2022-03-19 Sim_Eli/Simulatie_EliV2.R
1000 reps results

@@ -21,7 +21,7 @@ if(!all(versions == 0)) stop("Using the incorrect version of one or more package
 
 # set conditions for simulation
 hyper_parameters<-list(
-  ndataset = 1:100,            # number of replications per condition
+  ndataset = 1:1000,            # number of replications per condition
   es = c(0, 0.1, .2),        # true effect size = true correlation with outcome
   errorsd = c(0.81, .5, 0),   # corresponds to reliability of .6, .8, and 1
   n = c(20, 80, 200, 500),             # mean sample size per group
@@ -126,6 +126,7 @@ alg_names <- c("allsig", paste0(rep(algorithms, each = length(hyps)), hyps))
 
 # read in the simulation conditions 
 res <- readRDS("./Sim_Eli/summarydata.RData")
+setDT(res)
 conditions <- colnames(res)
 
 # make sure results are same length as conditions
@@ -148,3 +149,7 @@ f <- list.files("./Sim_Eli/Results", full.names = TRUE)
 file.remove(f)
 
 # END OF FILE
+tabres <- res[, lapply(.SD, function(x){mean(x > 3)}), .SDcols = alg_names, by = c("es", "errorsd", "n", "k")]
+write.csv(tabres, "tabres.csv", row.names = FALSE)
+saveRDS(tabres, "tabres.RData")
+#colMeans(res[, .SD > 3, .SDcols = alg_names, .gro])

### 2022-03-20 analysis.R
update 2022-03-20 09:57:44

@@ -1,0 +1,48 @@
+#dat <- readRDS("tabres.RData")
+dat <- readRDS(file.path("Sim_Eli", "sim_results_2022-03-19.RData"))
+lapply_at <- function(var, truees) {
+  results <- sapply(var, function(var) {
+    table(ordered(truees > .1, levels = c("FALSE", "TRUE")), ordered(var > 3, levels = c("FALSE", "TRUE")))
+  })
+  names(results) <- vapply(names(var), paste, c("TN", "FN", "FP", "TP"), sep = "_", 
+                           FUN.VALUE = character(4),
+                           USE.NAMES = FALSE)
+  as.list(results)
+}
+varsout <- c("allsig", "gpbf_ic", "gpbf_iu", "prodbf_ic", "prodbf_iu", "tbf_ic", "tbf_iu")
+varspred <- c("es", "errorsd", "n", "k", "hyp_val")
+
+# Example: per condition
+dat[!es < .1, lapply_at(.SD, truees = es), .SDcols = varsout, by = c("n", "errorsd")]
+
+# Overall
+#res <- dat[!es < .1, lapply_at(.SD, truees = es), .SDcols = varsout]
+res <- dat[!es < .1, sapply(.SD, function(var) {
+  table(ordered(es > .1, levels = c("FALSE", "TRUE")), ordered(var > 3, levels = c("FALSE", "TRUE")))
+}), .SDcols = varsout]
+rownames(res) <- c("TN", "FN", "FP", "TP")
+
+stats <- apply(res, 2, function(x){
+  attach(as.list(x))
+  alpha = FP / (FP + TN) #False positive rate 
+  beta = FN / (TP + FN) # False negative rate
+  sensitivity = 1-beta
+  specificity = 1-alpha
+  c(False_positive_rate_alpha = alpha,
+    False_negative_rate_beta = beta,
+    sensitivity = sensitivity,
+    specificity = specificity, 
+    pos_lr = sensitivity / (1 - specificity),
+    neg_lr = (1 - sensitivity) / specificity)
+})
+
+library(ggplot2)
+df_plot <- data.frame(Method = colnames(stats),
+        Sensitivity = stats["sensitivity", ],
+        Specificity = stats["specificity", ])
+        
+ggplot(df_plot, aes(x = Sensitivity, y = Specificity, colour = Method, shape = Method)) +
+  geom_point() +
+  scale_x_continuous(limits = c(0,1))+
+  scale_y_continuous(limits = c(0,1)) +
+  theme_bw()

### 2022-03-22 Sim_Eli/Simulatie_EliV2.R
Start Analysis
It took a little while to get used to data table commands, but i'm getting the hang of it. For now some marginal descriptives of the median BF are provided and some interpretation.

Also a start is made in analysing the confusion matrix metrics.

@@ -57,7 +57,7 @@ tab <- foreach(rownum = 1:nrow(summarydata), .options.snow = opts, .packages = c
     df <- rmvnorm(n, sigma = matrix(c(1, es, es, 1), nrow = 2))
     df + rnorm(2*n, sd = errorsd)
   })
-  
+
   # obtain estimates for correlations and their standard errors for every dataset
   res <- sapply(dfs, function(x){
     est <- cor(x)[2,1]                        #estimate for correlation between x and y
@@ -103,7 +103,7 @@ tab <- foreach(rownum = 1:nrow(summarydata), .options.snow = opts, .packages = c
   
   # returns in order: gpbf_ic, gpbf_iu, prodbf_ic, prodbf_iu, tbf_ic, tbf_iu
   c(rownum,
-    c(0,10)[classic_allsig+1],
+    c(0,10)[classic_allsig+1], #return 10 if classic_allsig = T, 0 if false
     gp_and_prod[1,],
     gp_and_prod[2,], 
     c(bf_together$fit$BF.c[1], bf_together$fit$BF.u[1]))

### 2022-03-22 analysis.R

@@ -1,5 +1,57 @@
 #dat <- readRDS("tabres.RData")
-dat <- readRDS(file.path("Sim_Eli", "sim_results_2022-03-19.RData"))
+
+library(data.table) # load in library
+dat <- readRDS(file.path("Sim_Eli", "sim_results_2022-03-19.RData")) # read in results
+
+# some preparation
+varsout <- c("allsig", "gpbf_ic", "gpbf_iu", "prodbf_ic", "prodbf_iu", "tbf_ic", "tbf_iu")
+varspred <- c("es", "errorsd", "n", "k", "hyp_val")
+
+#########################
+# MARGINAL DESCRIPTIVES #
+#########################
+
+# First we obtain some descriptives over the marginal conditions.
+descriptives <- function(datf){ 
+  descs <- as.matrix(
+    datf[, lapply(.SD, function(var){
+    medians <- median(var) # obtain overall medians per algorithm over medians per condition
+    mads <- mad(var)       # obtain Median absolute deviations per conditions
+    sums <- sum(var>3)     # obtain sums where the median is > 3
+    maxvals <- max(var)    # obtain the highest median
+    c(medians, mads, sums, maxvals)
+  }), .SDcols = varsout])
+  rownames(descs) <- c("median", "mad", "sum>3", "max")
+  descs
+}
+
+(marginal_descs <- descriptives(dat[ ,lapply(.SD, median), .SDcols = varsout, by = varspred]))
+# In theory, the algorithms should have 108/3 = 36 conditions where the median BF > 3. This is because BF > 3 only
+# in the conditions where true_es > hyp_val, which is one third of the time.
+# The marginal median for all algorithms is < 1, suggesting that the BF in most conditions is indeed < 3.
+# The Median Absolute Deviations are also not too severe, giving more proof for reliable estimates.
+# Interestingly, the prodbf_ic does have a notable higher mad.
+# The finding that the algorithms are rather conservative is further supported by the number of times
+# the median BF does exceed 3. Both gpbf algorithms count 0 conditions where median BF > 3, while for allsig this
+# number is 1. For the prodbf_iu and tbf algorithms 5 or 6 times is counted and only the prodbf seems to have a 
[…]

### 2022-03-23 Report.Rmd
start report

---
title: "Report"
author: "Eli Clapper"
date: "23/03/2022"
output: html_document
---

# Intro
Report on the results

call library and load in results
```{r, echo = F}
library(data.table) # load in libraries
library(DT)
dat <- readRDS(file.path("Sim_Eli", "sim_results_2022-03-19.RData")) #
```

some preparatory variables
```{r, echo = F}
varsout <- c("allsig", "gpbf_ic", "gpbf_iu", "prodbf_ic", "prodbf_iu", "tbf_ic", "tbf_iu")
varspred <- c("es", "errorsd", "n", "k", "hyp_val")
```

## Marginal Descriptives
Obtain marginal descriptives
```{r, echo = F}
descriptives <- function(datf){ 
  descs <- as.matrix(
    datf[, lapply(.SD, function(var){
    medians <- median(var) # obtain overall medians per algorithm over medians per condition
    mads <- mad(var)       # obtain Median absolute deviations per conditions
    sums <- sum(var>3)     # obtain sums where the median is > 3
    maxvals <- max(var)    # obtain the highest median
    c(medians, mads, sums, maxvals)
  }), .SDcols = varsout])
  rownames(descs) <- c("median", "mad", "sum>3", "max")
  descs
}
marginal_descs <- descriptives(dat[ ,lapply(.SD, median), .SDcols = varsout, by = varspred])
datatable(round(marginal_descs,5), options = list(dom = 't'))

```
In theory, the algorithms should have 108/3 = 36 conditions where the median BF > 3. This is because BF > 3 only
in the conditions where true_es > hyp_val, which is one third of the time.
The marginal median for all algorithms is < 1, suggesting that the BF in most conditions is indeed < 3.
The Median Absolute Deviations are also not too severe, giving more proof for reliable estimates.
Interestingly, the prodbf_ic does have a notable higher mad, which is likely due to its outliers.
The finding that the algorithms are rather conservative is further supported by the number of times
the median BF does exceed 3. Both gpbf algorithms count 0 conditions where median BF > 3, while for allsig this
number is 1. For the prodbf_iu and tbf algorithms 5 or 6 times is counted and only the prodbf seems to have a 
more notable number of times the median BF > 3, namely 16.

Now subset for when true BF > 3
```{r, echo = F}
es_gt_hyp_val <- descriptives(dat[es == .2 ,lapply(.SD, median), .SDcols = varsout, by = varspred])
datatable(round(es_gt_hyp_val,5),options = list(dom = "t"))
```
```{r}
all(es_gt_hyp_val['sum>3',] == marginal_descs['sum>3',]) 
```

Even when we subset for when es > hyp_val no median gets above 3. Also, all cases where the median BF > 3
is indeed when es > hyp_val. Again, the MADs are not too severe, except for the prodbf_ic, suggesting the algorithm is less stable when BF > 3.

When true BF < 3
```{r, echo =F}
es_st_hyp_val <- descriptives(dat[es < .2 ,lapply(.SD, median), .SDcols = varsout, by = varspred])
datatable(round(es_st_hyp_val,5), options = list(dom = 't'))
```
Also interesting is that the maximum median BF in the cases where es <= hyp_val all are ~1. 
this provides more support for conservative algorithms. Because the BFs has an interval of 0-$\infty$, it is expected that the mad for the cases where BF < 3 is especially low. But again, the MAD of the prodbf_ic is higher than its median, suggesting less stability even when BF < 3.

## Confusion matrix statistics
```{r, echo = F}
# obtain confusion matrix metrics
lapply_at <- function(var, truees) {
  results <- sapply(var, function(var) {
    table(ordered(truees > .1, levels = c("FALSE", "TRUE")), ordered(var > 3, levels = c("FALSE", "TRUE")))
  })
  names(results) <- vapply(names(var), paste, c("TN", "FN", "FP", "TP"), sep = "_", 
                           FUN.VALUE = character(4),
                           USE.NAMES = FALSE)
  as.list(results)
}

# in long format
#ORIGINALLY HAD datf[!es < .1], but I do not know why.
metrics <- function(datf){
  res <- datf[, sapply(.SD, function(var) {
    table(ordered(es > .1, levels = c("FALSE", "TRUE")), ordered(var > 3, levels = c("FALSE", "TRUE")))
  }), .SDcols = varsout]
  rownames(res) <- c("TN", "FN", "FP", "TP")
  return(res)
}

# obtain metrics per algorithm()
get_stats <- function(res){ 
  apply(res, 2, function(x){
    attach(as.list(x))
    alpha = FP / (FP + TN) #False positive rate 
    beta = FN / (TP + FN) # False negative rate
    sensitivity = 1-beta
    specificity = 1-alpha
    correct = (TN + TP) /(FP + FN + TN + TP)
    pos_lr = sensitivity / (1 - specificity)
    neg_lr = (1 - sensitivity) / specificity
    c(correct = correct,
      False_positive_rate_alpha = alpha,
      False_negative_rate_beta = beta,
      sensitivity = sensitivity,
      specificity = specificity, 
      pos_lr = ifelse(is.na(pos_lr), 0, pos_lr),
      neg_lr = ifelse(is.na(neg_lr), 0, neg_lr))
  })
}

library(ggplot2)
plot_df <-function(stats){ 
  data.frame(Method = colnames(stats),
        correct = stats["correct", ],
        Sensitivity = stats["sensitivity", ],
        Specificity = stats["specificity", ],
        FPR = stats["False_positive_rate_alpha", ],
        FNR = stats["False_negative_rate_beta", ],
        Pos_lr = stats["pos_lr", ],
        Neg_lr = stats["neg_lr", ])
}

#create plot df for subset of data
create_df_plot <- function(datf){
  metrics_long <- metrics(datf)
  stats <- get_stats(metrics_long)
  df_plot <- plot_df(stats)
  return(df_plot)
}
```

Marginal statistics
```{r, echo = F, message=F}
df_plot <- create_df_plot(dat)
marginal_metrics <- rbind(df_plot[,2:ncol(df_plot)], 
      mean = apply(df_plot[,2:ncol(df_plot)], 2 ,mean, na.rm = T))
datatable(round(marginal_metrics, 5), options = list(dom = 't'))
```
Firstly, on average, the algorithms were correct in 72\% of the cases. The prodbf_ic did best with 79\% identified correctly, closely followed by 74\% of the tbf_ic.The gbpf algorithms do overall worst with 67\% being identified correctly. Also interestingly, all _ic algorithms do better than their _iu counterparts.
Secondly, all measures suggest somewhat conservative algorithms, with a mild exception for prodbf_ic. The mean sensitivity of the algorithms is 0.17, meaning that on average, out of all cases where BF > 3, only 17\% is identified correctly. Interestingly, the gpbf algorithms are especially conservative, with the gpbf_iu not having a single case where its BF > 3. The tbf algorithms and prodbf_iu do slightly better in this regard. An obvious outlier is the prodbf_ic algorithm with 44\% correctly identified when BF > 3.
Where the algorithms make up is in the specificity, with a mean specificty of 0.98. This implies that out of all cases where BF < 3, 98\% is identified correctly on average. The prodbf_ic performs only slightly worse than the other algorithms, although the differences are close to minimal.

<!--add part about LR+ and LR- -->

<!-- below is in the make...
```{r, eval = F}
# conditional on n
ndf <- rbind(create_df_plot(dat[n==20]),
           create_df_plot(dat[n==80]),
           create_df_plot(dat[n==200]),
           create_df_plot(dat[n==500]))
ndf$cond <- c(rep("n = 20", 7), rep("n = 80", 7), rep("n = 200", 7), rep("n = 500", 7))

ggplot(ndf, aes(x = Sensitivity, y = Specificity, colour = Method, shape = cond)) +
  geom_point() +
  scale_x_continuous(limits = c(0,1))+
  scale_y_continuous(limits = c(0.75,1)) +
  theme_bw()

# conditional on errorsd
errorsddf <- rbind(create_df_plot(dat[errorsd == 0]),
             create_df_plot(dat[errorsd == 0.5]),
             create_df_plot(dat[errorsd == 0.81]))
errorsddf$cond <- c(rep("errorsd = 0", 7), rep("errorsd = 0.5", 7), rep("errorsd = 0.81", 7))

ggplot(errorsddf, aes(x = Sensitivity, y = Specificity, colour = Method, shape = cond)) +
  geom_point() +
  scale_x_continuous(limits = c(0,1))+
  scale_y_continuous(limits = c(0,1)) +
  theme_bw()

```
-->

### 2022-03-23 analysis.R

@@ -64,7 +64,7 @@ lapply_at <- function(var, truees) {
 
 # in long format
 metrics <- function(datf){
-  res <- datf[!es < .1, sapply(.SD, function(var) {
+  res <- datf[!es < .1,  sapply(.SD, function(var) {
     table(ordered(es > .1, levels = c("FALSE", "TRUE")), ordered(var > 3, levels = c("FALSE", "TRUE")))
   }), .SDcols = varsout]
   rownames(res) <- c("TN", "FN", "FP", "TP")
@@ -123,28 +123,3 @@ ggplot(df_plot, aes(x = Sensitivity, y = Specificity, colour = Method)) +
   scale_y_continuous(limits = c(0,1)) +
   theme_bw()
 
-# conditional on n
-ndf <- rbind(create_df_plot(dat[n==20]),
-           create_df_plot(dat[n==80]),
-           create_df_plot(dat[n==200]),
-           create_df_plot(dat[n==500]))
-ndf$cond <- c(rep("n = 20", 7), rep("n = 80", 7), rep("n = 200", 7), rep("n = 500", 7))
-
-ggplot(ndf, aes(x = Sensitivity, y = Specificity, colour = Method, shape = cond)) +
-  geom_point() +
-  scale_x_continuous(limits = c(0,1))+
-  scale_y_continuous(limits = c(0.75,1)) +
-  theme_bw()
-
-
-# conditional on errorsd
-errorsddf <- rbind(create_df_plot(dat[errorsd == 0]),
-             create_df_plot(dat[errorsd == 0.5]),
-             create_df_plot(dat[errorsd == 0.81]))
-errorsddf$cond <- c(rep("errorsd = 0", 7), rep("errorsd = 0.5", 7), rep("errorsd = 0.81", 7))
-
-ggplot(errorsddf, aes(x = Sensitivity, y = Specificity, colour = Method, shape = cond)) +
-  geom_point() +
-  scale_x_continuous(limits = c(0,1))+
-  scale_y_continuous(limits = c(0,1)) +
-  theme_bw()

### 2022-03-30 Report.Rmd
update report

@@
 
-call library and load in results
+<!--
+call library and load in results. The `dat` `data.table` is subsetted for all cases where $es \geq .1$ . This effectively removes 108/3 = 36 conditions from the results.
+-->
 ```{r, echo = F}
@@
 library(DT)
+library(ggplot2)
 dat <- readRDS(file.path("Sim_Eli", "sim_results_2022-03-19.RData")) #
+dat <- dat[!es < .1]
 ```
 
-some preparatory variables
+<!-- Some preparatory variables which are the algorithms (varsout) and the conditions (varspred) -->
 ```{r, echo = F}
@@
 ## Marginal Descriptives
-Obtain marginal descriptives
+<!--
+Obtain marginal descriptives for all algorithms separately.
+This function first aggregates all conditions by taking the median over all 1000 simulation runs on that condition, reducing the `data.table` to 72000/1000 = 72 medians for the different conditions.
+Finally it describes the distribution of medians of these 72 conditions for every algorithm.
+It takes some quantiles, the mean absolute deviations, the sum of where BF > 3 and the range.
+Later on, decision theory metrics such as Sensitivity and Positive Likelihood Ratio are calculated on the non-aggregated `data.table`.
+-->
 ```{r, echo = F}
@@
     datf[, lapply(.SD, function(var){
-    medians <- median(var) # obtain overall medians per algorithm over medians per condition
+    quants <- quantile(var, c(0.25, 0.50, 0.75, 0.90, 0.95))# obtain overall quantiles
     mads <- mad(var)       # obtain Median absolute deviations per conditions
@@
     maxvals <- max(var)    # obtain the highest median
-    c(medians, mads, sums, maxvals)
+    minvals <- min(var)    # obtain lowest median
+    c(quants[1], quants[2], quants[3], quants[4], quants[5], mads, sums, maxvals, minvals)
   }), .SDcols = varsout])
-  rownames(descs) <- c("median", "mad", "sum>3", "max")
+  rownames(descs) <- c('25 %', '50 %',  '75 %','90 %', '95 %', "mad", "sum>3", "max", "min")
   descs
@@
 datatable(round(marginal_descs,5), options = list(dom = 't'))
-
 ```
-In theory, the algorithms should have 108/3 = 36 conditions where the median BF > 3. This is because BF > 3 only
-in the conditions where true_es > hyp_val, which is one third of the time.
-The marginal median for all algorithms is < 1, suggesting that the BF in most conditions is indeed < 3.
-The Median Absolute Deviations are also not too severe, giving more proof for reliable estimates.
-Interestingly, the prodbf_ic does have a notable higher mad, which is likely due to its outliers.
+Optimally, the algorithms should have $\frac{72}{2} = 36$ conditions where the median BF > 3.
+The marginal median for all algorithms is $\approx 1.0$, however.
+Even at the $90^{th}$ percentile only the prodbf_ic algorithm has its median BF > 3.
+The Median Absolute Deviations are not too severe, providing evidence for reliable estimates.
+Prodbf_ic does have a notably higher mad, which is likely due to its large valued outliers.
 The finding that the algorithms are rather conservative is further supported by the number of times
@@
 more notable number of times the median BF > 3, namely 16. 
+Another observation is that the informative BFs against the unconstrained hypotheses seem to be a bit more conservative than their complement hypotheses counterparts. 
+Finally, it the columns for `prodbf_iu` and `tbf_iu` are exactly equal. It might be that these algorithms do exactly the same thing or there was an error in the simulation.
 
-Now subset for when true BF > 3
+Let us check it out
+```{r, echo=F}
+cor(dat[,..varsout])
+```
+`tbf_iu` and `prodbf_iu` are indeed perfectly correlated.
+Upon further inspection it does not seem to be a simulation error.
+It seems as if `tbf_iu` and `prodbf_iu` do the exact same thing.
+It would mean that the product of the unconstrained Bayes Factors is the same as the unconstrained Bayes Factor under the condition that the hypothesis holds for all groups $k$.
+
+With this in the back of our mind, let us continue.
+Now we subset for when true BF > 3 and describe the distribution for those conditional medians.
 ```{r, echo = F}
@@
 ```
+
 ```{r}
@@
 
-Even when we subset for when es > hyp_val no median gets above 3. Also, all cases where the median BF > 3
-is indeed when es > hyp_val. Again, the MADs are not too severe, except for the prodbf_ic, suggesting the algorithm is less stable when BF > 3.
+Even when the `data.table` is subsetted for when es > hyp_val no median BF gets above 3.
+Also, all cases where the median BF does exceed 3 is indeed only in the conditions when es > hyp_val, suggesting good algorithmic performance when in reality there is no effect.
+Again, the MADs are not too severe, except for the prodbf_ic, suggesting the prod_bf algorithm is less stable when BF > 3 and can calculate large estimates even when the true_es is just 0.1 point higher than the hypothesized value.
+Again, the unconstrained algorithms seem to generally be more conservative than the complementary.
 
@@
 ```{r, echo =F}
-es_st_hyp_val <- descriptives(dat[es < .2 ,lapply(.SD, median), .SDcols = varsout, by = varspred])
+es_st_hyp_val <- descriptives(dat[es == .1 ,lapply(.SD, median), .SDcols = varsout, by = varspred])
 datatable(round(es_st_hyp_val,5), options = list(dom = 't'))
 ```
-Also interesting is that the maximum median BF in the cases where es <= hyp_val all are ~1. 
-this provides more support for conservative algorithms.
-Because the BFs has an interval of 0-$\infty$, it is expected that the mad for the cases where BF < 3 is especially low.
-But again, the MAD of the prodbf_ic is higher than its median, suggesting less stability even when BF < 3.
+Further supporting the idea of conservative algorithms is the finding is that the maximum median BFs in the cases where es = hyp_val all are $\approx 1$ and  all $75^{th}$ percentiles are slightly below one.
+The algorithms seem to favour the hypotheses that are not in the hypothesized direction, although very moderately.
+The previously noticed effect of the unconstrained vs the complement algorithms is not immediately visible.
 
-## Confusion matrix statistics
+Correlation between algorithms, first for the dt with median BFs than for entire dt.
 ```{r, echo = F}
-# obtain confusion matrix metrics
+cormedian <- cor(dat[,lapply(.SD, median), .SDcols = varsout, by = varspred][,..varsout])
+datatable(round(cormedian,5), options = list(dom = 't'))
+```
+A first observation is that `allsig` is only correlated moderately with the `gpbf` algorithms and not at all with the others.
+Note that this correlation table consists of 72 values per algorithm.
+The other general trend is that the `gpbf` algorithms have the lowest correlation with `prodbf_ic` and `allsig` and the highest with each other.
+Finally, the `prodbf` algorithms correlate highly with `tbf` and fairly high with `gpbf`, but less with each other.
+The `tbf` algorithms do correlate highly together.
+
+For the entire dt
+```{r, echo = F}
+corcomplete <- cor(dat[,..varsout])
+datatable(round(corcomplete,5), options = list(dom = 't'))
+```
+We see roughly the same trends.
+The only real difference seems that all correlation with `prodbf_ic` have lowered quite a bit.
+This is likely due to the instable nature of the algorithm estimating very large estimates, while the other algorithms are more reserved.
+
+Inspect the conditions where median prod_bf > 3. Also ordered by k and prodbc_ic values.
+```{r, echo = F}
+prodbf_gt_3 <- dat[,lapply(.SD, median), .SDcols = varsout, by = varspred][prodbf_ic > 3][order(k, prodbf_ic, decreasing = T)]
+datatable(round(prodbf_gt_3,5), options = list(dom = 't'))
+```
+First of all, the conditions where the median BF exceeds three for `prodbf_ic`, are also the conditions where it exceeds three for all other algorithms, except `allsig.` The only condition where the median BF for `allsig` exceeded 3 was when $es = 0.2, errorsd = 0, n = 500, k = 1$, and $hyp\_val = 0.1$.
+It is evident that the higher values for the median BF occur for higher values of $k$, with ten out of sixteen times $k = 10$ and the rest when $k = 3$.
+This makes sense as the `prod_bf` algorithms take the product of the Bayes Factor for all $k$ groups.
+The likelihood of BF_ic and BF_iu being bigger than 1 or a specific group increases as $es > hyp\_val$, resulting in big numbers when all BFs are multiplied.
+The finding that the other algorithms are more conservative than the `prodbf_ic` is likely because they correct for the number of groups.
+For the `gpbf` algorithms by raising the `prodbf` to the power $\frac{1}{k}$.
+The `tbf` algorithms correct for $k$ by calculating the BFs under the condition that the effect in all $k$ groups together is bigger than the hypothesized value, rather than for all $k$ groups separately.
+The `prodbf_iu` algorithm does not necessarily correct for $k$, but it is not restricted to testing only against the complement of the hypothesized direction.
+It rather tests agains all hypotheses, except the informative.
+In other words, how much more likely is the informative hypothesis than all other hypotheses, rather than only the complement of the informative hypothesis.
+
+The other condition that seems to have some importance is $errorsd$, where most algorithms found an effect when $errorsd \neq 0.81$ and preferably when $errorsd = 0$.
+
+## Confusion matrix metrics
+```{r, echo = F}
+# obtain confusion matrix metrics with a dimension of 1 by alg*metrics (so 1x28 in this datatable)
 lapply_at <- function(var, truees) {
@@
 
-# in long format
-#ORIGINALLY HAD datf[!es < .1], but I do not know why.
+# in long format gives a 4x7 matrix with 4 metrics and 7 algorithms
 metrics <- function(datf){
@@
 
-# obtain metrics per algorithm()
+# obtain metrics per algorithm based on FP, TN, FN and TP() and collect in a matrix
 get_stats <- function(res){ 
@@
 
-library(ggplot2)
+# create df from stats matrix for plotting purposes
 plot_df <-function(stats){ 
@@
 
-#create plot df for subset of data
+# all in one function. Create plot df for a subset of the data table.
 create_df_plot <- function(datf){
@@
 marginal_metrics <- rbind(df_plot[,2:ncol(df_plot)], 
-      mean = apply(df_plot[,2:ncol(df_plot)], 2 ,mean, na.rm = T))
+      mean = apply(df_plot[,2:ncol(df_plot)], 2 ,mean))
 datatable(round(marginal_metrics, 5), options = list(dom = 't'))
 ```
-Firstly, on average, the algorithms were correct in 72\% of the cases.
-The prodbf_ic did best with 79\% identified correctly, closely followed by 74\% of the tbf_ic.The gbpf algorithms do overall worst with 67\% being identified correctly.
-Also interestingly, all _ic algorithms do better than their _iu counterparts.
-Secondly, all measures suggest somewhat conservative algorithms, with a mild exception for prodbf_ic.
+Firstly, on average, the algorithms were correct in 58\% of the cases.
+The `prodbf_ic` did best with 68\% identified correctly, followed by 62\% of the tbf_ic.The gbpf algorithms do overall worst with $\approx 50\%$ being identified correctly.
+Also interesting, all _ic algorithms do better than their _iu counterparts.
+Secondly, all measures also suggest somewhat conservative algorithms, with a mild exception for prodbf_ic.
 The mean sensitivity of the algorithms is 0.17, meaning that on average, out of all cases where BF > 3, only 17\% is identified correctly.
-Interestingly, the gpbf algorithms are especially conservative, with the gpbf_iu not having a single case where its BF > 3.
+Interestingly, the gpbf algorithms are especially conservative.
 The tbf algorithms and prodbf_iu do slightly better in this regard.
@@
 This implies that out of all cases where BF < 3, 98\% is identified correctly on average.
-The prodbf_ic performs only slightly worse than the other algorithms, although the differences are close to minimal.
+The prodbf_ic performs only slightly worse than the other algorithms, although the differences are not too severe.
 

### 2022-03-31 Report.Rmd
update report, add plots

@@
 
-## Marginal Descriptives
+### Marginal Descriptives
 <!--
@@
 The marginal median for all algorithms is $\approx 1.0$, however.
-Even at the $90^{th}$ percentile only the prodbf_ic algorithm has its median BF > 3.
-The Median Absolute Deviations are not too severe, providing evidence for reliable estimates.
-Prodbf_ic does have a notably higher mad, which is likely due to its large valued outliers.
+Even at the $90^{th}$ percentile only the `prodbf_ic` algorithm has its median BF > 3.
+The Median Absolute Deviations are not too severe, providing evidence for somewhat reliable estimates.
+`Prodbf_ic` does have a notably higher mad, which is likely due to its large valued outliers.
 The finding that the algorithms are rather conservative is further supported by the number of times
-the median BF does exceed 3. Both gpbf algorithms count 0 conditions where median BF > 3, while for allsig this
-number is 1. For the prodbf_iu and tbf algorithms 5 or 6 times is counted and only the prodbf seems to have a 
+the median BF does exceed 3. Both `gpbf` algorithms count 0 conditions where median BF > 3, while for `allsig` this
+number is 1. For the `prodbf_iu` and `tbf` algorithms 5 or 6 times is counted and only the `prodbf_ic` seems to have a 
 more notable number of times the median BF > 3, namely 16. 
 Another observation is that the informative BFs against the unconstrained hypotheses seem to be a bit more conservative than their complement hypotheses counterparts. 
-Finally, it the columns for `prodbf_iu` and `tbf_iu` are exactly equal. It might be that these algorithms do exactly the same thing or there was an error in the simulation.
+Finally, it the columns for `prodbf_iu` and `tbf_iu` are exactly equal. Lets check that out.
 
 Let us check it out
-```{r, echo=F}
-cor(dat[,..varsout])
+```{r}
+all.equal(dat$tbf_iu, dat$prodbf_iu)
+cor(dat$tbf_iu, dat$prodbf_iu)
 ```
-`tbf_iu` and `prodbf_iu` are indeed perfectly correlated.
+`tbf_iu` and `prodbf_iu` are indeed perfectly correlated and all values in both columns are equal.
 Upon further inspection it does not seem to be a simulation error.
 It seems as if `tbf_iu` and `prodbf_iu` do the exact same thing.
-It would mean that the product of the unconstrained Bayes Factors is the same as the unconstrained Bayes Factor under the condition that the hypothesis holds for all groups $k$.
+It would mean that the product of the unconstrained Bayes Factors is the same as the unconstrained Bayes Factor under the condition that the hypothesis holds for all $k$ groups.
 
-With this in the back of our mind, let us continue.
+#### Descriptives for es > .1
+With this in the back of our mind, let us continue. We can filter the tbf_iu out of the data.
 Now we subset for when true BF > 3 and describe the distribution for those conditional medians.
 ```{r, echo = F}
+dat <- dat[,tbf_iu:=NULL]
+varsout <- c("allsig", "gpbf_ic", "gpbf_iu", "prodbf_ic", "prodbf_iu", "tbf_ic")
+
 es_gt_hyp_val <- descriptives(dat[es == .2 ,lapply(.SD, median), .SDcols = varsout, by = varspred])
@@
 
-When true BF < 3
+#### Descriptives for es < .1
 ```{r, echo =F}
@@
 
-Correlation between algorithms, first for the dt with median BFs than for entire dt.
+#### correlation metrics between algorithms
+Also looking at the correlation between the algorithms, first for the dt with median BFs and afterwards for non-aggregated dt.
 ```{r, echo = F}
@@
 We see roughly the same trends.
-The only real difference seems that all correlation with `prodbf_ic` have lowered quite a bit.
-This is likely due to the instable nature of the algorithm estimating very large estimates, while the other algorithms are more reserved.
+The only real difference seems that all correlations with `prodbf_ic` have lowered quite a bit.
+This is likely due to the algorithm being able to calculate very large estimates, while the other algorithms are more reserved.
 
-Inspect the conditions where median prod_bf > 3. Also ordered by k and prodbc_ic values.
+### inspecting conditions when median prodbf_ic > 3
+Finally, before going to the non-aggregated `data.table`, I inspect the conditions where median prod_bf > 3. That filetered dt is then ordered by $k$ and $prodbc$_ic values.
 ```{r, echo = F}
 prodbf_gt_3 <- dat[,lapply(.SD, median), .SDcols = varsout, by = varspred][prodbf_ic > 3][order(k, prodbf_ic, decreasing = T)]
-datatable(round(prodbf_gt_3,5), options = list(dom = 't'))
+datatable(round(prodbf_gt_3,3), options = list(pageLength = 5))
 ```
@@
 The `tbf` algorithms correct for $k$ by calculating the BFs under the condition that the effect in all $k$ groups together is bigger than the hypothesized value, rather than for all $k$ groups separately.
-The `prodbf_iu` algorithm does not necessarily correct for $k$, but it is not restricted to testing only against the complement of the hypothesized direction.
-It rather tests agains all hypotheses, except the informative.
-In other words, how much more likely is the informative hypothesis than all other hypotheses, rather than only the complement of the informative hypothesis.
+The `prodbf_iu` algorithm does not directly correct for $k$, but it is not restricted to testing only against the complement of the hypothesized direction.
+It rather tests against the unconstrained hypotheses.
+In other words, how much more likely is the informative hypothesis than the hypothesis where there is no equality constrained at all.
 
@@
 
-## Confusion matrix metrics
+### Confusion matrix metrics
+
+<!-- defining functions below which are useful for later when plotting per condition -->
 ```{r, echo = F}
@@
 get_stats <- function(res){ 
+  suppressMessages(
   apply(res, 2, function(x){
@@
   })
+  )
 }
@@
 }
+
+#set colours to use for different algorithms
+colours <- c("black", "orange" , "red", "#79aaf7", "blue", "#00a62f" )
+
+#obtain numeric values conditions take on
+cond_num <- function(cond){return(as.numeric(names(table(dat[,..cond]))))}
+
+#obtain condition names
+cond_names <- function(cond){return(paste0(cond, " = ",names(table(dat[,..cond]))))}
 ```
@@
 ```
+
 Firstly, on average, the algorithms were correct in 58\% of the cases.
 The `prodbf_ic` did best with 68\% identified correctly, followed by 62\% of the tbf_ic.The gbpf algorithms do overall worst with $\approx 50\%$ being identified correctly.
-Also interesting, all _ic algorithms do better than their _iu counterparts.
-Secondly, all measures also suggest somewhat conservative algorithms, with a mild exception for prodbf_ic.
-The mean sensitivity of the algorithms is 0.17, meaning that on average, out of all cases where BF > 3, only 17\% is identified correctly.
-Interestingly, the gpbf algorithms are especially conservative.
-The tbf algorithms and prodbf_iu do slightly better in this regard.
-An obvious outlier is the prodbf_ic algorithm with 44\% correctly identified when BF > 3.
+Also interesting, all `_ic` algorithms do better than their `_iu` counterparts.
+
+Secondly, all measures also suggest somewhat conservative algorithms, with a mild exception for `prodbf_ic.` The mean sensitivity of the algorithms is 0.17, meaning that on average, out of all cases where BF > 3, only 17\% is identified correctly.
+Interestingly, the `gpbf` algorithms are especially conservative with an almost perfect specificity, but no ability to detect an effect.
+The `tbf` algorithms and `prodbf_iu` do slightly better in this regard.
+An obvious outlier is the `prodbf_ic` algorithm with 44\% correctly identified when BF > 3.
+
 Where the algorithms make up is in the specificity, with a mean specificty of 0.98.
@@
 The prodbf_ic performs only slightly worse than the other algorithms, although the differences are not too severe.
+The measure does not say much however in combination with the low sensitivity.
+If all cases are evaluated as having no effect anyway, it is no surprise or feat that the specificity is (close to) perfect.
+For a clearer picture, we look at the LR+ and LR-.
 
-<!--add part about LR+ and LR- -->
+Looking at the LR+, `prodbf_iu` has the best performance with 30.3, followed by `tbf_ic` with 15.3.
+It means that the `prodbf_iu`, out all its accredited effects, does best to separate it from actual no effects.
+The finding that this value is lower for `prodbf_ic` means that it has a higher likelihood of finding an effect when there is none.
 
-<!-- below is in the make...
-```{r, eval = F}
-# conditional on n
-ndf <- rbind(create_df_plot(dat[n==20]),
-           create_df_plot(dat[n==80]),
-           create_df_plot(dat[n==200]),
-           create_df_plot(dat[n==500]))
-ndf$cond <- c(rep("n = 20", 7), rep("n = 80", 7), rep("n = 200", 7), rep("n = 500", 7))
+The LR- corrects for the the finding that some algorithms' conservatism.
+For example, `gpbf_iu` has a value of 1, which means that out of all the tests it determined that there was no effect, 100\% of the times it was wrong when in reality there was an effect.
+`prodbf_ic` actually does best in this regard.
+Out of all the tests it determined that there was no effect, 60\% it was wrong when there in reality was an effect.
 
-ggplot(ndf, aes(x = Sensitivity, y = Specificity, colour = Method, shape = cond)) +
+A first conclusion would be that the `prodbf_ic` is the algorithm of choice.
+It is the only algorithm that has a somewhat higher ability to detect an effect when there is one, while still performing well when there is no effect.
+It does have a somewhat higher probability of finding an effect when there in reality is none which could be a valid concern.
+When the chance of a Type I error needs to be reduced as much as possible, it would be wiser to choose `prodbf_iu` or `tbf_ic` which have less ability to detect an effect, but are also less likely to call an effect when there in reality is none.
+
+### Plots per condition
+
+First for $errorsd$
+
+```{r, echo = F}
+sd_df <- do.call(rbind, lapply(cond_num('errorsd'), function(x){create_df_plot(dat[errorsd == x])})) 
+sd_df$condition <- rep(cond_names('errorsd'), each = length(varsout))
+
+ggplot(sd_df, aes(x = Sensitivity, y = Specificity, colour = Method, shape = condition)) +
   geom_point() +
   scale_x_continuous(limits = c(0,1))+
-  scale_y_continuous(limits = c(0.75,1)) +
+  scale_y_continuous(limits = c(0.85,1)) +
+  scale_color_manual(values = colours) +
   theme_bw()
+```
 
-# conditional on errorsd
-errorsddf <- rbind(create_df_plot(dat[errorsd == 0]),
-             create_df_plot(dat[errorsd == 0.5]),
-             create_df_plot(dat[errorsd == 0.81]))
-errorsddf$cond <- c(rep("errorsd = 0", 7), rep("errorsd = 0.5", 7), rep("errorsd = 0.81", 7))
+Then for $n$
 
-ggplot(errorsddf, aes(x = Sensitivity, y = Specificity, colour = Method, shape = cond)) +
+```{r, echo = F}
+n_df <- do.call(rbind, lapply(cond_num('n'), function(x){create_df_plot(dat[n == x])})) 
+n_df$condition <- rep(cond_names('n'), each = length(varsout))
+
+ggplot(n_df, aes(x = Sensitivity, y = Specificity, colour = Method, shape = condition)) +
   geom_point() +
   scale_x_continuous(limits = c(0,1))+
-  scale_y_continuous(limits = c(0,1)) +
+  scale_y_continuous(limits = c(0.80,1)) +
+  scale_color_manual(values = colours) +
   theme_bw()
+```
 
+Finally for $k$
+
+```{r, echo = F}
+k_df <- do.call(rbind, lapply(cond_num('k'), function(x){create_df_plot(dat[k == x])})) 
+k_df$condition <- rep(cond_names('k'), each = length(varsout))
+
+ggplot(k_df, aes(x = Sensitivity, y = Specificity, colour = Method, shape = condition)) +
+  geom_point() +
+  scale_x_continuous(limits = c(0,1))+
+  scale_y_continuous(limits = c(0.85,1)) +
+  scale_color_manual(values = colours) +
+  theme_bw()
 ```
--->

### 2022-04-07 Analysis/Report.Rmd
add interactions, finalize conclusions
The results and conclusions are as good as done, but the script needs to be optimized a bit.

--- 2022-03-31 Report.Rmd
@@
 <!--
-call library and load in results. The `dat` `data.table` is subsetted for all cases where $es \geq .1$ . This effectively removes 108/3 = 36 conditions from the results.
+call library and load in results. The `dat` `data.table` is subsetted for all cases where $es >= .1$ . This removes 108/3 = 36 conditions from the results.
 -->
@@
 It takes some quantiles, the mean absolute deviations, the sum of where BF > 3 and the range.
-Later on, decision theory metrics such as Sensitivity and Positive Likelihood Ratio are calculated on the non-aggregated `data.table`.
+Later on, confusion matrix metrics such as Sensitivity and Positive Likelihood Ratio are calculated on the non-aggregated `data.table`.
 -->
+
+*Table 1.* The marginal median Bayes Factors for all algorithms.
 ```{r, echo = F}
@@
 ```
+*Every condition has been simulation 1000 times. For every condition and algoritm, the Bayes Factors are aggregated by takin the median, leaving 72 median BFs per algorithm*
+
 Optimally, the algorithms should have $\frac{72}{2} = 36$ conditions where the median BF > 3.
@@
 `Prodbf_ic` does have a notably higher mad, which is likely due to its large valued outliers.
-The finding that the algorithms are rather conservative is further supported by the number of times
-the median BF does exceed 3. Both `gpbf` algorithms count 0 conditions where median BF > 3, while for `allsig` this
-number is 1. For the `prodbf_iu` and `tbf` algorithms 5 or 6 times is counted and only the `prodbf_ic` seems to have a 
-more notable number of times the median BF > 3, namely 16. 
+Because the BF is bound between 0 and $\infty$, outliers are plausible and can be influential on the MAD when, in this case, only 72 medians construct the distribution of conditional medians.
+The finding is that the algorithms are rather conservative, which is further supported by the number of times the median BF does exceed 3.
+Both `gpbf` algorithms count 0 conditions where median BF > 3, while for `allsig` this number is 1.
+For the `prodbf_iu` and `tbf` algorithms 5 or 6 times is counted and only the `prodbf_ic` seems to have a more notable number of times the median BF > 3, namely 16.
 Another observation is that the informative BFs against the unconstrained hypotheses seem to be a bit more conservative than their complement hypotheses counterparts.
-Finally, it the columns for `prodbf_iu` and `tbf_iu` are exactly equal. Lets check that out.
+Finally, it seems as if the columns for `prodbf_iu` and `tbf_iu` are exactly equal.
+Lets check that out.
 
-Let us check it out
 ```{r}
@@
 ```
-`tbf_iu` and `prodbf_iu` are indeed perfectly correlated and all values in both columns are equal.
+`tbf_iu` and `prodbf_iu` are indeed perfectly correlated and all values in both columns in the complete `data.table` are equal.
 Upon further inspection it does not seem to be a simulation error.
@@
 #### Descriptives for es > .1
-With this in the back of our mind, let us continue. We can filter the tbf_iu out of the data.
+
+With this in the back of our mind, let us continue. We can filter the tbf_iu and gpbf_iu out of the data.
 Now we subset for when true BF > 3 and describe the distribution for those conditional medians.
-```{r, echo = F}
+```{r}
 dat <- dat[,tbf_iu:=NULL]
 varsout <- c("allsig", "gpbf_ic", "gpbf_iu", "prodbf_ic", "prodbf_iu", "tbf_ic")
+```
 
+*Table 2.* The median Bayes Factors for all algorithms in the condition that es > .1.
+```{r, echo = F}
 es_gt_hyp_val <- descriptives(dat[es == .2 ,lapply(.SD, median), .SDcols = varsout, by = varspred])
@@
 ```{r}
-all(es_gt_hyp_val['sum>3',] == marginal_descs['sum>3',]) 
+all(es_gt_hyp_val['sum>3',varsout] == marginal_descs['sum>3',varsout]) 
 ```
@@
 #### Descriptives for es < .1
+
+*Table 2.* The median Bayes Factors for all algorithms in the condition that es < .1.
 ```{r, echo =F}
@@
 
-#### correlation metrics between algorithms
-Also looking at the correlation between the algorithms, first for the dt with median BFs and afterwards for non-aggregated dt.
-```{r, echo = F}
-cormedian <- cor(dat[,lapply(.SD, median), .SDcols = varsout, by = varspred][,..varsout])
-datatable(round(cormedian,5), options = list(dom = 't'))
-```
-A first observation is that `allsig` is only correlated moderately with the `gpbf` algorithms and not at all with the others.
-Note that this correlation table consists of 72 values per algorithm.
-The other general trend is that the `gpbf` algorithms have the lowest correlation with `prodbf_ic` and `allsig` and the highest with each other.
-Finally, the `prodbf` algorithms correlate highly with `tbf` and fairly high with `gpbf`, but less with each other.
-The `tbf` algorithms do correlate highly together.
-
-For the entire dt
-```{r, echo = F}
-corcomplete <- cor(dat[,..varsout])
-datatable(round(corcomplete,5), options = list(dom = 't'))
-```
-We see roughly the same trends.
-The only real difference seems that all correlations with `prodbf_ic` have lowered quite a bit.
-This is likely due to the algorithm being able to calculate very large estimates, while the other algorithms are more reserved.
-
 ### inspecting conditions when median prodbf_ic > 3
 Finally, before going to the non-aggregated `data.table`, I inspect the conditions where median prod_bf > 3. That filetered dt is then ordered by $k$ and $prodbc$_ic values.
+
+*Table 2.* The 16 conditions when the median BF for prodbf_ic > 3.
 ```{r, echo = F}
@@
 
-#obtain numeric values conditions take on
+# obtain numeric values each conditions take on
 cond_num <- function(cond){return(as.numeric(names(table(dat[,..cond]))))}
 
-#obtain condition names
+#obtain condition names. For k it returns c("k = 1", "k = 3", "k = 10")
 cond_names <- function(cond){return(paste0(cond, " = ",names(table(dat[,..cond]))))}
@@
 
-Marginal statistics
+*Table 4.* The marginal confusion matrix metrics.
 ```{r, echo = F, message=F}
@@
 It means that the `prodbf_iu`, out all its accredited effects, does best to separate it from actual no effects.
-The finding that this value is lower for `prodbf_ic` means that it has a higher likelihood of finding an effect when there is none.
+The finding that this value is lower for `prodbf_ic` means that `prodbf_ic` has a higher likelihood of finding an effect when there is none.
 
-The LR- corrects for the the finding that some algorithms' conservatism.
+The LR- corrects for algorithms' conservatism.
 For example, `gpbf_iu` has a value of 1, which means that out of all the tests it determined that there was no effect, 100\% of the times it was wrong when in reality there was an effect.
@@
 
-First for $errorsd$
+since gpbf_iu has not found an effect in any simulation run, it is not informative to include it in a plot
+```{r}
+all(dat$gpbf_iu < 3)
+```
 
+And thus it can be omitted during plotting of the confusion matrix metrics, because it is not informative
+```{r}
+dat <- dat[,gpbf_iu:=NULL]
+varsout <- c("allsig", "gpbf_ic", "prodbf_ic", "prodbf_iu", "tbf_ic")
+```
+
+First the relationship between sensitivity and specificity is plot for $errorsd$, then for $n$ and finally for $k$
+It was expected that there might have been an interaction between $errorsd$ and $k$ and so the final plot will show this interaction
+
+#### errorsd
+
+*Figure 1.* The effect of *errorsd* on the sensitivty and specificity of the algorithms.
 ```{r, echo = F}
@@
 ggplot(sd_df, aes(x = Sensitivity, y = Specificity, colour = Method, shape = condition)) +
-  geom_point() +
+  geom_point(size = 2.5) +
   scale_x_continuous(limits = c(0,1))+
@@
 
-Then for $n$
+There seems to be a moderate effect of $errorsd$ for `prodbf_ic`.
+It seems that the higher the reliability, the more power, but the more chance for Type I errors.
+Do note the y-axis representing specificity, however as it starts at 0.85.
+In the condition that $errorsd = 0$ it sensitivity is above 50\%, showing that with no reliability issues, it still has a great ablity to detect no effect (specificity = 0.88), and even an ability to detect an effect that is higher than chance (Sensitivity = 0.62).
+Do be aware that this means a 12\% chance of a Type I error and a 48\% of a Type II error.
 
+Conventionally, the probabilty of a Type I error is set at 5\%.
+In this is desired the `tbf_ic` is the superior algorithm.
+When $errorsd = 0$ its specificity still 0.98, while its sensitivity is 0.47.
+Still, a conventional desire for power is 80\% and no algorithm comes close to this number.
+The `gpbf_ic` and `allsig` algorithms do not seem to be affected much by the variation in $errrosd$.
+
+The relationship does make sense as the `tbf` algorithm seems to be more dependent on group consensus to find an effect.
+There is a high probability of consensus in the condition that $errorsd = 0$.
+For `prodbf_ic`, its BF more easily increases drastically if only a few groups find large effects, as the BF is bound between 0 and $\infty$.
+If the BF of only one group is, let's say, 2000, than multiplying with the other Bfs will likely still lead to a high `prodBF_ic`.
+
+It seems all in all for $n$ as if the performance difference between algorithms is larger than the performance difference between conditions.
+
+#### n
+
+*Figure 2.* The effect of *n* on the sensitivty and specificity of the algorithms.
 ```{r, echo = F}
@@
 ggplot(n_df, aes(x = Sensitivity, y = Specificity, colour = Method, shape = condition)) +
-  geom_point() +
+  geom_point(size = 2.5) +
   scale_x_continuous(limits = c(0,1))+
@@
 
-Finally for $k$
+There seems to be only a slight improvement of the algorithms as $n$ gets larger.
+Again, the y-axis starts at 0.80.
+For all algorithms except `prodbf_ic`, n has barely any influence.
+This means that the performance of these algorithms are not much dependent on the sample sizes within the dataframes.
+For `prodbc_ic` the effect of $n$ is more clearly visible with quite some difference in specificity when $n$ drops from 80 to 20.
+Sensitivity does seem to slightly improve as $n$ gets larger for all algorithms except `gpbf_ic`. When $n = 500$, there is almost no loss in specificity, while this loss gets larger as $n$ decreases.
 
+#### k
+*Figure 3.* The effect of *k* on the sensitivity and specificity of the algorithms.
 ```{r, echo = F}
@@
 ggplot(k_df, aes(x = Sensitivity, y = Specificity, colour = Method, shape = condition)) +
-  geom_point() +
+  geom_point(size = 2.5) +
   scale_x_continuous(limits = c(0,1))+
@@
 ```
+
+the general effect of $k$ seems to be that the higher $k$, the more balance between specificity and sensitivity.
+Especially `prodbf_ic` at $k = 10$ shows promising results with Specificity = 0.87 and Sensitivity = 0.78.
+Also the `tbf_ic` does well at $k = 10$, with a notably lower Sensitivity (0.52) than `prodbf_ic`, but a notably higher Specificity (0.98).
+
+### interactions
+<!-- this is not optimal and takes long time to run, thats why the outputs are saved in .RData files. -->
+```{r, eval = F, echo = F}
+kn_df <- do.call(rbind, lapply(cond_num('k'), function(x){
+  do.call(rbind, lapply(cond_num('n'), function(y){
+    temp <- create_df_plot(dat[k == x & n == y])
+    temp$condition1 <- paste0('k = ', x)
+    temp$condition2 <- paste0('n = ', y)
+    temp
+    }))
+  }))
+
+ke_df <- do.call(rbind, lapply(cond_num('k'), function(x){
+  do.call(rbind, lapply(cond_num('errorsd'), function(y){
+    temp <- create_df_plot(dat[k == x & errorsd == y])
+    temp$condition1 <- paste0('k = ', x)
+    temp$condition2 <- paste0('errorsd = ', y)
+    temp
+    }))
+  }))
+
+en_df <- do.call(rbind, lapply(cond_num('errorsd'), function(x){
+  do.call(rbind, lapply(cond_num('n'), function(y){
+    temp <- create_df_plot(dat[errorsd == x & n == y])
+    temp$condition1 <- paste0('errorsd = ', x)
+    temp$condition2 <- paste0('n = ', y)
+    temp
+    }))
+  }))
+
+saveRDS(kn_df, file = "./Analysis/plot_kn.RData")
+saveRDS(ke_df, file = "./Analysis/plot_ke.RData")
+saveRDS(en_df, file = "./Analysis/plot_en.RData")
+
+```
+
+<!-- read in interaction dfs to plot -->
+```{r, echo = F}
+kn_df <- readRDS("../Analysis/plot_kn.RData")
+ke_df <- readRDS("../Analysis/plot_ke.RData")
+en_df <- readRDS("../Analysis/plot_en.RData")
+
+```
+
+*Figure 4.* The effect of the interaction between *k* and *n* on the sensitivity and specificity of the algorithms.
+```{r, echo = F}
+ggplot(kn_df, aes(x = Sensitivity, y = Specificity, colour = Method, shape = condition1)) +
+  geom_point(size = 2.5) +
+  scale_x_continuous(limits = c(0,1))+
+  scale_y_continuous(limits = c(0.70,1)) +
+  scale_color_manual(values = colours) +
+  facet_wrap(~condition2) +
+  theme_bw()
+
+```
+
+The general trends of $k$ and $n$ seem to hold that the higher both $n$ and $k$, the better the performance.
+When $n = 80$ it becomes more clear that specificity is sacrificied for better sensitivity, but mostly only for `prodbf_ic`.
+Especially in the case where $n = 500$ and $k = 10$, `prodbf_ic` performs very well with Sensitivity = 0.88 and Specificity = 0.95.
+This suggests a Type I error of 5\% and a power of 88\%. But also `tbf_ic` stands out as both $k$ and $n$ get larger.
+
+*Figure 5.* The effect of the interaction between *k* and *errorsd* on the sensitivity and specificity of the algorithms.
+```{r, echo = F}
+ggplot(ke_df, aes(x = Sensitivity, y = Specificity, colour = Method, shape = condition1)) +
+  geom_point(size = 2.5) +
+  scale_x_continuous(limits = c(0,1))+
+  scale_y_continuous(limits = c(0.75,1)) +
+  scale_color_manual(values = colours) +
+  facet_wrap(~condition2) +
+  theme_bw()
+
+```
+
+It seems the the marginal effect of $errorsd$ gets larger for higher values of $k$.
+Meaning a higher rise in power, but drop in specificity for larger values of $k$.
+This is mainly seen for `prodbf_ic` and a bit for `tbf_ic`.
+Also interesting to note is that when $errorsd = 0$ and $k = 10$ the only known condition is where an algorithm has a higher Sensitivity than specificity.
+`prodbf_ic` scores 0.96 on sensitivty and 0.77 on specificity in this condition.
+
+*Figure 6.* The effect of the interaction between *n* and *errrorsd* on the sensitivity and specificity of the algorithms.
+```{r, echo = F}
+ggplot(en_df, aes(x = Sensitivity, y = Specificity, colour = Method, shape = condition1)) +
+  geom_point(size = 2.5) +
+  scale_x_continuous(limits = c(0,1))+
+  scale_y_continuous(limits = c(0.6,1)) +
+  scale_color_manual(values = colours) +
+  facet_wrap(~condition2) +
+  theme_bw()
+
+```
+
+It seems that in general, the effect of $errorsd$ is barely noticeable for different values of $n$. It only becomes for noticeable as $n < 80$ and only for `prodbf_ic`.
+
+# Conclusions
+All in all it can be said the the algorithms are conservative.
+In almost all conditions for all algorithms, the specificity was higher than the sensitivity.
+In fact the only condition and algorithm when this was reversed was for `prodbf_ic` when $k = 10$ and $errorsd = 0$.
+This conservatism can be interpreted as good, because the probability of making a Type I error is low.
+However, the finding that the sensitivity was rather low implies that there is almost no power to detect an effect.
+The current study only simulated conditions when the true effect size was .1 higher than the hypothesized value.
+Although, this is closer to reality as effect sizes are rarely large, its still interesting if this conservatism is retained by the algorithms as the differences between es and hyp_val gets larger.
+
+There were some exceptions that seemed promising.
+These only apply to the `prodbf_ic` and `tbf_ic` algorithms, however.
+These conditions were mainly if both $k$ and $n$ are large.
+Real life examples usually do not have data of 10 groups, each containing 500 observations, although within the Baysian Framework this is more plausible as more research is done in the field.
+
+The variation in conditions mainly affected `prodbf_ic`.
+The initial guess is that `prodbf_ic` can be unstable as it takes the product of the BFs for all $k$ groups.
+Suppose $k = 10$ and for all is found that $BF_{ic} = 2$, the `prodbf_ic` will output $2^{10} = 1024$, although no group on its own is convincingly in favour of any hypotheses.
+The `tbf` is more stable as it evaluates if all BFs together exceed 3.
+The `gpbf` is the `prodbf`$^{\frac{1}{k}}$ which for our particular example makes $2^{10*\frac{1}{10}} = 2^1 = 2$.
+This is the main reason that as $errorsd = 0$, the `gpbf_ic` gets more extreme as the probability of consensus between groups becomes larger.
+
+The reason that the `iu` algorithms are more conservative is because they are not constrained by testing against only the complement of the specified hypothesis.
+
+The overall conclusion is that the algorithms lack power to detect an effect and can not realistically be used.
+Only if much data is available on multiple groups do the `prodbf_ic` and `tbf_ic` become useful, Where the `prodbf_ic` sacrificies a bit more specificity than `tbf_ic` for more sensitivity.
+
+```{r, echo = F, eval = F}
+rm(list = ls())
+gc(full = T, reset = T)
+cat("\014")
+```

### 2022-04-07 Analysis/Report.R

@@ -1,0 +1,268 @@
+## ---- echo = F-------------------------------------------------------------------------------------------------------------------------------------
+library(data.table) # load in libraries
+library(DT)
+library(ggplot2)
+dat <- readRDS(file.path("Sim_Eli", "sim_results_2022-03-19.RData")) #
+dat <- dat[!es < .1]
+
+
+## ---- echo = F-------------------------------------------------------------------------------------------------------------------------------------
+varsout <- c("allsig", "gpbf_ic", "gpbf_iu", "prodbf_ic", "prodbf_iu", "tbf_ic", "tbf_iu")
+varspred <- c("es", "errorsd", "n", "k", "hyp_val")
+
+
+## ---- echo = F-------------------------------------------------------------------------------------------------------------------------------------
+descriptives <- function(datf){ 
+  descs <- as.matrix(
+    datf[, lapply(.SD, function(var){
+    quants <- quantile(var, c(0.25, 0.50, 0.75, 0.90, 0.95))# obtain overall quantiles
+    mads <- mad(var)       # obtain Median absolute deviations per conditions
+    sums <- sum(var>3)     # obtain sums where the median is > 3
+    maxvals <- max(var)    # obtain the highest median
+    minvals <- min(var)    # obtain lowest median
+    c(quants[1], quants[2], quants[3], quants[4], quants[5], mads, sums, maxvals, minvals)
+  }), .SDcols = varsout])
+  rownames(descs) <- c('25 %', '50 %',  '75 %','90 %', '95 %', "mad", "sum>3", "max", "min")
+  descs
+}
+marginal_descs <- descriptives(dat[ ,lapply(.SD, median), .SDcols = varsout, by = varspred])
+datatable(round(marginal_descs,5), options = list(dom = 't'))
+
+
+## --------------------------------------------------------------------------------------------------------------------------------------------------
+all.equal(dat$tbf_iu, dat$prodbf_iu)
+cor(dat$tbf_iu, dat$prodbf_iu)
+
+
+## --------------------------------------------------------------------------------------------------------------------------------------------------
+dat <- dat[,tbf_iu:=NULL]
+varsout <- c("allsig", "gpbf_ic", "gpbf_iu", "prodbf_ic", "prodbf_iu", "tbf_ic")
+
+
[…]

### 2022-04-07 Analysis/analysis.R

@@ -1,0 +1,48 @@
+#dat <- readRDS("tabres.RData")
+dat <- readRDS(file.path("Sim_Eli", "sim_results_2022-03-19.RData"))
+lapply_at <- function(var, truees) {
+  results <- sapply(var, function(var) {
+    table(ordered(truees > .1, levels = c("FALSE", "TRUE")), ordered(var > 3, levels = c("FALSE", "TRUE")))
+  })
+  names(results) <- vapply(names(var), paste, c("TN", "FN", "FP", "TP"), sep = "_", 
+                           FUN.VALUE = character(4),
+                           USE.NAMES = FALSE)
+  as.list(results)
+}
+varsout <- c("allsig", "gpbf_ic", "gpbf_iu", "prodbf_ic", "prodbf_iu", "tbf_ic", "tbf_iu")
+varspred <- c("es", "errorsd", "n", "k", "hyp_val")
+
+# Example: per condition
+dat[!es < .1, lapply_at(.SD, truees = es), .SDcols = varsout, by = c("n", "errorsd")]
+
+# Overall
+#res <- dat[!es < .1, lapply_at(.SD, truees = es), .SDcols = varsout]
+res <- dat[!es < .1, sapply(.SD, function(var) {
+  table(ordered(es > .1, levels = c("FALSE", "TRUE")), ordered(var > 3, levels = c("FALSE", "TRUE")))
+}), .SDcols = varsout]
+rownames(res) <- c("TN", "FN", "FP", "TP")
+
+stats <- apply(res, 2, function(x){
+  attach(as.list(x))
+  alpha = FP / (FP + TN) #False positive rate 
+  beta = FN / (TP + FN) # False negative rate
+  sensitivity = 1-beta
+  specificity = 1-alpha
+  c(False_positive_rate_alpha = alpha,
+    False_negative_rate_beta = beta,
+    sensitivity = sensitivity,
+    specificity = specificity, 
+    pos_lr = sensitivity / (1 - specificity),
+    neg_lr = (1 - sensitivity) / specificity)
+})
+
+library(ggplot2)
+df_plot <- data.frame(Method = colnames(stats),
+                      Sensitivity = stats["sensitivity", ],
+                      Specificity = stats["specificity", ])
+
+ggplot(df_plot, aes(x = Sensitivity, y = Specificity, colour = Method, shape = Method)) +
+  geom_point() +
+  scale_x_continuous(limits = c(0,1))+
+  scale_y_continuous(limits = c(0,1)) +
+  theme_bw()

### 2022-04-07 analysis.R

@@ -1,125 +1,0 @@
-#dat <- readRDS("tabres.RData")
-
-library(data.table) # load in library
-dat <- readRDS(file.path("Sim_Eli", "sim_results_2022-03-19.RData")) # read in results
-
-# some preparation
-varsout <- c("allsig", "gpbf_ic", "gpbf_iu", "prodbf_ic", "prodbf_iu", "tbf_ic", "tbf_iu")
-varspred <- c("es", "errorsd", "n", "k", "hyp_val")
-
-#########################
-# MARGINAL DESCRIPTIVES #
-#########################
-
-# First we obtain some descriptives over the marginal conditions.
-descriptives <- function(datf){ 
-  descs <- as.matrix(
-    datf[, lapply(.SD, function(var){
-    medians <- median(var) # obtain overall medians per algorithm over medians per condition
-    mads <- mad(var)       # obtain Median absolute deviations per conditions
-    sums <- sum(var>3)     # obtain sums where the median is > 3
-    maxvals <- max(var)    # obtain the highest median
-    c(medians, mads, sums, maxvals)
-  }), .SDcols = varsout])
-  rownames(descs) <- c("median", "mad", "sum>3", "max")
-  descs
-}
-
-(marginal_descs <- descriptives(dat[ ,lapply(.SD, median), .SDcols = varsout, by = varspred]))
-# In theory, the algorithms should have 108/3 = 36 conditions where the median BF > 3. This is because BF > 3 only
-# in the conditions where true_es > hyp_val, which is one third of the time.
-# The marginal median for all algorithms is < 1, suggesting that the BF in most conditions is indeed < 3.
-# The Median Absolute Deviations are also not too severe, giving more proof for reliable estimates.
-# Interestingly, the prodbf_ic does have a notable higher mad.
-# The finding that the algorithms are rather conservative is further supported by the number of times
-# the median BF does exceed 3. Both gpbf algorithms count 0 conditions where median BF > 3, while for allsig this
-# number is 1. For the prodbf_iu and tbf algorithms 5 or 6 times is counted and only the prodbf seems to have a 
-# more notable number of times the median BF > 3, namely 16.
-
-(es_gt_hyp_val <- descriptives(dat[es == .2 ,lapply(.SD, median), .SDcols = varsout, by = varspred])) # when true BF > 3
[…]

### 2022-04-07 Analysis/Report.Rmd
minor commit

@@
 author: "Eli Clapper"
-date: "23/03/2022"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
 output: html_document
@@
 -->
+
 ```{r, echo = F}
@@
 library(ggplot2)
-dat <- readRDS(file.path("Sim_Eli", "sim_results_2022-03-19.RData")) #
+dat <- readRDS(file.path("../Sim_Eli", "sim_results_2022-03-19.RData")) #
 dat <- dat[!es < .1]

### 2022-04-13 Analysis/Report.Rmd
Finalized Report V1
The report is done. The code can be optimized, but works. The three .RData files in the analysis folders are data.frames that can be used for plotting interactions. It took a while to run the code to generate those, so they were saved.

@@
 Finally it describes the distribution of medians of these 72 conditions for every algorithm.
-It takes some quantiles, the mean absolute deviations, the sum of where BF > 3 and the range.
+It takes some quantiles, the median absolute deviations from the overall median of that algorithm, the sum of where BF > 3 and the range.
 Later on, confusion matrix metrics such as Sensitivity and Positive Likelihood Ratio are calculated on the non-aggregated `data.table`.
@@
 ```
-*Every condition has been simulation 1000 times. For every condition and algoritm, the Bayes Factors are aggregated by takin the median, leaving 72 median BFs per algorithm*
+*Every condition has been simulated 1000 times. For every condition and algorithm, the Bayes Factors are aggregated by taking the median over all Bayes Factors, leaving 72 median BF's per algorithm*
 
-Optimally, the algorithms should have $\frac{72}{2} = 36$ conditions where the median BF > 3.
+Optimally, the algorithms should have $\frac{72}{2} = 36$ conditions where the median BF > 3 in the cases where $es$ > $hyp_val$.
 The marginal median for all algorithms is $\approx 1.0$, however.
@@
 Because the BF is bound between 0 and $\infty$, outliers are plausible and can be influential on the MAD when, in this case, only 72 medians construct the distribution of conditional medians.
+
 The finding is that the algorithms are rather conservative, which is further supported by the number of times the median BF does exceed 3.
@@
 ```
+
 `tbf_iu` and `prodbf_iu` are indeed perfectly correlated and all values in both columns in the complete `data.table` are equal.
@@
 
-With this in the back of our mind, let us continue. We can filter the tbf_iu and gpbf_iu out of the data.
+With this in mind, let us continue. We can filter the tbf_iu out of the data.
 Now we subset for when true BF > 3 and describe the distribution for those conditional medians.
@@
 Even when the `data.table` is subsetted for when es > hyp_val no median BF gets above 3.
-Also, all cases where the median BF does exceed 3 is indeed only in the conditions when es > hyp_val, suggesting good algorithmic performance when in reality there is no effect.
+Also, all cases where the median BF does exceed 3 is indeed only in the conditions when es > hyp_val, suggesting the algorithms do not accredit effects when in reality there is indeed no effect.
 Again, the MADs are not too severe, except for the prodbf_ic, suggesting the prod_bf algorithm is less stable when BF > 3 and can calculate large estimates even when the true_es is just 0.1 point higher than the hypothesized value.
@@
 ```
-Further supporting the idea of conservative algorithms is the finding is that the maximum median BFs in the cases where es = hyp_val all are $\approx 1$ and  all $75^{th}$ percentiles are slightly below one.
+Further supporting the idea of conservative algorithms is the finding is that the maximum median BFs in the cases where $es = hyp\_val$ all are $\approx 1$ and  all $75^{th}$ percentiles are slightly below one.
 The algorithms seem to favour the hypotheses that are not in the hypothesized direction, although very moderately.
-The previously noticed effect of the unconstrained vs the complement algorithms is not immediately visible.
+The previously noticed effect of the unconstrained vs.
+the complement algorithms is not immediately visible.
 
 ### inspecting conditions when median prodbf_ic > 3
-Finally, before going to the non-aggregated `data.table`, I inspect the conditions where median prod_bf > 3. That filetered dt is then ordered by $k$ and $prodbc$_ic values.
+Finally, before going to the non-aggregated `data.table`, the conditions are inspected where median `prodbf_ic` > 3. The dt is then ordered by $k$ and `prodbf_ic` values.
 
@@
 The other condition that seems to have some importance is $errorsd$, where most algorithms found an effect when $errorsd \neq 0.81$ and preferably when $errorsd = 0$.
+Further analysis will investigate this idea.
 
@@
 
-Secondly, all measures also suggest somewhat conservative algorithms, with a mild exception for `prodbf_ic.` The mean sensitivity of the algorithms is 0.17, meaning that on average, out of all cases where BF > 3, only 17\% is identified correctly.
-Interestingly, the `gpbf` algorithms are especially conservative with an almost perfect specificity, but no ability to detect an effect.
+Secondly, all measures also suggest conservative algorithms, with a mild exception for `prodbf_ic.` The mean Sensitivity of the algorithms is 0.17, meaning that on average, out of all cases where BF > 3, only 17\% is identified correctly.
+Interestingly, the `gpbf` algorithms are especially conservative with an almost perfect Specificity, but no ability to detect an effect.
 The `tbf` algorithms and `prodbf_iu` do slightly better in this regard.
@@
 
-Where the algorithms make up is in the specificity, with a mean specificty of 0.98.
+Where the algorithms make up is in the Specificity, with a mean specificty of 0.98.
 This implies that out of all cases where BF < 3, 98\% is identified correctly on average.
-The prodbf_ic performs only slightly worse than the other algorithms, although the differences are not too severe.
-The measure does not say much however in combination with the low sensitivity.
-If all cases are evaluated as having no effect anyway, it is no surprise or feat that the specificity is (close to) perfect.
+The `prodbf_ic` performs only slightly worse than the other algorithms, although the differences are not too severe.
+The measure does not say much however in combination with the low Sensitivity.
+If all cases are evaluated as having no effect anyway, it is no surprise or feat that the Specificity is (close to) perfect.
 For a clearer picture, we look at the LR+ and LR-.
@@
 Looking at the LR+, `prodbf_iu` has the best performance with 30.3, followed by `tbf_ic` with 15.3.
-It means that the `prodbf_iu`, out all its accredited effects, does best to separate it from actual no effects.
+It means that the `prodbf_iu`, out off all cases it accredited an effect, it does best to separate it from actual no effects.
 The finding that this value is lower for `prodbf_ic` means that `prodbf_ic` has a higher likelihood of finding an effect when there is none.
@@
 ### Plots per condition
-
-since gpbf_iu has not found an effect in any simulation run, it is not informative to include it in a plot
+since `gpbf_iu` has not found an effect in any simulation run, it is not informative to include it in a plot
 ```{r}
@@
 
-First the relationship between sensitivity and specificity is plot for $errorsd$, then for $n$ and finally for $k$
+First the relationship between Sensitivity and Specificity is plot for $errorsd$, then for $n$ and finally for $k$
 It was expected that there might have been an interaction between $errorsd$ and $k$ and so the final plot will show this interaction
@@
 
-*Figure 1.* The effect of *errorsd* on the sensitivty and specificity of the algorithms.
+*Figure 1.* The effect of *errorsd* on the sensitivty and Specificity of the algorithms.
 ```{r, echo = F}
@@
 It seems that the higher the reliability, the more power, but the more chance for Type I errors.
-Do note the y-axis representing specificity, however as it starts at 0.85.
-In the condition that $errorsd = 0$ it sensitivity is above 50\%, showing that with no reliability issues, it still has a great ablity to detect no effect (specificity = 0.88), and even an ability to detect an effect that is higher than chance (Sensitivity = 0.62).
+Do note that the y-axis representing Specificity starts at 0.85.
+In the condition that $errorsd = 0$, the Sensitivity of `prodbf_ic` is above 50\%, showing that with no reliability issues, it still has a great ability to detect no effect (Specificity = 0.88), and even an ability to detect an effect that is higher than chance (Sensitivity = 0.62).
 Do be aware that this means a 12\% chance of a Type I error and a 48\% of a Type II error.
 
-Conventionally, the probabilty of a Type I error is set at 5\%.
-In this is desired the `tbf_ic` is the superior algorithm.
-When $errorsd = 0$ its specificity still 0.98, while its sensitivity is 0.47.
+If a lower Type I error rate is desired, then `tbf_ic` is the superior algorithm.
+When $errorsd = 0$ its Specificity still 0.98, while its Sensitivity is 0.47.
 Still, a conventional desire for power is 80\% and no algorithm comes close to this number.
@@
 For `prodbf_ic`, its BF more easily increases drastically if only a few groups find large effects, as the BF is bound between 0 and $\infty$.
-If the BF of only one group is, let's say, 2000, than multiplying with the other Bfs will likely still lead to a high `prodBF_ic`.
+If the BF of only one group is, let's say, 2000, than multiplying with the other BF's will likely still lead to a high `prodBF_ic`.
 
-It seems all in all for $n$ as if the performance difference between algorithms is larger than the performance difference between conditions.
+All in all it seems as if if the performance difference between algorithms is larger than the performance difference between conditions.
 
@@
 
-*Figure 2.* The effect of *n* on the sensitivty and specificity of the algorithms.
+*Figure 2.* The effect of *n* on the Sensitivity and Specificity of the algorithms.
 ```{r, echo = F}
@@
 Again, the y-axis starts at 0.80.
-For all algorithms except `prodbf_ic`, n has barely any influence.
-This means that the performance of these algorithms are not much dependent on the sample sizes within the dataframes.
-For `prodbc_ic` the effect of $n$ is more clearly visible with quite some difference in specificity when $n$ drops from 80 to 20.
-Sensitivity does seem to slightly improve as $n$ gets larger for all algorithms except `gpbf_ic`. When $n = 500$, there is almost no loss in specificity, while this loss gets larger as $n$ decreases.
+For all algorithms except `prodbf_ic`, $n$ has barely any influence.
+This means that the performance of these algorithms are not much dependent on the sample sizes within groups $k$.
+For `prodbc_ic` the effect of $n$ is more clearly visible with quite some difference in Specificity when $n$ drops from 80 to 20.
+Sensitivity does seem to slightly improve as $n$ gets larger for all algorithms except `gpbf_ic`. When $n = 500$, there is almost no loss in Specificity, while this loss gets larger as $n$ decreases.
 
 #### k
-*Figure 3.* The effect of *k* on the sensitivity and specificity of the algorithms.
+*Figure 3.* The effect of *k* on the Sensitivity and Specificity of the algorithms.
 ```{r, echo = F}
@@
 
-the general effect of $k$ seems to be that the higher $k$, the more balance between specificity and sensitivity.
+the general effect of $k$ seems to be that the higher $k$, the more balance between Specificity and Sensitivity.
 Especially `prodbf_ic` at $k = 10$ shows promising results with Specificity = 0.87 and Sensitivity = 0.78.
@@
 ```{r, eval = F, echo = F}
-kn_df <- do.call(rbind, lapply(cond_num('k'), function(x){
-  do.call(rbind, lapply(cond_num('n'), function(y){
-    temp <- create_df_plot(dat[k == x & n == y])
-    temp$condition1 <- paste0('k = ', x)
-    temp$condition2 <- paste0('n = ', y)
-    temp
-    }))
-  }))
-
-ke_df <- do.call(rbind, lapply(cond_num('k'), function(x){
-  do.call(rbind, lapply(cond_num('errorsd'), function(y){
-    temp <- create_df_plot(dat[k == x & errorsd == y])
-    temp$condition1 <- paste0('k = ', x)
-    temp$condition2 <- paste0('errorsd = ', y)
-    temp
-    }))
-  }))
-
-en_df <- do.call(rbind, lapply(cond_num('errorsd'), function(x){
-  do.call(rbind, lapply(cond_num('n'), function(y){
-    temp <- create_df_plot(dat[errorsd == x & n == y])
-    temp$condition1 <- paste0('errorsd = ', x)
-    temp$condition2 <- paste0('n = ', y)
-    temp
-    }))
-  }))
-
-saveRDS(kn_df, file = "./Analysis/plot_kn.RData")
-saveRDS(ke_df, file = "./Analysis/plot_ke.RData")
-saveRDS(en_df, file = "./Analysis/plot_en.RData")
+# kn_df <- do.call(rbind, lapply(cond_num('k'), function(x){
+#   do.call(rbind, lapply(cond_num('n'), function(y){
+#     temp <- create_df_plot(dat[k == x & n == y])
+#     temp$condition1 <- paste0('k = ', x)
+#     temp$condition2 <- paste0('n = ', y)
+#     temp
+#     }))
+#   }))
+# 
+# ke_df <- do.call(rbind, lapply(cond_num('k'), function(x){
+#   do.call(rbind, lapply(cond_num('errorsd'), function(y){
+#     temp <- create_df_plot(dat[k == x & errorsd == y])
+#     temp$condition1 <- paste0('k = ', x)
+#     temp$condition2 <- paste0('errorsd = ', y)
+#     temp
+#     }))
+#   }))
+# 
+# en_df <- do.call(rbind, lapply(cond_num('errorsd'), function(x){
+#   do.call(rbind, lapply(cond_num('n'), function(y){
+#     temp <- create_df_plot(dat[errorsd == x & n == y])
+#     temp$condition1 <- paste0('errorsd = ', x)
+#     temp$condition2 <- paste0('n = ', y)
+#     temp
+#     }))
+#   }))
+# 
+# saveRDS(kn_df, file = "./Analysis/plot_kn.RData")
+# saveRDS(ke_df, file = "./Analysis/plot_ke.RData")
+# saveRDS(en_df, file = "./Analysis/plot_en.RData")
 
@@
 
-*Figure 4.* The effect of the interaction between *k* and *n* on the sensitivity and specificity of the algorithms.
+*Figure 4.* The effect of the interaction between *k* and *n* on the Sensitivity and Specificity of the algorithms.
 ```{r, echo = F}
@@
 The general trends of $k$ and $n$ seem to hold that the higher both $n$ and $k$, the better the performance.
-When $n = 80$ it becomes more clear that specificity is sacrificied for better sensitivity, but mostly only for `prodbf_ic`.
+When $n = 80$ it becomes more clear that Specificity is sacrificed for better Sensitivity, but mostly only for `prodbf_ic`.
 Especially in the case where $n = 500$ and $k = 10$, `prodbf_ic` performs very well with Sensitivity = 0.88 and Specificity = 0.95.
@@
 
-*Figure 5.* The effect of the interaction between *k* and *errorsd* on the sensitivity and specificity of the algorithms.
+*Figure 5.* The effect of the interaction between *k* and *errorsd* on the Sensitivity and Specificity of the algorithms.
 ```{r, echo = F}
@@
 It seems the the marginal effect of $errorsd$ gets larger for higher values of $k$.
-Meaning a higher rise in power, but drop in specificity for larger values of $k$.
+Meaning a higher rise in power, but drop in Specificity for larger values of $k$.
 This is mainly seen for `prodbf_ic` and a bit for `tbf_ic`.
-Also interesting to note is that when $errorsd = 0$ and $k = 10$ the only known condition is where an algorithm has a higher Sensitivity than specificity.
-`prodbf_ic` scores 0.96 on sensitivty and 0.77 on specificity in this condition.
+Again, this makes sense as `prodbf_ic` only needs one group with a large BF to be able to calculate a large overall estimate.
+The interaction that as the $errorsd$ gets lower and $k$ gets higher, the Specificity gets lower, maybe has to do with the bounds of the BF being $0 \to \infty$.
+It is likely easier to get a more extreme estimate towards $\infty$ than it is towards 0.
+If $errorsd = 0.81$, then it becomes easier to obtain groups with more extreme estimates towards 0, and the larger values will be less influential.
+Also interesting to note is that when $errorsd = 0$ and $k = 10$ the only known condition is where an algorithm has a higher Sensitivity than Specificity.
+`prodbf_ic` scores 0.96 on Sensitivity and 0.77 on Specificity in this condition.
 
-*Figure 6.* The effect of the interaction between *n* and *errrorsd* on the sensitivity and specificity of the algorithms.
+*Figure 6.* The effect of the interaction between *n* and *errrorsd* on the Sensitivity and Specificity of the algorithms.
 ```{r, echo = F}
@@
 All in all it can be said the the algorithms are conservative.
-In almost all conditions for all algorithms, the specificity was higher than the sensitivity.
-In fact the only condition and algorithm when this was reversed was for `prodbf_ic` when $k = 10$ and $errorsd = 0$.
+In almost all conditions for all algorithms, the Specificity was higher than the Sensitivity.
+In fact, the only condition and algorithm when this was reversed was for `prodbf_ic` when $k = 10$ and $errorsd = 0$.
 This conservatism can be interpreted as good, because the probability of making a Type I error is low.
-However, the finding that the sensitivity was rather low implies that there is almost no power to detect an effect.
+However, the finding that the Sensitivity was rather low implies that there is almost no power to detect an effect.
 The current study only simulated conditions when the true effect size was .1 higher than the hypothesized value.
-Although, this is closer to reality as effect sizes are rarely large, its still interesting if this conservatism is retained by the algorithms as the differences between es and hyp_val gets larger.
+Although, this is closer to reality as effect sizes in populations are rarely large.
+In smaller samples, larger effect sizes are more plausible however and its interesting to see if this conservatism is retained by the algorithms as the differences between $es$ and $hyp_val$ gets larger
 
-There were some exceptions that seemed promising.
+There were some exceptions that seemed more promising.
 These only apply to the `prodbf_ic` and `tbf_ic` algorithms, however.
 These conditions were mainly if both $k$ and $n$ are large.
-Real life examples usually do not have data of 10 groups, each containing 500 observations, although within the Baysian Framework this is more plausible as more research is done in the field.
+Real life examples usually do not have data of 10 groups, each containing 500 observations, although within the Baysian Framework this is more plausible with the use of historical datasets in current analyses.
 
@@
 The `gpbf` is the `prodbf`$^{\frac{1}{k}}$ which for our particular example makes $2^{10*\frac{1}{10}} = 2^1 = 2$.
-This is the main reason that as $errorsd = 0$, the `gpbf_ic` gets more extreme as the probability of consensus between groups becomes larger.
+This is the main reason that as $errorsd = 0$, the `prodbf_ic` gets more extreme as the probability of consensus between groups becomes larger when $es > hyp\_val$.
 
@@
 The overall conclusion is that the algorithms lack power to detect an effect and can not realistically be used.
-Only if much data is available on multiple groups do the `prodbf_ic` and `tbf_ic` become useful, Where the `prodbf_ic` sacrificies a bit more specificity than `tbf_ic` for more sensitivity.
+Only if much data is available on multiple groups do the `prodbf_ic` and `tbf_ic` become useful, Where the `prodbf_ic` sacrifices a bit more Specificity than `tbf_ic` for more Sensitivity.
 

### 2022-05-09 Analysis/Report.Rmd
update 2022-05-09 19:02:38

@@
 title: "Report"
-author: "Eli Clapper"
+author: "Caspar van Lissa & Eli Clapper"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 library(ggplot2)
-dat <- readRDS(file.path("../Sim_Eli", "sim_results_2022-03-19.RData")) #
-dat <- dat[!es < .1]
-```
-
-<!-- Some preparatory variables which are the algorithms (varsout) and the conditions (varspred) -->
-```{r, echo = F}
-varsout <- c("allsig", "gpbf_ic", "gpbf_iu", "prodbf_ic", "prodbf_iu", "tbf_ic", "tbf_iu")
-varspred <- c("es", "errorsd", "n", "k", "hyp_val")
-```
-
-### Marginal Descriptives
-<!--
-Obtain marginal descriptives for all algorithms separately.
-This function first aggregates all conditions by taking the median over all 1000 simulation runs on that condition, reducing the `data.table` to 72000/1000 = 72 medians for the different conditions.
-Finally it describes the distribution of medians of these 72 conditions for every algorithm.
-It takes some quantiles, the median absolute deviations from the overall median of that algorithm, the sum of where BF > 3 and the range.
-Later on, confusion matrix metrics such as Sensitivity and Positive Likelihood Ratio are calculated on the non-aggregated `data.table`.
--->
-
-*Table 1.* The marginal median Bayes Factors for all algorithms.
-```{r, echo = F}
-descriptives <- function(datf){ 
-  descs <- as.matrix(
-    datf[, lapply(.SD, function(var){
-    quants <- quantile(var, c(0.25, 0.50, 0.75, 0.90, 0.95))# obtain overall quantiles
-    mads <- mad(var)       # obtain Median absolute deviations per conditions
-    sums <- sum(var>3)     # obtain sums where the median is > 3
-    maxvals <- max(var)    # obtain the highest median
-    minvals <- min(var)    # obtain lowest median
-    c(quants[1], quants[2], quants[3], quants[4], quants[5], mads, sums, maxvals, minvals)
-  }), .SDcols = varsout])
-  rownames(descs) <- c('25 %', '50 %',  '75 %','90 %', '95 %', "mad", "sum>3", "max", "min")
-  descs
-}
-marginal_descs <- descriptives(dat[ ,lapply(.SD, median), .SDcols = varsout, by = varspred])
-datatable(round(marginal_descs,5), options = list(dom = 't'))
-```
-*Every condition has been simulated 1000 times. For every condition and algorithm, the Bayes Factors are aggregated by taking the median over all Bayes Factors, leaving 72 median BF's per algorithm*
-
-Optimally, the algorithms should have $\frac{72}{2} = 36$ conditions where the median BF > 3 in the cases where $es$ > $hyp_val$.
-The marginal median for all algorithms is $\approx 1.0$, however.
-Even at the $90^{th}$ percentile only the `prodbf_ic` algorithm has its median BF > 3.
-The Median Absolute Deviations are not too severe, providing evidence for somewhat reliable estimates.
-`Prodbf_ic` does have a notably higher mad, which is likely due to its large valued outliers.
-Because the BF is bound between 0 and $\infty$, outliers are plausible and can be influential on the MAD when, in this case, only 72 medians construct the distribution of conditional medians.
-
-The finding is that the algorithms are rather conservative, which is further supported by the number of times the median BF does exceed 3.
-Both `gpbf` algorithms count 0 conditions where median BF > 3, while for `allsig` this number is 1.
-For the `prodbf_iu` and `tbf` algorithms 5 or 6 times is counted and only the `prodbf_ic` seems to have a more notable number of times the median BF > 3, namely 16.
-Another observation is that the informative BFs against the unconstrained hypotheses seem to be a bit more conservative than their complement hypotheses counterparts.
-Finally, it seems as if the columns for `prodbf_iu` and `tbf_iu` are exactly equal.
-Lets check that out.
-
-```{r}
-all.equal(dat$tbf_iu, dat$prodbf_iu)
-cor(dat$tbf_iu, dat$prodbf_iu)
-```
-
-`tbf_iu` and `prodbf_iu` are indeed perfectly correlated and all values in both columns in the complete `data.table` are equal.
-Upon further inspection it does not seem to be a simulation error.
-It seems as if `tbf_iu` and `prodbf_iu` do the exact same thing.
-It would mean that the product of the unconstrained Bayes Factors is the same as the unconstrained Bayes Factor under the condition that the hypothesis holds for all $k$ groups.
-
-#### Descriptives for es > .1
-
-With this in mind, let us continue. We can filter the tbf_iu out of the data.
-Now we subset for when true BF > 3 and describe the distribution for those conditional medians.
-```{r}
-dat <- dat[,tbf_iu:=NULL]
-varsout <- c("allsig", "gpbf_ic", "gpbf_iu", "prodbf_ic", "prodbf_iu", "tbf_ic")
-```
-
-*Table 2.* The median Bayes Factors for all algorithms in the condition that es > .1.
-```{r, echo = F}
-es_gt_hyp_val <- descriptives(dat[es == .2 ,lapply(.SD, median), .SDcols = varsout, by = varspred])
-datatable(round(es_gt_hyp_val,5),options = list(dom = "t"))
-```
-
-```{r}
-all(es_gt_hyp_val['sum>3',varsout] == marginal_descs['sum>3',varsout]) 
-```
-
-Even when the `data.table` is subsetted for when es > hyp_val no median BF gets above 3.
-Also, all cases where the median BF does exceed 3 is indeed only in the conditions when es > hyp_val, suggesting the algorithms do not accredit effects when in reality there is indeed no effect.
-Again, the MADs are not too severe, except for the prodbf_ic, suggesting the prod_bf algorithm is less stable when BF > 3 and can calculate large estimates even when the true_es is just 0.1 point higher than the hypothesized value.
-Again, the unconstrained algorithms seem to generally be more conservative than the complementary.
-
-#### Descriptives for es < .1
-
-*Table 2.* The median Bayes Factors for all algorithms in the condition that es < .1.
-```{r, echo =F}
-es_st_hyp_val <- descriptives(dat[es == .1 ,lapply(.SD, median), .SDcols = varsout, by = varspred])
-datatable(round(es_st_hyp_val,5), options = list(dom = 't'))
-```
-Further supporting the idea of conservative algorithms is the finding is that the maximum median BFs in the cases where $es = hyp\_val$ all are $\approx 1$ and  all $75^{th}$ percentiles are slightly below one.
-The algorithms seem to favour the hypotheses that are not in the hypothesized direction, although very moderately.
-The previously noticed effect of the unconstrained vs.
-the complement algorithms is not immediately visible.
-
-### inspecting conditions when median prodbf_ic > 3
-Finally, before going to the non-aggregated `data.table`, the conditions are inspected where median `prodbf_ic` > 3. The dt is then ordered by $k$ and `prodbf_ic` values.
-
-*Table 2.* The 16 conditions when the median BF for prodbf_ic > 3.
-```{r, echo = F}
-prodbf_gt_3 <- dat[,lapply(.SD, median), .SDcols = varsout, by = varspred][prodbf_ic > 3][order(k, prodbf_ic, decreasing = T)]
-datatable(round(prodbf_gt_3,3), options = list(pageLength = 5))
-```
-First of all, the conditions where the median BF exceeds three for `prodbf_ic`, are also the conditions where it exceeds three for all other algorithms, except `allsig.` The only condition where the median BF for `allsig` exceeded 3 was when $es = 0.2, errorsd = 0, n = 500, k = 1$, and $hyp\_val = 0.1$.
-It is evident that the higher values for the median BF occur for higher values of $k$, with ten out of sixteen times $k = 10$ and the rest when $k = 3$.
-This makes sense as the `prod_bf` algorithms take the product of the Bayes Factor for all $k$ groups.
-The likelihood of BF_ic and BF_iu being bigger than 1 or a specific group increases as $es > hyp\_val$, resulting in big numbers when all BFs are multiplied.
-The finding that the other algorithms are more conservative than the `prodbf_ic` is likely because they correct for the number of groups.
-For the `gpbf` algorithms by raising the `prodbf` to the power $\frac{1}{k}$.
-The `tbf` algorithms correct for $k$ by calculating the BFs under the condition that the effect in all $k$ groups together is bigger than the hypothesized value, rather than for all $k$ groups separately.
-The `prodbf_iu` algorithm does not directly correct for $k$, but it is not restricted to testing only against the complement of the hypothesized direction.
-It rather tests against the unconstrained hypotheses.
-In other words, how much more likely is the informative hypothesis than the hypothesis where there is no equality constrained at all.
-
-The other condition that seems to have some importance is $errorsd$, where most algorithms found an effect when $errorsd \neq 0.81$ and preferably when $errorsd = 0$.
-Further analysis will investigate this idea.
-
-### Confusion matrix metrics
-
-<!-- defining functions below which are useful for later when plotting per condition -->
-```{r, echo = F}
-# obtain confusion matrix metrics with a dimension of 1 by alg*metrics (so 1x28 in this datatable)
-lapply_at <- function(var, truees) {
-  results <- sapply(var, function(var) {
-    table(ordered(truees > .1, levels = c("FALSE", "TRUE")), ordered(var > 3, levels = c("FALSE", "TRUE")))
-  })
-  names(results) <- vapply(names(var), paste, c("TN", "FN", "FP", "TP"), sep = "_", 
-                           FUN.VALUE = character(4),
-                           USE.NAMES = FALSE)
-  as.list(results)
-}
-
-# in long format gives a 4x7 matrix with 4 metrics and 7 algorithms
-metrics <- function(datf){
-  res <- datf[, sapply(.SD, function(var) {
-    table(ordered(es > .1, levels = c("FALSE", "TRUE")), ordered(var > 3, levels = c("FALSE", "TRUE")))
-  }), .SDcols = varsout]
-  rownames(res) <- c("TN", "FN", "FP", "TP")
-  return(res)
-}
-
-# obtain metrics per algorithm based on FP, TN, FN and TP() and collect in a matrix
-get_stats <- function(res){ 
-  suppressMessages(
-  apply(res, 2, function(x){
-    attach(as.list(x))
-    alpha = FP / (FP + TN) #False positive rate 
-    beta = FN / (TP + FN) # False negative rate
-    sensitivity = 1-beta
-    specificity = 1-alpha
-    correct = (TN + TP) /(FP + FN + TN + TP)
-    pos_lr = sensitivity / (1 - specificity)
-    neg_lr = (1 - sensitivity) / specificity
-    c(correct = correct,
-      False_positive_rate_alpha = alpha,
-      False_negative_rate_beta = beta,
-      sensitivity = sensitivity,
-      specificity = specificity, 
-      pos_lr = ifelse(is.na(pos_lr), 0, pos_lr),
-      neg_lr = ifelse(is.na(neg_lr), 0, neg_lr))
-  })
-  )
-}
-
-# create df from stats matrix for plotting purposes
-plot_df <-function(stats){ 
-  data.frame(Method = colnames(stats),
-        correct = stats["correct", ],
-        Sensitivity = stats["sensitivity", ],
-        Specificity = stats["specificity", ],
-        FPR = stats["False_positive_rate_alpha", ],
-        FNR = stats["False_negative_rate_beta", ],
-        Pos_lr = stats["pos_lr", ],
-        Neg_lr = stats["neg_lr", ])
-}
-
-# all in one function. Create plot df for a subset of the data table.
-create_df_plot <- function(datf){
-  metrics_long <- metrics(datf)
-  stats <- get_stats(metrics_long)
-  df_plot <- plot_df(stats)
-  return(df_plot)
-}
-
-#set colours to use for different algorithms
-colours <- c("black", "orange" , "red", "#79aaf7", "blue", "#00a62f" )
-
-# obtain numeric values each conditions take on
-cond_num <- function(cond){return(as.numeric(names(table(dat[,..cond]))))}
-
-#obtain condition names. For k it returns c("k = 1", "k = 3", "k = 10")
-cond_names <- function(cond){return(paste0(cond, " = ",names(table(dat[,..cond]))))}
-```
-
-*Table 4.* The marginal confusion matrix metrics.
-```{r, echo = F, message=F}
-df_plot <- create_df_plot(dat)
-marginal_metrics <- rbind(df_plot[,2:ncol(df_plot)], 
-      mean = apply(df_plot[,2:ncol(df_plot)], 2 ,mean))
-datatable(round(marginal_metrics, 5), options = list(dom = 't'))
-```
-
-Firstly, on average, the algorithms were correct in 58\% of the cases.
-The `prodbf_ic` did best with 68\% identified correctly, followed by 62\% of the tbf_ic.The gbpf algorithms do overall worst with $\approx 50\%$ being identified correctly.
-Also interesting, all `_ic` algorithms do better than their `_iu` counterparts.
-
-Secondly, all measures also suggest conservative algorithms, with a mild exception for `prodbf_ic.` The mean Sensitivity of the algorithms is 0.17, meaning that on average, out of all cases where BF > 3, only 17\% is identified correctly.
-Interestingly, the `gpbf` algorithms are especially conservative with an almost perfect Specificity, but no ability to detect an effect.
-The `tbf` algorithms and `prodbf_iu` do slightly better in this regard.
-An obvious outlier is the `prodbf_ic` algorithm with 44\% correctly identified when BF > 3.
-
-Where the algorithms make up is in the Specificity, with a mean specificty of 0.98.
-This implies that out of all cases where BF < 3, 98\% is identified correctly on average.
-The `prodbf_ic` performs only slightly worse than the other algorithms, although the differences are not too severe.
-The measure does not say much however in combination with the low Sensitivity.
-If all cases are evaluated as having no effect anyway, it is no surprise or feat that the Specificity is (close to) perfect.
-For a clearer picture, we look at the LR+ and LR-.
-
-Looking at the LR+, `prodbf_iu` has the best performance with 30.3, followed by `tbf_ic` with 15.3.
-It means that the `prodbf_iu`, out off all cases it accredited an effect, it does best to separate it from actual no effects.
-The finding that this value is lower for `prodbf_ic` means that `prodbf_ic` has a higher likelihood of finding an effect when there is none.
-
-The LR- corrects for algorithms' conservatism.
-For example, `gpbf_iu` has a value of 1, which means that out of all the tests it determined that there was no effect, 100\% of the times it was wrong when in reality there was an effect.
-`prodbf_ic` actually does best in this regard.
-Out of all the tests it determined that there was no effect, 60\% it was wrong when there in reality was an effect.
-
-A first conclusion would be that the `prodbf_ic` is the algorithm of choice.
-It is the only algorithm that has a somewhat higher ability to detect an effect when there is one, while still performing well when there is no effect.
-It does have a somewhat higher probability of finding an effect when there in reality is none which could be a valid concern.
-When the chance of a Type I error needs to be reduced as much as possible, it would be wiser to choose `prodbf_iu` or `tbf_ic` which have less ability to detect an effect, but are also less likely to call an effect when there in reality is none.
-
-### Plots per condition
-since `gpbf_iu` has not found an effect in any simulation run, it is not informative to include it in a plot
-```{r}
-all(dat$gpbf_iu < 3)
-```
-
-And thus it can be omitted during plotting of the confusion matrix metrics, because it is not informative
-```{r}
-dat <- dat[,gpbf_iu:=NULL]
-varsout <- c("allsig", "gpbf_ic", "prodbf_ic", "prodbf_iu", "tbf_ic")
-```
-
-First the relationship between Sensitivity and Specificity is plot for $errorsd$, then for $n$ and finally for $k$
-It was expected that there might have been an interaction between $errorsd$ and $k$ and so the final plot will show this interaction
-
-#### errorsd
-
-*Figure 1.* The effect of *errorsd* on the sensitivty and Specificity of the algorithms.
-```{r, echo = F}
-sd_df <- do.call(rbind, lapply(cond_num('errorsd'), function(x){create_df_plot(dat[errorsd == x])})) 
-sd_df$condition <- rep(cond_names('errorsd'), each = length(varsout))
-
-ggplot(sd_df, aes(x = Sensitivity, y = Specificity, colour = Method, shape = condition)) +
-  geom_point(size = 2.5) +
-  scale_x_continuous(limits = c(0,1))+
-  scale_y_continuous(limits = c(0.85,1)) +
-  scale_color_manual(values = colours) +
-  theme_bw()
-```
-
-There seems to be a moderate effect of $errorsd$ for `prodbf_ic`.
-It seems that the higher the reliability, the more power, but the more chance for Type I errors.
-Do note that the y-axis representing Specificity starts at 0.85.
-In the condition that $errorsd = 0$, the Sensitivity of `prodbf_ic` is above 50\%, showing that with no reliability issues, it still has a great ability to detect no effect (Specificity = 0.88), and even an ability to detect an effect that is higher than chance (Sensitivity = 0.62).
-Do be aware that this means a 12\% chance of a Type I error and a 48\% of a Type II error.
-
-If a lower Type I error rate is desired, then `tbf_ic` is the superior algorithm.
-When $errorsd = 0$ its Specificity still 0.98, while its Sensitivity is 0.47.
-Still, a conventional desire for power is 80\% and no algorithm comes close to this number.
-The `gpbf_ic` and `allsig` algorithms do not seem to be affected much by the variation in $errrosd$.
-
-The relationship does make sense as the `tbf` algorithm seems to be more dependent on group consensus to find an effect.
-There is a high probability of consensus in the condition that $errorsd = 0$.
-For `prodbf_ic`, its BF more easily increases drastically if only a few groups find large effects, as the BF is bound between 0 and $\infty$.
-If the BF of only one group is, let's say, 2000, than multiplying with the other BF's will likely still lead to a high `prodBF_ic`.
-
-All in all it seems as if if the performance difference between algorithms is larger than the performance difference between conditions.
-
-#### n
-
-*Figure 2.* The effect of *n* on the Sensitivity and Specificity of the algorithms.
-```{r, echo = F}
-n_df <- do.call(rbind, lapply(cond_num('n'), function(x){create_df_plot(dat[n == x])})) 
-n_df$condition <- rep(cond_names('n'), each = length(varsout))
-
-ggplot(n_df, aes(x = Sensitivity, y = Specificity, colour = Method, shape = condition)) +
-  geom_point(size = 2.5) +
-  scale_x_continuous(limits = c(0,1))+
-  scale_y_continuous(limits = c(0.80,1)) +
-  scale_color_manual(values = colours) +
-  theme_bw()
-```
-
-There seems to be only a slight improvement of the algorithms as $n$ gets larger.
-Again, the y-axis starts at 0.80.
-For all algorithms except `prodbf_ic`, $n$ has barely any influence.
-This means that the performance of these algorithms are not much dependent on the sample sizes within groups $k$.
-For `prodbc_ic` the effect of $n$ is more clearly visible with quite some difference in Specificity when $n$ drops from 80 to 20.
-Sensitivity does seem to slightly improve as $n$ gets larger for all algorithms except `gpbf_ic`. When $n = 500$, there is almost no loss in Specificity, while this loss gets larger as $n$ decreases.
-
-#### k
-*Figure 3.* The effect of *k* on the Sensitivity and Specificity of the algorithms.
-```{r, echo = F}
-k_df <- do.call(rbind, lapply(cond_num('k'), function(x){create_df_plot(dat[k == x])})) 
-k_df$condition <- rep(cond_names('k'), each = length(varsout))
-
-ggplot(k_df, aes(x = Sensitivity, y = Specificity, colour = Method, shape = condition)) +
-  geom_point(size = 2.5) +
-  scale_x_continuous(limits = c(0,1))+
-  scale_y_continuous(limits = c(0.85,1)) +
-  scale_color_manual(values = colours) +
-  theme_bw()
-```
-
-the general effect of $k$ seems to be that the higher $k$, the more balance between Specificity and Sensitivity.
-Especially `prodbf_ic` at $k = 10$ shows promising results with Specificity = 0.87 and Sensitivity = 0.78.
-Also the `tbf_ic` does well at $k = 10$, with a notably lower Sensitivity (0.52) than `prodbf_ic`, but a notably higher Specificity (0.98).
-
-### interactions
-<!-- this is not optimal and takes long time to run, thats why the outputs are saved in .RData files. -->
-```{r, eval = F, echo = F}
-# kn_df <- do.call(rbind, lapply(cond_num('k'), function(x){
-#   do.call(rbind, lapply(cond_num('n'), function(y){
-#     temp <- create_df_plot(dat[k == x & n == y])
-#     temp$condition1 <- paste0('k = ', x)
-#     temp$condition2 <- paste0('n = ', y)
-#     temp
-#     }))
-#   }))
-# 
-# ke_df <- do.call(rbind, lapply(cond_num('k'), function(x){
-#   do.call(rbind, lapply(cond_num('errorsd'), function(y){
-#     temp <- create_df_plot(dat[k == x & errorsd == y])
-#     temp$condition1 <- paste0('k = ', x)
-#     temp$condition2 <- paste0('errorsd = ', y)
-#     temp
-#     }))
-#   }))
-# 
-# en_df <- do.call(rbind, lapply(cond_num('errorsd'), function(x){
-#   do.call(rbind, lapply(cond_num('n'), function(y){
-#     temp <- create_df_plot(dat[errorsd == x & n == y])
-#     temp$condition1 <- paste0('errorsd = ', x)
-#     temp$condition2 <- paste0('n = ', y)
-#     temp
-#     }))
-#   }))
-# 
-# saveRDS(kn_df, file = "./Analysis/plot_kn.RData")
-# saveRDS(ke_df, file = "./Analysis/plot_ke.RData")
-# saveRDS(en_df, file = "./Analysis/plot_en.RData")
+rename <- c("prodbf" = "PBF", "gpbf" = "gPBF", "tbf" = "TBF", "ipd" = "IPD", "rma" = "RMA", "allsig" = "VC")
 
@@
 
-<!-- read in interaction dfs to plot -->
-```{r, echo = F}
-kn_df <- readRDS("../Analysis/plot_kn.RData")
-ke_df <- readRDS("../Analysis/plot_ke.RData")
-en_df <- readRDS("../Analysis/plot_en.RData")
+# Results
 
+For each algorithm, a confusion matrix was obtained by tabulating inferential decisions made using that algorithm against the population status of the hypothesis (true or false).
+For all BFs, the informative hypothesis was accepted if BF > 3.
+For the IPD and RMA meta-analysis, the null-hypothesis was rejected if the confidence interval for the overall effect size excluded the hypothesized value.
+Finally, for the vote count (VC) algorithm, the null was rejected if all individual group hypotheses were rejected.
+Table \@ref(tab:tabconf) shows the resulting confusion matrix metrics.
+The only algorithm that shows relatively high values for both sensitivity and specificity is the product Bayes factor (PBF).
+This is also reflected in the high accuracy, which represents the total proportion of correct (true positive/true negative) decisions.
+
+```{r tabconf, echo = F, message=F}
+tab <- read.csv("../confusion.csv", stringsAsFactors = FALSE)
+names(tab)[1] <- "Metric"
+tab$Metric <- gsub("^.+_", "", tab$Metric)
+tab <- tab[!tab$Metric == "lr", !grepl("iu", names(tab))]
+names(tab) <- gsub("_ic", "", names(tab))
+names(tab)[match(names(rename), names(tab))] <- rename
+tab <- tab[tab$Metric %in% c("sensitivity", "specificity", "accuracy"),]
+rownames(tab) <- NULL
+knitr::kable(tab, digits = 2, caption = "Marginal confusion matrix metrics.")
+```
+These results indicate that all algorithms except PBF had low sensitivity to detect a true effect.
+In contrast, specificity was very high for all algorithms except PBF.
+This suggests that the other algorithms classified most conditions as negatives (no effect found),
+regardless of the existence of a population effect.
+The PBF trades a loss of specificity for increased sensitivity,
+and average levels of both were approximately equal.
+As the PBF vastly outperformed other BFs, these were omitted from further analysis.
+
+## Effect of simulation conditions
+
+We examined the effect of simulation conditions on overall accuracy.
+
+```{r tabeffect, echo = FALSE}
+tab <- read.csv("../effect_of_conditions.csv", stringsAsFactors = FALSE)
+tab <- tab[, !grepl("(gpbf|tbf)", names(tab))]
+tab <- tab[, c("condition", "prodbf_ic", "rma", "ipd", "allsig", "prodbf_ic.vs..rma", "ipd.vs..prodbf_ic", "allsig.vs..prodbf_ic")]
+names(tab) <-  c("Factor", "PBF", "RMA", "IPD", "VC", "vs RMA", "vs IPD", "vs VC")
+
+knitr::kable(tab, digits = 2, caption = "Partial eta squared of the effect of each design factor on accuracy for each algorithm and for the difference between PBF and all other algorithms (e.g., vs RMA).")
+```
+PBF performance was most impacted by sample size $n$, followed by the number of groups $k$, and residual heterogeneity.
+
+### Effect of sample size
+
+```{r fign, fig.cap="Mean performance by sample size"}
+res <- readRDS("../confusion_by_cond.RData")
+res <- res[res$alg %in% c("ipd", "rma", "allsig", "prodbf_ic"), ]
+
+df_plot <- melt(res, measure.vars = c("sensitivity", "specificity"),
+             variable.name = "Outcome", value.name = "val")
+df_plot <- df_plot[ ,list(mean=mean(val)), by=c("n", "Outcome", "alg")]
+ggplot(df_plot, aes(x = n, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
+  geom_point() +
+  geom_line() +
+  theme_bw() +
+  labs(y = "Mean performance", x = "Sample size")
 ```
 
-*Figure 4.* The effect of the interaction between *k* and *n* on the Sensitivity and Specificity of the algorithms.
-```{r, echo = F}
-ggplot(kn_df, aes(x = Sensitivity, y = Specificity, colour = Method, shape = condition1)) +
-  geom_point(size = 2.5) +
-  scale_x_continuous(limits = c(0,1))+
-  scale_y_continuous(limits = c(0.70,1)) +
-  scale_color_manual(values = colours) +
-  facet_wrap(~condition2) +
-  theme_bw()
+Figure \@ref(fig:fign) indicates that for PBF, both sensitivity and specificity increase with sample size.
+The other algorithms also show increasing sensitivity, but not specificity, which is at a ceiling.
+This difference explains the effect of errorsd on the diference between algorithms (see Table \@ref(tab:tabeffect)).
 
+### Effect of number of groups
+
+```{r, fig.cap="Mean performance by number of groups"}
+df_plot <- melt(res, measure.vars = c("sensitivity", "specificity"),
+             variable.name = "Outcome", value.name = "val")
+df_plot <- df_plot[ ,list(mean=mean(val)), by=c("k", "Outcome", "alg")]
+ggplot(df_plot, aes(x = k, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
+  geom_point() +
+  geom_line() +
+  theme_bw() +
+  labs(y = "Mean performance", x = "Number of groups")
+```
+The figure indicates that, for PBF at higher levels of k, lower sensitivity is exchanged for greater specificity.
+The other algorithms do not show this pattern, as their specificity is at a ceiling.
+This difference in pattern of effects explains why number of groups has a moderate effect on the difference between algorithms (see Table \@ref(tab:tabeffect)).
+Only VC shows decreasing sensitivity with an increasing number of groups;
+this is because the probability of obtaining any false negatives increases with the number of groups.
+
+### Effect of residual heterogeneity
+
+```{r, fig.cap="Mean performance by errorsd"}
+
+# dat <- readRDS(file.path("../Sim", "sim_results_2022-05-07.RData"))
+# varsout <- c("ipd", "rma", "allsig", "prodbf_ic")
+# varspred <- c("errorsd", "n", "k", "hyp_val")
+# 
+# df_plot <- dat[!es < .1,  lapply(.SD, function(var) {
+#   as.vector(table(ordered(es > .1, levels = c("FALSE", "TRUE")), ordered(var > 3, levels = c("FALSE", "TRUE"))))
+# }), .SDcols = varsout, by = "errorsd"]
+# 
+# tabres[, metric := rep(c("TN", "FN", "FP", "TP"), nrow(tabres)/4)]
+# 
+# tabres <- melt(tabres, measure.vars = varsout,
+#              variable.name = "alg", value.name = "val")
+# tabres <- dcast(tabres, errorsd + n + k + hyp_val + alg ~ metric, value.var = "val")
+# tabres[, alpha := FP / (FP + TN)]
+# tabres[, beta := FN / (TP + FN)]
+# tabres[, sensitivity := 1-beta]
+# tabres[, specificity := 1-alpha]
+# tabres[, accuracy := (TP + TN) / (TP+TN+FP+FN)]
+
+df_plot <- melt(res, measure.vars = c("sensitivity", "specificity"),
+             variable.name = "Outcome", value.name = "val")
+df_plot <- df_plot[ ,list(mean=mean(val)), by=c("errorsd", "Outcome", "alg")]
+ggplot(df_plot, aes(x = errorsd, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
+  geom_point() +
+  geom_line() +
+  theme_bw() +
+  labs(y = "Mean performance")
 ```
 
-The general trends of $k$ and $n$ seem to hold that the higher both $n$ and $k$, the better the performance.
-When $n = 80$ it becomes more clear that Specificity is sacrificed for better Sensitivity, but mostly only for `prodbf_ic`.
-Especially in the case where $n = 500$ and $k = 10$, `prodbf_ic` performs very well with Sensitivity = 0.88 and Specificity = 0.95.
-This suggests a Type I error of 5\% and a power of 88\%. But also `tbf_ic` stands out as both $k$ and $n$ get larger.
+The figure indicates that for PBF, at higher levels of residual heterogeneity, lower sensitivity is exchanged for greater specificity.
+The other algorithms do not show this pattern, as their specificity is at a ceiling.
+Their sensitivity decreases with residual heterogeneity,
+and therefore, so does their overall performance.
+This difference in pattern of effects explains why residual heterogeneity has a moderate effect on the difference between algorithms (see Table \@ref(tab:tabeffect)).
+Note that, in contrast to other algorithms, the performance of PBF is less susceptible to residual heterogeneity.
 
-*Figure 5.* The effect of the interaction between *k* and *errorsd* on the Sensitivity and Specificity of the algorithms.
-```{r, echo = F}
-ggplot(ke_df, aes(x = Sensitivity, y = Specificity, colour = Method, shape = condition1)) +
-  geom_point(size = 2.5) +
-  scale_x_continuous(limits = c(0,1))+
-  scale_y_continuous(limits = c(0.75,1)) +
-  scale_color_manual(values = colours) +
-  facet_wrap(~condition2) +
-  theme_bw()
+# Discussion
 
-```
+Within the scope of this simulation, PBF had relatively better inferential properties than the other algorithms under consideration.
+Although other algorithms had superior specificity, PBF had the highest levels of sensitivity.
+Thus, PBF balanced the ability to accept an informative hypothesis in the presence of a true population effect with the ability to reject the hypothesis when there was no effect.
+<!-- Conventional threshold for type 1 error: 5% -->
+<!-- Conventional threshold for power: 80% -->
+<!-- J. Cohen, Statistical Power Analysis for the -->
+<!--  Behavioral Sciences, 2nd ed. (Erlbaum, Hillsdale, -->
+<!--  NJ, 1988). This is the source of the system of power -->
+<!--  analysis described here; the power values and sam -->
+<!--  ple sizes of the illustrations derive from this book's -->
+<!--  tables. -->
+<!--  2. J -->
 
-It seems the the marginal effect of $errorsd$ gets larger for higher values of $k$.
-Meaning a higher rise in power, but drop in Specificity for larger values of $k$.
-This is mainly seen for `prodbf_ic` and a bit for `tbf_ic`.
-Again, this makes sense as `prodbf_ic` only needs one group with a large BF to be able to calculate a large overall estimate.
-The interaction that as the $errorsd$ gets lower and $k$ gets higher, the Specificity gets lower, maybe has to do with the bounds of the BF being $0 \to \infty$.
-It is likely easier to get a more extreme estimate towards $\infty$ than it is towards 0.
-If $errorsd = 0.81$, then it becomes easier to obtain groups with more extreme estimates towards 0, and the larger values will be less influential.
-Also interesting to note is that when $errorsd = 0$ and $k = 10$ the only known condition is where an algorithm has a higher Sensitivity than Specificity.
-`prodbf_ic` scores 0.96 on Sensitivity and 0.77 on Specificity in this condition.
+An important caveat is that none of the algorithms met conventional criteria for power (80%) and type I error (5%).
+Overall, PBF had the best performance, with power of 76% and type I error of 24%.
+Thus, if researchers intend to synthesize evidence for an informative hypothesis across heterogeneous studies, PBF may be the most suitable method - but its limitations should be acknowledged in applied research.
 
-*Figure 6.* The effect of the interaction between *n* and *errrorsd* on the Sensitivity and Specificity of the algorithms.
-```{r, echo = F}
-ggplot(en_df, aes(x = Sensitivity, y = Specificity, colour = Method, shape = condition1)) +
-  geom_point(size = 2.5) +
-  scale_x_continuous(limits = c(0,1))+
-  scale_y_continuous(limits = c(0.6,1)) +
-  scale_color_manual(values = colours) +
-  facet_wrap(~condition2) +
-  theme_bw()
+The overall performance of PBF increased most with increasing sample size.
+With an increasing number of groups, slightly decreasing specificity was traded off for increasing sensitivity.
+With increasing residual heterogeneity, increasing specificity was traded off for decreasing sensitivity.
 
-```
+<!-- # Conclusions -->
 
-It seems that in general, the effect of $errorsd$ is barely noticeable for different values of $n$. It only becomes for noticeable as $n < 80$ and only for `prodbf_ic`.
-
-# Conclusions
-All in all it can be said the the algorithms are conservative.
+<!-- All in all it can be said the the algorithms are conservative.
 In almost all conditions for all algorithms, the Specificity was higher than the Sensitivity.
@@
 Although, this is closer to reality as effect sizes in populations are rarely large.
-In smaller samples, larger effect sizes are more plausible however and its interesting to see if this conservatism is retained by the algorithms as the differences between $es$ and $hyp_val$ gets larger
+In smaller samples, larger effect sizes are more plausible however and its interesting to see if this conservatism is retained by the algorithms as the differences between $es$ and $hyp_val$ gets larger -->
 
-There were some exceptions that seemed more promising.
+<!-- There were some exceptions that seemed more promising.
 These only apply to the `prodbf_ic` and `tbf_ic` algorithms, however.
@@
 Real life examples usually do not have data of 10 groups, each containing 500 observations, although within the Baysian Framework this is more plausible with the use of historical datasets in current analyses.
+-->
 
-The variation in conditions mainly affected `prodbf_ic`.
+<!-- The variation in conditions mainly affected `prodbf_ic`.
 The initial guess is that `prodbf_ic` can be unstable as it takes the product of the BFs for all $k$ groups.
@@
 This is the main reason that as $errorsd = 0$, the `prodbf_ic` gets more extreme as the probability of consensus between groups becomes larger when $es > hyp\_val$.
+-->
 
-The reason that the `iu` algorithms are more conservative is because they are not constrained by testing against only the complement of the specified hypothesis.
+<!-- The reason that the `iu` algorithms are more conservative is because they are not constrained by testing against only the complement of the specified hypothesis. -->
 
-The overall conclusion is that the algorithms lack power to detect an effect and can not realistically be used.
+<!-- The overall conclusion is that the algorithms lack power to detect an effect and can not realistically be used.
 Only if much data is available on multiple groups do the `prodbf_ic` and `tbf_ic` become useful, Where the `prodbf_ic` sacrifices a bit more Specificity than `tbf_ic` for more Sensitivity.
+-->
 
-```{r, echo = F, eval = F}
-rm(list = ls())
-gc(full = T, reset = T)
-cat("\014")
-```
+<!-- ```{r, echo = F, eval = F} -->
+<!-- rm(list = ls()) -->
+<!-- gc(full = T, reset = T) -->
+<!-- cat("\014") -->
+<!-- ``` -->

### 2022-05-09 Analysis/analysis.R

@@ -1,5 +1,6 @@
 #dat <- readRDS("tabres.RData")
-dat <- readRDS(file.path("Sim_Eli", "sim_results_2022-03-19.RData"))
+library(data.table)
+dat <- readRDS(file.path("Sim", "sim_results_2022-05-08.RData"))
 lapply_at <- function(var, truees) {
   results <- sapply(var, function(var) {
     table(ordered(truees > .1, levels = c("FALSE", "TRUE")), ordered(var > 3, levels = c("FALSE", "TRUE")))
@@ -9,8 +10,8 @@ lapply_at <- function(var, truees) {
                            USE.NAMES = FALSE)
   as.list(results)
 }
-varsout <- c("allsig", "gpbf_ic", "gpbf_iu", "prodbf_ic", "prodbf_iu", "tbf_ic", "tbf_iu")
-varspred <- c("es", "errorsd", "n", "k", "hyp_val")
+varsout <- c("ipd", "rma", "allsig", "gpbf_ic", "gpbf_iu", "prodbf_ic", "prodbf_iu", "tbf_ic", "tbf_iu")
+varspred <- c("errorsd", "n", "k", "hyp_val")
 
 # Example: per condition
 dat[!es < .1, lapply_at(.SD, truees = es), .SDcols = varsout, by = c("n", "errorsd")]
@@ -22,27 +23,103 @@ res <- dat[!es < .1, sapply(.SD, function(var) {
 }), .SDcols = varsout]
 rownames(res) <- c("TN", "FN", "FP", "TP")
 
+tabres <- dat[!es < .1,  lapply(.SD, function(var) {
+  as.vector(table(ordered(es > .1, levels = c("FALSE", "TRUE")), ordered(var > 3, levels = c("FALSE", "TRUE"))))
+}), .SDcols = varsout, by = varspred]
+tabres[, metric := rep(c("TN", "FN", "FP", "TP"), nrow(tabres)/4)]
+
+tabres <- melt(tabres, measure.vars = varsout,
+             variable.name = "alg", value.name = "val")
+tabres <- dcast(tabres, errorsd + n + k + hyp_val + alg ~ metric, value.var = "val")
+tabres[, alpha := FP / (FP + TN)]
+tabres[, beta := FN / (TP + FN)]
+tabres[, sensitivity := 1-beta]
+tabres[, specificity := 1-alpha]
+tabres[, accuracy := (TP + TN) / (TP+TN+FP+FN)]
+
[…]

### 2022-05-09 Sim/Simulatie_V2.R

@@ -1,0 +1,168 @@
+#clear console and env
+rm(list=ls(all.names = T))
+cat("\014")
+
+# Delete folder where previous results are in and create new 'Results' folder
+unlink("./sim/Results", recursive = T)
+dir.create("./sim/Results")
+
+#load necessary packages
+dependencies <- c('MASS', 'bain', 'metafor', 'lme4', 'mvtnorm')
+lapply(dependencies, function(x){library(x, character.only = T)})
+
+# Check package versions
+versions <- c(
+  compareVersion(as.character(packageVersion("bain")), "0.2.8"),
+  compareVersion(as.character(packageVersion("MASS")), "7.3.55"))
+if(!all(versions == 0)) stop("Using the incorrect version of one or more packages.")
+
+# Load simulation functions from source -----------------------------------
+# source('sim/functions.R')
+
+# set conditions for simulation
+hyper_parameters<-list(
+  ndataset = 1:1000,            # number of replications per condition
+  es = c(0, 0.1, .2),        # true effect size = true correlation with outcome
+  errorsd = c(0.81, .5, 0),   # corresponds to reliability of .6, .8, and 1
+  n = c(20, 80, 200, 500),             # mean sample size per group
+  k = c(2, 3, 10),               # number of groups
+  hyp_val = c(0.1)    # thresholds for informative hypotheses
+)
+
+# Create hypergrid with simulation parameters and save it as .RData file extension
+summarydata <- expand.grid(hyper_parameters, stringsAsFactors = FALSE)
+set.seed(6164900)
+summarydata$seed <- sample(1:.Machine$integer.max, nrow(summarydata))
+saveRDS(summarydata, file = "./sim/summarydata.RData")
+# summarydata<-readRDS("./sim/summarydata.RData")
+
+# prepare parallel processing
+library(doSNOW)
+nclust <- parallel::detectCores() 
+cl <- makeCluster(nclust) 
+registerDoSNOW(cl) 
+
+# add progression bar
+pb <- txtProgressBar(min = 0, max = nrow(summarydata), style = 3)
+opts <- list(progress = function(n) setTxtProgressBar(pb, n))
+
+# run simulation
+tab <- foreach(rownum = 1:nrow(summarydata), .options.snow = opts, .packages = c("bain", "mvtnorm", "lme4", "metafor"), .combine = rbind) %dopar% {
+  # Set seed
+  attach(summarydata[rownum, ])
+  set.seed(seed)
+  
[…]

### 2022-05-09 Sim/functions.R

@@ -1,0 +1,94 @@
+# 
+# 
+# 
+# BFs <- function(es, n, hyp_val, k, errorsd){
+#   
+#   # create dataframe for each group
+#   dfs <- lapply(1:k, function(i){
+#     df <- rmvnorm(n, sigma = matrix(c(1, es, es, 1), nrow = 2))
+#     df + rnorm(2*n, sd = errorsd)
+#   })
+#   
+#   # obtain estimates for correlations and their standard errors for every dataset
+#   res <- sapply(dfs, function(x){
+#     est <- cor(x)[2,1]                        #estimate for correlation between x and y
+#     se_est <- sqrt((1 - est^2)/(n - 2)) #estimate for standard error = sqrt((1-r^2)/df) with df = N - 2
+#     c(est, se_est)               #return estimate, se and sample size of particular set
+#   })
+#   
+#   # Classic approach
+#   classic_allsig <- all(apply(res, 2, function(x){
+#     (x[1]-hyp_val)/x[2] > 1.644854
+#   }))
+#   
+#   # necessary naming for bain and further preparing
+#   colnames(res) <- paste0('r', 1:k)
+#   sig <- lapply(res[2,], matrix)    # make list of covariance matrices for the datasets, as shown in the Bain vignette
+#   #ngroup <- rep(n, k)       # obtain sample size per group
+#   
+#   #run bf_individual to extract product bf and geometric product bf
+#   bf_individual <- lapply(paste0(colnames(res), ">", hyp_val), # for every group, we hypothesize that r_k > hyp_val
+#                           bain,                 # call bain
+#                           x = res[1,],          # estimates
+#                           Sigma = sig,          # all rho's are assumed to be independent
+#                           n = rep(n, k),           # pass the named vector of sample sizes per group to bain
+#                           group_parameters = 1, # every group k has 1 parameter which is rho_xy
+#                           joint_parameters = 0) # they do not share parameters 
+# 
+#   # extract BF_ic and BF_iu for the parameter of every group
+#   BFs <- t(sapply(bf_individual, function(x){
+#     c(x$fit$BF.c[1], x$fit$BF.u[1])
+#   }))
+# 
[…]

### 2022-05-09 Sim_Eli/Simulatie_EliV2.R

@@ -1,155 +1,0 @@
-#clear console and env
-rm(list=ls(all.names = T))
-cat("\014")
-
-# Delete folder where previous results are in and create new 'Results' folder
-unlink("./sim_Eli/Results", recursive = T)
-dir.create("./Sim_Eli/Results")
-
-#load necessary packages
-dependencies <- c('MASS', 'bain')
-lapply(dependencies, function(x){library(x, character.only = T)})
-
-# Check package versions
-versions <- c(
-  compareVersion(as.character(packageVersion("bain")), "0.2.8"),
-  compareVersion(as.character(packageVersion("MASS")), "7.3.55"))
-if(!all(versions == 0)) stop("Using the incorrect version of one or more packages.")
-
-# Load simulation functions from source -----------------------------------
-# source('Sim_Eli/functions.R')
-
-# set conditions for simulation
-hyper_parameters<-list(
-  ndataset = 1:1000,            # number of replications per condition
-  es = c(0, 0.1, .2),        # true effect size = true correlation with outcome
-  errorsd = c(0.81, .5, 0),   # corresponds to reliability of .6, .8, and 1
-  n = c(20, 80, 200, 500),             # mean sample size per group
-  k = c(1, 3, 10),               # number of groups
-  hyp_val = c(0.1)    # thresholds for informative hypotheses
-)
-
-# Create hypergrid with simulation parameters and save it as .RData file extension
-summarydata <- expand.grid(hyper_parameters, stringsAsFactors = FALSE)
-set.seed(6164900)
-summarydata$seed <- sample(1:.Machine$integer.max, nrow(summarydata))
-saveRDS(summarydata, file = "./Sim_Eli/summarydata.RData")
-# summarydata<-readRDS("./Sim_Eli/summarydata.RData")
-
-# prepare parallel processing
-library(doSNOW)
-nclust <- parallel::detectCores() 
-cl <- makeCluster(nclust) 
-registerDoSNOW(cl) 
-
-# add progression bar
-pb <- txtProgressBar(min = 0, max = nrow(summarydata), style = 3)
-opts <- list(progress = function(n) setTxtProgressBar(pb, n))
-
-# run simulation
-tab <- foreach(rownum = 1:nrow(summarydata), .options.snow = opts, .packages = c("bain", "mvtnorm"), .combine = rbind) %dopar% {
-  # Set seed
-  attach(summarydata[rownum, ])
-  set.seed(seed)
-  
-  # create dataframe for each group
[…]

### 2022-05-09 Sim_Eli/functions.R

@@ -1,94 +1,0 @@
-# 
-# 
-# 
-# BFs <- function(es, n, hyp_val, k, errorsd){
-#   
-#   # create dataframe for each group
-#   dfs <- lapply(1:k, function(i){
-#     df <- rmvnorm(n, sigma = matrix(c(1, es, es, 1), nrow = 2))
-#     df + rnorm(2*n, sd = errorsd)
-#   })
-#   
-#   # obtain estimates for correlations and their standard errors for every dataset
-#   res <- sapply(dfs, function(x){
-#     est <- cor(x)[2,1]                        #estimate for correlation between x and y
-#     se_est <- sqrt((1 - est^2)/(n - 2)) #estimate for standard error = sqrt((1-r^2)/df) with df = N - 2
-#     c(est, se_est)               #return estimate, se and sample size of particular set
-#   })
-#   
-#   # Classic approach
-#   classic_allsig <- all(apply(res, 2, function(x){
-#     (x[1]-hyp_val)/x[2] > 1.644854
-#   }))
-#   
-#   # necessary naming for bain and further preparing
-#   colnames(res) <- paste0('r', 1:k)
-#   sig <- lapply(res[2,], matrix)    # make list of covariance matrices for the datasets, as shown in the Bain vignette
-#   #ngroup <- rep(n, k)       # obtain sample size per group
-#   
-#   #run bf_individual to extract product bf and geometric product bf
-#   bf_individual <- lapply(paste0(colnames(res), ">", hyp_val), # for every group, we hypothesize that r_k > hyp_val
-#                           bain,                 # call bain
-#                           x = res[1,],          # estimates
-#                           Sigma = sig,          # all rho's are assumed to be independent
-#                           n = rep(n, k),           # pass the named vector of sample sizes per group to bain
-#                           group_parameters = 1, # every group k has 1 parameter which is rho_xy
-#                           joint_parameters = 0) # they do not share parameters 
-# 
-#   # extract BF_ic and BF_iu for the parameter of every group
-#   BFs <- t(sapply(bf_individual, function(x){
-#     c(x$fit$BF.c[1], x$fit$BF.u[1])
-#   }))
-# 
[…]

### 2022-05-09 Sim_Eli/preliminary_analysis.R

@@ -1,66 +1,0 @@
-########################
-# Preliminary Analyses #
-########################
-library(data.table)
-rm(list=ls())
-
-# algorithms we used
-algorithms <- c("gpbf", "prodbf", "tbf")
-hyps <- c("_ic", "_iu")
-alg_names <- paste0(rep(algorithms, each = length(hyps)), hyps)
-
-# some preparation
-summarydata <- readRDS("./Sim_Eli/summarydata.RData")
-conditions <- c("ndataset", "es", "tau2", "n", "k", "hyp_val")
-res <- readRDS(file.path("Sim_Eli", "sim_results_2022-03-14.RData"))
-lc <- length(conditions)
-
-# obtain unique conditions
-unique_conds <- unique(summarydata[,2:lc])
-unique_conds <- t(apply(unique_conds, 1, function(x){
-  paste(paste0(colnames(unique_conds), " == ", x), collapse = " & ")
-}))
-
-# convert res to datatable and make outcome values numeric
-res <- as.data.table(res)
-res[, c(lc:ncol(res)):=lapply(.SD, as.numeric), .SDcols=c(lc:ncol(res))]
-
-# we will sample 16 conditions to plot
-N <- 16
-colors <- c('#c15cf7', '#dda7fa', '#5c8df7', '#acc5fa', '#f5bd3b', '#ffe7b0')
-
-# Sample 16 integers from 1 to length(unique_conds)
-set.seed(6164900)
-sample_cond <- sample(1:length(unique_conds), 16)
-
-
-# function to plot histogram of BFs of a condition which is called in the next function
-# while subsetting the res datatable.
-plotcond <- function(outcome, name, condition, col){
-  condition <- unlist(strsplit(condition, "&"))
-  condition <- gsub("==", "=", condition)
-  condition <- paste(condition, collapse = " ")
-  
-  propBF3 <- sum(outcome > 3) / length(outcome)
-  hist(outcome, main = condition, col = col,
-       xlab = paste0("median: ", round(median(outcome),2),
-                     ";    sd: ", round(sd(outcome),2), 
-                     ";    prop BF > 3: ", round(propBF3,2)),
-       cex.lab = 1.3)
-}
-
-
-# Plot histogram s of the calculated BayesFactors from 16 random conditions in a single grid
-plot16 <- function(name, line = -1.5, colr){
-  par(mfrow = c(4,4))
-  for(i in seq_along(sample_cond)){
[…]

### 2022-05-11 manuscript.Rmd
First draft manuscript

---
title             : "Tutorial: Aggregate evidence from heterogeneous replication studies using the product Bayes factor"
shorttitle        : "PRODUCT BAYES FACTOR"

author:
  - name: "Caspar J. Van Lissa"
    affiliation: "1,2"
    corresponding: yes
    address: "Padualaan 14, 3584CH Utrecht, The Netherlands"
    email: "c.j.vanlissa@uu.nl"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Formal Analysis
      - Funding acquisition
      - Methodology
      - Project administration
      - Software
      - Supervision
      - Writing – original draft
      - Writing – review & editing
  - name: "Eli-Boaz Clapper"
    affiliation: "1"
    role:
      - Formal Analysis
      - Writing – original draft
      - Writing – review & editing
affiliation:
  - id            : "1"
    institution   : "Utrecht University, dept. Methodology & Statistics"
  - id            : "2"
    institution   : "Open Science Community Utrecht"

authornote: |
  This is a preprint paper, generated from Git Commit # `r substr(gert::git_commit_id(),1,7)`. This work was funded by a NWO Veni Grant (NWO Grant Number VI.Veni.191G.090), awarded to the lead author.

abstract: |
  TODO

keywords          : "bayes factor, evidence synthesis, bayesian, meta-analysis"
wordcount         : "5356"

bibliography      : ["references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
knit: worcs::cite_essential
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
run_everything <- FALSE
library("papaja")
library(tidySEM)
library(kableExtra)
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, warning = FALSE)
```

Recent years have seen a crisis of confidence over the reliability of published results in psychology, and science more broadly [REF].
Replication research has come into focus as a solution to this crisis
and a way to derive knowledge that will stand the test of time.
In step with this interest in replication research,
research synthesis methods have blossomed.
These methods aggregate the findings of multiple studies,
either quantitatively [REF] or qualitatively [REF TMSR].
Between-studies heterogeneity is a key challenge in research synthesis.
Heterogeneity appears when studies examine the same research question in different laboratories, use idiosyncratic methods, and sample from distinct populations.

Meta-analysis is the most common type of research synthesis.
In meta-analysis, heterogeneity can be accounted for in four ways.
First, if studies are exact replications, one can assume that no heterogeneity exists and conduct a fixed-effect meta-analysis to estimate their common population effect.
Second, when differences between studies can be assumed to be random, random-effects meta-analysis can be used to estimate the mean of a distribution of population effects.
Third, when there are a few systematic differences between studies, these can be accounted for using meta-regression.
And finally, when there are many potential differences and it is not known beforehand which are relevant,
exploratory techniques like random forest meta-analysis and penalized metaregression can be used to identify relevant moderators.
Each of these approaches requires making different assumptions about the nature of heterogeneity.

An alternative approach that does not impose such assumptions is Bayesian evidence synthesis [BES, REF Kuiper 2013].
BES focuses on the aggregation of evidence in favor of an informative hypothesis $H_i$ across studies.
The amount of evidence for this hypothesis is expressed as a Bayes factor, BF.
A Bayes factor can be interpreed as the ratio of evidence in favor of $H_i$ divided by evidence against it.
Thus, BF = 10 means that the data provide ten times more support in favor of the hypothesis than against it.
These Bayes factors can be synthesized across studies by taking their product.
The resulting product Bayes factor PBF summarizes the total evidence for the hypothesis.
Note that other approaches to BES exist; for an overview, see REF Klugkist.

This tutorial paper introduces the first implementation of BES in userfriendly open source software.
A function `pbf()` was contributed to the `bain` R-package for Bayesian informative hypothesis evaluation, version XXX.
This paper presents a simulation study to validate the method and benchmark it against alternative evidence synthesis methods,
and illustrates several use cases through reproducible examples.

# Simulation study

The present simulation study set out to validate the PBF algorithm.
For each iteration of the simulation, we simulated a correlation coefficient in multiple samples.
This correlation coefficient was either equal to or larger than an informative hypothesis value.
In this study, the informative hypothesis was kept constant, and specified as $H_i: \rho > .1$.
We then computed a PBF to determine the amount of evidence in favor of $H_i$ across these multiple samples.
A PBF > 3 was used as a decision criterion to conclude that $H_i$ was supported.
As a benchmark for comparison, we used several other algorithms that might feasibly be used by researchers who intend to examine whether a hypothesis is true across several independent samples.
The first was a naive "vote count" algorithm,
which used one-sided z-tests to examine whether a null hypothesis corresponding to the informative hypothesis was rejected in the majority of samples, i.e.: $H_0: \rho = .1$.
The second was a random-effects meta-analysis (RMA), which is the standard in the field.
For this algorithm, the null-hypothesis was rejected if a 90% confidence interval for the overall effect size excluded $H_0$.
Third was an individual participant data (IPD) meta-analysis;
a multilevel model clustered by sample.
This algorithm, too, was evaluated using a 90% confidence interval for the overall effect size.

## Performance indicators

For each algorithm, a confusion matrix was obtained by tabulating inferential decisions made using the criteria described above against the population status of the hypothesis (true or false).
This matrix gives the number of decisions that were true positives (TP), true negatives (TN), false positives (FP) and false negatives (FN).
These quantities were summarized as sensitivity, $1-\frac{FN}{TP + FN}$, the probability of concluding the hypothesis is false given that it was truly false in the population,
and specificity, $1-\frac{FP}{FP + TN}$, the probability of concluding the hypothesis is true given that it was true in the population.
The overall performance was captured by the accuracy, which represents the total proportion of correct (true positive/true negative) decisions, $\frac{TP + TN}{TP+TN+FP+FN}$.

## Design factors

To examine performance in a range of realistic scenarios,
several design factors were manipulated:
The population effect size $\rho \in (-0.1, 0, 0.1, 0.2)$,
the sample size per group $n \in (20, 80, 200, 500)$,
the number of groups $k \in (2, 3, 10)$, where 2 is the minimum number of groups and 10 is a very large number of groups,
and the reliability of the two correlated variables, $\alpha \in (0.6, 0.8, 1.0)$, where 0.6 is the lowest reliability conventionally considered to be acceptable, and 1 represents an assumption of perfect reliability, as is the case when analyzing correlations between observed items or mean scale scores.
The design factors combined to produce `r out$conditions` unique conditions.
For all simulation conditions, 1000 data sets were generated.

```{r, echo = F}
library(data.table) # load in libraries
library(DT)
library(ggplot2)
rename <- c("prodbf" = "PBF", "gpbf" = "gPBF", "tbf" = "TBF", "ipd" = "IPD", "rma" = "RMA", "allsig" = "VC")

```

# Results

```{r tabconf, echo = F, message=F}
tab <- read.csv("confusion.csv", stringsAsFactors = FALSE)
names(tab)[1] <- "Metric"
tab$Metric <- gsub("^.+_", "", tab$Metric)
tab <- tab[!tab$Metric == "lr", !grepl("iu", names(tab))]
names(tab) <- gsub("_ic", "", names(tab))
names(tab)[match(names(rename), names(tab))] <- rename
tab <- tab[tab$Metric %in% c("sensitivity", "specificity", "accuracy"),]
rownames(tab) <- NULL
knitr::kable(tab, digits = 2, caption = "Marginal confusion matrix metrics.")
```
These results indicate that all algorithms except PBF had low sensitivity to detect a true effect.
In contrast, specificity was very high for all algorithms except PBF.
This suggests that the other algorithms classified most conditions as negatives (no effect found),
regardless of the existence of a population effect.
The PBF trades a loss of specificity for increased sensitivity,
and average levels of both were approximately equal.
As the PBF vastly outperformed other BFs, these were omitted from further analysis.

## Effect of simulation conditions

We examined the effect of simulation conditions on overall accuracy.

```{r tabeffect, echo = FALSE}
tab <- read.csv("effect_of_conditions.csv", stringsAsFactors = FALSE)
tab <- tab[, !grepl("(gpbf|tbf)", names(tab))]
tab <- tab[, c("condition", "prodbf_ic", "rma", "ipd", "allsig", "prodbf_ic.vs..rma", "ipd.vs..prodbf_ic", "allsig.vs..prodbf_ic")]
names(tab) <-  c("Factor", "PBF", "RMA", "IPD", "VC", "vs RMA", "vs IPD", "vs VC")

knitr::kable(tab, digits = 2, caption = "Partial eta squared of the effect of each design factor on accuracy for each algorithm and for the difference between PBF and all other algorithms (e.g., vs RMA).")
```
PBF performance was most impacted by sample size $n$, followed by the number of groups $k$, and residual heterogeneity.

### Effect of sample size

```{r fign, fig.cap="Mean performance by sample size"}
res <- readRDS("confusion_by_cond.RData")
res <- res[res$alg %in% c("ipd", "rma", "allsig", "prodbf_ic"), ]

df_plot <- melt(res, measure.vars = c("sensitivity", "specificity"),
             variable.name = "Outcome", value.name = "val")
df_plot <- df_plot[ ,list(mean=mean(val)), by=c("n", "Outcome", "alg")]
ggplot(df_plot, aes(x = n, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  labs(y = "Mean performance", x = "Sample size")
```

Figure \@ref(fig:fign) indicates that for PBF, both sensitivity and specificity increase with sample size.
The other algorithms also show increasing sensitivity, but not specificity, which is at a ceiling.
This difference explains the effect of errorsd on the diference between algorithms (see Table \@ref(tab:tabeffect)).

### Effect of number of groups

```{r, fig.cap="Mean performance by number of groups"}
df_plot <- melt(res, measure.vars = c("sensitivity", "specificity"),
             variable.name = "Outcome", value.name = "val")
df_plot <- df_plot[ ,list(mean=mean(val)), by=c("k", "Outcome", "alg")]
ggplot(df_plot, aes(x = k, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  labs(y = "Mean performance", x = "Number of groups")
```
The figure indicates that, for PBF at higher levels of k, lower sensitivity is exchanged for greater specificity.
The other algorithms do not show this pattern, as their specificity is at a ceiling.
This difference in pattern of effects explains why number of groups has a moderate effect on the difference between algorithms (see Table \@ref(tab:tabeffect)).
Only VC shows decreasing sensitivity with an increasing number of groups;
this is because the probability of obtaining any false negatives increases with the number of groups.

### Effect of residual heterogeneity

```{r, fig.cap="Mean performance by errorsd"}

# dat <- readRDS(file.path("Sim", "sim_results_2022-05-07.RData"))
# varsout <- c("ipd", "rma", "allsig", "prodbf_ic")
# varspred <- c("errorsd", "n", "k", "hyp_val")
# 
# df_plot <- dat[!es < .1,  lapply(.SD, function(var) {
#   as.vector(table(ordered(es > .1, levels = c("FALSE", "TRUE")), ordered(var > 3, levels = c("FALSE", "TRUE"))))
# }), .SDcols = varsout, by = "errorsd"]
# 
# tabres[, metric := rep(c("TN", "FN", "FP", "TP"), nrow(tabres)/4)]
# 
# tabres <- melt(tabres, measure.vars = varsout,
#              variable.name = "alg", value.name = "val")
# tabres <- dcast(tabres, errorsd + n + k + hyp_val + alg ~ metric, value.var = "val")
# tabres[, alpha := FP / (FP + TN)]
# tabres[, beta := FN / (TP + FN)]
# tabres[, sensitivity := 1-beta]
# tabres[, specificity := 1-alpha]
# tabres[, accuracy := (TP + TN) / (TP+TN+FP+FN)]

df_plot <- melt(res, measure.vars = c("sensitivity", "specificity"),
             variable.name = "Outcome", value.name = "val")
df_plot <- df_plot[ ,list(mean=mean(val)), by=c("errorsd", "Outcome", "alg")]
ggplot(df_plot, aes(x = errorsd, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  labs(y = "Mean performance")
```

The figure indicates that for PBF, at higher levels of residual heterogeneity, lower sensitivity is exchanged for greater specificity.
The other algorithms do not show this pattern, as their specificity is at a ceiling.
Their sensitivity decreases with residual heterogeneity,
and therefore, so does their overall performance.
This difference in pattern of effects explains why residual heterogeneity has a moderate effect on the difference between algorithms (see Table \@ref(tab:tabeffect)).
Note that, in contrast to other algorithms, the performance of PBF is less susceptible to residual heterogeneity.

# Discussion

Within the scope of this simulation, PBF had relatively better inferential properties than the other algorithms under consideration.
Although other algorithms had superior specificity, PBF had the highest levels of sensitivity.
Thus, PBF balanced the ability to accept an informative hypothesis in the presence of a true population effect with the ability to reject the hypothesis when there was no effect.
<!-- Conventional threshold for type 1 error: 5% -->
<!-- Conventional threshold for power: 80% -->
<!-- J. Cohen, Statistical Power Analysis for the -->
<!--  Behavioral Sciences, 2nd ed. (Erlbaum, Hillsdale, -->
<!--  NJ, 1988). This is the source of the system of power -->
<!--  analysis described here; the power values and sam -->
<!--  ple sizes of the illustrations derive from this book's -->
<!--  tables. -->
<!--  2. J -->

An important caveat is that none of the algorithms met conventional criteria for power (80%) and type I error (5%).
Overall, PBF had the best performance, with power of 76% and type I error of 24%.
Thus, if researchers intend to synthesize evidence for an informative hypothesis across heterogeneous studies, PBF may be the most suitable method - but its limitations should be acknowledged in applied research.

The overall performance of PBF increased most with increasing sample size.
With an increasing number of groups, slightly decreasing specificity was traded off for increasing sensitivity.
With increasing residual heterogeneity, increasing specificity was traded off for decreasing sensitivity.

# Results

# Discussion

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup

### 2022-05-11 manuscript.Rmd
Tweaks

@@
 These methods aggregate the findings of multiple studies,
-either quantitatively [REF] or qualitatively [REF TMSR].
-Between-studies heterogeneity is a key challenge in research synthesis.
+and thus enable drawing overarching conclusions across multiple studies.
+Both quantitative [REF] and qualitative [REF TMSR] research synthesis methods exist, but the present paper focuses on quantitative methods.
+
+A key challenge in quantitative research synthesis is dealing with between-studies heterogeneity.
 Heterogeneity appears when studies examine the same research question in different laboratories, use idiosyncratic methods, and sample from distinct populations.
-
-Meta-analysis is the most common type of research synthesis.
+Meta-analysis is the most common research synthesis method.
 In meta-analysis, heterogeneity can be accounted for in four ways.
@@
 And finally, when there are many potential differences and it is not known beforehand which are relevant,
-exploratory techniques like random forest meta-analysis and penalized metaregression can be used to identify relevant moderators.
+exploratory techniques like random forest meta-analysis and penalized meta-regression can be used to identify relevant moderators.
 Each of these approaches requires making different assumptions about the nature of heterogeneity.
@@
 For this algorithm, the null-hypothesis was rejected if a 90% confidence interval for the overall effect size excluded $H_0$.
-Third was an individual participant data (IPD) meta-analysis;
-a multilevel model clustered by sample.
-This algorithm, too, was evaluated using a 90% confidence interval for the overall effect size.
+Third was an individual participant data (IPD) meta-analysis.
+Like classic meta-analysis, IPD is a multilevel model, clustered by sample.
+In contrast to classic meta-analysis, however, IPD freely estimates variance at the first level, because raw data are available.
+The PBF can be estimated using either sufficient statistics (as in meta-analysis) or using raw data (as in IPD).
+It is thus informative to compare the PBF to both these methods.
+IPD was also evaluated using a 90% confidence interval for the overall effect size.
 
@@
 ```
+First, we examined overall model performance across conditions.
 These results indicate that all algorithms except PBF had low sensitivity to detect a true effect.

### 2022-05-12 Analysis/analysis.R
update 2022-05-12 22:35:27

@@ -1,6 +1,7 @@
 #dat <- readRDS("tabres.RData")
 library(data.table)
-dat <- readRDS(file.path("Sim", "sim_results_2022-05-08.RData"))
+outlist <- list()
+dat <- readRDS(file.path("Sim", "sim_results_2022-05-12.RData"))
 lapply_at <- function(var, truees) {
   results <- sapply(var, function(var) {
     table(ordered(truees > .1, levels = c("FALSE", "TRUE")), ordered(var > 3, levels = c("FALSE", "TRUE")))
@@ -10,42 +11,48 @@ lapply_at <- function(var, truees) {
                            USE.NAMES = FALSE)
   as.list(results)
 }
-varsout <- c("ipd", "rma", "allsig", "gpbf_ic", "gpbf_iu", "prodbf_ic", "prodbf_iu", "tbf_ic", "tbf_iu")
-varspred <- c("errorsd", "n", "k", "hyp_val")
+names(dat) <- c("ndataset", "es", "reliability", "n", "k", "hyp_val", "seed", "IPD", 
+                "RMA", "VC", "gpbf_ic", "gpbf_iu", "PBF", "prodbf_iu", 
+                "tbf_ic", "tbf_iu")
 
-# Example: per condition
-dat[!es < .1, lapply_at(.SD, truees = es), .SDcols = varsout, by = c("n", "errorsd")]
+varsout <- c("IPD", "RMA", "VC", "PBF")
+varspred <- c("reliability", "n", "k", "hyp_val")
 
+# Drop cases where es < .1; these just inflate true negatives for all algorithms
+dat <- dat[!es < .1, ]
+outlist$simreps <- nrow(dat)
 # Overall
 #res <- dat[!es < .1, lapply_at(.SD, truees = es), .SDcols = varsout]
-res <- dat[!es < .1, sapply(.SD, function(var) {
-  table(ordered(es > .1, levels = c("FALSE", "TRUE")), ordered(var > 3, levels = c("FALSE", "TRUE")))
-}), .SDcols = varsout]
-rownames(res) <- c("TN", "FN", "FP", "TP")
 
-tabres <- dat[!es < .1,  lapply(.SD, function(var) {
-  as.vector(table(ordered(es > .1, levels = c("FALSE", "TRUE")), ordered(var > 3, levels = c("FALSE", "TRUE"))))
-}), .SDcols = varsout, by = varspred]
[…]

### 2022-05-12 Sim/Simulatie_V2.R

@@ -3,8 +3,8 @@ rm(list=ls(all.names = T))
 cat("\014")
 
 # Delete folder where previous results are in and create new 'Results' folder
-unlink("./sim/Results", recursive = T)
-dir.create("./sim/Results")
+unlink("results", recursive = T)
+dir.create("results")
 
 #load necessary packages
 dependencies <- c('MASS', 'bain', 'metafor', 'lme4', 'mvtnorm')
@@ -22,7 +22,7 @@ if(!all(versions == 0)) stop("Using the incorrect version of one or more package
 # set conditions for simulation
 hyper_parameters<-list(
   ndataset = 1:1000,            # number of replications per condition
-  es = c(0, 0.1, .2),        # true effect size = true correlation with outcome
+  es = c(0.1, .2, .3, .4),        # true effect size = true correlation with outcome
   errorsd = c(0.81, .5, 0),   # corresponds to reliability of .6, .8, and 1
   n = c(20, 80, 200, 500),             # mean sample size per group
   k = c(2, 3, 10),               # number of groups
@@ -63,7 +63,7 @@ tab <- foreach(rownum = 1:nrow(summarydata), .options.snow = opts, .packages = c
   df_join <- data.frame(cbind(df_join, rep(1:k, each = n)))
   res_ipd <- lmer(formula = X1 ~ X2 + (1|X3),
                              data    = df_join) #to run the model
-  ci_ipd <- confint(res_ipd)
+  ci_ipd <- confint(res_ipd, level = .90)
   
   # obtain estimates for correlations and their standard errors for every dataset
   res <- sapply(dfs, function(x){
@@ -74,12 +74,12 @@ tab <- foreach(rownum = 1:nrow(summarydata), .options.snow = opts, .packages = c
   
   df_meta <- data.frame(t(res))
   
-  res_met <- rma(yi = df_meta[,1], sei = df_meta[,2])
+  res_met <- rma(yi = df_meta[,1], sei = df_meta[,2], level = 90)
 
   # Classic approach
-  classic_allsig <- all(apply(res, 2, function(x){
+  classic_votecount <- round(mean(apply(res, 2, function(x){
     (x[1]-hyp_val)/x[2] > 1.644854
-  }))
+  })))
   
   # necessary naming for bain and further preparing
   colnames(res) <- paste0('r', 1:k)
[…]

### 2022-05-12 manuscript.Rmd

@@
 Replication research has come into focus as a solution to this crisis
-and a way to derive knowledge that will stand the test of time.
+and a way to derive knowledge that will stand the test of time [see @lavelleWhenCrisisBecomes2021].
 In step with this interest in replication research,
-research synthesis methods have blossomed.
+research synthesis methods have blossomed [REF].
 These methods aggregate the findings of multiple studies,
 and thus enable drawing overarching conclusions across multiple studies.
-Both quantitative [REF] and qualitative [REF TMSR] research synthesis methods exist, but the present paper focuses on quantitative methods.
+Both quantitative [e.g., @vanlissaSelectRelevantModerators2021] and qualitative [e.g., @vanlissaMappingPhenomenaRelevant2021] research synthesis methods exist.
+The present paper focuses on quantitative methods.
 
-A key challenge in quantitative research synthesis is dealing with between-studies heterogeneity.
-Heterogeneity appears when studies examine the same research question in different laboratories, use idiosyncratic methods, and sample from distinct populations.
-Meta-analysis is the most common research synthesis method.
-In meta-analysis, heterogeneity can be accounted for in four ways.
+A key challenge in quantitative research synthesis is dealing with between-studies heterogeneity [@higginsReevaluationRandomeffectsMetaanalysis2009].
+Heterogeneity appears when, for example, studies examine the same research question in different laboratories, use idiosyncratic methods, and sample from distinct populations.
+Meta-analysis is the most common research synthesis method [@borensteinIntroductionMetaAnalysis2009].
+In meta-analysis, heterogeneity can be accounted for in four ways [see @vanlissaSmallSampleMetaanalyses2020].
 First, if studies are exact replications, one can assume that no heterogeneity exists and conduct a fixed-effect meta-analysis to estimate their common population effect.
@@
 
-An alternative approach that does not impose such assumptions is Bayesian evidence synthesis [BES, REF Kuiper 2013].
+An alternative approach that does not impose such assumptions is Bayesian evidence synthesis [BES, @kuiperCombiningStatisticalEvidence2013].
 BES focuses on the aggregation of evidence in favor of an informative hypothesis $H_i$ across studies.
@@
 The resulting product Bayes factor PBF summarizes the total evidence for the hypothesis.
-Note that other approaches to BES exist; for an overview, see REF Klugkist.
+Note that other approaches to BES exist [see "Bayesian Evidence Synthesis" in @heckReviewApplicationsBayes2022].
 
 This tutorial paper introduces the first implementation of BES in userfriendly open source software.
-A function `pbf()` was contributed to the `bain` R-package for Bayesian informative hypothesis evaluation, version XXX.
+A function `pbf()` was contributed to the `bain` R-package for Bayesian informative hypothesis evaluation, version `r as.character(packageVersion("bain"))`.
 This paper presents a simulation study to validate the method and benchmark it against alternative evidence synthesis methods,
@@
 the number of groups $k \in (2, 3, 10)$, where 2 is the minimum number of groups and 10 is a very large number of groups,
-and the reliability of the two correlated variables, $\alpha \in (0.6, 0.8, 1.0)$, where 0.6 is the lowest reliability conventionally considered to be acceptable, and 1 represents an assumption of perfect reliability, as is the case when analyzing correlations between observed items or mean scale scores.
+and the reliability of the two correlated variables, $\alpha \in (0.6, 0.8, 1.0)$, where 0.6 is the lowest reliability conventionally considered to be acceptable, and 1 represents perfect reliability, as is the assumed when analyzing correlations between observed items or scale scores.
 The design factors combined to produce `r out$conditions` unique conditions.
@@
 ```
-PBF performance was most impacted by sample size $n$, followed by the number of groups $k$, and residual heterogeneity.
+PBF performance was most impacted by sample size $n$, followed by the number of groups $k$, and reliability.
 
@@
 The other algorithms also show increasing sensitivity, but not specificity, which is at a ceiling.
-This difference explains the effect of errorsd on the diference between algorithms (see Table \@ref(tab:tabeffect)).
+This difference explains the effect of reliability on the diference between algorithms (see Table \@ref(tab:tabeffect)).
 
@@
 
-### Effect of residual heterogeneity
+### Effect of reliability
 
-```{r, fig.cap="Mean performance by errorsd"}
+```{r, fig.cap="Mean performance by reliability"}
 
@@
   theme_bw() +
-  labs(y = "Mean performance")
+  labs(y = "Mean performance", x = "Reliability")
 ```
 
-The figure indicates that for PBF, at higher levels of residual heterogeneity, lower sensitivity is exchanged for greater specificity.
+The figure indicates that for PBF, at higher reliability, lower sensitivity is exchanged for greater specificity.
 The other algorithms do not show this pattern, as their specificity is at a ceiling.
-Their sensitivity decreases with residual heterogeneity,
+Their sensitivity decreases with lower reliability,
 and therefore, so does their overall performance.
-This difference in pattern of effects explains why residual heterogeneity has a moderate effect on the difference between algorithms (see Table \@ref(tab:tabeffect)).
-Note that, in contrast to other algorithms, the performance of PBF is less susceptible to residual heterogeneity.
+This difference in pattern of effects explains why reliability has a moderate effect on the difference between algorithms (see Table \@ref(tab:tabeffect)).
+Note that, in contrast to other algorithms, the performance of PBF is less susceptible to reliability.
 
@@
 With an increasing number of groups, slightly decreasing specificity was traded off for increasing sensitivity.
-With increasing residual heterogeneity, increasing specificity was traded off for decreasing sensitivity.
+With increasing reliability, increasing specificity was traded off for decreasing sensitivity.
 

### 2022-05-13 Analysis/Report.Rmd
update 2022-05-13 19:29:25

@@
 names(tab) <- gsub("_ic", "", names(tab))
-names(tab)[match(names(rename), names(tab))] <- rename
+# names(tab)[match(names(rename), names(tab))] <- rename
 tab <- tab[tab$Metric %in% c("sensitivity", "specificity", "accuracy"),]
@@
 tab <- tab[, !grepl("(gpbf|tbf)", names(tab))]
-tab <- tab[, c("condition", "prodbf_ic", "rma", "ipd", "allsig", "prodbf_ic.vs..rma", "ipd.vs..prodbf_ic", "allsig.vs..prodbf_ic")]
-names(tab) <-  c("Factor", "PBF", "RMA", "IPD", "VC", "vs RMA", "vs IPD", "vs VC")
-
+tab <- tab[, grepl("^\\w+$", names(tab)) | (grepl("\\.{2}", names(tab)) & grepl("PBF", names(tab)))]
+names(tab)[grepl("\\.{2}", names(tab))] <- paste0("vs ", gsub("(\\.|PBF|vs)", "", names(tab)[grepl("\\.{2}", names(tab))]))
 knitr::kable(tab, digits = 2, caption = "Partial eta squared of the effect of each design factor on accuracy for each algorithm and for the difference between PBF and all other algorithms (e.g., vs RMA).")
@@
 res <- readRDS("../confusion_by_cond.RData")
-res <- res[res$alg %in% c("ipd", "rma", "allsig", "prodbf_ic"), ]
+res <- res[res$alg %in% c("IPD", "RMA", "VC", "PBF"), ]
 
@@
 ```{r, fig.cap="Mean performance by errorsd"}
-
-# dat <- readRDS(file.path("../Sim", "sim_results_2022-05-07.RData"))
-# varsout <- c("ipd", "rma", "allsig", "prodbf_ic")
-# varspred <- c("errorsd", "n", "k", "hyp_val")
-# 
-# df_plot <- dat[!es < .1,  lapply(.SD, function(var) {
-#   as.vector(table(ordered(es > .1, levels = c("FALSE", "TRUE")), ordered(var > 3, levels = c("FALSE", "TRUE"))))
-# }), .SDcols = varsout, by = "errorsd"]
-# 
-# tabres[, metric := rep(c("TN", "FN", "FP", "TP"), nrow(tabres)/4)]
-# 
-# tabres <- melt(tabres, measure.vars = varsout,
-#              variable.name = "alg", value.name = "val")
-# tabres <- dcast(tabres, errorsd + n + k + hyp_val + alg ~ metric, value.var = "val")
-# tabres[, alpha := FP / (FP + TN)]
-# tabres[, beta := FN / (TP + FN)]
-# tabres[, sensitivity := 1-beta]
-# tabres[, specificity := 1-alpha]
-# tabres[, accuracy := (TP + TN) / (TP+TN+FP+FN)]
-
 df_plot <- melt(res, measure.vars = c("sensitivity", "specificity"),
              variable.name = "Outcome", value.name = "val")
-df_plot <- df_plot[ ,list(mean=mean(val)), by=c("errorsd", "Outcome", "alg")]
-ggplot(df_plot, aes(x = errorsd, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
+df_plot <- df_plot[ ,list(mean=mean(val)), by=c("reliability", "Outcome", "alg")]
+ggplot(df_plot, aes(x = reliability, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
   geom_point() +

### 2022-05-13 Analysis/analysis.R

@@ -1,7 +1,7 @@
 #dat <- readRDS("tabres.RData")
 library(data.table)
 outlist <- list()
-dat <- readRDS(file.path("Sim", "sim_results_2022-05-12.RData"))
+dat <- readRDS(file.path("Sim", "sim_results_2022-05-08.RData"))
 lapply_at <- function(var, truees) {
   results <- sapply(var, function(var) {
     table(ordered(truees > .1, levels = c("FALSE", "TRUE")), ordered(var > 3, levels = c("FALSE", "TRUE")))
@@ -23,9 +23,9 @@ dat <- dat[!es < .1, ]
 outlist$simreps <- nrow(dat)
 # Overall
 #res <- dat[!es < .1, lapply_at(.SD, truees = es), .SDcols = varsout]
-
-
-tabres <- lapply(c(.2, .3), function(thises){
+esvals <- unique(dat$es)
+esvals <- esvals[!esvals == .1]
+tabres <- lapply(esvals, function(thises){
   out <- dat[es %in% c(.1, thises),  lapply(.SD, function(var) {
     as.vector(table(ordered(es > .1, levels = c("FALSE", "TRUE")), ordered(var > 3, levels = c("FALSE", "TRUE"))))
   }), .SDcols = varsout, by = varspred]

### 2022-05-13 Sim/Simulatie_V2.R
update 2022-05-13 19:33:50

@@ -22,7 +22,7 @@ if(!all(versions == 0)) stop("Using the incorrect version of one or more package
 # set conditions for simulation
 hyper_parameters<-list(
   ndataset = 1:1000,            # number of replications per condition
-  es = c(0.1, .2, .3, .4),        # true effect size = true correlation with outcome
+  es = c(0.1, .2),        # true effect size = true correlation with outcome
   errorsd = c(0.81, .5, 0),   # corresponds to reliability of .6, .8, and 1
   n = c(20, 80, 200, 500),             # mean sample size per group
   k = c(2, 3, 10),               # number of groups

### 2022-05-13 manuscript.Rmd
update 2022-05-13 20:25:20

@@
 r_refs("r-references.bib")
+out <- readRDS("out.RData")
 ```
@@
 For each iteration of the simulation, we simulated a correlation coefficient in multiple samples.
-This correlation coefficient was either equal to or larger than an informative hypothesis value.
-In this study, the informative hypothesis was kept constant, and specified as $H_i: \rho > .1$.
-We then computed a PBF to determine the amount of evidence in favor of $H_i$ across these multiple samples.
+The informative hypothesis was kept constant at $H_i: \rho > .1$.
+All simulation conditions were evaluated once in the presence of a true population effect,
+defined as $\rho = .2$,
+and once in the presence of a null effect, defined as $\rho = .1$.
 A PBF > 3 was used as a decision criterion to conclude that $H_i$ was supported.
 As a benchmark for comparison, we used several other algorithms that might feasibly be used by researchers who intend to examine whether a hypothesis is true across several independent samples.
-The first was a naive "vote count" algorithm,
-which used one-sided z-tests to examine whether a null hypothesis corresponding to the informative hypothesis was rejected in the majority of samples, i.e.: $H_0: \rho = .1$.
+The first was a "vote count" algorithm,
+which is common but not considered to be good practice [REF].
+This approach used one-sided z-tests to examine whether a null hypothesis corresponding to the informative hypothesis was rejected in the majority of samples, i.e.: $H_0: \rho = .1$.
 The second was a random-effects meta-analysis (RMA), which is the standard in the field.
@@
 several design factors were manipulated:
-The population effect size $\rho \in (-0.1, 0, 0.1, 0.2)$,
-the sample size per group $n \in (20, 80, 200, 500)$,
+The sample size per group $n \in (20, 80, 200, 500)$,
 the number of groups $k \in (2, 3, 10)$, where 2 is the minimum number of groups and 10 is a very large number of groups,
 and the reliability of the two correlated variables, $\alpha \in (0.6, 0.8, 1.0)$, where 0.6 is the lowest reliability conventionally considered to be acceptable, and 1 represents perfect reliability, as is the assumed when analyzing correlations between observed items or scale scores.
-The design factors combined to produce `r out$conditions` unique conditions.
+The design factors combined to produce `r out$simreps/1000` unique conditions.
 For all simulation conditions, 1000 data sets were generated.
@@
 library(ggplot2)
-rename <- c("prodbf" = "PBF", "gpbf" = "gPBF", "tbf" = "TBF", "ipd" = "IPD", "rma" = "RMA", "allsig" = "VC")
+# rename <- c("prodbf" = "PBF", "gpbf" = "gPBF", "tbf" = "TBF", "ipd" = "IPD", "rma" = "RMA", "allsig" = "VC")
 
@@
 We examined the effect of simulation conditions on overall accuracy.
-
 ```{r tabeffect, echo = FALSE}
@@
 res <- readRDS("confusion_by_cond.RData")
-res <- res[res$alg %in% c("ipd", "rma", "allsig", "prodbf_ic"), ]
 
@@
              variable.name = "Outcome", value.name = "val")
-df_plot <- df_plot[ ,list(mean=mean(val)), by=c("n", "Outcome", "alg")]
+df_plot <- as.data.frame(df_plot)
+df_plot <- aggregate(df_plot$val, df_plot[c("n", "Outcome", "alg")], FUN=mean)
+names(df_plot)[ncol(df_plot)] <- "mean"
+# df_plot <- df_plot[ ,list(mean=mean(val)), by=c("n", "Outcome", "alg")]
 ggplot(df_plot, aes(x = n, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
@@
 ```{r, fig.cap="Mean performance by number of groups"}
-df_plot <- melt(res, measure.vars = c("sensitivity", "specificity"),
+df_plot <- data.table::melt(res, measure.vars = c("sensitivity", "specificity"),
              variable.name = "Outcome", value.name = "val")
-df_plot <- df_plot[ ,list(mean=mean(val)), by=c("k", "Outcome", "alg")]
+# df_plot <- df_plot[ ,list(mean=mean(val)), by=c("k", "Outcome", "alg")]
+df_plot <- as.data.frame(df_plot)
+df_plot <- aggregate(df_plot$val, df_plot[c("k", "Outcome", "alg")], FUN=mean)
+names(df_plot)[ncol(df_plot)] <- "mean"
 ggplot(df_plot, aes(x = k, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
@@
 ```{r, fig.cap="Mean performance by reliability"}
-
-# dat <- readRDS(file.path("Sim", "sim_results_2022-05-07.RData"))
-# varsout <- c("ipd", "rma", "allsig", "prodbf_ic")
-# varspred <- c("errorsd", "n", "k", "hyp_val")
-# 
-# df_plot <- dat[!es < .1,  lapply(.SD, function(var) {
-#   as.vector(table(ordered(es > .1, levels = c("FALSE", "TRUE")), ordered(var > 3, levels = c("FALSE", "TRUE"))))
-# }), .SDcols = varsout, by = "errorsd"]
-# 
-# tabres[, metric := rep(c("TN", "FN", "FP", "TP"), nrow(tabres)/4)]
-# 
-# tabres <- melt(tabres, measure.vars = varsout,
-#              variable.name = "alg", value.name = "val")
-# tabres <- dcast(tabres, errorsd + n + k + hyp_val + alg ~ metric, value.var = "val")
-# tabres[, alpha := FP / (FP + TN)]
-# tabres[, beta := FN / (TP + FN)]
-# tabres[, sensitivity := 1-beta]
-# tabres[, specificity := 1-alpha]
-# tabres[, accuracy := (TP + TN) / (TP+TN+FP+FN)]
-
 df_plot <- melt(res, measure.vars = c("sensitivity", "specificity"),
              variable.name = "Outcome", value.name = "val")
-df_plot <- df_plot[ ,list(mean=mean(val)), by=c("errorsd", "Outcome", "alg")]
-ggplot(df_plot, aes(x = errorsd, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
+# df_plot <- df_plot[ ,list(mean=mean(val)), by=c("reliability", "Outcome", "alg")]
+df_plot <- as.data.frame(df_plot)
+df_plot <- aggregate(df_plot$val, df_plot[c("reliability", "Outcome", "alg")], FUN=mean)
+names(df_plot)[ncol(df_plot)] <- "mean"
+ggplot(df_plot, aes(x = reliability, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
   geom_point() +

### 2022-05-13 Analysis/analysis.R

@@ -118,6 +118,7 @@ stats <- apply(res, 2, function(x){
 stats <- stats[, order(stats["accuracy",], decreasing = T)]
 write.csv(stats, "confusion.csv")
 
+saveRDS(out, "out.RData")
 # library(ggplot2)
 # df_plot <- data.frame(Method = colnames(stats),
 #                       Sensitivity = stats["sensitivity", ],

### 2022-05-24 manuscript.Rmd
update 2022-05-14 10:04:28
Start contribution to manuscript
Added some information in the introduction and in the simulation study. Added a few comments as well.

--- 2022-05-12 manuscript.Rmd
@@
 
-Recent years have seen a crisis of confidence over the reliability of published results in psychology, and science more broadly [REF].
-Replication research has come into focus as a solution to this crisis
-and a way to derive knowledge that will stand the test of time [see @lavelleWhenCrisisBecomes2021].
-In step with this interest in replication research,
-research synthesis methods have blossomed [REF].
-These methods aggregate the findings of multiple studies,
-and thus enable drawing overarching conclusions across multiple studies.
-Both quantitative [e.g., @vanlissaSelectRelevantModerators2021] and qualitative [e.g., @vanlissaMappingPhenomenaRelevant2021] research synthesis methods exist.
-The present paper focuses on quantitative methods.
+Recent years have seen a crisis of confidence over the reliability of published results in psychology, and science more broadly [@brembs2018prestigious].
+Replication research has come into focus as a solution to this crisis and as a way to derive knowledge that will stand the test of time [see @lavelleWhenCrisisBecomes2021]. 
+For many, highly regarded published studies however, results could not be replicated [@brembs2018prestigious].
+A lack of validating replication research partly strengthened the belief that single studies results should solely be viewed as part of many studies studying the same research question [@asendorpf2016recommendations].
+Evidence over multiple studies must be aggregated to derive more reliable conclusions and within this paradigm, research synthesis methods have blossomed [@shadish2015meta; @nakagawa2019research].
+The present paper focuses on quantitative methods for research synthesis [e.g., @vanlissaSelectRelevantModerators2021], although qualitative [e.g., @vanlissaMappingPhenomenaRelevant2021] research synthesis methods also exist.
 
-A key challenge in quantitative research synthesis is dealing with between-studies heterogeneity [@higginsReevaluationRandomeffectsMetaanalysis2009].
-Heterogeneity appears when, for example, studies examine the same research question in different laboratories, use idiosyncratic methods, and sample from distinct populations.
-Meta-analysis is the most common research synthesis method [@borensteinIntroductionMetaAnalysis2009].
+The most common of the quantitative methods is the meta-analysis in which results over different studies are aggregated to conclude one overarching effect [@borensteinIntroductionMetaAnalysis2009].
+A key-challenge in meta-analysis is dealing with between-study heterogeneity [@higginsReevaluationRandomeffectsMetaanalysis2009]. 
+Whether replications are conceptual or exact in nature, differences in outcomes can be due to the research question being examined in different laboratories, using idiosyncratic methods or sampling from distinct populations.
+Such differences due to virtue of not being the exact same study lead to variability in the outcome that can both be attributed to random and systematic measurement error.
+
 In meta-analysis, heterogeneity can be accounted for in four ways [see @vanlissaSmallSampleMetaanalyses2020].
-First, if studies are exact replications, one can assume that no heterogeneity exists and conduct a fixed-effect meta-analysis to estimate their common population effect.
-Second, when differences between studies can be assumed to be random, random-effects meta-analysis can be used to estimate the mean of a distribution of population effects.
-Third, when there are a few systematic differences between studies, these can be accounted for using meta-regression.
-And finally, when there are many potential differences and it is not known beforehand which are relevant,
-exploratory techniques like random forest meta-analysis and penalized meta-regression can be used to identify relevant moderators.
-Each of these approaches requires making different assumptions about the nature of heterogeneity.
+First, if studies are exact replications, one may assume that all heterogeneity in the outcome exists because of random measurement error. 
+A fixed-effect meta-analysis can then conducted to estimate the common population effect. 
+Second, when heterogeneity between studies can be assumed to be random, random-effects meta-analysis can be used to estimate the mean of a distribution of population effects. 
+Third, when there are a few systematic differences between studies, meta-regression can be performed in which the variables that cause systematic differences in outcomes are controlled for. 
+And finally, when there are many potential variables that cause systematic differences and it is not known beforehand which are relevant, exploratory techniques like random forest meta-analysis and penalized meta-regression can be used to identify relevant moderators.
 
-An alternative approach that does not impose such assumptions is Bayesian evidence synthesis [BES, @kuiperCombiningStatisticalEvidence2013].
-BES focuses on the aggregation of evidence in favor of an informative hypothesis $H_i$ across studies.
-The amount of evidence for this hypothesis is expressed as a Bayes factor, BF.
-A Bayes factor can be interpreed as the ratio of evidence in favor of $H_i$ divided by evidence against it.
+<!-- maybe provide more examples of why hetergeneity can be an issue in meta-analysis.
+For example that the aggregation of model parameters is not as straightforward because they can be in different units.while BF has no unit and can thus be easily aggregated -->
+
+Although each approach has its own specific assumptions, they are all regarding the nature of heterogeneity. 
+An alternative approach for evidence synthesis that does not impose such assumptions is Bayesian evidence synthesis (BES) [@kuiperCombiningStatisticalEvidence2013]. 
+Instead of aggregating parameter estimates of an effect over studies, which, due to heterogeneity, is not a straightforward endeavour, BES focuses on the aggregation of evidence in favor of an informative hypothesis $H_i$ across studies.
+The only assumptions BES has is that all study-specific hypotheses aim to provide evidence for the same underlying theoretical relationship, although the direction of effect, if any, is arbitrary.
+The amount of evidence for this hypothesis in each study and over all studies is expressed as a Bayes factor (BF). 
+A Bayes factor can be interpreted as the ratio of evidence in favor of $H_i$ divided by evidence against it. 
 Thus, BF = 10 means that the data provide ten times more support in favor of the hypothesis than against it. 
-These Bayes factors can be synthesized across studies by taking their product.
-The resulting product Bayes factor PBF summarizes the total evidence for the hypothesis.
-Note that other approaches to BES exist [see "Bayesian Evidence Synthesis" in @heckReviewApplicationsBayes2022].
+The BFs in each study are synthesized to one value by taking their product. 
+This value is known as the Product Bayes Factor (PBF) and it summarizes the total evidence for the hypothesis. 
+<!--BES thus requires its users to formulate an informative hypothesis about a theoretical relationship and evaluate it using the data in studies that examined that relationship -->. 
+In summary, BES a three stage method; 1.) formulate a informative hypothesis 2.) calculate the BF in each study examining the theoretical relationship and 3.) take the product of all BFs to obtain the PBF.
+Do note that other approaches to BES exist [see "Bayesian Evidence Synthesis" in @heckReviewApplicationsBayes2022].
 
-This tutorial paper introduces the first implementation of BES in userfriendly open source software.
+This tutorial paper introduces the first implementation of BES in user-friendly open source software.
 A function `pbf()` was contributed to the `bain` R-package for Bayesian informative hypothesis evaluation, version `r as.character(packageVersion("bain"))`.
@@
 The present simulation study set out to validate the PBF algorithm. 
-For each iteration of the simulation, we simulated a correlation coefficient in multiple samples.
-This correlation coefficient was either equal to or larger than an informative hypothesis value.
-In this study, the informative hypothesis was kept constant, and specified as $H_i: \rho > .1$.
-We then computed a PBF to determine the amount of evidence in favor of $H_i$ across these multiple samples.
+108 different plausible simulation conditions were specified, where each condition was repeated 1000 times, each time with different simulated data.
+For each of the 108 x 1000 iterations of the simulation, we simulated a correlation coefficient in multiple groups representing different studies examining a theoretical relationship.
+This correlation coefficient was either equal to or .1 units larger than an informative hypothesis value.
+This informative hypothesis was kept constant over all iterations, and was specified as $H_i: \rho > .1$.
+For every group in the iteration, the $H_i$ was tested against its complementary hypothesis: $H_c: \rho \leq .1$. 
+This allowed the calculation of the $BF_{ic}$ for all groups.
+The product was taken of the $BF_{ic}$ over all groups to compute the PBF and determine the amount of evidence in favor of $H_i$ across these multiple samples.
 A PBF > 3 was used as a decision criterion to conclude that $H_i$ was supported.
-As a benchmark for comparison, we used several other algorithms that might feasibly be used by researchers who intend to examine whether a hypothesis is true across several independent samples.
-The first was a naive "vote count" algorithm,
-which used one-sided z-tests to examine whether a null hypothesis corresponding to the informative hypothesis was rejected in the majority of samples, i.e.: $H_0: \rho = .1$.
+As a benchmark for comparing the PBF, we used several other algorithms that might feasibly be used by researchers who intend to examine whether a hypothesis is true across several independent studies.
+Among these were two other Bayesian aggregation methods:
+the gPBF, which takes the geometric mean of the $BF_{ic}$ over all groups, defined as $\sqrt[k]{PBF}$,
+and the Together Bayes Factor (TBF), defined as <!-- *formula for the TBF* -->
+Other algorithms included a naive "vote count" algorithm,
+which used one-sided z-tests to examine whether a null hypothesis corresponding to the informative hypothesis was rejected in the majority of groups, i.e.: $H_0: \rho = .1$.
 The second was a random-effects meta-analysis (RMA), which is the standard in the field.
@@
 Third was an individual participant data (IPD) meta-analysis.
-Like classic meta-analysis, IPD is a multilevel model, clustered by sample.
+Like classic meta-analysis, IPD is a multilevel model, clustered by group.
 In contrast to classic meta-analysis, however, IPD freely estimates variance at the first level, because raw data are available.
@@
 This matrix gives the number of decisions that were true positives (TP), true negatives (TN), false positives (FP) and false negatives (FN).
-These quantities were summarized as sensitivity, $1-\frac{FN}{TP + FN}$, the probability of concluding the hypothesis is false given that it was truly false in the population,
-and specificity, $1-\frac{FP}{FP + TN}$, the probability of concluding the hypothesis is true given that it was true in the population.
+These quantities were summarized as sensitivity, $1-\frac{FN}{TP + FN}$, the probability of concluding the hypothesis is false given that it was truly false in the population, <!-- i think sensitivity is the ability to detect an effect, given that the effect is truly there.
+I found that its formula is usually presented as TP/(TP+FN)-->
+and specificity, $1-\frac{FP}{FP + TN}$, the probability of concluding the hypothesis is true given that it was also true in the population.
+<!-- and specificity is the ability to not acknowledge an effect, given that there indeed is none-->
 The overall performance was captured by the accuracy, which represents the total proportion of correct (true positive/true negative) decisions, $\frac{TP + TN}{TP+TN+FP+FN}$.
@@
 the number of groups $k \in (2, 3, 10)$, where 2 is the minimum number of groups and 10 is a very large number of groups,
-and the reliability of the two correlated variables, $\alpha \in (0.6, 0.8, 1.0)$, where 0.6 is the lowest reliability conventionally considered to be acceptable, and 1 represents perfect reliability, as is the assumed when analyzing correlations between observed items or scale scores.
+and the reliability of the two correlated variables, $\alpha \in (0.6, 0.8, 1.0)$, where 0.6 is the lowest reliability conventionally considered to be acceptable, and 1 represents perfect reliability, as is assumed when analyzing correlations between observed items or scale scores.
 The design factors combined to produce `r out$conditions` unique conditions.
@@
 PBF performance was most impacted by sample size $n$, followed by the number of groups $k$, and reliability.
+<!-- should we also test for interaction effects? I seem to remember that there were some interesting effects of errorsd*k and n*k -->
 

### 2022-05-24 Sim/Simulatie_V2.R

@@ -3,8 +3,8 @@ rm(list=ls(all.names = T))
 cat("\014")
 
 # Delete folder where previous results are in and create new 'Results' folder
-unlink("./sim/Results", recursive = T)
-dir.create("./sim/Results")
+# unlink("./sim/Results", recursive = T)
+# dir.create("./sim/Results")
 
 #load necessary packages
 dependencies <- c('MASS', 'bain', 'metafor', 'lme4', 'mvtnorm')

### 2022-05-28 manuscript.Rmd
include conceptual difference BES and meta analysis
also, make start on pbf() function

@@
 In meta-analysis, heterogeneity can be accounted for in four ways [see @vanlissaSmallSampleMetaanalyses2020].
-First, if studies are exact replications, one may assume that all heterogeneity in the outcome exists because of random measurement error. 
-A fixed-effect meta-analysis can then conducted to estimate the common population effect. 
-Second, when heterogeneity between studies can be assumed to be random, random-effects meta-analysis can be used to estimate the mean of a distribution of population effects. 
+First, if studies are exact replications, one may assume that no heterogeneity in the outcome exists and a fixed-effect meta-analysis can be conducted to estimate the common population effect. 
+Second, when heterogeneity between studies can be assumed to be because of random measurement error, random-effects meta-analysis can be used to estimate the mean of a distribution of population effects.
 Third, when there are a few systematic differences between studies, meta-regression can be performed in which the variables that cause systematic differences in outcomes are controlled for. 
@@
 <!-- maybe provide more examples of why hetergeneity can be an issue in meta-analysis.
-For example that the aggregation of model parameters is not as straightforward because they can be in different units.while BF has no unit and can thus be easily aggregated -->
+For example that the aggregation of model parameters is not as straightforward because they can be in different units (e.g.
+correlation vs odds ratios).
+while BF has no unit and can thus be easily aggregated.
+Caspar, do you agree that such examples may strengthen the paper?
+-->
 
-Although each approach has its own specific assumptions, they are all regarding the nature of heterogeneity. 
+Although each approach has its own specific assumptions, they all are regarding the nature of heterogeneity. 
 An alternative approach for evidence synthesis that does not impose such assumptions is Bayesian evidence synthesis (BES) [@kuiperCombiningStatisticalEvidence2013]. 
 Instead of aggregating parameter estimates of an effect over studies, which, due to heterogeneity, is not a straightforward endeavour, BES focuses on the aggregation of evidence in favor of an informative hypothesis $H_i$ across studies.
+The evidence in favor of $H_i$ is relative to some alternative hypothesis $H_a$.
 The only assumptions BES has is that all study-specific hypotheses aim to provide evidence for the same underlying theoretical relationship, although the direction of effect, if any, is arbitrary.
-The amount of evidence for this hypothesis in each study and over all studies is expressed as a Bayes factor (BF). 
-A Bayes factor can be interpreted as the ratio of evidence in favor of $H_i$ divided by evidence against it. 
-Thus, BF = 10 means that the data provide ten times more support in favor of the hypothesis than against it. 
+The amount of evidence for this $H_i$ against an $H_a$ in each study and over all studies is expressed as a Bayes factor (BF). 
+A Bayes factor can be interpreted as the ratio of evidence in favor of $H_i$ divided by evidence in favor of $H_a$. 
+Thus, $BF_{ia}$ = 10 means that the data provide ten times more support in favor of $H_i$ than $H_a$. 
 The BFs in each study are synthesized to one value by taking their product. 
-This value is known as the Product Bayes Factor (PBF) and it summarizes the total evidence for the hypothesis. 
+This value is known as the Product Bayes Factor (PBF) and it summarizes the total evidence for the hypothesis against an alternative. 
 <!--BES thus requires its users to formulate an informative hypothesis about a theoretical relationship and evaluate it using the data in studies that examined that relationship -->. 
-In summary, BES a three stage method; 1.) formulate a informative hypothesis 2.) calculate the BF in each study examining the theoretical relationship and 3.) take the product of all BFs to obtain the PBF.
+In summary, BES is a three stage method where first both an $H_i$ and $H_a$ are formulated, second a $BF_{ia}$ is calculated in each study under consideration and, finally the product of all $BF_{ia}$'s is taken to obtain the PBF reflecting the overall support for $H_i$ over $H_a$.
 Do note that other approaches to BES exist [see "Bayesian Evidence Synthesis" in @heckReviewApplicationsBayes2022].
+
+What also should be kept in mind is that meta-analysis and BES do answer different research questions.
+Meta-analysis evaluates a certain theoretical relationship between variables. 
+It pools parameter estimates that evaluate this relationship over multiple studies to obtain a combined effect. 
+BES on the other hand compares a specific $H_i$ against some $H_a$.
+Aggregating support for this $H_i$ against $H_a$ over studies shows how robust the evaluation is against heterogeneity.
+The different answer from both methods thus make for complementary information.
+Meta-analysis provides support for or against a theoretical relationship in general, where BES provides support for a hypothesis about this theoretical relationship against some alternative.
+Whether only one of the methods or both should be used depends on the goal of the evidence synthesis.
 
@@
 108 different plausible simulation conditions were specified, where each condition was repeated 1000 times, each time with different simulated data.
-For each of the 108 x 1000 iterations of the simulation, we simulated a correlation coefficient in multiple groups representing different studies examining a theoretical relationship.
-This correlation coefficient was either equal to or .1 units larger than an informative hypothesis value.
+For each of the 108 x 1000 iterations of the simulation, we simulated a correlation coefficient, $\rho$, in multiple groups representing different studies examining a theoretical relationship.
+$\rho$ was either equal to or .1 units larger than an informative hypothesis value.
 This informative hypothesis was kept constant over all iterations, and was specified as $H_i: \rho > .1$.
@@
 This allowed the calculation of the $BF_{ic}$ for all groups.
-The product was taken of the $BF_{ic}$ over all groups to compute the PBF and determine the amount of evidence in favor of $H_i$ across these multiple samples.
+The product was taken of the $BF_{ic}$ over all groups to compute the PBF and determine the amount of evidence in favor of $H_i$ over $H_c$ across the multiple groups
 A PBF > 3 was used as a decision criterion to conclude that $H_i$ was supported.
@@
 and the Together Bayes Factor (TBF), defined as <!-- *formula for the TBF* -->
-Other algorithms included a naive "vote count" algorithm,
-which used one-sided z-tests to examine whether a null hypothesis corresponding to the informative hypothesis was rejected in the majority of groups, i.e.: $H_0: \rho = .1$.
+Other algorithms included a naive "vote count" algorithm, which used one-sided z-tests to examine whether a null hypothesis corresponding to the informative hypothesis was rejected in the majority of groups, i.e.: $H_0: \rho = .1$.
 The second was a random-effects meta-analysis (RMA), which is the standard in the field.

### 2022-05-28 UI Function/agg_bf.R

@@ -62,3 +62,25 @@ agg_bf(y~x, grouping = 'k', data = M, algorithm = 'tbf_ic', hypothesis = '= -0.2
 
 
 
+# pbf <- function(x, hypothesis, ...){
+#   cl <- match.call()
+#   
+#   if(all(sapply(x, inherits, what = "bain"))){ #if every object of x is a bain object...
+#     
+#     hypotheses <- sapply(x, function(y){y$hypotheses}) # check if hypotheses are equal
+#     if(length(unique(hypotheses)) != 1){
+#       stop("hypotheses of bain objects are not all equal")}
+#     
+#     BFs <- sapply(x, function(y){y$fit$BF.c[1]})
+#     res <- list(BFs = BFs, pbf = prod(BFs)) # obtain pbf ic, might need to change dependent on alternative hyp
+#     
+#   } else {                                  # if objects are not bain objects
+#     cl[[1L]] <- quote(bain)                 # change function call to bain
+#     
+#     cl[[x]] <- lapply(x, eval.parent, cl)   # run bain(x[[x], hypothesis]) op alle objecten in x
+#     cl[[1L]] <- quote(pbf)                  # change function call back to pbf
+#     eval.parent(cl)                         # call function pbf again, with x now being list of bain objects
+#   }
+#   return(res)
+# }
+# pbf(dat, "x=y")

### 2022-06-01 manuscript.Rmd
add pbf function

@@
 And finally, when there are many potential variables that cause systematic differences and it is not known beforehand which are relevant, exploratory techniques like random forest meta-analysis and penalized meta-regression can be used to identify relevant moderators.
-
-<!-- maybe provide more examples of why hetergeneity can be an issue in meta-analysis.
-For example that the aggregation of model parameters is not as straightforward because they can be in different units (e.g.
-correlation vs odds ratios).
-while BF has no unit and can thus be easily aggregated.
-Caspar, do you agree that such examples may strengthen the paper?
--->
 

### 2022-06-01 UI Function/agg_bf.R

@@ -60,27 +60,3 @@ agg_bf(y~x, grouping = 'k', data = M, algorithm = 'gpbf_ic', hypothesis = '= 0.2
 agg_bf(y~x, grouping = 'k', data = M, algorithm = 'prodbf_iu', hypothesis = '> -0.2')
 agg_bf(y~x, grouping = 'k', data = M, algorithm = 'tbf_ic', hypothesis = '= -0.2')
 
-
-
-# pbf <- function(x, hypothesis, ...){
-#   cl <- match.call()
-#   
-#   if(all(sapply(x, inherits, what = "bain"))){ #if every object of x is a bain object...
-#     
-#     hypotheses <- sapply(x, function(y){y$hypotheses}) # check if hypotheses are equal
-#     if(length(unique(hypotheses)) != 1){
-#       stop("hypotheses of bain objects are not all equal")}
-#     
-#     BFs <- sapply(x, function(y){y$fit$BF.c[1]})
-#     res <- list(BFs = BFs, pbf = prod(BFs)) # obtain pbf ic, might need to change dependent on alternative hyp
-#     
-#   } else {                                  # if objects are not bain objects
-#     cl[[1L]] <- quote(bain)                 # change function call to bain
-#     
-#     cl[[x]] <- lapply(x, eval.parent, cl)   # run bain(x[[x], hypothesis]) op alle objecten in x
-#     cl[[1L]] <- quote(pbf)                  # change function call back to pbf
-#     eval.parent(cl)                         # call function pbf again, with x now being list of bain objects
-#   }
-#   return(res)
-# }
-# pbf(dat, "x=y")

### 2022-06-14 manuscript.Rmd
Update manuscript.Rmd

--- 2022-05-24 manuscript.Rmd
@@
 Replication research has come into focus as a solution to this crisis and as a way to derive knowledge that will stand the test of time [see @lavelleWhenCrisisBecomes2021]. 
-For many, highly regarded published studies however, results could not be replicated [@brembs2018prestigious].
-A lack of validating replication research partly strengthened the belief that single studies results should solely be viewed as part of many studies studying the same research question [@asendorpf2016recommendations].
-Evidence over multiple studies must be aggregated to derive more reliable conclusions and within this paradigm, research synthesis methods have blossomed [@shadish2015meta; @nakagawa2019research].
+<!--For many, highly regarded published studies however, results could not be replicated [@brembs2018prestigious].-->
+<!--A lack of validating replication research partly strengthened the belief that single studies results should solely be viewed as part of many studies studying the same research question [@asendorpf2016recommendations].
+-->
+<!--Evidence over multiple studies must be aggregated to derive more reliable conclusions and within this paradigm, research synthesis methods have blossomed [@shadish2015meta; @nakagawa2019research].-->
+In step with this interest in replication research,
+research synthesis methods have blossomed [REF].
+These methods aggregate the findings of multiple studies,
+and thus enable drawing overarching conclusions across multiple studies.
 The present paper focuses on quantitative methods for research synthesis [e.g., @vanlissaSelectRelevantModerators2021], although qualitative [e.g., @vanlissaMappingPhenomenaRelevant2021] research synthesis methods also exist.
 
-The most common of the quantitative methods is the meta-analysis in which results over different studies are aggregated to conclude one overarching effect [@borensteinIntroductionMetaAnalysis2009].
-A key-challenge in meta-analysis is dealing with between-study heterogeneity [@higginsReevaluationRandomeffectsMetaanalysis2009]. 
-Whether replications are conceptual or exact in nature, differences in outcomes can be due to the research question being examined in different laboratories, using idiosyncratic methods or sampling from distinct populations.
-Such differences due to virtue of not being the exact same study lead to variability in the outcome that can both be attributed to random and systematic measurement error.
+A key challenge in quantitative research synthesis is dealing with between-studies heterogeneity [@higginsReevaluationRandomeffectsMetaanalysis2009].
+Heterogeneity appears when, for example, studies examine the same research question in different laboratories, use idiosyncratic methods, and sample from distinct populations.
+The most common quantitative research synthesis method is meta-analysis,
+in which results of different studies are aggregated to estimate an aggregate effect size [@borensteinIntroductionMetaAnalysis2009].
+In meta-analysis, heterogeneity can be accounted for in four ways [see @vanlissaSmallSampleMetaanalyses2020].
+First, if studies are exact replications, one may assume that no heterogeneity in the outcome exists and a fixed-effect meta-analysis can be conducted to estimate the common population effect. 
+Second, when heterogeneity between studies can be assumed to be because of random measurement error, random-effects meta-analysis can be used to estimate the mean of a distribution of population effects.
+Third, when there are a few systematic differences between studies, these can be accounted for using meta-regression.
+And finally, when there are many potential variables that cause systematic differences and it is not known beforehand which are relevant, exploratory techniques like random forest meta-analysis and penalized meta-regression can be used to identify relevant moderators.
+Each of these approaches requires making different assumptions about the nature of heterogeneity.
 
-In meta-analysis, heterogeneity can be accounted for in four ways [see @vanlissaSmallSampleMetaanalyses2020].
-First, if studies are exact replications, one may assume that all heterogeneity in the outcome exists because of random measurement error. 
-A fixed-effect meta-analysis can then conducted to estimate the common population effect. 
-Second, when heterogeneity between studies can be assumed to be random, random-effects meta-analysis can be used to estimate the mean of a distribution of population effects. 
-Third, when there are a few systematic differences between studies, meta-regression can be performed in which the variables that cause systematic differences in outcomes are controlled for. 
-And finally, when there are many potential variables that cause systematic differences and it is not known beforehand which are relevant, exploratory techniques like random forest meta-analysis and penalized meta-regression can be used to identify relevant moderators.
+An alternative approach for evidence synthesis that does not impose such assumptions is Bayesian evidence synthesis (BES) [@kuiperCombiningStatisticalEvidence2013]. 
+Instead of aggregating effect sizes, BES aggregates the evidence in favor of an informative hypothesis $H_i$ across studies.
+The amount of evidence for this hypothesis is expressed as a Bayes factor, BF.
+A Bayes factor can be interpreed as the ratio of evidence in favor of $H_i$ divided by evidence against it.
+Thus, BF = 10 means that the data provide ten times more support in favor of the hypothesis than against it.
+These Bayes factors can be synthesized across studies by taking their product.
+The resulting product Bayes factor (PBF) summarizes the total evidence for the hypothesis.
+Note that other approaches to BES exist [see "Bayesian Evidence Synthesis" in @heckReviewApplicationsBayes2022].
+The only assumption of BES is that all study-specific hypotheses provide evidence about the same underlying theoretical relationship.
 
-<!-- maybe provide more examples of why hetergeneity can be an issue in meta-analysis.
-For example that the aggregation of model parameters is not as straightforward because they can be in different units.while BF has no unit and can thus be easily aggregated -->
-
-Although each approach has its own specific assumptions, they are all regarding the nature of heterogeneity. 
-An alternative approach for evidence synthesis that does not impose such assumptions is Bayesian evidence synthesis (BES) [@kuiperCombiningStatisticalEvidence2013]. 
-Instead of aggregating parameter estimates of an effect over studies, which, due to heterogeneity, is not a straightforward endeavour, BES focuses on the aggregation of evidence in favor of an informative hypothesis $H_i$ across studies.
-The only assumptions BES has is that all study-specific hypotheses aim to provide evidence for the same underlying theoretical relationship, although the direction of effect, if any, is arbitrary.
-The amount of evidence for this hypothesis in each study and over all studies is expressed as a Bayes factor (BF). 
-A Bayes factor can be interpreted as the ratio of evidence in favor of $H_i$ divided by evidence against it. 
-Thus, BF = 10 means that the data provide ten times more support in favor of the hypothesis than against it. 
-The BFs in each study are synthesized to one value by taking their product. 
-This value is known as the Product Bayes Factor (PBF) and it summarizes the total evidence for the hypothesis. 
-<!--BES thus requires its users to formulate an informative hypothesis about a theoretical relationship and evaluate it using the data in studies that examined that relationship -->. 
-In summary, BES a three stage method; 1.) formulate a informative hypothesis 2.) calculate the BF in each study examining the theoretical relationship and 3.) take the product of all BFs to obtain the PBF.
-Do note that other approaches to BES exist [see "Bayesian Evidence Synthesis" in @heckReviewApplicationsBayes2022].
+Although meta-analysis and BES are both research synthesis methods, they answer different research questions.
+Meta-analysis estimates the value or distribution of a population effect size. 
+It pools estimates of this effect size across multiple studies to obtain an overall estimate of the effect size.
+It thus answers the question: Givern certain assumptions about between-studies heterogeneity, what is the average population effect size?
+BES, on the other hand, aggregates evidence in favor of an informative hypothesis across multiple studies.
+It thus answers the question: Do all these studies support the informative hypothesis?
+Both methods thus provide complementary information.
 
@@
 The present simulation study set out to validate the PBF algorithm.
-108 different plausible simulation conditions were specified, where each condition was repeated 1000 times, each time with different simulated data.
-For each of the 108 x 1000 iterations of the simulation, we simulated a correlation coefficient in multiple groups representing different studies examining a theoretical relationship.
-This correlation coefficient was either equal to or .1 units larger than an informative hypothesis value.
-This informative hypothesis was kept constant over all iterations, and was specified as $H_i: \rho > .1$.
-For every group in the iteration, the $H_i$ was tested against its complementary hypothesis: $H_c: \rho \leq .1$. 
-This allowed the calculation of the $BF_{ic}$ for all groups.
-The product was taken of the $BF_{ic}$ over all groups to compute the PBF and determine the amount of evidence in favor of $H_i$ across these multiple samples.
+For each iteration of the simulation, we simulated a correlation coefficient in multiple samples.
+The informative hypothesis was kept constant at $H_i: \rho > .1$.
+All simulation conditions were evaluated once in the presence of a true population effect,
+defined as $\rho = .2$,
+and once in the presence of a null effect, defined as $\rho = .1$.
 A PBF > 3 was used as a decision criterion to conclude that $H_i$ was supported.
-As a benchmark for comparing the PBF, we used several other algorithms that might feasibly be used by researchers who intend to examine whether a hypothesis is true across several independent studies.
-Among these were two other Bayesian aggregation methods:
-the gPBF, which takes the geometric mean of the $BF_{ic}$ over all groups, defined as $\sqrt[k]{PBF}$,
-and the Together Bayes Factor (TBF), defined as <!-- *formula for the TBF* -->
-Other algorithms included a naive "vote count" algorithm,
-which used one-sided z-tests to examine whether a null hypothesis corresponding to the informative hypothesis was rejected in the majority of groups, i.e.: $H_0: \rho = .1$.
+As a benchmark for comparison, we used several other algorithms that might feasibly be used by researchers who intend to examine whether a hypothesis is true across several independent samples.
+The first was a "vote count" algorithm,
+which is common but not considered to be good practice [REF].
+This approach used one-sided z-tests to examine whether a null hypothesis corresponding to the informative hypothesis was rejected in the majority of samples, i.e.: $H_0: \rho = .1$.
 The second was a random-effects meta-analysis (RMA), which is the standard in the field.
@@
 Third was an individual participant data (IPD) meta-analysis.
-Like classic meta-analysis, IPD is a multilevel model, clustered by group.
+Like classic meta-analysis, IPD is a multilevel model, clustered by sample.
 In contrast to classic meta-analysis, however, IPD freely estimates variance at the first level, because raw data are available.
@@
 I found that its formula is usually presented as TP/(TP+FN)-->
+<!--CJ: This is all the same, see https://en.wikipedia.org/wiki/Sensitivity_and_specificity -->
 and specificity, $1-\frac{FP}{FP + TN}$, the probability of concluding the hypothesis is true given that it was also true in the population.
@@
 <!-- should we also test for interaction effects? I seem to remember that there were some interesting effects of errorsd*k and n*k -->
+<!-- CJ: I don't think so, that might be more relevant for one of the papers that focused on the simulation study. For this tutorial, a simpler sim is fine. -->
 

### 2022-06-14 manuscript.Rmd

@@
 
-bibliography      : ["references.bib"]
+bibliography      : ["product_bayes.bib"]
 
@@
 several design factors were manipulated:
-The population effect size $\rho \in (-0.1, 0, 0.1, 0.2)$,
-the sample size per group $n \in (20, 80, 200, 500)$,
+The sample size per group $n \in (20, 80, 200, 500)$,
 the number of groups $k \in (2, 3, 10)$, where 2 is the minimum number of groups and 10 is a very large number of groups,

### 2022-06-14 UI Function/pbf.R
Implement more robust common hypothesis check

@@ -3,8 +3,7 @@ library(bain)
 # pbf moet ook werken voor meerdere hypotheses en de posterior model probabilities
 # moeten gereturned worden.
 # zodra functie compleet is, fork bain en implementeer
-pbf <- function(x, hypothesis, ...){
-  
+pbf <- function(x, ...){
   if(!all(sapply(x, inherits, what = "bain"))){ 
     cl <- match.call()
     cl[[1L]] <- quote(bain)  
@@ -17,11 +16,22 @@ pbf <- function(x, hypothesis, ...){
     cl[[1L]] <- quote(pbf)
     eval.parent(cl)
   }
-  
-  hypotheses <- sapply(x, function(y){y$hypotheses}) # check if hypotheses are equal
-  if(length(unique(hypotheses)) != 1){
-    stop("hypotheses of bain objects are not all equal")}
-  BFs <- sapply(x, function(y){y$fit$BF.c[1]})
+  # Merge the hypotheses from list item 1 and 2 into object merged
+  if(length(x) > 1){
+    hyps <- x[[1]]$hypotheses
+    for(i in length(x)-1){
+      browser()
+      hyps <- c(hyps, x[[i+1]]$hypotheses)
+      # Drop all non-duplicated hypotheses from merged
+      hyps <- hyps[duplicated(hyps)]
+      # If merged now has length 0, throw error
+      if(length(hyps) == 0){
+        stop("The objects passed to pbf() have no hypotheses in common.")
+      }
+      # Else, go back to step 1, but now merge merged with list item 3
+    }
+  }
+  BFs <- sapply(x, function(y){y$fit$BF.c[which(y$hypotheses %in% hyps)]})
   res <- list(BFs = BFs, pbf = prod(BFs)) # obtain pbf ic, might need to change dependent on alternative hyp
   return(res)
 }

### 2022-06-14 UI Function/pbf.R
pbf() does not require hypothesis when called on bain objects

@@ -20,7 +20,6 @@ pbf <- function(x, ...){
   if(length(x) > 1){
     hyps <- x[[1]]$hypotheses
     for(i in length(x)-1){
-      browser()
       hyps <- c(hyps, x[[i+1]]$hypotheses)
       # Drop all non-duplicated hypotheses from merged
       hyps <- hyps[duplicated(hyps)]
@@ -44,5 +43,6 @@ ttests <- lapply(1:4, function(i){
 pbf(ttests, "x=y")
 
 # and with bain-objects
+# CJ: This does not require specifying a hypothesis in pbf() call
 bains <- lapply(ttests,bain, hypothesis = "x=y")
-pbf(bains, "x=y")
+pbf(bains)

### 2022-06-14 UI Function/pbf.R
pbf() works for multiple hyps

@@ -30,8 +30,8 @@ pbf <- function(x, ...){
       # Else, go back to step 1, but now merge merged with list item 3
     }
   }
-  BFs <- sapply(x, function(y){y$fit$BF.c[which(y$hypotheses %in% hyps)]})
-  res <- list(BFs = BFs, pbf = prod(BFs)) # obtain pbf ic, might need to change dependent on alternative hyp
+  BFs <- do.call(cbind, lapply(x, function(y){y$fit$BF.c[which(y$hypotheses %in% hyps)]}))
+  res <- list(BFs = BFs, pbf = apply(BFs, 1, prod)) # obtain pbf ic, might need to change dependent on alternative hyp
   return(res)
 }
 
@@ -46,3 +46,24 @@ pbf(ttests, "x=y")
 # CJ: This does not require specifying a hypothesis in pbf() call
 bains <- lapply(ttests,bain, hypothesis = "x=y")
 pbf(bains)
+
+
+# Evaluation of informative hypotheses for an ANOVA
+# make a factor of variable site
+sesamesim$site <- as.factor(sesamesim$site)
+# execute an analysis of variance using lm() which, due to the -1, returns
+# estimates of the means of postnumb per group
+anov <- lm(postnumb~site-1,sesamesim[1:75,])
+# take a look at the estimated means and their names
+coef(anov)
+# set a seed value
+set.seed(100)
+# use the names to formulate and test hypotheses with bain
+results <- bain(anov, "site2=site1=site3=site4=site5; site2>site1=site3=site4=site5; site2>site5>site1>site3>site4")
+
+anov <- lm(postnumb~site-1,sesamesim[76:150,])
+results2 <- bain(anov, "site2=site1=site3=site4=site5; site2=site1=site3>site4=site5; site2>site5>site1>site3>site4")
+anov <- lm(postnumb~site-1,sesamesim[151:nrow(sesamesim),])
+results3 <- bain(anov, "site2=site1=site3=site4=site5; site2=site1=site3=site4>site5; site2>site5>site1>site3>site4")
+
+pbf(list(results, results2, results3))

### 2022-06-14 UI Function/pbf.R
add example with one hyp

@@ -48,19 +48,19 @@ bains <- lapply(ttests,bain, hypothesis = "x=y")
 pbf(bains)
 
 
-# Evaluation of informative hypotheses for an ANOVA
-# make a factor of variable site
 sesamesim$site <- as.factor(sesamesim$site)
-# execute an analysis of variance using lm() which, due to the -1, returns
-# estimates of the means of postnumb per group
-anov <- lm(postnumb~site-1,sesamesim[1:75,])
-# take a look at the estimated means and their names
-coef(anov)
-# set a seed value
-set.seed(100)
-# use the names to formulate and test hypotheses with bain
-results <- bain(anov, "site2=site1=site3=site4=site5; site2>site1=site3=site4=site5; site2>site5>site1>site3>site4")
 
+anov <- lm(postnumb~site-1,sesamesim[1:75,])
+results <- bain(anov, "site2=site1=site3=site4=site5; site2>site1=site3=site4=site5; site2>site5>site1>site3>site4")
+anov <- lm(postnumb~site-1,sesamesim[76:150,])
+results2 <- bain(anov, "site2=site1=site3=site4=site5; site2=site1=site3>site4=site5; site2>site5>site1>site3>site4")
+anov <- lm(postnumb~site-1,sesamesim[151:nrow(sesamesim),])
+results3 <- bain(anov, "site2=site1=site3=site4=site5; site2=site1=site3=site4>site5; site2>site5>site1>site3>site4")
+
+pbf(list(results, results2, results3))
+
+anov <- lm(postnumb~site-1,sesamesim[1:75,])
+results <- bain(anov, "site2=site1>site3=site4=site5; site2>site1=site3=site4=site5; site2>site5>site1>site3>site4")
 anov <- lm(postnumb~site-1,sesamesim[76:150,])
 results2 <- bain(anov, "site2=site1=site3=site4=site5; site2=site1=site3>site4=site5; site2>site5>site1>site3>site4")
 anov <- lm(postnumb~site-1,sesamesim[151:nrow(sesamesim),])

### 2022-06-15 pbfImprovement.R
Create pbfImprovement.R

@@ -1,0 +1,106 @@
+# Maybe I am thinking to hard about this, but it seems that the function does not yet deal with
+# hypotheses in bain objects potentially having different locations.
+# I'll show an example
+
+library(bain)
+
+pbf <- function(x, ...){
+  if(!all(sapply(x, inherits, what = "bain"))){ 
+    cl <- match.call()
+    cl[[1L]] <- quote(bain)  
+    for(i in (1:length(x))){
+      cl[['x']] <- x[[i]]
+      x[[i]] <- eval.parent(cl)
+    }
+    # cl[['x']] <- lapply(x, eval.parent, cl)
+    cl[['x']] <- x
+    cl[[1L]] <- quote(pbf)
+    eval.parent(cl)
+  }
+  # Merge the hypotheses from list item 1 and 2 into object merged
+  if(length(x) > 1){
+    hyps <- x[[1]]$hypotheses
+    for(i in length(x)-1){
+      hyps <- c(hyps, x[[i+1]]$hypotheses)
+      # Drop all non-duplicated hypotheses from merged
+      hyps <- hyps[duplicated(hyps)]
+      # If merged now has length 0, throw error
+      if(length(hyps) == 0){
+        stop("The objects passed to pbf() have no hypotheses in common.")
+      }
+      # Else, go back to step 1, but now merge merged with list item 3
+    }
+  }
+  BFs <- do.call(cbind, lapply(x, function(y){y$fit$BF.c[which(y$hypotheses %in% hyps)]}))
+  res <- list(BFs = BFs, pbf = apply(BFs, 1, prod)) # obtain pbf ic, might need to change dependent on alternative hyp
+  return(res)
+}
+
+
+# we create four lm objects
+set.seed(6164)
+lms <- lapply(1:4, function(i){
+  tt = as.data.frame(cbind(y = rnorm(1000,0,1), x = rnorm(1000, 0.2,1), z=rnorm(1000, 0.2,1)))
+  lm(tt$y ~ tt$z + tt$x)
+})
+pbf(lms, "x=z;z<0") #pbfs of 32957 and 529
+
+
+
+
+# now I call bain on the lm objects, but the change the order of the two hypotheses for the first and last two lm objects
+bains_order1 <- lapply(lms[1:2], bain, hypothesis = 'x=z;z<0')
+bains_order2 <- lapply(lms[3:4], bain, hypothesis = 'z<0;x=z')
+bains <- c(bains_order1,bains_order2)
+
+
+
+
[…]

### 2022-06-15 UI Function/pbf.R
update pbf(), correct ordering in hypothesis extraction

@@ -30,7 +30,14 @@ pbf <- function(x, ...){
       # Else, go back to step 1, but now merge merged with list item 3
     }
   }
-  BFs <- do.call(cbind, lapply(x, function(y){y$fit$BF.c[which(y$hypotheses %in% hyps)]}))
+  
+  BFs <- do.call(cbind, lapply(x, function(y){
+    hyps_in <- which(y$hypotheses %in% hyps)           # find which hypotheses exist in y
+    hyps_indices <- match(hyps[hyps_in], y$hypotheses) # find index of those hypotheses
+    y$fit$BF.c[hyps_indices]                           # get hypotheses in correct order
+  }))
+  rownames(BFs) <- paste0(sprintf('H%d: ', 1:length(hyps)),hyps) # give names
+  
   res <- list(BFs = BFs, pbf = apply(BFs, 1, prod)) # obtain pbf ic, might need to change dependent on alternative hyp
   return(res)
 }

### 2022-06-15 pbfImprovement.R

@@ -1,106 +1,0 @@
-# Maybe I am thinking to hard about this, but it seems that the function does not yet deal with
-# hypotheses in bain objects potentially having different locations.
-# I'll show an example
-
-library(bain)
-
-pbf <- function(x, ...){
-  if(!all(sapply(x, inherits, what = "bain"))){ 
-    cl <- match.call()
-    cl[[1L]] <- quote(bain)  
-    for(i in (1:length(x))){
-      cl[['x']] <- x[[i]]
-      x[[i]] <- eval.parent(cl)
-    }
-    # cl[['x']] <- lapply(x, eval.parent, cl)
-    cl[['x']] <- x
-    cl[[1L]] <- quote(pbf)
-    eval.parent(cl)
-  }
-  # Merge the hypotheses from list item 1 and 2 into object merged
-  if(length(x) > 1){
-    hyps <- x[[1]]$hypotheses
-    for(i in length(x)-1){
-      hyps <- c(hyps, x[[i+1]]$hypotheses)
-      # Drop all non-duplicated hypotheses from merged
-      hyps <- hyps[duplicated(hyps)]
-      # If merged now has length 0, throw error
-      if(length(hyps) == 0){
-        stop("The objects passed to pbf() have no hypotheses in common.")
-      }
-      # Else, go back to step 1, but now merge merged with list item 3
-    }
-  }
-  BFs <- do.call(cbind, lapply(x, function(y){y$fit$BF.c[which(y$hypotheses %in% hyps)]}))
-  res <- list(BFs = BFs, pbf = apply(BFs, 1, prod)) # obtain pbf ic, might need to change dependent on alternative hyp
-  return(res)
-}
-
-
-# we create four lm objects
-set.seed(6164)
-lms <- lapply(1:4, function(i){
-  tt = as.data.frame(cbind(y = rnorm(1000,0,1), x = rnorm(1000, 0.2,1), z=rnorm(1000, 0.2,1)))
-  lm(tt$y ~ tt$z + tt$x)
-})
-pbf(lms, "x=z;z<0") #pbfs of 32957 and 529
-
-
-
-
-# now I call bain on the lm objects, but the change the order of the two hypotheses for the first and last two lm objects
-bains_order1 <- lapply(lms[1:2], bain, hypothesis = 'x=z;z<0')
-bains_order2 <- lapply(lms[3:4], bain, hypothesis = 'z<0;x=z')
-bains <- c(bains_order1,bains_order2)
-
-
-
-
[…]

### 2022-06-15 UI Function/pbf.R
put calculation of BFs on single line

@@ -31,11 +31,7 @@ pbf <- function(x, ...){
     }
   }
   
-  BFs <- do.call(cbind, lapply(x, function(y){
-    hyps_in <- which(y$hypotheses %in% hyps)           # find which hypotheses exist in y
-    hyps_indices <- match(hyps[hyps_in], y$hypotheses) # find index of those hypotheses
-    y$fit$BF.c[hyps_indices]                           # get hypotheses in correct order
-  }))
+  BFs <- do.call(cbind, lapply(x, function(y){y$fit$BF.c[match(hyps[which(y$hypotheses %in% hyps)], y$hypotheses)]}))
   rownames(BFs) <- paste0(sprintf('H%d: ', 1:length(hyps)),hyps) # give names
   
   res <- list(BFs = BFs, pbf = apply(BFs, 1, prod)) # obtain pbf ic, might need to change dependent on alternative hyp

### 2022-06-15 UI Function/pbf.R
rewrite BFs calculation with match()

@@ -30,13 +30,13 @@ pbf <- function(x, ...){
       # Else, go back to step 1, but now merge merged with list item 3
     }
   }
-  
-  BFs <- do.call(cbind, lapply(x, function(y){y$fit$BF.c[match(hyps[which(y$hypotheses %in% hyps)], y$hypotheses)]}))
+  BFs <- do.call(cbind, lapply(x, function(y){y$fit$BF.c[match(hyps, y$hypotheses)]}))
   rownames(BFs) <- paste0(sprintf('H%d: ', 1:length(hyps)),hyps) # give names
   
   res <- list(BFs = BFs, pbf = apply(BFs, 1, prod)) # obtain pbf ic, might need to change dependent on alternative hyp
   return(res)
 }
+
 
 # try out with non-bain objects
 ttests <- lapply(1:4, function(i){

### 2022-06-16 UI Function/pbf.R
pbf returns data.frame with class pbf

@@ -31,9 +31,10 @@ pbf <- function(x, ...){
     }
   }
   BFs <- do.call(cbind, lapply(x, function(y){y$fit$BF.c[match(hyps, y$hypotheses)]}))
-  rownames(BFs) <- paste0(sprintf('H%d: ', 1:length(hyps)),hyps) # give names
-  
-  res <- list(BFs = BFs, pbf = apply(BFs, 1, prod)) # obtain pbf ic, might need to change dependent on alternative hyp
+  colnames(BFs) <- paste0("Sample ", 1:ncol(BFs))
+  res <- data.frame(PBF = apply(BFs, 1, prod), BFs)# obtain pbf ic, might need to change dependent on alternative hyp
+  rownames(res) <- paste0(sprintf('H%d: ', 1:length(hyps)),hyps) # give names
+  class(res) <- c("pbf", class(res))
   return(res)
 }
 
@@ -60,7 +61,7 @@ results2 <- bain(anov, "site2=site1=site3=site4=site5; site2=site1=site3>site4=s
 anov <- lm(postnumb~site-1,sesamesim[151:nrow(sesamesim),])
 results3 <- bain(anov, "site2=site1=site3=site4=site5; site2=site1=site3=site4>site5; site2>site5>site1>site3>site4")
 
-pbf(list(results, results2, results3))
+res_base <- pbf(list(results, results2, results3))
 
 anov <- lm(postnumb~site-1,sesamesim[1:75,])
 results <- bain(anov, "site2=site1>site3=site4=site5; site2>site1=site3=site4=site5; site2>site5>site1>site3>site4")
@@ -70,3 +71,15 @@ anov <- lm(postnumb~site-1,sesamesim[151:nrow(sesamesim),])
 results3 <- bain(anov, "site2=site1=site3=site4=site5; site2=site1=site3=site4>site5; site2>site5>site1>site3>site4")
 
 pbf(list(results, results2, results3))
+
+anov <- lm(postnumb~site-1,sesamesim[1:75,])
+results <- bain(anov, "site2=site1=site3=site4=site5; site2>site1=site3=site4=site5; site2>site5>site1>site3>site4")
+anov <- lm(postnumb~site-1,sesamesim[76:150,])
+results2 <- bain(anov, "site2=site1=site3>site4=site5; site2=site1=site3=site4=site5; site2>site5>site1>site3>site4")
+anov <- lm(postnumb~site-1,sesamesim[151:nrow(sesamesim),])
[…]

### 2022-06-16 UI Function/pbf.R
make beginning of test suite (Eli, please do the other ones!)

@@ -38,13 +38,18 @@ pbf <- function(x, ...){
   return(res)
 }
 
-
+library(testthat)
 # try out with non-bain objects
+set.seed(100)
 ttests <- lapply(1:4, function(i){
   tt = as.data.frame(cbind(y = rnorm(1000,0,1), x = rnorm(1000, 0.2,1)))
   t_test(tt$y, tt$x)
 })
-pbf(ttests, "x=y")
+test_that("pbf works for t_tests", {
+  expect_error({pbf(ttests, "x=y")}, NA)
+  res <- pbf(ttests, "x=y")
+  expect_equivalent(res[1,1], 1.644e-14)
+})
 
 # and with bain-objects
 # CJ: This does not require specifying a hypothesis in pbf() call

### 2022-06-16 manuscript.Rmd
Start Tutorial

@@
 
-# Results
+# Tutorial
+This tutorial about Bayesian Evidence Synthesis (BES) requires installation of `r R.Version()$version.string`, Rstudio version `r RStudio.Version()$version` and package `bain` version `r packageVersion('bain')`.
+This tutorial focusses on synthesizing evidence for an informative hypotheses using the Product Bayes Factor (pbf).
+To illustrate the method, the `sesamesim` dataset will be used which is part of `bain`.
+For a detailed description of the dataset, run `?bain::sesamesim`.
 
-# Discussion
+<!-- skeleton
+1.) explain calculation of single BF and what it means, refer to previous tutorial on the bayes factor
+2.) introduce the case where there are multiple studies and we want to evaluate a single hypothesis
+3.) introduce pbf(), its uses, outputs and interpretation
+-->
+
+### The Bayes Factor using bain.
+In conventional Null-hypothesis significance testing (NHST), as the name implies, data is analysed to generate evidence for or against the null-hypothesis.
+This method can only evaluate a single hypothesis, the null-hypothesis, and its output is dichotomous in nature: whether a theoretical relationship is probable or improbable to exist.
+However, there are many more hypotheses that can theoretically be tested and researchers may not be interested in the null-hypothesis.
+They may also not only be interested in whether the hypothesis is probable, but also by how much.
+In the Bayesian framework, the Bayes Factor (BF) is a test statistic that gives the amount of support for any possible desired hypothesis.
+Because the focus of this tutorial is on BES, this tutorial will only briefly introduce how to evaluate an hypothesis using the R package `bain` and interpret its output.
+For an extensive tutorial on how to evaluate informative hypotheses using `bain` see [@hoijtink2019tutorial].
+
+Let's say we are interested in whether the status of the site children grew up in (`site`) affects their score on a numbers test after watching Sesame Street for a year (`postnumb`).
+`postnumb` is a continuous variable which ranges from 0-63.
+`site` is a nominal variable indicating group membership with possible values 1 = disadvantaged inner city, 2 = advantaged suburban, 3 = advantaged rural, 4 = disadvantaged rural and 5 = disadvantaged Spanish speaking.
+In NHST we specify the following null-hypothesis:
+
+$H_0: \mu_{site1} =\mu_{site2} = \mu_{site3} = \mu_{site4} =\mu_{site5}$
+
+However, we may expect that the children growing up in an advantaged site will have a higher score than children growing up in a disdvantaged site. Our informative hypothesis is:
+
+$H_A: \mu_{site2},\mu_{site3} > \mu_{site1},\mu_{site4},\mu_{site5}$
+
+We can test the $H_0$ and $H_A$ simultaneously by running a linear regression on the data and passing the `lm` object with the specific hypotheses to bain.
+```{r}
+# load bain package
+library(bain)
+
+# convert `site` to nominal variable
+sesamesim$site <- as.factor(sesamesim$site)
+
+# specify model, allow meaningful zero-point
+fit <- lm(postnumb ~ site-1, data = sesamesim)
+
+fit_bain <- bain(fit, hypothesis = "site1=site2=site3=site4=site5; site2&site3>site1&site4&site5" )
+print(fit_bain)
+
+```
+<!-- Both BF.c and BF.u are close to zero, the probability of the HA being more probable than the Hc or Hu is small. <further interpretation>-->
+
+### Single vs multiple studies.
+In the case of multiple studies, it is possible to evaluate the hypotheses under consideration by synthesizing the evidence over all studies.
+In the example we split the `sesamesim` dataset up into three equal parts, to simulate three studies into the same subject.
+In this example, each result has the $H_0$ and $H_A$ in common, but contains one result-specific hypothesis. The `pbf()` function only gives output for the hypotheses that are present in all results.
+```{r}
+anov <- lm(postnumb~site-1,sesamesim[1:75,])
+results <- bain(anov, "site2=site1=site3=site4=site5; site2>site1=site3=site4=site5; site2&site3>site1&site4&site5")
+anov2 <- lm(postnumb~site-1,sesamesim[76:150,])
+results2 <- bain(anov, "site2=site1=site3=site4=site5; site2=site1=site3>site4=site5; site2&site3>site1&site4&site5")
+anov3 <- lm(postnumb~site-1,sesamesim[151:nrow(sesamesim),])
+results3 <- bain(anov, "site2=site1=site3=site4=site5; site2=site1=site3=site4>site5; site2&site3>site1&site4&site5")
+
+pbf(list(results, results2, results3))
+
+```
+The function returns the individual BF.c's of each study, as well as their product.
 

### 2022-06-27 manuscript.Rmd
expand tutorial

@@
 In conventional Null-hypothesis significance testing (NHST), as the name implies, data is analysed to generate evidence for or against the null-hypothesis.
-This method can only evaluate a single hypothesis, the null-hypothesis, and its output is dichotomous in nature: whether a theoretical relationship is probable or improbable to exist.
+This method can only evaluate a single hypothesis, the null-hypothesis, and its output is dichotomous in nature: whether a theoretical relationship is probable or improbable to exist given the data.
 However, there are many more hypotheses that can theoretically be tested and researchers may not be interested in the null-hypothesis.
@@
 `site` is a nominal variable indicating group membership with possible values 1 = disadvantaged inner city, 2 = advantaged suburban, 3 = advantaged rural, 4 = disadvantaged rural and 5 = disadvantaged Spanish speaking.
-In NHST we specify the following null-hypothesis:
+In NHST we specify the null-hypothesis that the mean `postnumb` is equal for all groups in  site`:
 
@@
 
-However, we may expect that the children growing up in an advantaged site will have a higher score than children growing up in a disdvantaged site. Our informative hypothesis is:
+However, we may have prior expectations that the children growing up in an advantaged site will have a higher score than children growing up in a disadvantaged site. Our informative hypothesis is:
 
-$H_A: \mu_{site2},\mu_{site3} > \mu_{site1},\mu_{site4},\mu_{site5}$
+$H_i: \mu_{site2},\mu_{site3} > \mu_{site1},\mu_{site4},\mu_{site5}$
 
@@
 
-fit_bain <- bain(fit, hypothesis = "site1=site2=site3=site4=site5; site2&site3>site1&site4&site5" )
+#specify Hypotheses to evaluate
+hypotheses <- "site2=site1=site3=site4=site5; site2&site3>site1&site4&site5"
+hypotheses <- "site2=site1=site3=site4=site5"
+
+fit_bain <- bain(fit, hypothesis = hypotheses)
 print(fit_bain)
@@
 ```
-<!-- Both BF.c and BF.u are close to zero, the probability of the HA being more probable than the Hc or Hu is small. <further interpretation>-->
+The output of `bain` seems to support neither hypothesis.
+The `fit` for both hypotheses is approximately zero which means that almost none of the data are in agreement with the restrictions the specified hypotheses pose.
+The `BF.c`, which is the Bayes Factor of the hypothesis versus its complement, is for both hypotheses close to zero.
+This implies that the complement of both hypotheses is $\frac{1}{BF.c}$ times more likely than the passed hypotheses.
+Finally, `PMPc`, the Posterior Model Probabilities of the specified hypotheses with $H_c$ included, is also close to zero for both specified hypothesis.
+This means that *out of* the specified hypotheses and their complement, the probability of $H_c$ being true is 99.9%.
 
 ### Single vs multiple studies.
-In the case of multiple studies, it is possible to evaluate the hypotheses under consideration by synthesizing the evidence over all studies.
-In the example we split the `sesamesim` dataset up into three equal parts, to simulate three studies into the same subject.
-In this example, each result has the $H_0$ and $H_A$ in common, but contains one result-specific hypothesis. The `pbf()` function only gives output for the hypotheses that are present in all results.
+Sometimes there is access to the data of not only a single, but multiple studies that all contain the necessary information to evaluate our hypothesis $H_i$.
+To evaluate $H_i$, it is possible to synthesize the support for $H_i$ over all studies using the PBF.
+In this example we split the `sesamesim` dataset up into three equal parts, to simulate three studies into the same topic.
+
 ```{r}
-anov <- lm(postnumb~site-1,sesamesim[1:75,])
-results <- bain(anov, "site2=site1=site3=site4=site5; site2>site1=site3=site4=site5; site2&site3>site1&site4&site5")
-anov2 <- lm(postnumb~site-1,sesamesim[76:150,])
-results2 <- bain(anov, "site2=site1=site3=site4=site5; site2=site1=site3>site4=site5; site2&site3>site1&site4&site5")
-anov3 <- lm(postnumb~site-1,sesamesim[151:nrow(sesamesim),])
-results3 <- bain(anov, "site2=site1=site3=site4=site5; site2=site1=site3=site4>site5; site2&site3>site1&site4&site5")
-
-pbf(list(results, results2, results3))
+sample1 <- sesamesim[1:75,]
+sample2 <- sesamesim[76:150,]
+sample3 <- sesamesim[151:nrow(sesamesim),]
 
 ```
-The function returns the individual BF.c's of each study, as well as their product.
+
+In the first use-case of `pbf()`, an `lm()` is ran for each sample. All `lm` objects are stored in a list which is passed to `pbf()`.
+`pbf()` supports lists of all objects that `bain()` also supports. All supported types are specified in bain vignette *REF*. One exception is the use of named vectors, which will be covered later.
+```{r}
+lm_sample1 <- lm(postnumb~site-1, sample1)
+lm_sample2 <- lm(postnumb~site-1, sample2)
+lm_sample3 <- lm(postnumb~site-1, sample3)
+
+lm_allsamples <- list(lm_sample1, lm_sample2, lm_sample3)
+
+pbf(lm_allsamples, hypotheses)
+```
+
+The function returns a `data.frame` containing the passed hypotheses in the rows and in the columns the PBF and the individual Bayes Factor for each sample individually.
+The `PBF` of both hypotheses is close to zero, meaning that the combined evidence for both hypotheses over all studies is low.
+When looking at the individual Bayes Factors, we also see that the PBF is lower than all of the Bayes Factors individually.
+For sample 3 the specified hypotheses are even more likely than their complement.
+This example illustrates another strength of the PBF; being able to evaluate a hypothesis more extremely than the most extreme individual evaluation.
+If all samples do not support a hypothesis, than the number of times the hypothesis is not supported should be included in its evaluation.
+This leads to a more extreme combined evaluation than any individual evaluation.
+Likewise, if there is very strong support in only a small minority of the samples, while the majority of samples shows no support at all, than it is likely that the hypothesis is not true.
+
+### using bain objects
+The `pbf()` function also works when objects are first converted to `bain` objects.
+In this specific example, each `bain` object corresponding to one of the three samples has the $H_0$ and $H_A$ in common, but contains one sample-specific hypothesis.
+The `pbf()` function only returns output for the hypotheses that are present in all results.
+When `bain` objects are passed to the `pbf()`, no additional hypotheses need to be passed to `pbf()` as they are already saved in the objects.
+
+```{r}
+bain_sample1 <- bain(lm_sample1, "site2=site1=site3=site4=site5; site2>site1=site3=site4=site5; site2&site3>site1&site4&site5")
+bain_sample2 <- bain(lm_sample2, "site2=site1=site3=site4=site5; site2=site1=site3>site4=site5; site2&site3>site1&site4&site5")
+bain_sample3 <- bain(lm_sample3, "site2=site1=site3=site4=site5; site2=site1=site3=site4>site5; site2&site3>site1&site4&site5")
+
+bain_allsamples <- list(bain_sample1, bain_sample2, bain_sample3)
+
+pbf(bain_allsamples)
+
+```
+
+As can be seen, the results are identical to the results in the previous example.
+This means that the sample-specific hypotheses have been left out and only the hypotheses that each sample has in common are retained.
+
+When there is no hypothesis that is shared by each sample, `pbf()` throws an error.
+```{r}
+bains_different_hyps <- list(
+  bain(lm_sample1, "site2=site1=site3=site4=site5"),
+  bain(lm_sample2, "site2=site1=site3>site4=site5"),
+  bain(lm_sample3, "site2&site3>site1&site4&site5")
+)
+
+tryCatch({
+  pbf(bains_different_hyps)},
+  error = function(e){message(e)}
+      )
+
+```
+The function returns an error stating that the samples have no hypothesis in common and thus the PBF cannot be calculated.
+
+### using named vectors
+Another specific use-case is when the raw data from different samples is not available.
+In this case, the researcher only has access to the parameter-values that are provided in the research papers.
+Because the `pbf()` function inherently converts objects to `bain` objects, each sample requires three pieces of information in order to be usable: the named parameter estimates, a covariance matrix of their standard errors and the sample size.
+First these values are extracted from all samples.
+A function is created that extracts is information which is then used to extract the necessary information from all samples.
+
+```{r}
+get_estimates <- function(sample){
+  number_sites <- length(levels(sesamesim$site))
+
+  sample_size <- nrow(sample)
+
+  site_estimates <- sapply(1:number_sites, function(i){
+    rbind(
+      mean(sample[sample['site'] == i, "postnumb"]),
+      sd(sample[sample['site'] == i, "postnumb"])
+      )
+    })
+
+  est_names <- sprintf("site%d", 1:number_sites)
+  site_means <- site_estimates[1,]
+  names(site_means) <- est_names
+
+  covmat <- diag((site_estimates[2,] / sqrt(sample_size))^2)
+  dimnames(covmat) <- list(est_names, est_names)
+  return(list(means = site_means, covmat = covmat, sample_size = sample_size))
+}
+
+ests_sample1 <- get_estimates(sample1)
+ests_sample2 <- get_estimates(sample2)
+ests_sample3 <- get_estimates(sample3)
+
+```
+
+Then, each object is converted to a `bain` object with these estimates
+```{r}
+bain_sample1_ests <- bain(x = ests_sample1[['means']], 
+                          Sigma = ests_sample1[['covmat']], 
+                          n = ests_sample1[['sample_size']], 
+                          hypothesis = hypotheses)
+
+bain_sample2_ests <- bain(x = ests_sample2[['means']], 
+                          Sigma = ests_sample2[['covmat']], 
+                          n = ests_sample2[['sample_size']], 
+                          hypothesis = hypotheses)
+
+bain_sample3_ests <- bain(x = ests_sample3[['means']], 
+                          Sigma = ests_sample3[['covmat']], 
+                          n = ests_sample3[['sample_size']], 
+                          hypothesis = hypotheses)
+
+bain_allsamples_ests = list(bain_sample1_ests, bain_sample2_ests, bain_sample3_ests)
+```
+
+And finally, on these objects, `pbf()` can be called.
+```{r}
+pbf(bain_allsamples_ests)
+```
+<!-- It is interesting that these estimates are not identical to the other examples... Is that because we do not have ALL information available in this use-case, or is something off in the script? -->
 

### 2022-06-27 manuscript.Rmd
minor changes

@@
 `site` is a nominal variable indicating group membership with possible values 1 = disadvantaged inner city, 2 = advantaged suburban, 3 = advantaged rural, 4 = disadvantaged rural and 5 = disadvantaged Spanish speaking.
-In NHST we specify the null-hypothesis that the mean `postnumb` is equal for all groups in  site`:
+In NHST we specify the null-hypothesis that the mean `postnumb` is equal for all groups in  `site`:
 
@@
 hypotheses <- "site2=site1=site3=site4=site5; site2&site3>site1&site4&site5"
-hypotheses <- "site2=site1=site3=site4=site5"
 
@@
 This leads to a more extreme combined evaluation than any individual evaluation.
-Likewise, if there is very strong support in only a small minority of the samples, while the majority of samples shows no support at all, than it is likely that the hypothesis is not true.
+Likewise, if there is very strong support in only a small minority of the samples, while the majority of samples shows no support at all, than it is likely that the hypothesis is not true and PBF should be reflective of this.
 
@@
 The `pbf()` function only returns output for the hypotheses that are present in all results.
-When `bain` objects are passed to the `pbf()`, no additional hypotheses need to be passed to `pbf()` as they are already saved in the objects.
+When `bain` objects are passed to the `pbf()`, no additional hypotheses need to be passed to `pbf()` as they are already saved in the `bain` objects.
 
@@
   bain(lm_sample2, "site2=site1=site3>site4=site5"),
-  bain(lm_sample3, "site2&site3>site1&site4&site5")
+  bain(lm_sample3, "site2&site3<site1&site4&site5")
 )
@@
 First these values are extracted from all samples.
-A function is created that extracts is information which is then used to extract the necessary information from all samples.
+A function is created that extracts this information which is then used to extract the necessary information from all samples.
 
@@
 get_estimates <- function(sample){
+
+  # number of groups in the 'site' variable
   number_sites <- length(levels(sesamesim$site))
 
+  # sample size of each sample
   sample_size <- nrow(sample)
 
+  # obtains mean and standard deviation of postnumb for each group
   site_estimates <- sapply(1:number_sites, function(i){
@@
 
+  # get the estimates of the mean postnumb for each group and give them appropriate names
   est_names <- sprintf("site%d", 1:number_sites)
@@
 
+  # obtain covariance matrix of the estimates, samples are assumed to be independent.
   covmat <- diag((site_estimates[2,] / sqrt(sample_size))^2)
   dimnames(covmat) <- list(est_names, est_names)
+
   return(list(means = site_means, covmat = covmat, sample_size = sample_size))
@@
 
-Then, each object is converted to a `bain` object with these estimates
+Then, each object is converted to a `bain` object using the extracted information and combined in a list.
 ```{r}

### 2022-06-27 UI Function/pbf.R

@@ -31,9 +31,10 @@ pbf <- function(x, ...){
     }
   }
   BFs <- do.call(cbind, lapply(x, function(y){y$fit$BF.c[match(hyps, y$hypotheses)]}))
-  rownames(BFs) <- paste0(sprintf('H%d: ', 1:length(hyps)),hyps) # give names
-  
-  res <- list(BFs = BFs, pbf = apply(BFs, 1, prod)) # obtain pbf ic, might need to change dependent on alternative hyp
+  colnames(BFs) <- paste0("Sample ", 1:ncol(BFs))
+  res <- data.frame(PBF = apply(BFs, 1, prod), BFs)# obtain pbf ic, might need to change dependent on alternative hyp
+  rownames(res) <- paste0(sprintf('H%d: ', 1:length(hyps)),hyps) # give names
+  class(res) <- c("pbf", class(res))
   return(res)
 }
 

### 2022-06-27 manuscript.Rmd

--- 2022-06-27 manuscript.Rmd
@@
 ### using bain objects
-The `pbf()` function also works when objects are first converted to `bain` objects.
-In this specific example, each `bain` object corresponding to one of the three samples has the $H_0$ and $H_A$ in common, but contains one sample-specific hypothesis.
-The `pbf()` function only returns output for the hypotheses that are present in all results.
-When `bain` objects are passed to the `pbf()`, no additional hypotheses need to be passed to `pbf()` as they are already saved in the objects.
+
+The `pbf()` function also accepts multiple `bain` objects.
+This makes it possible to, for example, evaluate different sets of hypotheses for multiple data sets before using the resulting `bain` objects to aggregate the evidence for all hypotheses those datasets share in common.
+The example below illustrates this princuple.
+Each analysis has two hypotheses in common, but also contains one sample-specific hypothesis.
+The `pbf()` function is called on the resulting `bain` object and returns output for the two common hypotheses.
+Note that, in this case, `pbf()` does not require an argument `hypotheses`, as the hypotheses are contained in the `bain` objects.
 
@@
 This means that the sample-specific hypotheses have been left out and only the hypotheses that each sample has in common are retained.
+When there are no common hypotheses, `pbf()` throws an error.
 
-When there is no hypothesis that is shared by each sample, `pbf()` throws an error.
-```{r}
+```{r, echo = FALSE, eval = FALSE}
 bains_different_hyps <- list(
@@
 ```
-The function returns an error stating that the samples have no hypothesis in common and thus the PBF cannot be calculated.
+<!--The function returns an error stating that the samples have no hypothesis in common and thus the PBF cannot be calculated.-->
 
-### using named vectors
-Another specific use-case is when the raw data from different samples is not available.
-In this case, the researcher only has access to the parameter-values that are provided in the research papers.
-Because the `pbf()` function inherently converts objects to `bain` objects, each sample requires three pieces of information in order to be usable: the named parameter estimates, a covariance matrix of their standard errors and the sample size.
-First these values are extracted from all samples.
-A function is created that extracts is information which is then used to extract the necessary information from all samples.
+### Using sufficient statistics
 
-```{r}
+A third use-case occurs when the raw data from different samples is not available.
+This may happen, for example, when aggregating findings from the published literature (similar to meta-analysis).
+In this case, one can use the default interface of `bain`, as explained in [REFERENCE BAIN PAPER Hoijtink, Mulder, van Lissa, & Gu, (2019) <doi:10.31234/osf.io/v3shc>].
+This function requires four arguments: A named vector of parameter estimates, their asymptotic covariance matrix, the original sample size, and the number of within-group and between-group parameters.
+These arguments may be derived in different ways for distinct samples.
+The example below illustrates how to aggregate evidence in favor of one hypotheses across three studies that each used different methods.
+<!-- CJ: What I'd like to do here is show how to aggregate a t-test, a regression coefficient and a correlation coefficient -->
+
+```{r, eval = FALSE, echo = FALSE}
 get_estimates <- function(sample){
@@
 <!-- It is interesting that these estimates are not identical to the other examples... Is that because we do not have ALL information available in this use-case, or is something off in the script? -->
+<!-- CJ: We should ask Herbert about this, but this example won't make it into this paper. If you want, you can send Herbert a reproducible example of the 'problem'!-->
 

### 2022-07-15 manuscript.Rmd
start with moral politics data in tutorial

@@
 This tutorial focusses on synthesizing evidence for an informative hypotheses using the Product Bayes Factor (pbf).
-To illustrate the method, the `sesamesim` dataset will be used which is part of `bain`.
-For a detailed description of the dataset, run `?bain::sesamesim`.
+To illustrate the method, a simulated counterpart of the data used in [REF Van Leeuwen et al.
+2022] will be used avalable in `bain`.
+The data is used to investigate the relationship between political orientation and moral dispositions in three countries: the United states of America, Denmark and the Netherlands.
+A more detailed description of the dataset is either provided in [REF Van Leeuwen et al.
+2022] or by running `?bain::synthetic_us`, `?bain::synthetic_dk` or `?bain::synthetic_nl`.
 
@@
 
-Let's say we are interested in whether the status of the site children grew up in (`site`) affects their score on a numbers test after watching Sesame Street for a year (`postnumb`).
+Suppose we are interested in whether someone's socio-political orientation, i.e.
+liberal-conservative, is associated with the moral value of 'group loyalty'.
+Participants in all countries completed two scales that measure either socio-political orientation or group loyalty.
+Because each scale consists of multiple items of which some are reverse coded, single factors scores for both scales are first created for each country.
+Because the scales may differ per country, they also need to be standardized
+
+```{r}
+# reverse code
+# create factor scores
+
+# I create simple average scores for now
+aggregate_scores <- function(country){
+  df <- as.data.frame(cbind(
+    scale(apply(country[,grepl("soc_", colnames(country))],1, mean)),
+    apply(country[,grepl("grp_", colnames(country))],1, mean),
+    apply(country[,grepl("her_", colnames(country))],1, mean)
+    ))
+  colnames(df) <- c('soc','grp', 'her')
+  return(df)
+  }
+
+NLa <- aggregate_scores(NL)
+USa <- aggregate_scores(US)
+DKa <- aggregate_scores(DK)
+
+```
+
+Socio-political orientation and group loyalty have been aggregated to variables `soc` and `grp` respectively.
+Both are continuous variables with higher scores indicating conservatism or higher group loyalty respectively.
+In NHST we specify the null hypothesis that there is no association between `soc` and `grp`.
+We can test this hypothesis in each country by calculating a standardized regressoin coefficient.
+
+$H_0: \beta_{grp} = 0$.
+
+However, we may have prior expectations that higher group loyalty is somewhat related to  higher conservatism. An informative hypothesis could be:
+
+$H_i: \beta_{grp} > .1$
+
+<!-- SESAMESIM DATA: Let's say we are interested in whether the status of the site children grew up in (`site`) affects their score on a numbers test after watching Sesame Street for a year (`postnumb`).
 `postnumb` is a continuous variable which ranges from 0-63.
@@
 
-$H_i: \mu_{site2},\mu_{site3} > \mu_{site1},\mu_{site4},\mu_{site5}$
+$H_i: \mu_{site2},\mu_{site3} > \mu_{site1},\mu_{site4},\mu_{site5}$ -->
 
-We can test the $H_0$ and $H_A$ simultaneously by running a linear regression on the data and passing the `lm` object with the specific hypotheses to bain.
+Because it is very improbable that $\beta_{grp}$ is exactly equal to 0, more information can be won by testing $H_i$.
+We can test $H_i$ by running a linear regression on the data and passing the `lm` object with the specific hypotheses to `bain`. For the Netherlands, the Bayes Factor would be calculated as follows:
+
 ```{r}
@@
 
-# convert `site` to nominal variable
-sesamesim$site <- as.factor(sesamesim$site)
-
 # specify model, allow meaningful zero-point
-fit <- lm(postnumb ~ site-1, data = sesamesim)
+fit <- lm(soc ~ grp, data = NLa)
 
 #specify Hypotheses to evaluate
-hypotheses <- "site2=site1=site3=site4=site5; site2&site3>site1&site4&site5"
-hypotheses <- "site2=site1=site3=site4=site5"
+hypotheses <- "grp > .1"
 
@@
 ```
-The output of `bain` seems to support neither hypothesis.
+<!-- SESAMESIM DATA: The output of `bain` seems to support neither hypothesis.
 The `fit` for both hypotheses is approximately zero which means that almost none of the data are in agreement with the restrictions the specified hypotheses pose.
@@
 This means that *out of* the specified hypotheses and their complement, the probability of $H_c$ being true is 99.9%.
+-->
 
 ### Single vs multiple studies.
-Sometimes there is access to the data of not only a single, but multiple studies that all contain the necessary information to evaluate our hypothesis $H_i$.
+<!-- SESAMESIM DATA: Sometimes there is access to the data of not only a single, but multiple studies that all contain the necessary information to evaluate our hypothesis $H_i$.
 To evaluate $H_i$, it is possible to synthesize the support for $H_i$ over all studies using the PBF.
@@
 Likewise, if there is very strong support in only a small minority of the samples, while the majority of samples shows no support at all, than it is likely that the hypothesis is not true.
+-->
+
+Sometimes there is access to the data of not only a single, but multiple studies that all contain the necessary information to evaluate our hypothesis $H_i$.
+To evaluate $H_i$, it is possible to synthesize the support for $H_i$ over all studies using the PBF calculated with `pbf()`.
+In this example, data to evaluate $H_i$ is available for three countries.
+
+In the first use-case of `pbf()`, an `lm()` is ran for each country All `lm` objects are stored in a list which is passed to `pbf()`.
+`pbf()` supports lists of all objects that `bain()` also supports. All supported types are specified in bain vignette *REF*. One exception is the use of named vectors, which will be covered later.
+```{r}
+fit_US <- lm(soc ~ grp + her, data = USa)
+fit_DK <- lm(soc ~ grp + her, data = DKa)
+fit_NL <- lm(soc ~ grp + her, data = NLa)
+
+all_fits <- list(fit_US, fit_DK, fit_NL)
+
+pbf(all_fits, hypotheses)
+```
+
+<!-- SESAME DATASET: The function returns a `data.frame` containing the passed hypotheses in the rows and in the columns the PBF and the individual Bayes Factor for each sample individually.
+The `PBF` of both hypotheses is close to zero, meaning that the combined evidence for both hypotheses over all studies is low.
+When looking at the individual Bayes Factors, we also see that the PBF is lower than all of the Bayes Factors individually.
+For sample 3 the specified hypotheses are even more likely than their complement.
+This example illustrates another strength of the PBF; being able to evaluate a hypothesis more extremely than the most extreme individual evaluation.
+If all samples do not support a hypothesis, than the number of times the hypothesis is not supported should be included in its evaluation.
+This leads to a more extreme combined evaluation than any individual evaluation.
+Likewise, if there is very strong support in only a small minority of the samples, while the majority of samples shows no support at all, than it is likely that the hypothesis is not true.
+-->
 
@@
 This makes it possible to, for example, evaluate different sets of hypotheses for multiple data sets before using the resulting `bain` objects to aggregate the evidence for all hypotheses those datasets share in common.
-The example below illustrates this princuple.
-Each analysis has two hypotheses in common, but also contains one sample-specific hypothesis.
-The `pbf()` function is called on the resulting `bain` object and returns output for the two common hypotheses.
+The example below illustrates this principle.
+The analyses have one hypotheses in common ($H_i: \beta_{grp} > .1$), but also contains one sample-specific hypothesis regarding $\beta_{her}$.
+The `pbf()` function is called on the list of  resulting `bain` objects and returns output for the common hypotheses.
 Note that, in this case, `pbf()` does not require an argument `hypotheses`, as the hypotheses are contained in the `bain` objects.
@@
 ```{r}
-bain_sample1 <- bain(lm_sample1, "site2=site1=site3=site4=site5; site2>site1=site3=site4=site5; site2&site3>site1&site4&site5")
-bain_sample2 <- bain(lm_sample2, "site2=site1=site3=site4=site5; site2=site1=site3>site4=site5; site2&site3>site1&site4&site5")
-bain_sample3 <- bain(lm_sample3, "site2=site1=site3=site4=site5; site2=site1=site3=site4>site5; site2&site3>site1&site4&site5")
+bain_US <- bain(fit_US, "grp > .1; her = 0")
+bain_DK <- bain(fit_DK, "grp > .1; her > 0")
+bain_NL <- bain(fit_NL, "grp > .1; her < 0")
 
-bain_allsamples <- list(bain_sample1, bain_sample2, bain_sample3)
+bain_all <- list(bain_US, bain_DK, bain_NL)
 
-pbf(bain_allsamples)
+pbf(bain_all)
 
@@
 bains_different_hyps <- list(
-  bain(lm_sample1, "site2=site1=site3=site4=site5"),
-  bain(lm_sample2, "site2=site1=site3>site4=site5"),
-  bain(lm_sample3, "site2&site3>site1&site4&site5")
+  bain(fit_US, "grp = 0"),
+  bain(fit_DK, "grp > .1"),
+  bain(fit_NL, "grp < .1")
 )
@@
 <!-- CJ: What I'd like to do here is show how to aggregate a t-test, a regression coefficient and a correlation coefficient -->
+
+```{r}
+library(bain)
+
+### THESE DATA WILL BE USED (READ CASPARS WHATSAPP MESSAGES)
+DK <- bain::synthetic_dk
+US <- bain::synthetic_us
+NL <- bain::synthetic_nl
+?bain::synthetic_dk
+
+# dat <- as.data.frame(cbind(
+#   X = c(rnorm(1000,0,1), rnorm(1000, 0.2, 1)),
+#   group = rep(0:1, each = 1000)
+#   ))
+
+# dat$group <- as.factor(dat$group)
+
+# x <- dat[dat$group == 0, 1]
+# y <- dat[dat$group == 1, 1]
+# 
+# ttest <- t_test(x,y, var.equal = TRUE, paired = FALSE)
+# res <- bain(ttest, hypothesis = "y > x")
+# 
+# 
+# fit_lm <- lm(X ~ group-1, data = dat)
+# res_lm <- bain(fit_lm, hypothesis = "group1 > group0")
+# 
+# cor.test(y = dat$X, x = rep(0:1, each = 1000))
+```
 

### 2022-07-19 manuscript.Rmd

@@
 The amount of evidence for this hypothesis is expressed as a Bayes factor, BF.
-A Bayes factor can be interpreed as the ratio of evidence in favor of $H_i$ divided by evidence against it.
+A Bayes factor can be interpreted as the ratio of evidence in favor of $H_i$ divided by evidence against it.
 Thus, BF = 10 means that the data provide ten times more support in favor of the hypothesis than against it.
@@
 It pools estimates of this effect size across multiple studies to obtain an overall estimate of the effect size.
-It thus answers the question: Givern certain assumptions about between-studies heterogeneity, what is the average population effect size?
+It thus answers the question: Given certain assumptions about between-studies heterogeneity, what is the average population effect size?
 BES, on the other hand, aggregates evidence in favor of an informative hypothesis across multiple studies.
@@
 This tutorial focusses on synthesizing evidence for an informative hypotheses using the Product Bayes Factor (pbf). 
-To illustrate the method, a simulated counterpart of the data used in [REF Van Leeuwen et al.
-2022] will be used avalable in `bain`.
+To illustrate the method, a simulated counterpart of the data used in [@vanleeuwen2022morality] will be used available in `bain`. 
 The data is used to investigate the relationship between political orientation and moral dispositions in three countries: the United states of America, Denmark and the Netherlands. 
-A more detailed description of the dataset is either provided in [REF Van Leeuwen et al.
-2022] or by running `?bain::synthetic_us`, `?bain::synthetic_dk` or `?bain::synthetic_nl`.
+A more detailed description of the dataset is either provided in [@vanleeuwen2022morality] or by running `?bain::synthetic_us`, `?bain::synthetic_dk` or `?bain::synthetic_nl`.
+
+```{r}
+US <- bain::synthetic_us
+DK <- bain::synthetic_dk
+NL <- bain::synthetic_nl
+```
 
@@
 In conventional Null-hypothesis significance testing (NHST), as the name implies, data is analysed to generate evidence for or against the null-hypothesis. 
-This method can only evaluate a single hypothesis, the null-hypothesis, and its output is dichotomous in nature: whether a theoretical relationship is probable or improbable to exist given the data.
+This method evaluates a single hypothesis, the null-hypothesis, and its output is dichotomous in nature: whether a theoretical relationship is probable or improbable to exist given the data. 
 However, there are many more hypotheses that can theoretically be tested and researchers may not be interested in the null-hypothesis.
@@
 
-Suppose we are interested in whether someone's socio-political orientation, i.e.
-liberal-conservative, is associated with the moral value of 'group loyalty'.
-Participants in all countries completed two scales that measure either socio-political orientation or group loyalty.
+Suppose we are interested in whether socio-political orientation, i.e. liberal-conservative, is associated with the subjective importance of the moral value 'group loyalty'. 
+Participants in the three countries completed two scales of which one measured their socio-political orientation and the other their group loyalty. 
 Because each scale consists of multiple items of which some are reverse coded, single factors scores for both scales are first created for each country. 
-Because the scales may differ per country, they also need to be standardized
+The scales also differ per country, and thus are standardized.
+
+<!--EB: For simplicities sake, I for now create simple average scores without taking into account reverse coding. This should likely be improved with factor scores -->
 
 ```{r}
-# reverse code
-# create factor scores
-
 # I create simple average scores for now
@@
 Socio-political orientation and group loyalty have been aggregated to variables `soc` and `grp` respectively. 
-Both are continuous variables with higher scores indicating conservatism or higher group loyalty respectively.
+Both are continuous variables with higher scores indicating either conservatism or higher group loyalty. 
 In NHST we specify the null hypothesis that there is no association between `soc` and `grp`. 
-We can test this hypothesis in each country by calculating a standardized regressoin coefficient.
+One method to test the hypothesis is to calculate a standardized regression coefficient.
 
@@
 
-However, we may have prior expectations that higher group loyalty is somewhat related to  higher conservatism. An informative hypothesis could be:
+However, we may have prior expectations that higher group loyalty is somewhat related to  higher conservatism. 
+An informative hypothesis could be:
 
-$H_i: \beta_{grp} > .1$
+$H_i: \beta_{grp} > 0$
 
@@
 Because it is very improbable that $\beta_{grp}$ is exactly equal to 0, more information can be won by testing $H_i$.
-We can test $H_i$ by running a linear regression on the data and passing the `lm` object with the specific hypotheses to `bain`. For the Netherlands, the Bayes Factor would be calculated as follows:
+We can test $H_i$ by running a linear regression on the data and passing the `lm` object with the specific hypotheses to `bain`. 
+For a single study, such as the study that investigated the Netherlands, the Bayes factor would be obtained as follows:
 
@@
 #specify Hypotheses to evaluate
-hypotheses <- "grp > .1"
+hypotheses <- "grp > 0"
 
@@
 Sometimes there is access to the data of not only a single, but multiple studies that all contain the necessary information to evaluate our hypothesis $H_i$. 
-To evaluate $H_i$, it is possible to synthesize the support for $H_i$ over all studies using the PBF calculated with `pbf()`.
+To evaluate $H_i$, it is possible to synthesize the support for $H_i$ over all studies using the pbf calculated with `pbf()`. 
 In this example, data to evaluate $H_i$ is available for three countries.
 
-In the first use-case of `pbf()`, an `lm()` is ran for each country All `lm` objects are stored in a list which is passed to `pbf()`.
-`pbf()` supports lists of all objects that `bain()` also supports. All supported types are specified in bain vignette *REF*. One exception is the use of named vectors, which will be covered later.
+In the first use-case of `pbf()`, an `lm()` is ran for each country. 
+All `lm` objects are stored in a list which is passed to `pbf()`.
+`pbf()` supports lists of all objects that `bain()` also supports. 
+All supported types are specified in the `bain` vignette. 
+One exception is the use of named vectors, which will be covered later.
 ```{r}
@@
 The example below illustrates this principle.
-The analyses have one hypotheses in common ($H_i: \beta_{grp} > .1$), but also contains one sample-specific hypothesis regarding $\beta_{her}$.
-The `pbf()` function is called on the list of  resulting `bain` objects and returns output for the common hypotheses.
-Note that, in this case, `pbf()` does not require an argument `hypotheses`, as the hypotheses are contained in the `bain` objects.
+The analyses have one hypotheses in common ($H_i: \beta_{grp} > 0$), but also contains one sample-specific hypothesis regarding $\beta_{her}$.
+The `pbf()` function is called on the list of the resulting `bain` objects and returns output for the common hypotheses.
+Note that, in this case, `pbf()` does not require an argument `hypothesiss`, as the hypotheses are contained in the `bain` objects.
 
 ```{r}
-bain_US <- bain(fit_US, "grp > .1; her = 0")
-bain_DK <- bain(fit_DK, "grp > .1; her > 0")
-bain_NL <- bain(fit_NL, "grp > .1; her < 0")
+bain_US <- bain(fit_US, "grp > 0; her = 0")
+bain_DK <- bain(fit_DK, "grp > 0; her > 0")
+bain_NL <- bain(fit_NL, "grp > 0; her < 0")
 
@@
 This means that the sample-specific hypotheses have been left out and only the hypotheses that each sample has in common are retained.
-When there are no common hypotheses, `pbf()` throws an error.
+When there are no hypotheses at all that are common in all objects, `pbf()` throws an error.
 
@@
 This may happen, for example, when aggregating findings from the published literature (similar to meta-analysis).
-In this case, one can use the default interface of `bain`, as explained in [REFERENCE BAIN PAPER Hoijtink, Mulder, van Lissa, & Gu, (2019) <doi:10.31234/osf.io/v3shc>].
+In this case, one can use the default interface of `bain`, as explained in [@hoijtink2019tutorial].
 This function requires four arguments: A named vector of parameter estimates, their asymptotic covariance matrix, the original sample size, and the number of within-group and between-group parameters.
@@
 
+In this specific use-case, let's suppose three different statistics have been calculated to analyse the data of the three countries. 
+Each statistic was calculated to establish the relationship between `grp` and `soc`:
+1. A correlation coefficient was calculated using the US data.
+2. A standardized regression coefficient for `grp` was calculated using the DK data.
+3. A t-test was performed using the NL data. 
+To simulate groups to perform a t-test on, the `grp` variable in the NL data will be recoded to 0 and 1 which represent a score below or above the median of `grp` respectively.
+
+The current $H_i: \beta_{grp} > 0$ effectively tests the following:
+*There is a positive association between political conservatism and subjective importance of group loyalty*
+But $\beta_{grp}$ is a regression-specific parameter which was not calculated in two out of three studies.
+Therefore, the three sample-specific hypotheses that answer the same underlying $H_i$ are:
+- 1. $H_i^{US}: \rho_{grp,soc} > 0$
+- 2. $H_i^{DK}: \beta_{grp} > 0$
+- 3. $H_i^{NL}: \mu_{soc}^{conservative} > \mu_{soc}^{liberal}$
+
+These statistics to test the hypotheses along with their standard error can be passed to `bain` and do not require the raw data. 
+The pbf can then be calculated and a single score can be synthesized to evaluate $H_i$. 
+This despite different statistical methods being used to establish the relationship between `grp` and `soc`.
+
 ```{r}
-library(bain)
+US_est <- c(grp = cor(USa$soc, USa$grp))
+US_se <- sqrt(1-US_est^2) / sqrt(nrow(USa)-2) # formula for SE of correlation coefficient: sqrt(1-r^2)/sqrt(n-2)
 
-### THESE DATA WILL BE USED (READ CASPARS WHATSAPP MESSAGES)
-DK <- bain::synthetic_dk
-US <- bain::synthetic_us
-NL <- bain::synthetic_nl
-?bain::synthetic_dk
+DK_fit <- summary(lm(soc~grp, data = DKa))
+DK_est <- c(grp = DK_fit$coefficients['grp', 'Estimate'])
+DK_se <- DK_fit$coefficients['grp', 'Std. Error']
 
-# dat <- as.data.frame(cbind(
-#   X = c(rnorm(1000,0,1), rnorm(1000, 0.2, 1)),
-#   group = rep(0:1, each = 1000)
-#   ))
+### EB: I have a question When calculating the estimate for NL.
+I assume I need the effect size of the t-test, rather than the t-statistic itself.
+This requires Cohen's d.
+Is this correct?
+In this case I do not even need to do a t-test, I only need the descriptive statistics of the two samples
 
-# dat$group <- as.factor(dat$group)
+NLa$grp_recoded <- as.factor(sapply(NLa$grp,function(i){ ifelse(i < median(NLa$grp), 0, 1) }))
+liberals <- NLa$soc[NLa$grp_recoded == 0]
+conservatives <- NLa$soc[NLa$grp_recoded == 1]
 
-# x <- dat[dat$group == 0, 1]
-# y <- dat[dat$group == 1, 1]
-# 
-# ttest <- t_test(x,y, var.equal = TRUE, paired = FALSE)
-# res <- bain(ttest, hypothesis = "y > x")
-# 
-# 
-# fit_lm <- lm(X ~ group-1, data = dat)
-# res_lm <- bain(fit_lm, hypothesis = "group1 > group0")
-# 
-# cor.test(y = dat$X, x = rep(0:1, each = 1000))
+nlib <- length(liberals)
+ncon <- length(conservatives)
+pooled_sd <- sqrt(((ncon-1)*sd(conservatives) + (nlib-1)*sd(liberals)) / (nlib+ncon-2) )
+
+NL_est <- c(grp = (mean(conservatives) - mean(liberals)) / pooled_sd)
+NL_se <- sqrt(((ncon+nlib)/(ncon*nlib)) + NL_est^2/(2*(ncon+nlib))) # formula for SE of cohen's D retrieved from https://stats.stackexchange.com/questions/495015/what-is-the-formula-for-the-standard-error-of-cohens-d#:~:text=Cohen's%20d%20is%20defined%20as,%E2%88%922%20(see%20here).
+
+```
+The values are then converted to `bain` objects, put in a list, and passed to `pbf()`. 
+```{r}
+US_bain <- bain(x = US_est, Sigma = matrix(US_se, 1, 1), n = nrow(USa), hypothesis = hypotheses)
+DK_bain <- bain(x = DK_est, Sigma = matrix(DK_se, 1, 1), n = nrow(DKa), hypothesis = hypotheses)
+NL_bain <- bain(x = NL_est, Sigma = matrix(NL_se, 1, 1), n = nrow(NLa), hypothesis = hypotheses)
+
+pbf(list(US_bain, DK_bain, NL_bain))
 ```
 
+The output provides evidence for or against the hypothesis: *there is a positive association between political conservatism and subjective importance of group loyalty*.
+Although each study used a different method to operationalize the evidence for that hypothesis, their evidence can be synthesized to a single value using the pbf.
+
+<!--
 ```{r, eval = FALSE, echo = FALSE}
@@
 ```
+-->
 <!-- It is interesting that these estimates are not identical to the other examples... Is that because we do not have ALL information available in this use-case, or is something off in the script? -->

### 2022-07-19 manuscript.Rmd
Rewrite opening paragraph of tutorial

@@
 # Tutorial
-This tutorial about Bayesian Evidence Synthesis (BES) requires installation of `r R.Version()$version.string`, Rstudio version `r RStudio.Version()$version` and package `bain` version `r packageVersion('bain')`.
+
+This tutorial about Bayesian Evidence Synthesis (BES) requires R-package `bain` version `r packageVersion('bain')` or later.
 This tutorial focusses on synthesizing evidence for an informative hypotheses using the Product Bayes Factor (pbf). 
-To illustrate the method, a simulated counterpart of the data used in [@vanleeuwen2022morality] will be used available in `bain`. 
-The data is used to investigate the relationship between political orientation and moral dispositions in three countries: the United states of America, Denmark and the Netherlands. 
+The data used to illustrate the method have been simulated based on the data presented in [@vanleeuwen2022morality].
+These simulated data are available in the `bain` package. 
 A more detailed description of the dataset is either provided in [@vanleeuwen2022morality] or by running `?bain::synthetic_us`, `?bain::synthetic_dk` or `?bain::synthetic_nl`.
+The research question is whether there is an association between political orientation and moral dispositions.
+Suitable data were collected in three countries: the United states of America, Denmark and the Netherlands.
+Each sample contains multiple measures of political orientation and moral dispositions.
+In the original publication, the PBF was used to aggregate evidence across scales and countries to obtain an overall measure of support for the central hypothesis.
+This tutorial follows the same rationale, but uses only one effect size per sample,
+and varies the way this effect size is computed to illustrate different use cases.
 

### 2022-07-20 manuscript.Rmd
Caspar's edits of the tutorial

@@
 ```{r}
+NL <- bain::synthetic_nl
+DK <- bain::synthetic_dk
 US <- bain::synthetic_us
-DK <- bain::synthetic_dk
-NL <- bain::synthetic_nl
 ```
@@
 
-### The Bayes Factor using bain.
-In conventional Null-hypothesis significance testing (NHST), as the name implies, data is analysed to generate evidence for or against the null-hypothesis. 
-This method evaluates a single hypothesis, the null-hypothesis, and its output is dichotomous in nature: whether a theoretical relationship is probable or improbable to exist given the data. 
-However, there are many more hypotheses that can theoretically be tested and researchers may not be interested in the null-hypothesis.
+### How to use bain
+
+<!-- CJ: Is dit niet teveel herhaling van wat we in de inleiding zeggen?? -->
+<!-- In conventional Null-hypothesis significance testing (NHST), as the name implies, data is analysed to generate evidence for or against the null-hypothesis.  -->
+<!-- This method evaluates a single hypothesis, the null-hypothesis, and its output is dichotomous in nature: whether a theoretical relationship is probable or improbable to exist given the data.  -->
+<!-- However, there are many more hypotheses that can theoretically be tested and researchers may not be interested in the null-hypothesis.
 They may also not only be interested in whether the hypothesis is probable, but also by how much.
-In the Bayesian framework, the Bayes Factor (BF) is a test statistic that gives the amount of support for any possible desired hypothesis. 
-Because the focus of this tutorial is on BES, this tutorial will only briefly introduce how to evaluate an hypothesis using the R package `bain` and interpret its output. 
-For an extensive tutorial on how to evaluate informative hypotheses using `bain` see [@hoijtink2019tutorial].
-
-Suppose we are interested in whether socio-political orientation, i.e. liberal-conservative, is associated with the subjective importance of the moral value 'group loyalty'. 
-Participants in the three countries completed two scales of which one measured their socio-political orientation and the other their group loyalty. 
-Because each scale consists of multiple items of which some are reverse coded, single factors scores for both scales are first created for each country. 
-The scales also differ per country, and thus are standardized.
+-->
+<!-- In the Bayesian framework, the Bayes Factor (BF) is a test statistic that gives the amount of support for any possible desired hypothesis.  -->
+<!-- Because the focus of this tutorial is on BES,  -->
+First, we briefly introduce the basic use of `bain::bain()`, and how to interpret its output. 
+Suppose we are interested in whether the self-reported importance of family morality positively predicts a conservative socio-political orientation.
+Because both scales consist of multiple items, we use structural equation modeling (SEM) to perform latent variable regression.
+Our hypothesis is $H_i: \beta > 0$, where $\beta$ (beta) is the standardized regression coefficient.
+Note that, for the sake of simplicity, this hypothesis corresponds to a traditional one-sided null-hypothesis test.
+The original publication instead used a minimal effect size of interest, and specified $H_i: \beta > .1$.
+The code below illustrates the analysis for a single sample.
+For a more in-depth tutorial on `bain::bain()`, see @hoijtink2019tutorial, and for further guidance on the use of `bain::bain()` in SEM, see @REF VAN LISSA TUTORIAL.
 
 <!--EB: For simplicities sake, I for now create simple average scores without taking into account reverse coding. This should likely be improved with factor scores -->
+<!-- CJ: Keep the code SUPER simple. No functions etc -->
 
-```{r}
-# I create simple average scores for now
-aggregate_scores <- function(country){
-  df <- as.data.frame(cbind(
-    scale(apply(country[,grepl("soc_", colnames(country))],1, mean)),
-    apply(country[,grepl("grp_", colnames(country))],1, mean),
-    apply(country[,grepl("her_", colnames(country))],1, mean)
-    ))
-  colnames(df) <- c('soc','grp', 'her')
-  return(df)
-  }
+```{r echo = TRUE, eval = FALSE}
+# Load relevant packages
+library(bain)
+library(lavaan)
 
-NLa <- aggregate_scores(NL)
-USa <- aggregate_scores(US)
-DKa <- aggregate_scores(DK)
+# Specify a SEM-model with two latent variables and latent regression
+model_nl <- "
+family =~ fam_1 + fam_2 + fam_3
+conservative =~ sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
+                sepa_eco_1 + sepa_eco_2 + sepa_eco_3 + sepa_eco_4 + sepa_eco_5
+conservative ~ beta * family"
 
+# Estimate the model in lavaan
+results_nl <- sem(model = model_nl, data = NL)
+
+# Test the hypothesis that the effect size labeled 'beta' is positive
+# The argument standardize = TRUE uses the standardized model parameters
+bain(results_nl, hypothesis = "beta > 0", standardize = TRUE)
+```
+```{r echo = FALSE, eval = TRUE}
+# Load relevant packages
+library(bain)
+library(lavaan)
+
+# Specify a SEM-model with two latent variables and latent regression
+model_nl <- "
+family =~ fam_1 + fam_2 + fam_3
+conservative =~ sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
+                sepa_eco_1 + sepa_eco_2 + sepa_eco_3 + sepa_eco_4 + sepa_eco_5
+conservative ~ beta * family"
+
+# Estimate the model in lavaan
+results_nl <- sem(model = model_nl, data = NL)
+
+# Test the hypothesis that the effect size labeled 'beta' is positive
+# The argument standardize = TRUE uses the standardized model parameters
+bf_nl <- bain(results_nl, hypothesis = "beta > 0", standardize = TRUE)
+bf_nl
 ```
 
-Socio-political orientation and group loyalty have been aggregated to variables `soc` and `grp` respectively. 
-Both are continuous variables with higher scores indicating either conservatism or higher group loyalty. 
-In NHST we specify the null hypothesis that there is no association between `soc` and `grp`. 
-One method to test the hypothesis is to calculate a standardized regression coefficient.
+<!-- Socio-political orientation and group loyalty have been aggregated to variables `soc` and `grp` respectively.  -->
+<!-- Both are continuous variables with higher scores indicating either conservatism or higher group loyalty.  -->
+<!-- In NHST we specify the null hypothesis that there is no association between `soc` and `grp`.  -->
+<!-- One method to test the hypothesis is to calculate a standardized regression coefficient. -->
 
-$H_0: \beta_{grp} = 0$.
+<!-- $H_0: \beta_{grp} = 0$. -->
 
-However, we may have prior expectations that higher group loyalty is somewhat related to  higher conservatism. 
-An informative hypothesis could be:
+<!-- However, we may have prior expectations that higher group loyalty is somewhat related to  higher conservatism.  -->
+<!-- An informative hypothesis could be: -->
 
-$H_i: \beta_{grp} > 0$
+<!-- $H_i: \beta_{grp} > 0$ -->
 
@@
 
-Because it is very improbable that $\beta_{grp}$ is exactly equal to 0, more information can be won by testing $H_i$.
-We can test $H_i$ by running a linear regression on the data and passing the `lm` object with the specific hypotheses to `bain`. 
-For a single study, such as the study that investigated the Netherlands, the Bayes factor would be obtained as follows:
+<!-- Because it is very improbable that $\beta_{grp}$ is exactly equal to 0, more information can be won by testing $H_i$. -->
+<!-- We can test $H_i$ by running a linear regression on the data and passing the `lm` object with the specific hypotheses to `bain`.  -->
+<!-- For a single study, such as the study that investigated the Netherlands, the Bayes factor would be obtained as follows: -->
 
-```{r}
+```{r, echo = FALSE, eval = FALSE}
 # load bain package
@@
 
-### Single vs multiple studies.
-<!-- SESAMESIM DATA: Sometimes there is access to the data of not only a single, but multiple studies that all contain the necessary information to evaluate our hypothesis $H_i$.
-To evaluate $H_i$, it is possible to synthesize the support for $H_i$ over all studies using the PBF.
-In this example we split the `sesamesim` dataset up into three equal parts, to simulate three studies into the same topic.
+Note that the hypothesis is strongly supported when compared to its complement, $BF_c `report(bf_nl$fit$BF.c[1])`$.
 
-```{r}
-sample1 <- sesamesim[1:75,]
-sample2 <- sesamesim[76:150,]
-sample3 <- sesamesim[151:nrow(sesamesim),]
+### Aggregating evidence across studies
 
+As mentioned before, suitable data were collected to test the substantive hypothesis in three countries.
+There are differences between countries that prevent analyzing these data as a multilevel model, however.
+For instance, conservatism was measured using different scales.
+This is an appropriate situation to use the PBF to aggregate evidence across countries.
+First, we estimate a latent regression model for the remaining two countries, taking care to label the parameter of interest with the same label in all samples.
+Then, we bind all three SEM-models in a list, and call PBF to evaluate the hypothesis of interest on all models and aggregate the evidence.
+
+```{r, eval = FALSE, echo = TRUE}
+# Specify the models for DK and US
+model_dk <- "
+family =~ fam_1 + fam_2 + fam_3
+conservative =~ sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
+                sepa_eco_1 + sepa_eco_2 + sepa_eco_3 + sepa_eco_4 + sepa_eco_5
+conservative ~ beta * family"
+model_us <- "
+family =~ fam_1 + fam_2 + fam_3
+conservative =~ secs_soc_1 + secs_soc_2 + secs_soc_3 + secs_soc_4 + secs_soc_5 +
+                secs_soc_6 + secs_soc_7 + secs_eco_1 + secs_eco_2 + secs_eco_3 +
+                secs_eco_4 + secs_eco_5
+conservative ~ beta * family"
+
+# Estimate the model in lavaan
+results_dk <- sem(model = model_dk, data = DK)
+results_us <- sem(model = model_us, data = US)
+
+# Bind the models into a list
+results <- list(results_nl, results_dk, results_us)
+# Test the hypothesis that the effect size labeled 'beta' is positive
+pbf(results, hypothesis = "beta > 0", standardize = TRUE)
+```
+```{r, eval = TRUE, echo = FALSE}
+# Specify the models for DK and US
+model_dk <- "
+family =~ fam_1 + fam_2 + fam_3
+conservative =~ sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
+                sepa_eco_1 + sepa_eco_2 + sepa_eco_3 + sepa_eco_4 + sepa_eco_5
+conservative ~ beta * family"
+model_us <- "
+family =~ fam_1 + fam_2 + fam_3
+conservative =~ secs_soc_1 + secs_soc_2 + secs_soc_3 + secs_soc_4 + secs_soc_5 +
+                secs_soc_6 + secs_soc_7 + secs_eco_1 + secs_eco_2 + secs_eco_3 +
+                secs_eco_4 + secs_eco_5
+conservative ~ beta * family"
+
+# Estimate the model in lavaan
+results_dk <- sem(model = model_dk, data = DK)
+results_us <- sem(model = model_us, data = US)
+
+# Bind the models into a list
+results <- list(results_nl, results_dk, results_us)
+# Test the hypothesis that the effect size labeled 'beta' is positive
+bf_tot <- pbf(results, hypothesis = "beta > 0", standardize = TRUE)
+bf_tot
 ```
 
-In the first use-case of `pbf()`, an `lm()` is ran for each sample. All `lm` objects are stored in a list which is passed to `pbf()`.
-`pbf()` supports lists of all objects that `bain()` also supports. All supported types are specified in bain vignette *REF*. One exception is the use of named vectors, which will be covered later.
-```{r}
-lm_sample1 <- lm(postnumb~site-1, sample1)
-lm_sample2 <- lm(postnumb~site-1, sample2)
-lm_sample3 <- lm(postnumb~site-1, sample3)
+As the BF in all three samples is positive, the resulting PBF is very large, $PBF `report(bf_tot[1,1])`$.
+We can thus conclude that the central hypothesis receives overwhelming evidence across samples.
+<!-- In the first use-case of `pbf()`, an `lm()` is ran for each sample. All `lm` objects are stored in a list which is passed to `pbf()`. -->
+<!-- `pbf()` supports lists of all objects that `bain()` also supports.
+All supported types are specified in bain vignette *REF*.
+One exception is the use of named vectors, which will be covered later.
+-->
+<!-- ```{r} -->
+<!-- lm_sample1 <- lm(postnumb~site-1, sample1) -->
+<!-- lm_sample2 <- lm(postnumb~site-1, sample2) -->
+<!-- lm_sample3 <- lm(postnumb~site-1, sample3) -->
 
-lm_allsamples <- list(lm_sample1, lm_sample2, lm_sample3)
+<!-- lm_allsamples <- list(lm_sample1, lm_sample2, lm_sample3) -->
 
-pbf(lm_allsamples, hypotheses)
-```
+<!-- pbf(lm_allsamples, hypotheses) -->
+<!-- ``` -->
 
-The function returns a `data.frame` containing the passed hypotheses in the rows and in the columns the PBF and the individual Bayes Factor for each sample individually.
+<!-- The function returns a `data.frame` containing the passed hypotheses in the rows and in the columns the PBF and the individual Bayes Factor for each sample individually.
 The `PBF` of both hypotheses is close to zero, meaning that the combined evidence for both hypotheses over all studies is low.
@@
 
-Sometimes there is access to the data of not only a single, but multiple studies that all contain the necessary information to evaluate our hypothesis $H_i$. 
-To evaluate $H_i$, it is possible to synthesize the support for $H_i$ over all studies using the pbf calculated with `pbf()`. 
-In this example, data to evaluate $H_i$ is available for three countries.
+<!-- Sometimes there is access to the data of not only a single, but multiple studies that all contain the necessary information to evaluate our hypothesis $H_i$.  -->
+<!-- To evaluate $H_i$, it is possible to synthesize the support for $H_i$ over all studies using the pbf calculated with `pbf()`.  -->
+<!-- In this example, data to evaluate $H_i$ is available for three countries. -->
 
-In the first use-case of `pbf()`, an `lm()` is ran for each country. 
-All `lm` objects are stored in a list which is passed to `pbf()`.
-`pbf()` supports lists of all objects that `bain()` also supports. 
-All supported types are specified in the `bain` vignette. 
-One exception is the use of named vectors, which will be covered later.
-```{r}
-fit_US <- lm(soc ~ grp + her, data = USa)
-fit_DK <- lm(soc ~ grp + her, data = DKa)
-fit_NL <- lm(soc ~ grp + her, data = NLa)
+<!-- In the first use-case of `pbf()`, an `lm()` is ran for each country.  -->
+<!-- All `lm` objects are stored in a list which is passed to `pbf()`. -->
+<!-- `pbf()` supports lists of all objects that `bain()` also supports.  -->
+<!-- All supported types are specified in the `bain` vignette.  -->
+<!-- One exception is the use of named vectors, which will be covered later. -->
+<!-- ```{r} -->
+<!-- fit_US <- lm(soc ~ grp + her, data = USa) -->
+<!-- fit_DK <- lm(soc ~ grp + her, data = DKa) -->
+<!-- fit_NL <- lm(soc ~ grp + her, data = NLa) -->
 
-all_fits <- list(fit_US, fit_DK, fit_NL)
+<!-- all_fits <- list(fit_US, fit_DK, fit_NL) -->
 
-pbf(all_fits, hypotheses)
+<!-- pbf(all_fits, hypotheses) -->
+<!-- ``` -->
+
+### Using bain objects
+
+The `pbf()` function also accepts multiple `bain` objects.
+This makes it possible to, for example, evaluate different sets of hypotheses on different data sets before using the resulting `bain` objects to aggregate the evidence for all common hypotheses across datasets.
+The example below illustrates this use case.
+As before, all analyses share one hypotheses in common ($H_i: \beta_{fam} > 0$),
+but the Dutch sample now contains a sample-specific hypothesis regarding the effect of group morality, namely that $\beta_{grp} < 0$.
+The `pbf()` function is called on a list of `bain` objects.
+Note that, in this case, `pbf()` does not require an argument `hypothesis`, as the hypotheses are contained in the `bain` objects.
+
+```{r, echo = TRUE, eval = FALSE}
+# Add the additional predictor to the model, label the effect beta2
+model_nl <- paste(model_nl, "group =~ grp_1 + grp_2 + grp_3
+                             conservative ~ beta2 * group", sep = "\n")
+
+# Estimate the model in lavaan
+results_nl <- sem(model = model_nl, data = NL)
+
+# Obtain BF for each sample; note that the Dutch sample has two hypotheses
+bf_nl <- bain(results_nl, hypothesis = "beta > 0;beta2 < 0", standardize = TRUE)
+bf_dk <- bain(results_dk, "beta > 0")
+bf_us <- bain(results_us, "beta > 0")
+
+# Bind bain objects into a list
+bfs <- list(bf_nl, bf_dk, bf_us)
+
+# Call pbf on that list
+pbf(bfs)
 ```
 
-<!-- SESAME DATASET: The function returns a `data.frame` containing the passed hypotheses in the rows and in the columns the PBF and the individual Bayes Factor for each sample individually.
-The `PBF` of both hypotheses is close to zero, meaning that the combined evidence for both hypotheses over all studies is low.
-When looking at the individual Bayes Factors, we also see that the PBF is lower than all of the Bayes Factors individually.
-For sample 3 the specified hypotheses are even more likely than their complement.
-This example illustrates another strength of the PBF; being able to evaluate a hypothesis more extremely than the most extreme individual evaluation.
-If all samples do not support a hypothesis, than the number of times the hypothesis is not supported should be included in its evaluation.
-This leads to a more extreme combined evaluation than any individual evaluation.
-Likewise, if there is very strong support in only a small minority of the samples, while the majority of samples shows no support at all, than it is likely that the hypothesis is not true.
--->
-
-### using bain objects
-
-The `pbf()` function also accepts multiple `bain` objects.
-This makes it possible to, for example, evaluate different sets of hypotheses for multiple data sets before using the resulting `bain` objects to aggregate the evidence for all hypotheses those datasets share in common.
-The example below illustrates this principle.
-The analyses have one hypotheses in common ($H_i: \beta_{grp} > 0$), but also contains one sample-specific hypothesis regarding $\beta_{her}$.
-The `pbf()` function is called on the list of the resulting `bain` objects and returns output for the common hypotheses.
-Note that, in this case, `pbf()` does not require an argument `hypothesiss`, as the hypotheses are contained in the `bain` objects.
-
-```{r}
-bain_US <- bain(fit_US, "grp > 0; her = 0")
-bain_DK <- bain(fit_DK, "grp > 0; her > 0")
-bain_NL <- bain(fit_NL, "grp > 0; her < 0")
-
-bain_all <- list(bain_US, bain_DK, bain_NL)
-
-pbf(bain_all)
-
-```
-
-As can be seen, the results are identical to the results in the previous example.
-This means that the sample-specific hypotheses have been left out and only the hypotheses that each sample has in common are retained.
-When there are no hypotheses at all that are common in all objects, `pbf()` throws an error.
+As can be seen, the results are equivalent to the results in the previous example.
+The sample-specific hypothesis has been left out, and common hypotheses are retained and aggregated.
+If there are no common hypotheses across all objects, `pbf()` throws an error.
 
@@
 This function requires four arguments: A named vector of parameter estimates, their asymptotic covariance matrix, the original sample size, and the number of within-group and between-group parameters.
-These arguments may be derived in different ways for distinct samples.
+Note that, when analyzing a single parameter per sample, the standard error is sufficient to construct the asymptotic covariance matrix.
+Thus, this method is suitable for data that have been prepared for meta-analysis.
+Importantly, unlike meta-analysis, the present method is suitable for conceptual replications.
+It does not require uniform effect size measures across studies.
 The example below illustrates how to aggregate evidence in favor of one hypotheses across three studies that each used different methods.
@@
 
-In this specific use-case, let's suppose three different statistics have been calculated to analyse the data of the three countries. 
-Each statistic was calculated to establish the relationship between `grp` and `soc`:
-1. A correlation coefficient was calculated using the US data.
-2. A standardized regression coefficient for `grp` was calculated using the DK data.
-3. A t-test was performed using the NL data. 
-To simulate groups to perform a t-test on, the `grp` variable in the NL data will be recoded to 0 and 1 which represent a score below or above the median of `grp` respectively.
+The present use case effectively tests the following hypothesis:
+*There is a positive association between family morality and political conservatism*.
+This conceptual hypothesis is tested differently in the three samples, resulting in three different types of statistics and distinct sample-specific hypotheses:
 
-The current $H_i: \beta_{grp} > 0$ effectively tests the following:
-*There is a positive association between political conservatism and subjective importance of group loyalty*
-But $\beta_{grp}$ is a regression-specific parameter which was not calculated in two out of three studies.
-Therefore, the three sample-specific hypotheses that answer the same underlying $H_i$ are:
-- 1. $H_i^{US}: \rho_{grp,soc} > 0$
-- 2. $H_i^{DK}: \beta_{grp} > 0$
-- 3. $H_i^{NL}: \mu_{soc}^{conservative} > \mu_{soc}^{liberal}$
+1. A t-test was performed using the NL data; using Cohen's D gives $H_i^{NL}: \delta_{conservative > liberal} > 0$
+1. A bivariate regression coefficient for `grp` was calculated using the DK data, giving $H_i^{DK}: b_{fam} > 0$
+1. A correlation coefficient was calculated using the US data, giving $H_i^{US}: \rho_{fam,con} > 0$
 
-These statistics to test the hypotheses along with their standard error can be passed to `bain` and do not require the raw data. 
-The pbf can then be calculated and a single score can be synthesized to evaluate $H_i$. 
-This despite different statistical methods being used to establish the relationship between `grp` and `soc`.
+Note that we intentionally manipulate the data to fit these types of analyses;
+for example, we compute mean scale scores and dichotomize a continuous variable to conduct a t-test.
+We do not advocate these practices for applied research.
+
+First we obtain the relevant parameter estimates and their sampling variances,
+which allows us to test the specific hypotheses in `bain`:
+
+```{r eval = TRUE, echo = TRUE}
+# Create mean scale scores
+NL <- data.frame(
+  family = rowMeans(NL[c("fam_1", "fam_2", "fam_3")]),
+  conservative = rowMeans(NL[c("sepa_soc_1", "sepa_soc_2", "sepa_soc_3",
+                               "sepa_soc_4", "sepa_soc_5", "sepa_eco_1",
+                               "sepa_eco_2", "sepa_eco_3", "sepa_eco_4",
+                               "sepa_eco_5")]))
+DK <- data.frame(
+  family = rowMeans(DK[c("fam_1", "fam_2", "fam_3")]),
+  conservative = rowMeans(DK[c("sepa_soc_1", "sepa_soc_2", "sepa_soc_3",
+                               "sepa_soc_4", "sepa_soc_5", "sepa_eco_1",
+                               "sepa_eco_2", "sepa_eco_3", "sepa_eco_4",
+                               "sepa_eco_5")]))
+
+US <- data.frame(
+  family = rowMeans(US[c("fam_1", "fam_2", "fam_3")]),
+  conservative = rowMeans(US[c("secs_soc_1", "secs_soc_2", "secs_soc_3",
+                               "secs_soc_4", "secs_soc_5", "secs_soc_6",
+                               "secs_soc_7", "secs_eco_1", "secs_eco_2",
+                               "secs_eco_3", "secs_eco_4", "secs_eco_5")]))
+
+# NL: Conduct t-test using Cohen's D
+NL$group <- cut(NL$conservative, breaks = 2, labels = c("liberal", "conservative"))
+sample_sizes <- table(NL$group)
+sds <- tapply(NL$family, NL$group, sd)
+pooled_sd <- sqrt(sum((sample_sizes - 1) * sds) / (sum(sample_sizes) - 2))
+NL_est <- diff(tapply(NL$family, NL$group, mean)) / pooled_sd
+NL_var <- (sum(sample_sizes) / prod(sample_sizes)) + (NL_est^2 / (2*sum(sample_sizes)))
+
+# DK: Conduct bivariate regression
+DK_fit <- lm(conservative ~ family, data = DK)
+DK_est <- coef(DK_fit)["family"]
+DK_var <- vcov(DK_fit)["family", "family"]
+
+# US: Correlation coefficient
+US_est <- cor(US)[1, 2]
+US_var <- (1 - US_est^2)^2 / (nrow(US) - 1)
+
+# Name the estimates so hypotheses will be the same
+names(NL_est) <- names(DK_est) <- names(US_est) <- "parameter"
+```
+
+Then, we use `bain.default()` to evaluate the central hypothesis on each parameter estimate. 
+The pbf can then be calculated as in the previous use case.
 
 ```{r}
-US_est <- c(grp = cor(USa$soc, USa$grp))
-US_se <- sqrt(1-US_est^2) / sqrt(nrow(USa)-2) # formula for SE of correlation coefficient: sqrt(1-r^2)/sqrt(n-2)
+# Use bain.default() to obtain BF for the central hypothesis 
+NL_bain <- bain(x = NL_est, 
+                Sigma = matrix(NL_var, 1, 1),
+                n = nrow(NL),
+                hypothesis = "parameter > 0",
+                joint_parameters = 1)
+DK_bain <- bain(x = DK_est,
+                Sigma = matrix(DK_var, 1, 1),
+                n = nrow(DK),
+                hypothesis = "parameter > 0",
+                joint_parameters = 1)
+US_bain <- bain(x = US_est,
+                Sigma = matrix(US_var, 1, 1),
+                n = nrow(US),
+                hypothesis = "parameter > 0",
+                joint_parameters = 1)
 
-DK_fit <- summary(lm(soc~grp, data = DKa))
-DK_est <- c(grp = DK_fit$coefficients['grp', 'Estimate'])
-DK_se <- DK_fit$coefficients['grp', 'Std. Error']
-
-### EB: I have a question When calculating the estimate for NL.
-I assume I need the effect size of the t-test, rather than the t-statistic itself.
-This requires Cohen's d.
-Is this correct?
-In this case I do not even need to do a t-test, I only need the descriptive statistics of the two samples
-
-NLa$grp_recoded <- as.factor(sapply(NLa$grp,function(i){ ifelse(i < median(NLa$grp), 0, 1) }))
-liberals <- NLa$soc[NLa$grp_recoded == 0]
-conservatives <- NLa$soc[NLa$grp_recoded == 1]
-
-nlib <- length(liberals)
-ncon <- length(conservatives)
-pooled_sd <- sqrt(((ncon-1)*sd(conservatives) + (nlib-1)*sd(liberals)) / (nlib+ncon-2) )
-
-NL_est <- c(grp = (mean(conservatives) - mean(liberals)) / pooled_sd)
-NL_se <- sqrt(((ncon+nlib)/(ncon*nlib)) + NL_est^2/(2*(ncon+nlib))) # formula for SE of cohen's D retrieved from https://stats.stackexchange.com/questions/495015/what-is-the-formula-for-the-standard-error-of-cohens-d#:~:text=Cohen's%20d%20is%20defined%20as,%E2%88%922%20(see%20here).
-
-```
-The values are then converted to `bain` objects, put in a list, and passed to `pbf()`. 
-```{r}
-US_bain <- bain(x = US_est, Sigma = matrix(US_se, 1, 1), n = nrow(USa), hypothesis = hypotheses)
-DK_bain <- bain(x = DK_est, Sigma = matrix(DK_se, 1, 1), n = nrow(DKa), hypothesis = hypotheses)
-NL_bain <- bain(x = NL_est, Sigma = matrix(NL_se, 1, 1), n = nrow(NLa), hypothesis = hypotheses)
-
+# Aggregate evidence using pbf()
 pbf(list(US_bain, DK_bain, NL_bain))
@@
 
-The output provides evidence for or against the hypothesis: *there is a positive association between political conservatism and subjective importance of group loyalty*.
-Although each study used a different method to operationalize the evidence for that hypothesis, their evidence can be synthesized to a single value using the pbf.
+The results suggest substantial evidence in favor of the hypothesis that there is a positive association between family morality and political conservatism.
+Although each study used a different method to assess this hypothesis,
+their evidence can be synthesized using the pbf.
 

### 2022-07-20 manuscript.Rmd
begin addressing rebecca's comments

@@
 Third, when there are a few systematic differences between studies, these can be accounted for using meta-regression.
-And finally, when there are many potential variables that cause systematic differences and it is not known beforehand which are relevant, exploratory techniques like random forest meta-analysis and penalized meta-regression can be used to identify relevant moderators.
-Each of these approaches requires making different assumptions about the nature of heterogeneity.
+Finally, when there are many potential variables that cause systematic differences and it is not known beforehand which are relevant, exploratory techniques like random forest meta-analysis and penalized meta-regression can be used to identify relevant moderators.
+Each of these approaches requires making different assumptions about the nature of heterogeneity [see "Models for meta-analysis" in @vanlissaSmallSampleMetaanalyses2020].
 
-An alternative approach for evidence synthesis that does not impose such assumptions is Bayesian evidence synthesis (BES) [@kuiperCombiningStatisticalEvidence2013]. 
+An alternative approach for evidence synthesis that does not impose assumptions about heterogeneity is Bayesian evidence synthesis (BES) [@kuiperCombiningStatisticalEvidence2013]. 
 Instead of aggregating effect sizes, BES aggregates the evidence in favor of an informative hypothesis $H_i$ across studies.
 The amount of evidence for this hypothesis is expressed as a Bayes factor, BF.
-A Bayes factor can be interpreted as the ratio of evidence in favor of $H_i$ divided by evidence against it.
+A Bayes factor can be interpreted as the ratio of evidence in favor of one hypothesis $H_1$ relative to another hypothesis $H_2$, so $BF_{12} = \frac{H_1}{H_2}$.
+For the purpose of this paper, we will assume that all Bayes factors are the Bayes factors of an informative hypothesis $H_i$ against its complement $H_{!i}$, $BF_{c} = \frac{H_i}{H_!i}$. 
+In this context, the subscript $_!$ represents the negation operator; in other words, $H_{!i}$ means "not $H_i$".
+Consequently, $BF_c$ represents the ratio of evidence in favor of $H_i$ divided by evidence against it.
 Thus, BF = 10 means that the data provide ten times more support in favor of the hypothesis than against it.
-These Bayes factors can be synthesized across studies by taking their product.
+When multiple studies each provide evidence for $H_i$ in the form of complement Bayes factors,
+these Bayes factors can be synthesized across studies by taking their product.
 The resulting product Bayes factor (PBF) summarizes the total evidence for the hypothesis.
@@
 For each iteration of the simulation, we simulated a correlation coefficient in multiple samples.
+<!-- CJ: Hier staat multiple samples; elders multiple groups / studies / etc. Kan je die terminologie gelijktrekken? -->
 The informative hypothesis was kept constant at $H_i: \rho > .1$.
@@
 and once in the presence of a null effect, defined as $\rho = .1$.
-A PBF > 3 was used as a decision criterion to conclude that $H_i$ was supported.
+A PBF > 3 was used as a decision criterion to conclude that $H_i$ was supported over its complement.
+Note that this threshold is subjective but rooted in conventional standards for interpreting Bayes factors [@jeffreys1998theory].
+<!-- Rebecca: -->
+<!-- waar is > 3 op gebaseerd? -->
+<!-- Kass&Raftery?
+Als ja, bedenk wel dat dat om equaility restricties gaat (bij mijn weten).
+Je kan het natuurlijk gebruiken, maar plaats dan wel de kanttekening, maar is een uitgangspunt natuurlijk.
+-->
+<!-- Of vertel over betting odds... -->
+<!-- Maar het blijft een dichitome beslissing zo.... -->
+<!-- Laat je het gemiddelde PBF met range ofzo zien.... zo niet, dan zou dat toevoegen. Dan zie je meer dan een dichitome beslissing. -->
+<!-- En ms kan je ook ene keer 1 en 2 gebruiken, gewoon om te zien wat er gebeurt? -->
 As a benchmark for comparison, we used several other algorithms that might feasibly be used by researchers who intend to examine whether a hypothesis is true across several independent samples.
-The first was a "vote count" algorithm,
-which is common but not considered to be good practice [REF].
-This approach used one-sided z-tests to examine whether a null hypothesis corresponding to the informative hypothesis was rejected in the majority of samples, i.e.: $H_0: \rho = .1$.
+The first was vote counting: counting the number of significant effects across samples.
+Although this method is in use for aggregating conceptual replications,
+it should not be considered good practice.
+Three disadvantages are that vote counting disregards sample size,
+reduces statistical power,
+and does not quantify the strength of the evidence [(Scheerens, 2015):].
+Our vote counting algorithm used one-sided z-tests to examine whether a null hypothesis corresponding to the informative hypothesis was rejected in the majority of samples, i.e.: $H_0: \rho = .1$, and $H_a: \rho > .1$, which corresponds to $H_i$.
+Thus, for example, if $H_a$ was supported in three out of five samples, our vote counting algorithm would find overall support for $H_a$ (and, by extention, $H_i$).
 The second was a random-effects meta-analysis (RMA), which is the standard in the field.
@@
 
+# Discussion
+
+<!-- Rebecca: -->
+<!-- Ws niet een punt voor hier, maar wel in de discussie ofzo: -->
+<!-- It renders support for the hypothesis being the best in all studies. It, thus, does not render the support for the hypothesis being on average the best. -->
+
+<!-- Hier ook dan artikel van BEM noemen, maar oook zeggen dat het daar om null hypotheses gaat. -->
+
 \newpage

### 2022-07-20 manuscript.Rmd
update 2022-07-20 15:02:05

@@
 
+<!-- Rebecca: -->
+<!-- ms meer over het model vertellen. -->
+<!-- Is het voor de lezer duidelijk waar ineens deze groepen vandaan komen? Zijn dat je aantal studies die je gaat samennemen? -->
+<!-- Dus bijv: -->
+<!-- In de populatie zijn er twee variabelen, die cooreleren met de waarde rho.
+We hebben k studies die elk een sample van n hebben getrokken/verzameld.
+Dan vertelen wat de verschillende algrotithms doen met de verschillende samples (meeste als niet alle: In elke sample wordt de sample correlatie bepaald; dan estimates samenemen en null hypo test of in each sample bewijs voor hypo en dat samennemen).
+Dan iets over de verschillende waardes die je bekijkt en dat elke combi een simulatie conditie is.
+En dan iets als: Dit doen we 1000 keer per set van condities.
+-->
+<!-- En dan nog iets over reliability ook. -->
 The present simulation study set out to validate the PBF algorithm.
@@
 A PBF > 3 was used as a decision criterion to conclude that $H_i$ was supported over its complement.
-Note that this threshold is subjective but rooted in conventional standards for interpreting Bayes factors [@jeffreys1998theory].
+This threshold is subjective but rooted in conventional standards for interpreting Bayes factors [@jeffreys1998theory].
 <!-- Rebecca: -->
@@
 <!-- En ms kan je ook ene keer 1 en 2 gebruiken, gewoon om te zien wat er gebeurt? -->
-As a benchmark for comparison, we used several other algorithms that might feasibly be used by researchers who intend to examine whether a hypothesis is true across several independent samples.
-The first was vote counting: counting the number of significant effects across samples.
+<!-- Rebecca: -->
+<!-- Now, you will say that selecting > .1 with rho .1 is false, but it is not false. It is on the border, so Hi is true.  -->
+<!-- In practice it may be true half of the time, because of sampling, but still. -->
+<!-- This should be reflected by BF is approx 1. -->
+<!-- In your simulation, you use PBF > 3, so this may be hard to reach when the truth is on the border, but that is due to the subjective choice of 3... -->
+<!-- So, I already thought that using a smaller population rho value would be of added value, but now I believe you should really include this. -->
+<!-- I do see that this would end up in multiple results (once using .1 and 0.2, once (say) 0 or 0.05 and .2; but also when using other cut-off values for PBF). -->
+To examine performance in a range of realistic scenarios,
+several design factors were manipulated:
+The number of observations per sample $n \in (50, 200, 500, 800)$,
+which (assuming a known effect of $\rho = .1$) provides statistical power to reject a false null hypothesis at $\alpha = .05$ of
+$(`r paste0(formatC(pwr.r.test(c(50, 200, 500, 800), .1)$power, digits = 2, format=  "f"), collapse = ", ")`)$, respectively.
+the number of samples $k \in (2, 3, 10)$, where 2 is the minimum number of samples possible and 10 is a relatively large number, considering that replications are not very common.
+The reliability of the two correlated variables, $\alpha \in (0.6, 0.8, 1.0)$, where 0.6 is the lowest reliability conventionally considered to be acceptable, and 1 represents perfect reliability, as is
+assumed when analyzing correlations between observed items or scale scores.
+For all unique combinations of the design factors, 1000 data sets were generated.
+
+## Algorithms
+
+The main algorithm of interest was the PBF.
+As a benchmark for comparison, we included several other algorithms that might feasibly be used by researchers who intend to examine whether a hypothesis is true across several independent samples.
+The first was vote counting: counting the number of significant effects.
 Although this method is in use for aggregating conceptual replications,
@@
 For this algorithm, the null-hypothesis was rejected if a 90% confidence interval for the overall effect size excluded $H_0$.
-Third was an individual participant data (IPD) meta-analysis.
+The third algorithm was individual participant data (IPD) meta-analysis.
 Like classic meta-analysis, IPD is a multilevel model, clustered by sample.
-In contrast to classic meta-analysis, however, IPD freely estimates variance at the first level, because raw data are available.
+In contrast to classic meta-analysis, IPD freely estimates variance at the first level, because raw data are available.
 The PBF can be estimated using either sufficient statistics (as in meta-analysis) or using raw data (as in IPD).
@@
 IPD was also evaluated using a 90% confidence interval for the overall effect size.
+<!-- Rebecca: -->
+<!-- zou je dit vooraf moeten zeggen? -->
+<!-- Dan zeggen dat ILP dit ook kan en dan zeggen dus dat ook als vergelijking meenemen? -->
+<!-- Ms zelfs in intro dit al noemoen, dan zeggen dat er twee meta-an equivalenten zijn, waarin je null hypo testing kan doen. En we dat gaan meenemen om te vergelijken. -->
 
@@
 
-For each algorithm, a confusion matrix was obtained by tabulating inferential decisions made using the criteria described above against the population status of the hypothesis (true or false).
-This matrix gives the number of decisions that were true positives (TP), true negatives (TN), false positives (FP) and false negatives (FN).
-These quantities were summarized as sensitivity, $1-\frac{FN}{TP + FN}$, the probability of concluding the hypothesis is false given that it was truly false in the population, <!-- i think sensitivity is the ability to detect an effect, given that the effect is truly there.
+For each algorithm, inferential decisions made using the criteria described above were compared to the population status of the hypothesis (true or false).
+This results in a so-called "confusion matrix",
+which gives the number of decisions that were true positives (TP), true negatives (TN), false positives (FP) and false negatives (FN).
+These quantities were summarized as sensitivity, ${TP}{TP + FN}$, the probability of making the inferential decision that the informative hypothesis is supported, given that it was indeed true in the population, <!-- i think sensitivity is the ability to detect an effect, given that the effect is truly there.
 I found that its formula is usually presented as TP/(TP+FN)-->
 <!--CJ: This is all the same, see https://en.wikipedia.org/wiki/Sensitivity_and_specificity -->
-and specificity, $1-\frac{FP}{FP + TN}$, the probability of concluding the hypothesis is true given that it was also true in the population.
+and specificity, $\frac{TN}{TN+FP}$, the probability of concluding the informative hypothesis is not supported, given that it was indeed false in the population.
 <!-- and specificity is the ability to not acknowledge an effect, given that there indeed is none-->
-The overall performance was captured by the accuracy, which represents the total proportion of correct (true positive/true negative) decisions, $\frac{TP + TN}{TP+TN+FP+FN}$.
-
-## Design factors
-
-To examine performance in a range of realistic scenarios,
-several design factors were manipulated:
-The sample size per group $n \in (20, 80, 200, 500)$,
-the number of groups $k \in (2, 3, 10)$, where 2 is the minimum number of groups and 10 is a very large number of groups,
-and the reliability of the two correlated variables, $\alpha \in (0.6, 0.8, 1.0)$, where 0.6 is the lowest reliability conventionally considered to be acceptable, and 1 represents perfect reliability, as is assumed when analyzing correlations between observed items or scale scores.
-The design factors combined to produce `r out$conditions` unique conditions.
-For all simulation conditions, 1000 data sets were generated.
+The overall performance was captured by the accuracy, which represents the total proportion of correct (true positive and true negative) decisions, $\frac{TP + TN}{TP+TN+FP+FN}$.
 
@@
 First, we examined overall model performance across conditions.
-These results indicate that all algorithms except PBF had low sensitivity to detect a true effect.
+Recall that sensitivity is the probability of concluding that the informative hypothesis is supported, given that it was indeed true in the population,
+and specificity, is the probability of concluding that it is not supported, given that it was false in the population.
+The results indicate that all algorithms except PBF had low sensitivity to detect a true effect.
 In contrast, specificity was very high for all algorithms except PBF.
@@
 regardless of the existence of a population effect.
-The PBF trades a loss of specificity for increased sensitivity,
-and average levels of both were approximately equal.
-As the PBF vastly outperformed other BFs, these were omitted from further analysis.
+The PBF trades a loss of specificity for increased sensitivity.
+At the same time, the average specificity and sensitivity were approximately equal, see Table \@ref(tab:tabconf).
 
@@
 PBF performance was most impacted by sample size $n$, followed by the number of groups $k$, and reliability.
-<!-- should we also test for interaction effects? I seem to remember that there were some interesting effects of errorsd*k and n*k -->
-<!-- CJ: I don't think so, that might be more relevant for one of the papers that focused on the simulation study. For this tutorial, a simpler sim is fine. -->
+<!-- Rebecca: Nu niet compleet? -->
+<!-- Gek om steeds ene paragraaf van 1 zin te hebben iig. -->
+<!-- Verder mist er denk ik info. Ik zou graag eerst meer uitleg zien. Hoe heb je dat examine, wat ga je laten zien (niet het resultaat an sich, snap je wat ik bedoel...). -->
+<!-- Verder zou ik denk ik niet hier al weggeven wat je hierna gaat laten zien (en anders dat explicieteren dat dat is wat je doet). -->
 
@@
 
+<!-- Rebecca: -->
+<!-- ook iets zeggen over vergelijkbaarheid? -->
+<!-- Ws hierboven al iets over zeggen: -->
+<!-- In klassieke methoden heb je equality restrictie. -->
+<!-- Als rho = .1 dan null waar en moet je hem niet verwerpen, als .2 dan H0 niet waar en moet je H0 wel verwerpen. -->
+<!-- Bij BF heb je Hi en als .1 dan eigenlijk zelfs Hi waar, maar op border/boundary, dus ws dan BF = ong 1. Als .2 dan Hi waar en moet je BF > 1 vinden (evt dus > 3). -->
+
 \newpage

### 2022-07-20 manuscript.Rmd
update 2022-07-20 15:17:16

@@
 library(kableExtra)
+library(pwr)
 r_refs("r-references.bib")

### 2022-07-29 manuscript.Rmd
Start implementation of conditions into discussion

@@
 Finally, when there are many potential variables that cause systematic differences and it is not known beforehand which are relevant, exploratory techniques like random forest meta-analysis and penalized meta-regression can be used to identify relevant moderators.
-Each of these approaches requires making different assumptions about the nature of heterogeneity [see "Models for meta-analysis" in @vanlissaSmallSampleMetaanalyses2020].
+Each of these approaches require making different assumptions about the nature of heterogeneity [see "Models for meta-analysis" in @vanlissaSmallSampleMetaanalyses2020].
 
@@
 
-This tutorial paper introduces the first implementation of BES in user-friendly open source software.
+This tutorial paper introduces the first implementation of BES in useruse-friendly open source software.
 A function `pbf()` was contributed to the `bain` R-package for Bayesian informative hypothesis evaluation, version `r as.character(packageVersion("bain"))`.
@@
 Our vote counting algorithm used one-sided z-tests to examine whether a null hypothesis corresponding to the informative hypothesis was rejected in the majority of samples, i.e.: $H_0: \rho = .1$, and $H_a: \rho > .1$, which corresponds to $H_i$.
-Thus, for example, if $H_a$ was supported in three out of five samples, our vote counting algorithm would find overall support for $H_a$ (and, by extention, $H_i$).
+Thus, for example, if $H_a$ was supported in three out of five samples, our vote counting algorithm would find overall support for $H_a$ (and, by extension, $H_i$).
 The second was a random-effects meta-analysis (RMA), which is the standard in the field.
@@
 The other algorithms also show increasing sensitivity, but not specificity, which is at a ceiling.
-This difference explains the effect of reliability on the diference between algorithms (see Table \@ref(tab:tabeffect)).
+This difference explains the effect of reliability on the difference between algorithms (see Table \@ref(tab:tabeffect)).
 
@@
 
-Within the scope of this simulation, PBF had relatively better inferential properties than the other algorithms under consideration.
+This simulation study validated the product Bayes factor as an alternative method to synthesize evidence across studies.
+The PBF was benchmarked against other common methods for evidence synthesis.
+This was done by calculating the accuracy, sensitivity and specificity of the methods in different plausible varying conditions.
+
+Within the scope of this simulation, PBF had relatively better inferential properties <!-- EB: higher accuracy --> than the other algorithms under consideration.
 Although other algorithms had superior specificity, PBF had the highest levels of sensitivity.
 Thus, PBF balanced the ability to accept an informative hypothesis in the presence of a true population effect with the ability to reject the hypothesis when there was no effect.
+
+An important caveat is that none of the algorithms met conventional criteria for power (80%) and type I error (5%).
+Overall, PBF had the best performance, with a power of 76% and type I error of 24%.
+Thus, if researchers intend to synthesize evidence for an informative hypothesis across heterogeneous studies, PBF may be the most suitable method - but its limitations <!-- higher probability for type 1 errors --> should be acknowledged in applied research.
+
 <!-- Conventional threshold for type 1 error: 5% -->
@@
 
-An important caveat is that none of the algorithms met conventional criteria for power (80%) and type I error (5%).
-Overall, PBF had the best performance, with power of 76% and type I error of 24%.
-Thus, if researchers intend to synthesize evidence for an informative hypothesis across heterogeneous studies, PBF may be the most suitable method - but its limitations should be acknowledged in applied research.
+<!-- Rebecca: -->
+<!-- Ws niet een punt voor hier, maar wel in de discussie ofzo: -->
+<!-- It renders support for the hypothesis being the best in all studies. It, thus, does not render the support for the hypothesis being on average the best. -->
+
+<!-- Hier ook dan artikel van BEM noemen, maar ook zeggen dat het daar om null hypotheses gaat. -->
+
+<!-- Rebecca: -->
+<!-- ook iets zeggen over vergelijkbaarheid? -->
+<!-- Ws hierboven al iets over zeggen: -->
+<!-- In klassieke methoden heb je equality restrictie. -->
+<!-- Als rho = .1 dan null waar en moet je hem niet verwerpen, als .2 dan H0 niet waar en moet je H0 wel verwerpen. -->
+<!-- Bij BF heb je Hi en als .1 dan eigenlijk zelfs Hi waar, maar op border/boundary, dus ws dan BF = ong 1. Als .2 dan Hi waar en moet je BF > 1 vinden (evt dus > 3). -->
 
 The overall performance of PBF increased most with increasing sample size.
-With an increasing number of groups, slightly decreasing specificity was traded off for increasing sensitivity.
+PBF's sensitivity was .57 when a sample size of 20 per sample was simulated, while for all other algorithms this was below .11.
+Even at a sample size of 500, the sensitivity of the other algorithms only rose slightly above .60, 
+while for PBF this was .90.
+This provides evidence that state-of-the-art evidence synthesis methods have little power to detect a true effect.
+PBF on the other hand presents itself as a small sample size solution for evidence synthesis methods to detect true effects.
+This is reflected in the accuracy of the algorithms at small sample sizes.
+The accuracy at a sample size of 20, was around .50 for RMA, IPD and VC.
+This can be attributed to their low sensitivity and their specificity that is at ceiling level. 
+By consistently not finding an effect, the algorithms are correct 50\% of the time.
+At all sample sizes, the PBF was the most accurate algorithm,
+although it should be noted that even at a sample size of 500, PBFs specificity was .12 percentage points too low for conventional acceptable type I error rates.
+
+With an increasing number of groups, slightly decreasing specificity was traded off for increasing sensitivity.<!-- I would argue that the decrease is negligable, 0.767 at k =2, 0.763 at k = 10 -->
+With an increasing number of groups, specificity remained around .76 for PBF, while sensitivity increased to .87.
+Again, the specificity remained at ceiling level for the other algorithms.
+The sensitivity for RMA and IPD increased, but stayed below .51.
+In a similar way, this supports the use of the PBF as a small sample size solution.
+In this case when there are little available studies to synthesize evidence over.
+Even with as much as ten available studies, PBF was the only algorithm that has a conventionally acceptable power.
+
 With increasing reliability, increasing specificity was traded off for decreasing sensitivity.
+
+There was no marginal condition in which the PBF had a conventional acceptable type I error rate.
+Its mean was .76 which is almost .20 percentage points below .95.
+However, the other algorithms had no marginal condition where an acceptable power was achieved.
+The PBF can be used as a small sample size solution,
+although its increased type I error rate must be taken into account.
+It implies that PBF works best as a complementary method for evidence synthesis, rather than a replacement.
 
@@
 
-# Discussion
-
-<!-- Rebecca: -->
-<!-- Ws niet een punt voor hier, maar wel in de discussie ofzo: -->
-<!-- It renders support for the hypothesis being the best in all studies. It, thus, does not render the support for the hypothesis being on average the best. -->
-
-<!-- Hier ook dan artikel van BEM noemen, maar oook zeggen dat het daar om null hypotheses gaat. -->
-
-<!-- Rebecca: -->
-<!-- ook iets zeggen over vergelijkbaarheid? -->
-<!-- Ws hierboven al iets over zeggen: -->
-<!-- In klassieke methoden heb je equality restrictie. -->
-<!-- Als rho = .1 dan null waar en moet je hem niet verwerpen, als .2 dan H0 niet waar en moet je H0 wel verwerpen. -->
-<!-- Bij BF heb je Hi en als .1 dan eigenlijk zelfs Hi waar, maar op border/boundary, dus ws dan BF = ong 1. Als .2 dan Hi waar en moet je BF > 1 vinden (evt dus > 3). -->
-
 \newpage

### 2022-08-01 manuscript.Rmd
elaborate further on effect of design factors

@@
 
+<!--EB: I think reliability plot is a bit confusing (The one that is saved under ./manuscript_files/figure-latex/unnamed-chunk-3-1.png).
+That is because the x-axis is labeled 'reliability', but the values are the values of the 'errorsd' condition.
+This makes that the reliability decreases, rather than increases as you move away from the origin on the x-axis -->
 ```{r, fig.cap="Mean performance by reliability"}
@@
 This was done by calculating the accuracy, sensitivity and specificity of the methods in different plausible varying conditions.
+During the simulation the hypotheses remained constant $H_i: \rho > .1$, $H_0: \rho = .1$.
+To be able to calculate both sensitivity and specificity, half of the simulated conditions contained an effect, $\rho = .2$, while the other half did not, $\rho = .1$.
+<!-- EB: I don't know why we did not use H0:r=0 and Hi:r>0 as constant hypotheses. It seems more intuitive -->
+Sensitivity and specificity can be interpreted as Type I error rate and statistical power respectively and are used interchangeably.
+ <!-- Because Type 1 error rate and power seem to be key terms in the discussion, we may be able to include it in the introduction: 'How does the PBF compare to other state-of-the-art algorithms with regards to Type 1 error rate, power and overall accuracy?' -->
 
+The overall performance of PBF increased most with increasing sample size.
+ <!-- EB: also, shall we rename 'performance' to 'accuracy', I know its not the same, but performance seems a bit too subjective.
+After all, we do not have a clear cut decision when sensitivity and specificity is balanced 'correctly' -->
+PBF's sensitivity was .57 when a sample size of 20 per sample was simulated, while for all other algorithms this was below .11.
+It should be noted that in the condition where sensitivity could be calculated, the true effect $\rho$ was .2 and only .1 points above the boundary value for $H_i > .1$ and $H_0 = .1$.
+It is no surprise that at smaller samples, the small difference in effect size between null/informative hypothesis and true value of .1, is unlikely to be detected.
+Still, even at a relatively large sample size of 500 per sample, the sensitivity of the other algorithms only rose slightly above .60, 
+while for PBF this was .90.
+This provides evidence that state-of-the-art evidence synthesis methods have little power to detect a true, but small effect,
+even when the sample size per sample was substantial.
+PBF on the other hand presents itself as a small sample size solution for evidence synthesis methods to detect true effects,
+even if the effect itself can be considered small.
+This is also reflected in the accuracy of the algorithms at small sample sizes, which, at a sample size of 20, was around .50 for RMA, IPD and VC.
+This can be attributed to their low sensitivity and their specificity that is at ceiling level. 
+By consistently not finding an effect, the algorithms are correct 50\% of the time.
+At all sample sizes, the PBF was the most accurate algorithm,
+although it should be noted that even at a sample size of 500, PBFs specificity was .12 percentage points too high for conventional acceptable type I error rates of .05.
+
+With an increasing number of groups, slightly decreasing specificity was traded off for increasing sensitivity.<!--EB: I would argue that the decrease is negligible, 0.767 at k =2, 0.763 at k = 10 -->
+With an increasing number of groups, specificity remained around .76 for PBF, while sensitivity increased to .87.
+Again, the specificity remained at ceiling level for the other algorithms.
+The sensitivity for RMA and IPD increased, but stayed below .51.
+In a similar way, this supports the use of the PBF as a small sample size solution.
+In this case when there are little available studies to synthesize evidence over.
+It could be argued that when there are only 2 or 3 studies available to synthesize evidence over, that if a trade-off need be made, specificity is preferred over sensitivity.
+It seems unwise to more easily acknowledge an effect if there is little data available.
+However, the quality of the included studies should be considered before synthesizing evidence in any way.
+Moreover, when the researcher performs some form of evidence synthesis, then the ability of the method to find an effect is a necessary and fair component, even if little studies are available.
+Also, even with as much as ten available studies, PBF was the only algorithm that had a conventionally acceptable power.
+
+With increasing reliability, increasing specificity was traded off for decreasing sensitivity for PBF.
+<!--EB: I think its the other way around; with DECREASING reliability, specificity increases and sensitivity decreases, see my comment at the section: 'Effect of reliability' -->
+With increasing reliability, decreasing specificity was traded off for increasing sensitivity for the PBF.
+Also for the other algorithms, the sensitivity increased as the reliability increased.
+It implies that the PBF gets more conservative whenever reliability decreases.
+This can be considered beneficial as the PBF is more hesitant to acknowledge an effect as the measurements become less stable.
+Even at the smallest acceptable reliability, $\alpha = .6$, PBFs sensitivity was .57, where it was below .10 for the other algorithms.
+However, the specificity of the PBF when there is perfect reliability $\alpha=1$ was .63, were it was almost 1 for the other algorithms.
+
+Overall, there was only a single marginal condition where the accuracy of PBF was lower than that of any other algorithm.
+PBF scored .01-.02 percentage points lower accuracy than IPD and RMA when there was perfect reliability.
+Looking at the specificity and sensitivity however, this is explained by the relatively low specificity of PBF and the very high specificity of RMA and IPD.
+Thus,
 Within the scope of this simulation, PBF had relatively better inferential properties <!-- EB: higher accuracy --> than the other algorithms under consideration.
@@
 
-An important caveat is that none of the algorithms met conventional criteria for power (80%) and type I error (5%).
+An important caveat is that none of the algorithms met conventional criteria for power (80%) and type I error (5%) [@cohen2013statistical].
 Overall, PBF had the best performance, with a power of 76% and type I error of 24%.
 Thus, if researchers intend to synthesize evidence for an informative hypothesis across heterogeneous studies, PBF may be the most suitable method - but its limitations <!-- higher probability for type 1 errors --> should be acknowledged in applied research.
+
+As stated earlier, the PBF can be used as a small sample size solution,
+although its increased type I error rate must be taken into account.
+It implies that PBF works best as a complementary method for evidence synthesis, rather than a replacement.
 
@@
 <!-- Bij BF heb je Hi en als .1 dan eigenlijk zelfs Hi waar, maar op border/boundary, dus ws dan BF = ong 1. Als .2 dan Hi waar en moet je BF > 1 vinden (evt dus > 3). -->
-
-The overall performance of PBF increased most with increasing sample size.
-PBF's sensitivity was .57 when a sample size of 20 per sample was simulated, while for all other algorithms this was below .11.
-Even at a sample size of 500, the sensitivity of the other algorithms only rose slightly above .60, 
-while for PBF this was .90.
-This provides evidence that state-of-the-art evidence synthesis methods have little power to detect a true effect.
-PBF on the other hand presents itself as a small sample size solution for evidence synthesis methods to detect true effects.
-This is reflected in the accuracy of the algorithms at small sample sizes.
-The accuracy at a sample size of 20, was around .50 for RMA, IPD and VC.
-This can be attributed to their low sensitivity and their specificity that is at ceiling level. 
-By consistently not finding an effect, the algorithms are correct 50\% of the time.
-At all sample sizes, the PBF was the most accurate algorithm,
-although it should be noted that even at a sample size of 500, PBFs specificity was .12 percentage points too low for conventional acceptable type I error rates.
-
-With an increasing number of groups, slightly decreasing specificity was traded off for increasing sensitivity.<!-- I would argue that the decrease is negligable, 0.767 at k =2, 0.763 at k = 10 -->
-With an increasing number of groups, specificity remained around .76 for PBF, while sensitivity increased to .87.
-Again, the specificity remained at ceiling level for the other algorithms.
-The sensitivity for RMA and IPD increased, but stayed below .51.
-In a similar way, this supports the use of the PBF as a small sample size solution.
-In this case when there are little available studies to synthesize evidence over.
-Even with as much as ten available studies, PBF was the only algorithm that has a conventionally acceptable power.
-
-With increasing reliability, increasing specificity was traded off for decreasing sensitivity.
-
-There was no marginal condition in which the PBF had a conventional acceptable type I error rate.
-Its mean was .76 which is almost .20 percentage points below .95.
-However, the other algorithms had no marginal condition where an acceptable power was achieved.
-The PBF can be used as a small sample size solution,
-although its increased type I error rate must be taken into account.
-It implies that PBF works best as a complementary method for evidence synthesis, rather than a replacement.
 

### 2022-08-10 manuscript.Rmd
Read bem articles, add comments

@@
 To be able to calculate both sensitivity and specificity, half of the simulated conditions contained an effect, $\rho = .2$, while the other half did not, $\rho = .1$.
-<!-- EB: I don't know why we did not use H0:r=0 and Hi:r>0 as constant hypotheses. It seems more intuitive -->
+<!-- EB: Why did we not use H0:r=0 and Hi:r>0 as constant hypotheses? It seems more intuitive -->
 Sensitivity and specificity can be interpreted as Type I error rate and statistical power respectively and are used interchangeably.
@@
 It should be noted that in the condition where sensitivity could be calculated, the true effect $\rho$ was .2 and only .1 points above the boundary value for $H_i > .1$ and $H_0 = .1$.
-It is no surprise that at smaller samples, the small difference in effect size between null/informative hypothesis and true value of .1, is unlikely to be detected.
+It is no surprise that at smaller samples the small difference in effect size between null/informative hypothesis and true value of .1 is unlikely to be detected.
 Still, even at a relatively large sample size of 500 per sample, the sensitivity of the other algorithms only rose slightly above .60, 
@@
 
-<!-- Hier ook dan artikel van BEM noemen, maar ook zeggen dat het daar om null hypotheses gaat. -->
+<!-- EB: I assume Rebecca means that 'the hypothesis being on average the best' is the hypothesis that reflects the truth.
+While 'the hypothesis that is best in all studies' is the hypothesis that is closest to the truth compared to the other hypotheses that were tested.
+It does not say anything necessarily about whether the tested hypotheses in general are good reflections of the truth.
+
+<!--Rebecca: Hier ook dan artikel van BEM noemen, maar ook zeggen dat het daar om null hypotheses gaat. -->
+
+<!--EB: dit refereert naar Rouder, J. N., & Morey, R. D. (2011). A Bayes factor meta-analysis of Bem’s ESP claim. Psychonomic Bulletin & Review, 18(4), 682-689.
+
+The argument from Rouder and Morey (2011) is that we cannot take the PBF as a valid measure for meta-analysis.
+They show that calculating the BF in multiple samples and taking their product results in a different PBF than when all data is pooled and the BF is calculated after pooling.
+However, they assume a constant effect where all data is drawn from a common distribution The idiosyncratic differences between studies makes that pooling data is not a straightforward endeavour, because of the heterogeneity of the effect.
+Rouder and Morey do acknowledge their choice for a constant effect to be debatable.
+
+Quote from the Rouder and Morey (2011) paper:
+"It is reasonable to wonder whether the constant effect
+size model underlying the meta-analytic Bayes factor is
+warranted. We chose this approach because it is tractable
+when researchers have access to the test statistics, rather
+than the raw data."
+
+Interesting statement.
+Intuitively I do not see how having only the test statistics compared to raw data makes the assumption of a constant effect more probable.
+After all, the test statistic is derived from the raw data...
+
+Rebecca noticed that Rouder and Morey specifically talk about PBF not being a good measurement in the context of testing H0, which is something that we are not doing.
+However, the only relevant difference I see in their vs our method is that their H0 contains a single equality constraint, while we used inequality constraints.
+Rouder and Morey do not explicitly point this difference out and I dont see how the PBF would be a bad or good idea depending on type of constraints.
+-->
 

### 2022-08-19 manuscript.Rmd
further elaborate on Rebeccas comments

@@
 
-This simulation study validated the product Bayes factor as an alternative method to synthesize evidence across studies.
+This simulation study validated the product Bayes factor as an alternative or complementary method to synthesize evidence across studies.
 The PBF was benchmarked against other common methods for evidence synthesis.
@@
  <!-- Because Type 1 error rate and power seem to be key terms in the discussion, we may be able to include it in the introduction: 'How does the PBF compare to other state-of-the-art algorithms with regards to Type 1 error rate, power and overall accuracy?' -->
+The effect of sample size per study, number of studies and reliability of measurement on the performance of the algorithms was investigated.
 
-The overall performance of PBF increased most with increasing sample size.
+The overall performance of the PBF increased most with increasing sample size.
  <!-- EB: also, shall we rename 'performance' to 'accuracy', I know its not the same, but performance seems a bit too subjective.
@@
 Overall, there was only a single marginal condition where the accuracy of PBF was lower than that of any other algorithm.
-PBF scored .01-.02 percentage points lower accuracy than IPD and RMA when there was perfect reliability.
+PBF scored 1-2 percentage points lower accuracy than IPD and RMA when there was perfect reliability.
 Looking at the specificity and sensitivity however, this is explained by the relatively low specificity of PBF and the very high specificity of RMA and IPD.
@@
 
-<!-- EB: I assume Rebecca means that 'the hypothesis being on average the best' is the hypothesis that reflects the truth.
+<!-- EB: I assume Rebecca means that we should note that 'the hypothesis being on average the best' is the hypothesis that reflects the truth.
 While 'the hypothesis that is best in all studies' is the hypothesis that is closest to the truth compared to the other hypotheses that were tested.
 It does not say anything necessarily about whether the tested hypotheses in general are good reflections of the truth.
+E.g, if we only test whether the earth is either flat-shaped or triangular and not acknowledge other options, then we would incorrectly conclude that the earth is flat.-->
+The interpretation of the PBF should also be considered.
+The PBF renders support for some hypothesis being the best in all studies, but only compared to other hypotheses it has been tested against.
+This means that the tested hypothesis with most support is not necessarily the hypothesis that reflects the truth most closely.
+This is because there are many other plausible hypotheses that have not been tested.
+For example, testing whether the earth is flat or triangular excludes the possibility of a spherical earth.
 
@@
 They show that calculating the BF in multiple samples and taking their product results in a different PBF than when all data is pooled and the BF is calculated after pooling.
-However, they assume a constant effect where all data is drawn from a common distribution The idiosyncratic differences between studies makes that pooling data is not a straightforward endeavour, because of the heterogeneity of the effect.
+They also show another way of calculating a meta-analytic BF and that this pooled BF comes closer to that meta-analytic BF.
+However, this calculation assume a constant population effect and the same type of statistical model across studies
+.
+where all data is drawn from a common distribution The idiosyncratic differences between studies makes that pooling data is not a straightforward endeavour, because of the heterogeneity of the effect.
 Rouder and Morey do acknowledge their choice for a constant effect to be debatable.
@@
 However, the only relevant difference I see in their vs our method is that their H0 contains a single equality constraint, while we used inequality constraints.
-Rouder and Morey do not explicitly point this difference out and I dont see how the PBF would be a bad or good idea depending on type of constraints.
--->
+Rouder and Morey do not explicitly point this difference out and I don't see how the PBF would be a bad or good idea depending on type of constraints.
+
+They also do it for a single parameter t, but it would not work if there are multiple different units.-->
+
+The article of @rouder2011bayes states that a meta-analytic Bayes factor is not as straightforward as taking the product of $k$ individual Bayes factors.
+Rather, they provide a way of calculating more valid a meta-analytic Bayes factor.
+However, only under the assumption that the population effect is constant and the test statistic is from the same type of statistical model.
+Using this form of the PBF, they show that pooling the data and calculating a single Bayes factor on the pooled data comes closer to the meta-analytic Bayes factor than taking the product of each individual study.
+However, as stated earlier, one strength of the PBF lies in its ability to synthesize evidence for a hypothesis, rather than an estimate.
+This means that there is no assumption of a constant population effect and/or equal types of test statistics across studies.
+Furthermore, as the article of @rouder2011bayes states itself, pooling data is valid if all data is independent and drawn from a common distribution.
 
@@
 <!-- Bij BF heb je Hi en als .1 dan eigenlijk zelfs Hi waar, maar op border/boundary, dus ws dan BF = ong 1. Als .2 dan Hi waar en moet je BF > 1 vinden (evt dus > 3). -->
+
+<!--EB: What I think Rebecca means here is the following: 
+In classical NHST, there is an equality resitriction.
+we either do or do not reject H0, so either rho = x or rho != x.
+Rejecting H0, means that the data suggests that rho != .1, but rather something else than .1 (so .2, .9, -.9 etc.)
+With RMA and IPD we effectively say, rho = .1 or rho != .1 which is true or not (even if the test statistic = -.8,than .1 is not in the 90% CI and we still H0, even though its in the opposite direction.)
+
+with PBF we use inequality restrictions, because an equality restriction is too specific and is almost always an inferior hypothesis.
+With PBF we check if rho > .1 or rho <= .1, this means that if rho = -.8, we accept Hc and not Hi.
+It is thus hard to compare the RMA/IPD and PBF decisions to reject or accept H0.
+
+if rho = .1 and Rho = .1, than rejecting H0 is wrong. 
+if rho = .2, and Rho = .1 than we must reject H0. -->
 

### 2023-01-23 manuscript.Rmd

@@
 The PBF was benchmarked against other common methods for evidence synthesis.
-This was done by calculating the accuracy, sensitivity and specificity of the methods in different plausible varying conditions.
-During the simulation the hypotheses remained constant $H_i: \rho > .1$, $H_0: \rho = .1$.
-To be able to calculate both sensitivity and specificity, half of the simulated conditions contained an effect, $\rho = .2$, while the other half did not, $\rho = .1$.
+This was done by calculating a correlation coefficient and determining the accuracy, sensitivity and specificity as metrics of performance for the methods in different plausible varying conditions.
+During the simulation, the hypotheses that were tested remained constant $H_i: \rho > .1$, $H_0: \rho = .1$.
+To be able to calculate the performance metrics, half of the simulated conditions contained an effect, $\rho = .2$, while the other half did not, $\rho = .1$.
 <!-- EB: Why did we not use H0:r=0 and Hi:r>0 as constant hypotheses? It seems more intuitive -->
@@
  <!-- Because Type 1 error rate and power seem to be key terms in the discussion, we may be able to include it in the introduction: 'How does the PBF compare to other state-of-the-art algorithms with regards to Type 1 error rate, power and overall accuracy?' -->
-The effect of sample size per study, number of studies and reliability of measurement on the performance of the algorithms was investigated.
+The effect of sample size per study, number of studies to synthesize evidence over and reliability of measurement on the performance of the algorithms was investigated.
 
@@
 
-With an increasing number of groups, slightly decreasing specificity was traded off for increasing sensitivity.<!--EB: I would argue that the decrease is negligible, 0.767 at k =2, 0.763 at k = 10 -->
-With an increasing number of groups, specificity remained around .76 for PBF, while sensitivity increased to .87.
+With an increasing number of groups, slightly decreasing specificity was traded off for increasing sensitivity.<!--EB: I would argue that the decrease is negligible, 0.767 at k = 2, 0.763 at k = 10 -->
+With an increasing number of groups, specificity remained around .76-.77 for PBF, while sensitivity increased to .87.
 Again, the specificity remained at ceiling level for the other algorithms.
@@
 Moreover, when the researcher performs some form of evidence synthesis, then the ability of the method to find an effect is a necessary and fair component, even if little studies are available.
+<!--EB: for clarification: I mean to say that research should not be done if there is effectively no power to detect an effect anyway, because the outcome is in such a case practically pre-determined.
+-->
 Also, even with as much as ten available studies, PBF was the only algorithm that had a conventionally acceptable power.
@@
 PBF scored 1-2 percentage points lower accuracy than IPD and RMA when there was perfect reliability.
-Looking at the specificity and sensitivity however, this is explained by the relatively low specificity of PBF and the very high specificity of RMA and IPD.
+Looking at the specificity and sensitivity however, this is mostly explained by the specificity at ceiling level of RMA and IPD.
 Thus,
@@
 
-As stated earlier, the PBF can be used as a small sample size solution,
+Thus, the PBF can be used as a small sample size solution,
 although its increased type I error rate must be taken into account.
@@
 For example, testing whether the earth is flat or triangular excludes the possibility of a spherical earth.
+Still, because a Bayes factor is a continuous, real numbered quantity, it is more informative than the conclusion from a null-hypothesis significance test that only has the ability to dichotomously evaluate the null-hypothesis.
+The main benefit is the diversity of hypotheses that can be tested using the PBF.
 
@@
 They also show another way of calculating a meta-analytic BF and that this pooled BF comes closer to that meta-analytic BF.
-However, this calculation assume a constant population effect and the same type of statistical model across studies
-.
-where all data is drawn from a common distribution The idiosyncratic differences between studies makes that pooling data is not a straightforward endeavour, because of the heterogeneity of the effect.
+However, this calculation assumes a constant population effect and the use of the same type of statistical model across studies.
+Also, pooling the data assumes that all data is drawn from a common distribution.
+The idiosyncratic differences between studies makes that pooling data is not a straightforward endeavour, because of the heterogeneity of the effect.
 Rouder and Morey do acknowledge their choice for a constant effect to be debatable.
@@
 Rouder and Morey do not explicitly point this difference out and I don't see how the PBF would be a bad or good idea depending on type of constraints.
+-->
 
-They also do it for a single parameter t, but it would not work if there are multiple different units.-->
-
-The article of @rouder2011bayes states that a meta-analytic Bayes factor is not as straightforward as taking the product of $k$ individual Bayes factors.
+One possible issue with the PBF as defined in this research comes from an article of @rouder2011bayes. 
+The article states that a meta-analytic Bayes factor is not as straightforward as taking the product of $k$ individual Bayes factors.
+<!--EB: Their reason seems to be that Bayes factors take sample size into account, effectively reducing the ability of a Bayes factor to recognize an effect in small sample.
+I dont see how this invalidates the PBF.
+If there is an effect, then in 5 small studies, it may be true that a BF is calculated that barely support the hypothesis of an effect.
+However, taking the product of 5 BFs that are all 1.25 will result in a PBF of about 3.-->
 Rather, they provide a way of calculating more valid a meta-analytic Bayes factor.
@@
 This means that there is no assumption of a constant population effect and/or equal types of test statistics across studies.
-Furthermore, as the article of @rouder2011bayes states itself, pooling data is valid if all data is independent and drawn from a common distribution.
+Furthermore, as the article of @rouder2011bayes states itself, pooling data is preferred if all data is independent and drawn from a common distribution. 
+The study idiosyncrasies encountered in evidence synthesis should wary the researcher in blindly making the assumption of a common distribution.
 
@@
 <!--EB: What I think Rebecca means here is the following: 
-In classical NHST, there is an equality resitriction.
-we either do or do not reject H0, so either rho = x or rho != x.
-Rejecting H0, means that the data suggests that rho != .1, but rather something else than .1 (so .2, .9, -.9 etc.)
-With RMA and IPD we effectively say, rho = .1 or rho != .1 which is true or not (even if the test statistic = -.8,than .1 is not in the 90% CI and we still H0, even though its in the opposite direction.)
+In classical NHST, there is an equality restriction. we either do (r = .2) or do not (r = .1) reject H0. 
+Rejecting H0 means that the data suggests that rho != .1, but rather something else than .1 (so .2, .9, -.9 etc.)
+With RMA and IPD we effectively say, rho = .1 or rho != .1 which is true or not (even if the test statistic = -.8,than .1 is not in the 90% CI and we still reject  H0, even though the found effect is in the opposite direction of what we hypothesized.)
 
@@
 
-if rho = .1 and Rho = .1, than rejecting H0 is wrong. 
-if rho = .2, and Rho = .1 than we must reject H0. -->
+This would imply that there should be more cases where the frequentist methods acknowledge an effect.
+However, we see a very low sensitivity due to the small effect, so its not really a concern in this paper-->
+A final remark is that of performance comparison between PBF and the other algorithms which has to do with the differences in hypothesis specification. 
+For the frequentist and PBF algorithms we specified $H_0: \rho = .1$ and $H_c: \rho \leq .1$ respectively.
+This means, hypothetically, that if a strong effect in the opposite direction was found, let's say $\rho = -.6$, the frequentist algorithms would reject $H_0$ while the PBF would accept $H_c$.
+This implies that there should have been a higher power to detect an effect in the frequentist framework, as negative effects would reject $H_0$.
+This is contrary to what was observed in this research, however, as the frequentist approaches had a low sensitivity.
+Again, this has to do with the small simulated difference between the true effect and the boundary value.
 

### 2023-03-20 manuscript.Rmd
update 2023-03-20 14:39:27

--- 2022-08-10 manuscript.Rmd
@@
 ---
-title             : "Tutorial: Aggregate evidence from heterogeneous replication studies using the product Bayes factor"
+title             : "Aggregating evidence from conceptual replication studies using the product Bayes factor"
 shorttitle        : "PRODUCT BAYES FACTOR"
@@
   - name: "Eli-Boaz Clapper"
-    affiliation: "1"
+    affiliation: "2"
     role:
@@
       - Writing – review & editing
+  - name: "Rebecca Kuiper"
+    affiliation: "2"
+    role:
+      - Conceptualization
+      - Writing – review & editing
 affiliation:
   - id            : "1"
+    institution   : "Tilburg University, dept. Methodology & Statistics"
+  - id            : "2"
     institution   : "Utrecht University, dept. Methodology & Statistics"
-  - id            : "2"
-    institution   : "Open Science Community Utrecht"
 
@@
 abstract: |
-  TODO
+  The product Bayes factor (PBF) can synthesize evidence in favor of an informative hypothesis across heterogeneous replication studies.
+It is particularly useful when the number of studies is relatively low and conventional assumptions about between-studies heterogeneity are likely violated.
+The present paper introduces a user-friendly implementation of the PBF in the bain R-package, and demonstrates its use in several tutorial examples.
+The method was validated in a simulation study that manipulated sample size, number of replication samples, and reliability.
+PBF had a higher overall accuracy when benchmarked against other evidence synthesis methods, including random-effects meta-analysis (RMA).
+This was primarily due to PBF’s greater sensitivity in detecting a true effect.
+However, PBF had relatively lower specificity.
+The PBF showed increasing sensitivity and specificity with increasing sample size.
+With an increasing number of samples, lower sensitivity was traded for greater specificity.
+Although PBF's overall performance was less susceptible to reliability than the other algorithms, this masked a trade-off between reliability and specificity.
+PBF thus appears to be a promising method for meta-analysis of heterogeneous conceptual replication studies.
+Nonetheless, users should be aware of its lower specificity, and the fact that it takes a Bayesian approach to inference and addresses a qualitatively different research question than other evidence synthesis methods.
 
@@
 
-bibliography      : ["product_bayes.bib"]
+bibliography      : ["product_bayes.bib", "references.bib"]
 
@@
 ```{r setup, include = FALSE}
-run_everything <- FALSE
+run_everything <- TRUE
 library("papaja")
@@
 library(pwr)
+library(bain)
 r_refs("r-references.bib")
+source("UI function/pbf.R")
 out <- readRDS("out.RData")
@@
 Recent years have seen a crisis of confidence over the reliability of published results in psychology, and science more broadly [@brembs2018prestigious].
-Replication research has come into focus as a solution to this crisis and as a way to derive knowledge that will stand the test of time [see @lavelleWhenCrisisBecomes2021]. 
-<!--For many, highly regarded published studies however, results could not be replicated [@brembs2018prestigious].-->
-<!--A lack of validating replication research partly strengthened the belief that single studies results should solely be viewed as part of many studies studying the same research question [@asendorpf2016recommendations].
--->
-<!--Evidence over multiple studies must be aggregated to derive more reliable conclusions and within this paradigm, research synthesis methods have blossomed [@shadish2015meta; @nakagawa2019research].-->
+Replication research has emerged as one potential way to address this crisis and derive knowledge that will stand the test of time [see @lavelleWhenCrisisBecomes2021]. 
 In step with this interest in replication research,
-research synthesis methods have blossomed [REF].
-These methods aggregate the findings of multiple studies,
-and thus enable drawing overarching conclusions across multiple studies.
-The present paper focuses on quantitative methods for research synthesis [e.g., @vanlissaSelectRelevantModerators2021], although qualitative [e.g., @vanlissaMappingPhenomenaRelevant2021] research synthesis methods also exist.
+research synthesis methods have become increasingly popular.
+These methods aggregate research findings,
+and thus enable drawing overarching conclusions across multiple (replication) studies.
+In this paper, we introduce a new research synthesis method that uses Bayesian evidence synthesis to aggregate the evidence in favor of an informative hypothesis, quantified by the Bayes factor,
+across multiple studies.
+This method has the potential to provide a more comprehensive and accurate picture of the state of the literature, and to identify areas of consensus and disagreement among studies.
+We describe the method in detail and demonstrate its application through a tutorial example analysis.
 
 A key challenge in quantitative research synthesis is dealing with between-studies heterogeneity [@higginsReevaluationRandomeffectsMetaanalysis2009].
-Heterogeneity appears when, for example, studies examine the same research question in different laboratories, use idiosyncratic methods, and sample from distinct populations.
+When studies examine the same research question in different laboratories, use idiosyncratic methods, and sample from distinct populations, these between-study differences can introduce heterogeneity in findings.
 The most common quantitative research synthesis method is meta-analysis,
@@
 In meta-analysis, heterogeneity can be accounted for in four ways [see @vanlissaSmallSampleMetaanalyses2020].
-First, if studies are exact replications, one may assume that no heterogeneity in the outcome exists and a fixed-effect meta-analysis can be conducted to estimate the common population effect. 
-Second, when heterogeneity between studies can be assumed to be because of random measurement error, random-effects meta-analysis can be used to estimate the mean of a distribution of population effects.
+First, if studies are exact replications,
+one may assume that no heterogeneity in the outcome exists and a fixed-effect meta-analysis can be conducted to estimate the common population effect. 
+Second, when heterogeneity between studies can be assumed to be random,
+random-effects meta-analysis can be used to estimate the mean of a distribution of population effects. 
 Third, when there are a few systematic differences between studies, these can be accounted for using meta-regression.
-Finally, when there are many potential variables that cause systematic differences and it is not known beforehand which are relevant, exploratory techniques like random forest meta-analysis and penalized meta-regression can be used to identify relevant moderators.
-Each of these approaches require making different assumptions about the nature of heterogeneity [see "Models for meta-analysis" in @vanlissaSmallSampleMetaanalyses2020].
+Finally, when there are many potential variables that cause systematic differences and it is not known beforehand which are relevant, exploratory techniques like random forest meta-analysis and penalized meta-regression can be used to identify relevant moderators [@vanlissaSelectRelevantModerators2021].
+However, accounting for moderators requires a relatively high number of observations per moderator,
+which may not be available.
+Each of these approaches thus makes different assumptions about the nature of heterogeneity [see "Models for meta-analysis" in @vanlissaSmallSampleMetaanalyses2020].
 
-An alternative approach for evidence synthesis that does not impose assumptions about heterogeneity is Bayesian evidence synthesis (BES) [@kuiperCombiningStatisticalEvidence2013]. 
-Instead of aggregating effect sizes, BES aggregates the evidence in favor of an informative hypothesis $H_i$ across studies.
-The amount of evidence for this hypothesis is expressed as a Bayes factor, BF.
-A Bayes factor can be interpreted as the ratio of evidence in favor of one hypothesis $H_1$ relative to another hypothesis $H_2$, so $BF_{12} = \frac{H_1}{H_2}$.
-For the purpose of this paper, we will assume that all Bayes factors are the Bayes factors of an informative hypothesis $H_i$ against its complement $H_{!i}$, $BF_{c} = \frac{H_i}{H_!i}$. 
-In this context, the subscript $_!$ represents the negation operator; in other words, $H_{!i}$ means "not $H_i$".
+A crucial shortcoming of existing research synthesis methods is that the aforementioned assumptions
+may not be tenable when meta-analyzing studies that investigate the same informative hypothesis,
+but are otherwise very heterogeneous.
+The situation may arise where each study is uniquely identified by a combination of linearly dependent moderators.
+In this case, it is no longer possible to synthesize *effect sizes* while accounting for heterogeneity using statistical methods.
+<!-- even machine learning based methods cannot identify relevant moderators. -->
+It is still possible, however, to quantify the support these studs provide for the underlying informative hypothesis.
+To this end, Bayesian evidence synthesis (BES) aggregates the evidence in favor of an informative hypothesis $H_i$ across studies,
+without imposing assumptions about heterogeneity [@kuiperCombiningStatisticalEvidence2013].
+
+The amount of evidence for a hypothesis can be expressed as a Bayes factor, or BF.
+The BF can be interpreted as the ratio of evidence in favor of one hypothesis $H_1$ relative to another hypothesis $H_2$, so $BF_{12} = \frac{H_1}{H_2}$.
+Within the scope of this paper,
+all Bayes factors are the ratio of evidence in favor of an informative hypothesis $H_i$ relative to its complement $H_{!i}$, $BF_{c} = \frac{H_i}{H_!i}$. 
+The subscript $_!$ represents the negation operator;
+in other words, $H_{!i}$ means "not $H_i$".
 Consequently, $BF_c$ represents the ratio of evidence in favor of $H_i$ divided by evidence against it.
-Thus, BF = 10 means that the data provide ten times more support in favor of the hypothesis than against it.
+A value of $BF_c = 10$ means that the data provide ten times more support in favor of the hypothesis than against it.
+
 When multiple studies each provide evidence for $H_i$ in the form of complement Bayes factors,
-these Bayes factors can be synthesized across studies by taking their product.
+these Bayes factors can be synthesized across studies by taking their product [@kuiperCombiningStatisticalEvidence2013].
 The resulting product Bayes factor (PBF) summarizes the total evidence for the hypothesis.
-Note that other approaches to BES exist [see "Bayesian Evidence Synthesis" in @heckReviewApplicationsBayes2022].
-The only assumption of BES is that all study-specific hypotheses provide evidence about the same underlying theoretical relationship.
+The only assumption of the PBF is that all study-specific hypotheses provide evidence about the same underlying theoretical relationship.
+Note that other approaches to BES exist;
+for instance, it is possible to use the posterior of one study as the prior for a replication study, and thus accumulate evidence across studies [see @heckReviewApplicationsBayes2022].
+Such applications are out of scope of the present paper,
+which addresses the PBF approach to BES.
 
 Although meta-analysis and BES are both research synthesis methods, they answer different research questions.
-Meta-analysis estimates the value or distribution of a population effect size. 
+Meta-analysis estimates the point estimate or distribution of a population effect size. 
 It pools estimates of this effect size across multiple studies to obtain an overall estimate of the effect size.
-It thus answers the question: Given certain assumptions about between-studies heterogeneity, what is the average population effect size?
+It thus answers questions like:
+Given certain assumptions about between-studies heterogeneity, what is the average population effect size?
 BES, on the other hand, aggregates evidence in favor of an informative hypothesis across multiple studies.
-It thus answers the question: Do all these studies support the informative hypothesis?
-Both methods thus provide complementary information.
+It thus answers the question: Do all these studies support the hypothesis of interest?
+Both methods are appropriate for different research questions, and provide complementary information.
 
-This tutorial paper introduces the first implementation of BES in useruse-friendly open source software.
-A function `pbf()` was contributed to the `bain` R-package for Bayesian informative hypothesis evaluation, version `r as.character(packageVersion("bain"))`.
+This paper introduces the first implementation of BES in user-friendly free open source software (FOSS).
+A function `pbf()` was contributed to the `bain` R-package for Bayesian informative hypothesis evaluation, version `0.2.9`.
+<!-- `r as.character(packageVersion("bain"))`. -->
 This paper presents a simulation study to validate the method and benchmark it against alternative evidence synthesis methods,
@@
 <!-- En dan nog iets over reliability ook. -->
-The present simulation study set out to validate the PBF algorithm.
-For each iteration of the simulation, we simulated a correlation coefficient in multiple samples.
-<!-- CJ: Hier staat multiple samples; elders multiple groups / studies / etc. Kan je die terminologie gelijktrekken? -->
-The informative hypothesis was kept constant at $H_i: \rho > .1$.
-All simulation conditions were evaluated once in the presence of a true population effect,
-defined as $\rho = .2$,
-and once in the presence of a null effect, defined as $\rho = .1$.
-A PBF > 3 was used as a decision criterion to conclude that $H_i$ was supported over its complement.
-This threshold is subjective but rooted in conventional standards for interpreting Bayes factors [@jeffreys1998theory].
+The present simulation study set out to validate the PBF algorithm and benchmark it against other evidence synthesis methods.
+We simulated a scenario where an informative hypothesis about a correlation between two variables was measured across several independent samples,
+and the resulting evidence was synthesized across samples using multiple methods.
 <!-- Rebecca: -->
@@
 <!-- I do see that this would end up in multiple results (once using .1 and 0.2, once (say) 0 or 0.05 and .2; but also when using other cut-off values for PBF). -->
-To examine performance in a range of realistic scenarios,
-several design factors were manipulated:
-The number of observations per sample $n \in (50, 200, 500, 800)$,
-which (assuming a known effect of $\rho = .1$) provides statistical power to reject a false null hypothesis at $\alpha = .05$ of
-$(`r paste0(formatC(pwr.r.test(c(50, 200, 500, 800), .1)$power, digits = 2, format=  "f"), collapse = ", ")`)$, respectively.
-the number of samples $k \in (2, 3, 10)$, where 2 is the minimum number of samples possible and 10 is a relatively large number, considering that replications are not very common.
-The reliability of the two correlated variables, $\alpha \in (0.6, 0.8, 1.0)$, where 0.6 is the lowest reliability conventionally considered to be acceptable, and 1 represents perfect reliability, as is
+The informative hypothesis was kept constant at $H_i: \rho > .1$.
+To examine the performance of the different evidence synthesis methods in a range of scenarios,
+several design factors were manipulated.
+First was the presence or absence of a true population effect.
+A present effect was defined as $\rho = .2$;
+a null effect was defined as $\rho = .1$.
+The second design factor was the number of observations per sample $n \in (50, 200, 500, 800)$.
+These values were chosen because they correspond to a statistical
+<!-- pwr.r.test(c(50, 200, 500, 800), .1) -->
+power to reject a false null hypothesis of $\beta \in (.10, .30, .60, .80)$ power, respectively,
+assuming $\alpha = .05$ and a known effect size of $\rho = .1$.
+Third, we manipulated the number of independent samples (or: replication studies), $k \in (2, 3, 10)$.
+Fourth, the reliability of the two correlated variables was varied between $\alpha \in (0.6, 0.8, 1.0)$,
+where 0.6 is the lowest reliability conventionally considered to be acceptable, and 1 represents perfect reliability, as is
 assumed when analyzing correlations between observed items or scale scores.
-For all unique combinations of the design factors, 1000 data sets were generated.
+For all unique combinations of these design factors,
+the simulation was repeated 1000 times.
 
@@
 The main algorithm of interest was the PBF.
+As a decision criterion to conclude that $H_i$ was supported over its complement,
+we used $PBF > 3$ - a conventional threshold for inference using Bayes factors [@jeffreys1998theory].
+
 As a benchmark for comparison, we included several other algorithms that might feasibly be used by researchers who intend to examine whether a hypothesis is true across several independent samples.
-The first was vote counting: counting the number of significant effects.
-Although this method is in use for aggregating conceptual replications,
-it should not be considered good practice.
+The first was *vote counting*: counting the number of significant effects.
+Although this method is still in use for aggregating conceptual replications,
+it is considered bad practice.
 Three disadvantages are that vote counting disregards sample size,
 reduces statistical power,
-and does not quantify the strength of the evidence [(Scheerens, 2015):].
-Our vote counting algorithm used one-sided z-tests to examine whether a null hypothesis corresponding to the informative hypothesis was rejected in the majority of samples, i.e.: $H_0: \rho = .1$, and $H_a: \rho > .1$, which corresponds to $H_i$.
+and does not quantify the strength of the evidence [@hedgesVotecountingMethodsResearch1980].
+Our vote counting algorithm summed the number of one-sided z-tests of a null hypothesis corresponding to the informative hypothesis,
+so $H_0: \rho = .1$ and $H_a: \rho > .1$, which corresponds to $H_i$.
+The decision criterion was that the hypothesis was supported in the majority of samples.
 Thus, for example, if $H_a$ was supported in three out of five samples, our vote counting algorithm would find overall support for $H_a$ (and, by extension, $H_i$).
-The second was a random-effects meta-analysis (RMA), which is the standard in the field.
-For this algorithm, the null-hypothesis was rejected if a 90% confidence interval for the overall effect size excluded $H_0$.
-The third algorithm was individual participant data (IPD) meta-analysis.
-Like classic meta-analysis, IPD is a multilevel model, clustered by sample.
-In contrast to classic meta-analysis, IPD freely estimates variance at the first level, because raw data are available.
-The PBF can be estimated using either sufficient statistics (as in meta-analysis) or using raw data (as in IPD).
-It is thus informative to compare the PBF to both these methods.
-IPD was also evaluated using a 90% confidence interval for the overall effect size.
+
+The third algorithm was *random-effects meta-analysis* (RMA),
+which is the current gold standard for evidence synthesis [@viechtbauerConductingMetaanalysesMetafor2010].
+For this algorithm, the null-hypothesis was rejected if the lower bound of a 90% confidence interval for the overall effect size excluded $H_i$.
+Note that this corresponds to a test at $\alpha = .05$,
+because all effects in the simulation are directional.
+
+The fourth algorithm was *individual participant data* (IPD) meta-analysis [@rileyMetaanalysisIndividualParticipant2010].
+Like classic meta-analysis (RMA), IPD is a multilevel model, clustered by sample.
+IPD uses the raw data, which makes it possible to estimate variance at the first level.
+By contrast, RMA treats the first-level variance as known.
+Note that the PBF can be estimated using either sufficient statistics (as in meta-analysis) or using raw data (as in IPD).
+With this in mind, it is informative to benchmark it against both of these methods.
+IPD was also evaluated using the lower bound of a 90% confidence interval for the overall effect size.
 <!-- Rebecca: -->
@@
 For each algorithm, inferential decisions made using the criteria described above were compared to the population status of the hypothesis (true or false).
-This results in a so-called "confusion matrix",
-which gives the number of decisions that were true positives (TP), true negatives (TN), false positives (FP) and false negatives (FN).
-These quantities were summarized as sensitivity, ${TP}{TP + FN}$, the probability of making the inferential decision that the informative hypothesis is supported, given that it was indeed true in the population, <!-- i think sensitivity is the ability to detect an effect, given that the effect is truly there.
-I found that its formula is usually presented as TP/(TP+FN)-->
+The resulting confusion matrix gives the number of decisions that were true positives (TP), true negatives (TN), false positives (FP) and false negatives (FN).
+These quantities were summarized as sensitivity, $\frac{TP}{TP + FN}$,
+the ability to detect an effect given that it was indeed true in the population,
+<!-- i think sensitivity is the ability to detect an effect, given that the effect is truly there. I found that its formula is usually presented as TP/(TP+FN)-->
 <!--CJ: This is all the same, see https://en.wikipedia.org/wiki/Sensitivity_and_specificity -->
-and specificity, $\frac{TN}{TN+FP}$, the probability of concluding the informative hypothesis is not supported, given that it was indeed false in the population.
-<!-- and specificity is the ability to not acknowledge an effect, given that there indeed is none-->
+and specificity, $\frac{TN}{TN+FP}$,
+the ability to correctly conclude that the informative hypothesis is not supported,
+given that it was indeed false in the population. <!-- and specificity is the ability to not acknowledge an effect, given that there indeed is none-->
 The overall performance was captured by the accuracy, which represents the total proportion of correct (true positive and true negative) decisions, $\frac{TP + TN}{TP+TN+FP+FN}$.
@@
 First, we examined overall model performance across conditions.
-Recall that sensitivity is the probability of concluding that the informative hypothesis is supported, given that it was indeed true in the population,
-and specificity, is the probability of concluding that it is not supported, given that it was false in the population.
-The results indicate that all algorithms except PBF had low sensitivity to detect a true effect.
-In contrast, specificity was very high for all algorithms except PBF.
-This suggests that the other algorithms classified most conditions as negatives (no effect found),
-regardless of the existence of a population effect.
-The PBF trades a loss of specificity for increased sensitivity.
-At the same time, the average specificity and sensitivity were approximately equal, see Table \@ref(tab:tabconf).
+PBF had a higher overall accuracy than other algorithms
+followed by IPD, then RMA and finally VC,
+see Table \@ref(tab:tabconf).
+This higher accuracy was primarily driven by PBF's greater sensitivity to detect a true effect compared to other algorithms.
+However, PBF had a lower specificity compared to all other algorithms.
+This suggests that the PBF trades a loss of specificity for increased sensitivity.
 
@@
 
-We examined the effect of simulation conditions on overall accuracy.
 ```{r tabeffect, echo = FALSE}
@@
 ```
-PBF performance was most impacted by sample size $n$, followed by the number of groups $k$, and reliability.
-<!-- Rebecca: Nu niet compleet? -->
-<!-- Gek om steeds ene paragraaf van 1 zin te hebben iig. -->
-<!-- Verder mist er denk ik info. Ik zou graag eerst meer uitleg zien. Hoe heb je dat examine, wat ga je laten zien (niet het resultaat an sich, snap je wat ik bedoel...). -->
-<!-- Verder zou ik denk ik niet hier al weggeven wat je hierna gaat laten zien (en anders dat explicieteren dat dat is wat je doet). -->
+We used ANOVAs to examine the effect of simulation conditions on overall accuracy.
+The differences between algorithms were analyzed in analyses that included two-way interactions between design factors and algorithm.
+As the sample size was very large,
+significance tests were uninformative.
+We thus focused on interpreting the effect size of the effects of design factors.
+The performance of PBF was most impacted by sample size $n$, followed by the number of groups $k$, and reliability.
+The differences in the effects of sample size and number of groups were relatively small between PBF and the two best practice algorithms, RMA and IPD -
+but substantial between PBF and the suboptimal VC algorithm.
+The reverse pattern occurred for reliability, however:
+it showed a substantial difference in effect between PBF and the two best practice algorithms only.
 
@@
 
-```{r fign, fig.cap="Mean performance by sample size"}
+```{r}
 res <- readRDS("confusion_by_cond.RData")
@@
 # df_plot <- df_plot[ ,list(mean=mean(val)), by=c("n", "Outcome", "alg")]
-ggplot(df_plot, aes(x = n, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
+library(ggplot2)
+p <- ggplot(df_plot, aes(x = n, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
   geom_point() +
   geom_line() +
+  scale_linetype(guide = guide_legend(reverse = TRUE) )+
   theme_bw() +
-  labs(y = "Mean performance", x = "Sample size")
+  labs(y = "Mean performance", x = "Sample size")+
+  labs(shape = "Algorithm")
+ggsave("sample_size.pdf", p, device = "pdf", width = 140, height = 100, units = "mm")
+```
+```{r fign, fig.cap="Mean performance by sample size"}
+knitr::include_graphics("sample_size.pdf")
 ```
 
-Figure \@ref(fig:fign) indicates that for PBF, both sensitivity and specificity increase with sample size.
-The other algorithms also show increasing sensitivity, but not specificity, which is at a ceiling.
-This difference explains the effect of reliability on the difference between algorithms (see Table \@ref(tab:tabeffect)).
+Figure \@ref(fig:fign) indicates that for PBF, both sensitivity and specificity increased with sample size.
+The other algorithms showed only increasing sensitivity; 
+specificity was limited by a ceiling effect.
+This difference explains the effect of sample size on the difference between algorithms (see Table \@ref(tab:tabeffect)).
 
-### Effect of number of groups
+### Effect of number of samples
 
-```{r, fig.cap="Mean performance by number of groups"}
+```{r}
 df_plot <- data.table::melt(res, measure.vars = c("sensitivity", "specificity"),
@@
 names(df_plot)[ncol(df_plot)] <- "mean"
-ggplot(df_plot, aes(x = k, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
+p <- ggplot(df_plot, aes(x = k, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
   geom_point() +
   geom_line() +
+  scale_linetype(guide = guide_legend(reverse = TRUE) )+
   theme_bw() +
-  labs(y = "Mean performance", x = "Number of groups")
+  labs(y = "Mean performance", x = "Number of groups")+
+  labs(shape = "Algorithm")
+ggsave("groups.pdf", p, device = "pdf", width = 140, height = 100, units = "mm")
 ```
-The figure indicates that, for PBF at higher levels of k, lower sensitivity is exchanged for greater specificity.
-The other algorithms do not show this pattern, as their specificity is at a ceiling.
-This difference in pattern of effects explains why number of groups has a moderate effect on the difference between algorithms (see Table \@ref(tab:tabeffect)).
-Only VC shows decreasing sensitivity with an increasing number of groups;
-this is because the probability of obtaining any false negatives increases with the number of groups.
+```{r figk, fig.cap="Mean performance by number of groups"}
+knitr::include_graphics("groups.pdf")
+```
+
+The figure indicates that, for PBF at higher levels of $k$,
+lower sensitivity was traded for greater specificity.
+The other algorithms did not show this pattern, as their specificity was at a ceiling.
+This difference in pattern of effects explains why number of samples had a moderate effect on the difference between algorithms (see Table \@ref(tab:tabeffect)).
+Only VC showed decreasing sensitivity with an increasing number of groups;
+this is because the probability of obtaining false negatives increases with the number of groups [as also noted by @hedgesVotecountingMethodsResearch1980].
 
@@
 This makes that the reliability decreases, rather than increases as you move away from the origin on the x-axis -->
-```{r, fig.cap="Mean performance by reliability"}
+```{r}
 df_plot <- melt(res, measure.vars = c("sensitivity", "specificity"),
              variable.name = "Outcome", value.name = "val")
+df_plot$reliability[df_plot$reliability == 0] <- 1
+df_plot$reliability[df_plot$reliability == .5] <- .8
+df_plot$reliability[df_plot$reliability == .81] <- .6
+
 # df_plot <- df_plot[ ,list(mean=mean(val)), by=c("reliability", "Outcome", "alg")]
@@
 names(df_plot)[ncol(df_plot)] <- "mean"
-ggplot(df_plot, aes(x = reliability, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
+p <- ggplot(df_plot, aes(x = reliability, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
   geom_point() +
   geom_line() +
+  scale_linetype(guide = guide_legend(reverse = TRUE) )+
   theme_bw() +
-  labs(y = "Mean performance", x = "Reliability")
+  labs(y = "Mean performance", x = "Reliability")+
+  labs(shape = "Algorithm")
+ggsave("reliability.pdf", p, device = "pdf", width = 140, height = 100, units = "mm")
+```
+```{r figrel, fig.cap="Mean performance by reliability"}
+knitr::include_graphics("reliability.pdf")
 ```
 
-The figure indicates that for PBF, at higher reliability, lower sensitivity is exchanged for greater specificity.
-The other algorithms do not show this pattern, as their specificity is at a ceiling.
-Their sensitivity decreases with lower reliability,
-and therefore, so does their overall performance.
-This difference in pattern of effects explains why reliability has a moderate effect on the difference between algorithms (see Table \@ref(tab:tabeffect)).
-Note that, in contrast to other algorithms, the performance of PBF is less susceptible to reliability.
+Figure \@ref(fig:figrel) indicates that the PBF traded sensitivity for specificity.
+At low levels of reliability, specificity exceeded sensitivity;
+at high levels of reliability, this pattern was reversed.
+The other algorithms did not show this pattern, as their specificity was at a ceiling.
+Their sensitivity increased with higher reliability, however,
+and therefore, so did their overall performance.
+This difference in pattern of effects explains why reliability had a moderate effect on the difference between algorithms (see Table \@ref(tab:tabeffect)).
+Note that, whereas the overall accuracy of the PBF was found to be less susceptible to reliability as compared to other algorithms,
+this finding masked the trade-off between reliability and specificity.
 
@@
 
-This simulation study validated the product Bayes factor as an alternative method to synthesize evidence across studies.
-The PBF was benchmarked against other common methods for evidence synthesis.
-This was done by calculating the accuracy, sensitivity and specificity of the methods in different plausible varying conditions.
-During the simulation the hypotheses remained constant $H_i: \rho > .1$, $H_0: \rho = .1$.
-To be able to calculate both sensitivity and specificity, half of the simulated conditions contained an effect, $\rho = .2$, while the other half did not, $\rho = .1$.
-<!-- EB: Why did we not use H0:r=0 and Hi:r>0 as constant hypotheses? It seems more intuitive -->
-Sensitivity and specificity can be interpreted as Type I error rate and statistical power respectively and are used interchangeably.
- <!-- Because Type 1 error rate and power seem to be key terms in the discussion, we may be able to include it in the introduction: 'How does the PBF compare to other state-of-the-art algorithms with regards to Type 1 error rate, power and overall accuracy?' -->
+This simulation study examined the performance of the product Bayes factor (PBF) as compared against three other common methods for evidence synthesis (IPD, RMA, and VC).
+The results showed that PBF had a higher overall accuracy than the other algorithms,
+primarily due to its greater sensitivity to detect a true effect.
+PBF had lower specificity, however,
+suggesting that it trades specificity for increased sensitivity.
+<!-- PBF's performance was most impacted by sample size, both in terms of sensitivity and specificity.
+The other algorithms showed less improvement in accuracy with increasing sample size due to a ceiling effect for specificity.
+-->
+<!-- The number of groups also had a moderate effect on PBF's performance, -->
+<!-- increasing sensitivity while specificity remained approximately stable. -->
+<!-- Reliability had a moderate effect on the performance of all algorithms, with overall performance increasing as reliability increased. -->
+<!-- However, the effect of reliability on PBF was different than the other algorithms, as PBF traded increasing sensitivity for decreasing specificity as reliability increased, -->
+<!-- whereas the other algorithms showed only an increase in sensitivity and approximately stable specificity. -->
+PBF's performance was most impacted by sample size,
+followed by the number of groups and reliability.
+Even at the smallest sample size of $n = 20$,
+PBF had a superior sensitivity to the other algorithms.
+This suggests that PBF is more suitable than other methods as a small sample solution.
+When the number of samples increased,
+most algorithms showed approximately stable specificity
+and increasing sensitivity.
+A notable exception was vote counting;
+its sensitivity decreased with an increasing number of samples,
+as has been previously documented [@hedgesVotecountingMethodsResearch1980].
+With increasing reliability,
+PBF showed a substantially different pattern of results than the two other best practice algorithms (RMA and IPD).
+Although the overall effect of reliability on overall accuracy was smaller for PBF than for other algorithms,
+this masks a trade-off between reliability and specificity.
+Whereas most algorithms showed near-stable specificity and increasing sensitivity,
+PBF showed a clear trade-off between decreasing specificity and increasing sensitivity.
+This implies that the PBF is more conservative - less likely to detect an effect - in the presence of increasing measurement error.
 
-The overall performance of PBF increased most with increasing sample size.
- <!-- EB: also, shall we rename 'performance' to 'accuracy', I know its not the same, but performance seems a bit too subjective.
-After all, we do not have a clear cut decision when sensitivity and specificity is balanced 'correctly' -->
-PBF's sensitivity was .57 when a sample size of 20 per sample was simulated, while for all other algorithms this was below .11.
-It should be noted that in the condition where sensitivity could be calculated, the true effect $\rho$ was .2 and only .1 points above the boundary value for $H_i > .1$ and $H_0 = .1$.
-It is no surprise that at smaller samples the small difference in effect size between null/informative hypothesis and true value of .1 is unlikely to be detected.
-Still, even at a relatively large sample size of 500 per sample, the sensitivity of the other algorithms only rose slightly above .60, 
-while for PBF this was .90.
-This provides evidence that state-of-the-art evidence synthesis methods have little power to detect a true, but small effect,
-even when the sample size per sample was substantial.
-PBF on the other hand presents itself as a small sample size solution for evidence synthesis methods to detect true effects,
-even if the effect itself can be considered small.
-This is also reflected in the accuracy of the algorithms at small sample sizes, which, at a sample size of 20, was around .50 for RMA, IPD and VC.
-This can be attributed to their low sensitivity and their specificity that is at ceiling level. 
-By consistently not finding an effect, the algorithms are correct 50\% of the time.
-At all sample sizes, the PBF was the most accurate algorithm,
-although it should be noted that even at a sample size of 500, PBFs specificity was .12 percentage points too high for conventional acceptable type I error rates of .05.
+These results have important implications for applied evidence synthesis.
+For example, the finding that PBF had a higher overall accuracy due to greater sensitivity suggests that it may be a better choice than the other algorithms,
+particularly when detecting true effects has high priority.
+However, researchers should be aware that this increased sensitivity comes at a loss of specificity,
+which incurs a greater risk of false positive results.
+If specificity is a higher priority,
+other algorithms such as IPD or RMA may thus be more appropriate.
 
-With an increasing number of groups, slightly decreasing specificity was traded off for increasing sensitivity.<!--EB: I would argue that the decrease is negligible, 0.767 at k =2, 0.763 at k = 10 -->
-With an increasing number of groups, specificity remained around .76 for PBF, while sensitivity increased to .87.
-Again, the specificity remained at ceiling level for the other algorithms.
-The sensitivity for RMA and IPD increased, but stayed below .51.
-In a similar way, this supports the use of the PBF as a small sample size solution.
-In this case when there are little available studies to synthesize evidence over.
-It could be argued that when there are only 2 or 3 studies available to synthesize evidence over, that if a trade-off need be made, specificity is preferred over sensitivity.
-It seems unwise to more easily acknowledge an effect if there is little data available.
-However, the quality of the included studies should be considered before synthesizing evidence in any way.
-Moreover, when the researcher performs some form of evidence synthesis, then the ability of the method to find an effect is a necessary and fair component, even if little studies are available.
-Also, even with as much as ten available studies, PBF was the only algorithm that had a conventionally acceptable power.
+The present study also has some limitations.
+First, the simulation study makes specific assumptions that may not generalize to all real-world applications.
+A second important caveat is that most of the algorithms did not reach
+a level of sensitivity that is considered acceptable from a perspective of statistical power [i.e., greater than .80, @cohen2013statistical].
+PBF performed notably better than other algorithms,
+but its sensitivity still fell below .80 in many conditions.
+Low power increases the risk of false negatives,
+or failing to detect a true effect.
+One reason power was low is that,
+in conditions where a true effect was present,
+its value only exceeded the boundary value of the informative hypothesis by .1 points.
+Such small effects are hard to detect.
+All algorithms will likely perform better when the true effect is larger.
+The low sensitivity of all algorithms highlights the importance of reticence when interpreting evidence syntheses of studies with small samples and small effect sizes.
+It may be prudent to avoid generalizing such results to the population,
+and instead consider them as merely descriptive of the published research.
+Additionally, sensitivity analyses can be used to assess the robustness of the results to different modeling assumptions and methods.
 
-With increasing reliability, increasing specificity was traded off for decreasing sensitivity for PBF.
-<!--EB: I think its the other way around; with DECREASING reliability, specificity increases and sensitivity decreases, see my comment at the section: 'Effect of reliability' -->
-With increasing reliability, decreasing specificity was traded off for increasing sensitivity for the PBF.
-Also for the other algorithms, the sensitivity increased as the reliability increased.
-It implies that the PBF gets more conservative whenever reliability decreases.
-This can be considered beneficial as the PBF is more hesitant to acknowledge an effect as the measurements become less stable.
-Even at the smallest acceptable reliability, $\alpha = .6$, PBFs sensitivity was .57, where it was below .10 for the other algorithms.
-However, the specificity of the PBF when there is perfect reliability $\alpha=1$ was .63, were it was almost 1 for the other algorithms.
+A third limitation is that the evidence synthesis methods compared here represent different approaches to inference and answer different research questions.
+<!-- Their inferential properties are thus not directly comparable. -->
+<!-- PBF assesses whether an informative hypothesis is supported across all replication studies; -->
+<!-- RMA estimates an average effect size from summary statistics; -->
+<!-- IPD estimates an effect size from individual data, accounting for clustering within studies; -->
+<!-- and VC counts the number of times a hypothesis was supported. -->
+Since each of these methods is optimized for a different purpose,
+the present study should not be considered as a comprehensive assessment of their strengths and weaknesses.
+<!-- For example, PBF takes a Bayesian approach to inference, and a PBF of 10 in our study means that the (subjective) probability that our informative hypothesis is true is ten times larger than the probability that it is not true.
+-->
+<!-- By contrast, IPD takes a frequentist approach to inference, -->
+<!-- and a p-value of $p = .02$ means that there is a 2% probability of observing an effect at least as large as the one found, if the null-hypothesis were true. -->
+We nonetheless compare them because of their similar usage in evidence synthesis.
+It is up to individual researchers to choose an appropriate method,
+guided by the research question and the available information.
+For instance, when raw data is unavailable, IPD cannot be used,
+and when effect sizes are not available, only VC can be used.
 
-Overall, there was only a single marginal condition where the accuracy of PBF was lower than that of any other algorithm.
-PBF scored .01-.02 percentage points lower accuracy than IPD and RMA when there was perfect reliability.
-Looking at the specificity and sensitivity however, this is explained by the relatively low specificity of PBF and the very high specificity of RMA and IPD.
-Thus,
-Within the scope of this simulation, PBF had relatively better inferential properties <!-- EB: higher accuracy --> than the other algorithms under consideration.
-Although other algorithms had superior specificity, PBF had the highest levels of sensitivity.
-Thus, PBF balanced the ability to accept an informative hypothesis in the presence of a true population effect with the ability to reject the hypothesis when there was no effect.
+Aside from the aforementioned fact that the PBF answers a different research question than the other algorithms,
+it is worth noting limitations of the interpretation of the PBF.
+The PBF renders support for one specific informative hypothesis versus its complement.
+If the informative hypothesis is supported, this does not necessarily mean that it is also true.
+Consider the hypothetical example that a test of the informative hypothesis that the earth is flat received support with $BF = 3.01$.
+Although the data support this hypothesis over its complement, the hypothesis is clearly wrong (the earth is spherical).
+If we would have tested another hypothesis, e.g., the earth is shaped like an American football, it would have received much more support, e.g. $BF = 1000$, even though it is also wrong.
+A high Bayes factor thus does not mean that the hypothesis is true.
+Conversely, a low Bayes factor merely indicates that the informative hypothesis is not supported,
+and does not provide information about the true state of affairs.
 
-An important caveat is that none of the algorithms met conventional criteria for power (80%) and type I error (5%) [@cohen2013statistical].
-Overall, PBF had the best performance, with a power of 76% and type I error of 24%.
-Thus, if researchers intend to synthesize evidence for an informative hypothesis across heterogeneous studies, PBF may be the most suitable method - but its limitations <!-- higher probability for type 1 errors --> should be acknowledged in applied research.
+<!-- In sum, this simulation study suggest that PBF may be a good choice when sensitivity is of utmost importance, but researchers should be mindful of the trade-off between specificity and sensitivity.
+-->
+<!-- The results also highlight the importance of considering sample size, number of samples, -->
+<!-- and reliability when designing studies and conducting meta-analyses.
+Future research should examine the performance of additional algorithms and the impact of other design factors on algorithm performance.
+-->
+<!-- The results indicated that, compared to the other algorithms, -->
+<!-- PBF had a higher overall accuracy, -->
+<!-- which could be attributed to PBF's relatively greater sensitivity to detect a true effect. -->
+<!-- The greater sensitivity of PBF came at a loss of specificity compared to other algorithms. -->
+<!-- Specifically, all other algorithms showed ceiling effects for specificity and low base rates for sensitivity. -->
 
-As stated earlier, the PBF can be used as a small sample size solution,
-although its increased type I error rate must be taken into account.
-It implies that PBF works best as a complementary method for evidence synthesis, rather than a replacement.
+<!-- The overall accuracy of the PBF was most affected by the sample size of the underlying studies. -->
+<!-- This effect was partly driven by the absence of a ceiling effect of specificity for the PBF, compared to the other algorithms. -->
+<!-- Increasing sample size increased sensitivity and specificity for all algorithms. -->
+<!-- Only for PBF, however, did sensitivity surpass specificity at large ($\geq 200$) sample sizes. -->
+<!-- Combined with the higher overall level of specificity, this suggests that PBF may be preferable when detecting true effects has high priority. -->
 
-<!-- Conventional threshold for type 1 error: 5% -->
-<!-- Conventional threshold for power: 80% -->
-<!-- J. Cohen, Statistical Power Analysis for the -->
-<!--  Behavioral Sciences, 2nd ed. (Erlbaum, Hillsdale, -->
-<!--  NJ, 1988). This is the source of the system of power -->
-<!--  analysis described here; the power values and sam -->
-<!--  ple sizes of the illustrations derive from this book's -->
-<!--  tables. -->
-<!--  2. J -->
+<!-- <!-- Small samples -->
+<!-- At the smallest sample size of $n = 20$, the PBF had a better balance between sensitivity and specificity than the other algorithms. -->
+<!-- This suggests that PBF is more suitable than other methods as a small sample size solution for evidence synthesis. -->
 
-<!-- Rebecca: -->
-<!-- Ws niet een punt voor hier, maar wel in de discussie ofzo: -->
-<!-- It renders support for the hypothesis being the best in all studies. It, thus, does not render the support for the hypothesis being on average the best. -->
+<!-- that increased with increasing sample size, number of samples, and reliability. -->
+<!-- PBF, by contrast, showed lower base rates of specificity and higher base rates of sensitivity. -->
 
-<!-- EB: I assume Rebecca means that 'the hypothesis being on average the best' is the hypothesis that reflects the truth.
-While 'the hypothesis that is best in all studies' is the hypothesis that is closest to the truth compared to the other hypotheses that were tested.
-It does not say anything necessarily about whether the tested hypotheses in general are good reflections of the truth.
+<!-- With increasing sample size, number of samples, and reliability, PBF showed a trade-off between decreasing specificity and increasing sensitivity. -->
 
-<!--Rebecca: Hier ook dan artikel van BEM noemen, maar ook zeggen dat het daar om null hypotheses gaat. -->
+<!-- Overall, these results suggest that PBF had relatively better overall accuracy than the other algorithms under consideration. -->
+<!-- Although other algorithms had superior specificity, PBF had the highest levels of sensitivity. -->
+<!-- PBF thus traded off an increased ability to correctly accept informative hypotheses in the presence of true population effects for a decreased ability to reject the hypothesis when there was no effect.
+-->
+<!-- Users of the PBF should assess whether this trade-off is warranted. -->
+<!-- If it is not, more stringent criteria for inference or more conservative methods may be preferred. -->
 
-<!--EB: dit refereert naar Rouder, J. N., & Morey, R. D. (2011). A Bayes factor meta-analysis of Bem’s ESP claim. Psychonomic Bulletin & Review, 18(4), 682-689.
-
-The argument from Rouder and Morey (2011) is that we cannot take the PBF as a valid measure for meta-analysis.
-They show that calculating the BF in multiple samples and taking their product results in a different PBF than when all data is pooled and the BF is calculated after pooling.
-However, they assume a constant effect where all data is drawn from a common distribution The idiosyncratic differences between studies makes that pooling data is not a straightforward endeavour, because of the heterogeneity of the effect.
-Rouder and Morey do acknowledge their choice for a constant effect to be debatable.
-
-Quote from the Rouder and Morey (2011) paper:
-"It is reasonable to wonder whether the constant effect
-size model underlying the meta-analytic Bayes factor is
-warranted. We chose this approach because it is tractable
-when researchers have access to the test statistics, rather
-than the raw data."
-
-Interesting statement.
-Intuitively I do not see how having only the test statistics compared to raw data makes the assumption of a constant effect more probable.
-After all, the test statistic is derived from the raw data...
-
-Rebecca noticed that Rouder and Morey specifically talk about PBF not being a good measurement in the context of testing H0, which is something that we are not doing.
-However, the only relevant difference I see in their vs our method is that their H0 contains a single equality constraint, while we used inequality constraints.
-Rouder and Morey do not explicitly point this difference out and I dont see how the PBF would be a bad or good idea depending on type of constraints.
+<!-- OTHER PBF APPROACHES -->
+<!-- One possible issue with the PBF as defined in this research comes from an article of @rouder2011bayes.  -->
+<!-- The article states that a meta-analytic Bayes factor is not as straightforward as taking the product of $k$ individual Bayes factors. -->
+<!-- <!--EB: Their reason seems to be that Bayes factors take sample size into account, effectively reducing the ability of a Bayes factor to recognize an effect in small sample.
+I dont see how this invalidates the PBF.
+If there is an effect, then in 5 small studies, it may be true that a BF is calculated that barely support the hypothesis of an effect.
+However, taking the product of 5 BFs that are all 1.25 will result in a PBF of about 3.-->
+<!-- Rather, they provide a way of calculating more valid a meta-analytic Bayes factor. -->
+<!-- However, only under the assumption that the population effect is constant and the test statistic is from the same type of statistical model. -->
+<!-- Using this form of the PBF, they show that pooling the data and calculating a single Bayes factor on the pooled data comes closer to the meta-analytic Bayes factor than taking the product of each individual study.
 -->
-
-<!-- Rebecca: -->
-<!-- ook iets zeggen over vergelijkbaarheid? -->
-<!-- Ws hierboven al iets over zeggen: -->
-<!-- In klassieke methoden heb je equality restrictie. -->
-<!-- Als rho = .1 dan null waar en moet je hem niet verwerpen, als .2 dan H0 niet waar en moet je H0 wel verwerpen. -->
-<!-- Bij BF heb je Hi en als .1 dan eigenlijk zelfs Hi waar, maar op border/boundary, dus ws dan BF = ong 1. Als .2 dan Hi waar en moet je BF > 1 vinden (evt dus > 3). -->
+<!-- However, as stated earlier, one strength of the PBF lies in its ability to synthesize evidence for a hypothesis, rather than an estimate. -->
+<!-- This means that there is no assumption of a constant population effect and/or equal types of test statistics across studies. -->
+<!-- Furthermore, as the article of @rouder2011bayes states itself, pooling data is preferred if all data is independent and drawn from a common distribution.  -->
+<!-- The study idiosyncrasies encountered in evidence synthesis should wary the researcher in blindly making the assumption of a common distribution. -->
+<!-- A final remark is that of performance comparison between PBF and the other algorithms which has to do with the differences in hypothesis specification.  -->
+<!-- For the frequentist and PBF algorithms we specified $H_0: \rho = .1$ and $H_c: \rho \leq .1$ respectively. -->
+<!-- This means, hypothetically, that if a strong effect in the opposite direction was found, let's say $\rho = -.6$, the frequentist algorithms would reject $H_0$ while the PBF would accept $H_c$.
+-->
+<!-- This implies that there should have been a higher power to detect an effect in the frequentist framework, as negative effects would reject $H_0$. -->
+<!-- This is contrary to what was observed in this research, however, as the frequentist approaches had a low sensitivity. -->
+<!-- Again, this has to do with the small simulated difference between the true effect and the boundary value. -->
 
@@
 
-This tutorial about Bayesian Evidence Synthesis (BES) requires R-package `bain` version `r packageVersion('bain')` or later.
-This tutorial focusses on synthesizing evidence for an informative hypotheses using the Product Bayes Factor (pbf). 
-The data used to illustrate the method have been simulated based on the data presented in [@vanleeuwen2022morality].
-These simulated data are available in the `bain` package. 
-A more detailed description of the dataset is either provided in [@vanleeuwen2022morality] or by running `?bain::synthetic_us`, `?bain::synthetic_dk` or `?bain::synthetic_nl`.
-The research question is whether there is an association between political orientation and moral dispositions.
+This tutorial demonstrates how to synthesize evidence for an informative hypothesis across heterogeneous replications using the Product Bayes Factor (PBF). 
+We assume that users have installed the free open source statistical programming language R [@R-base].
+The R-package `bain` version `0.2.9` or later is required, which can be installed by running `install.packages("bain")` in the R console.
+The data used in this tutorial are included in the `bain` package,
+and have been simulated based on the data presented in [@vanleeuwen2022morality].
+A more detailed description of the datasets is found in [@vanleeuwen2022morality];
+additionally, the dataset documentation is accessed by running `?synthetic_us`, `?synthetic_dk` or `?synthetic_nl` in the R console.
+Van Leeuwen and colleagues conducted a theory-driven, preregistered study to address the research question whether political orientation and moral dispositions are associated.
 Suitable data were collected in three countries: the United states of America, Denmark and the Netherlands.
-Each sample contains multiple measures of political orientation and moral dispositions.
+Each sample contained multiple measures of political orientation and moral dispositions.
 In the original publication, the PBF was used to aggregate evidence across scales and countries to obtain an overall measure of support for the central hypothesis.
 This tutorial follows the same rationale, but uses only one effect size per sample,
-and varies the way this effect size is computed to illustrate different use cases.
+and varies the way this effect size is computed to illustrate the more typical use case where the same informative hypothesis has been studied in different ways in multiple studies.
+We will examine the informative hypothesis that self-reported importance of family morality is positively associated with a conservative socio-political orientation.
+First, we load the `bain` library and assign the data to three objects with convenient names:
 
-```{r}
-NL <- bain::synthetic_nl
-DK <- bain::synthetic_dk
-US <- bain::synthetic_us
+```{r eval = TRUE}
+library(bain)
+NL <- synthetic_nl
+DK <- synthetic_dk
+US <- synthetic_us
 ```
@@
 <!-- Because the focus of this tutorial is on BES,  -->
-First, we briefly introduce the basic use of `bain::bain()`, and how to interpret its output. 
-Suppose we are interested in whether the self-reported importance of family morality positively predicts a conservative socio-political orientation.
-Because both scales consist of multiple items, we use structural equation modeling (SEM) to perform latent variable regression.
-Our hypothesis is $H_i: \beta > 0$, where $\beta$ (beta) is the standardized regression coefficient.
-Note that, for the sake of simplicity, this hypothesis corresponds to a traditional one-sided null-hypothesis test.
-The original publication instead used a minimal effect size of interest, and specified $H_i: \beta > .1$.
-The code below illustrates the analysis for a single sample.
-For a more in-depth tutorial on `bain::bain()`, see @hoijtink2019tutorial, and for further guidance on the use of `bain::bain()` in SEM, see @REF VAN LISSA TUTORIAL.
-
+First, we briefly introduce the basic use of the `bain()` function, and how to interpret its output.
+We must estimate a model suitable for testing our informative hypothesis.
+Because both scales consist of multiple items, we can use structural equation modeling (SEM) to perform latent variable regression [see @vanlissaTeacherCornerEvaluating2020]:
 <!--EB: For simplicities sake, I for now create simple average scores without taking into account reverse coding. This should likely be improved with factor scores -->
@@
 
-```{r echo = TRUE, eval = FALSE}
+```{r echo = TRUE, eval = TRUE}
+# Load lavaan package for SEM
+library(lavaan)
+
+# Specify SEM-model for latent variable regression
+model_nl <- "
+fam =~ fam_1 + fam_2 + fam_3
+con =~ sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
+       sepa_eco_1 + sepa_eco_2 + sepa_eco_3 + sepa_eco_4 + sepa_eco_5
+con ~ beta * fam"
+
+# Estimate the model in lavaan
+results_nl <- sem(model = model_nl, data = NL)
+```
+```{r echo = FALSE, eval = FALSE}
 # Load relevant packages
@@
 model_nl <- "
-family =~ fam_1 + fam_2 + fam_3
-conservative =~ sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
+fam =~ fam_1 + fam_2 + fam_3
+con =~ sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
                 sepa_eco_1 + sepa_eco_2 + sepa_eco_3 + sepa_eco_4 + sepa_eco_5
-conservative ~ beta * family"
+con ~ beta * fam"
 
@@
 # The argument standardize = TRUE uses the standardized model parameters
-bain(results_nl, hypothesis = "beta > 0", standardize = TRUE)
+bf_nl <- bain(results_nl, hypothesis = "beta > 0", standardize = TRUE)
 ```
-```{r echo = FALSE, eval = TRUE}
-# Load relevant packages
-library(bain)
-library(lavaan)
 
-# Specify a SEM-model with two latent variables and latent regression
-model_nl <- "
-family =~ fam_1 + fam_2 + fam_3
-conservative =~ sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
-                sepa_eco_1 + sepa_eco_2 + sepa_eco_3 + sepa_eco_4 + sepa_eco_5
-conservative ~ beta * family"
+The informative hypothesis in this tutorial is $H_i: \beta > 0$, where $\beta$ (beta) is the standardized regression coefficient.
+This corresponds to a traditional one-sided null-hypothesis test.
+Note that this deviates from the original publication
+which used a minimal effect size of interest and specified $H_i: \beta > .1$.
+The code below illustrates how to obtain a Bayes factor for this informative hypothesis,
+using the output of the SEM analysis above.
+We can refer to the parameter `beta` by name because we labeled it in the `lavaan` syntax;
+if we had not done so, we could find the names of all model parameters by running `get_estimates(results_nl, standardize = TRUE)`.
+When evaluating the code below,
+note that the hypothesis is strongly supported when compared to its complement.
+<!-- , $BF_c `report(bf_nl$fit$BF.c[1])`$. -->
+For a more in-depth tutorial on `bain()`, see @hoijtink2019tutorial, and for further guidance on the use of `bain()` for SEM, see @vanlissaTeacherCornerEvaluating2020.
 
-# Estimate the model in lavaan
-results_nl <- sem(model = model_nl, data = NL)
-
-# Test the hypothesis that the effect size labeled 'beta' is positive
-# The argument standardize = TRUE uses the standardized model parameters
+```{r echo = TRUE, eval = TRUE, results='hide'}
+# Test that the effect labeled 'beta' is positive
 bf_nl <- bain(results_nl, hypothesis = "beta > 0", standardize = TRUE)
@@
 
-Note that the hypothesis is strongly supported when compared to its complement, $BF_c `report(bf_nl$fit$BF.c[1])`$.
+### Aggregating evidence across studies
 
-### Aggregating evidence across studies
+```{r, eval = FALSE, echo = FALSE, results="hide"}
+# Specify the models for DK and US
+model_dk <- "
+fam =~ fam_1 + fam_2 + fam_3
+con =~ sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
+sepa_eco_1 + sepa_eco_2 + sepa_eco_3 + sepa_eco_4 + sepa_eco_5
+con ~ beta * fam"
+model_us <- "
+fam =~ fam_1 + fam_2 + fam_3
+con =~
+secs_soc_1 + secs_soc_2 + secs_soc_3 + secs_soc_4 + secs_soc_5 + secs_soc_6 + secs_soc_7 +
+secs_eco_1 + secs_eco_2 + secs_eco_3 + secs_eco_4 + secs_eco_5
+con ~ beta * fam"
+
+# Estimate the model in lavaan
+results_dk <- sem(model = model_dk, data = DK)
+results_us <- sem(model = model_us, data = US)
+
+# Bind the models into a list
+results <- list(results_nl, results_dk, results_us)
+# Test the hypothesis that the effect size labeled 'beta' is positive
+bf_tot <- pbf(results, hypothesis = "beta > 0", standardize = TRUE)
+```
 
@@
 This is an appropriate situation to use the PBF to aggregate evidence across countries.
-First, we estimate a latent regression model for the remaining two countries, taking care to label the parameter of interest with the same label in all samples.
+First, we estimate a latent regression model for the remaining two countries, taking care to use the same label for the parameter of interest in all samples.
 Then, we bind all three SEM-models in a list, and call PBF to evaluate the hypothesis of interest on all models and aggregate the evidence.
+As the BF in all three samples is positive, the resulting PBF is very large.
+We can thus conclude that the central hypothesis receives overwhelming evidence across samples.
 
-```{r, eval = FALSE, echo = TRUE}
+```{r, eval = TRUE, echo = TRUE}
 # Specify the models for DK and US
 model_dk <- "
-family =~ fam_1 + fam_2 + fam_3
-conservative =~ sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
+fam =~ fam_1 + fam_2 + fam_3
+con =~
+sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
 sepa_eco_1 + sepa_eco_2 + sepa_eco_3 + sepa_eco_4 + sepa_eco_5
-conservative ~ beta * family"
+con ~ beta * fam"
 model_us <- "
-family =~ fam_1 + fam_2 + fam_3
-conservative =~ secs_soc_1 + secs_soc_2 + secs_soc_3 + secs_soc_4 + secs_soc_5 +
-                secs_soc_6 + secs_soc_7 + secs_eco_1 + secs_eco_2 + secs_eco_3 +
-                secs_eco_4 + secs_eco_5
-conservative ~ beta * family"
+fam =~ fam_1 + fam_2 + fam_3
+con =~
+secs_soc_1 + secs_soc_2 + secs_soc_3 + secs_soc_4 + secs_soc_5 +
+secs_soc_6 + secs_soc_7 +
+secs_eco_1 + secs_eco_2 + secs_eco_3 + secs_eco_4 + secs_eco_5
+con ~ beta * fam"
 
@@
 ```
-```{r, eval = TRUE, echo = FALSE}
-# Specify the models for DK and US
-model_dk <- "
-family =~ fam_1 + fam_2 + fam_3
-conservative =~ sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
-                sepa_eco_1 + sepa_eco_2 + sepa_eco_3 + sepa_eco_4 + sepa_eco_5
-conservative ~ beta * family"
-model_us <- "
-family =~ fam_1 + fam_2 + fam_3
-conservative =~ secs_soc_1 + secs_soc_2 + secs_soc_3 + secs_soc_4 + secs_soc_5 +
-                secs_soc_6 + secs_soc_7 + secs_eco_1 + secs_eco_2 + secs_eco_3 +
-                secs_eco_4 + secs_eco_5
-conservative ~ beta * family"
 
-# Estimate the model in lavaan
-results_dk <- sem(model = model_dk, data = DK)
-results_us <- sem(model = model_us, data = US)
-
-# Bind the models into a list
-results <- list(results_nl, results_dk, results_us)
-# Test the hypothesis that the effect size labeled 'beta' is positive
-bf_tot <- pbf(results, hypothesis = "beta > 0", standardize = TRUE)
-bf_tot
-```
-
-As the BF in all three samples is positive, the resulting PBF is very large, $PBF `report(bf_tot[1,1])`$.
-We can thus conclude that the central hypothesis receives overwhelming evidence across samples.
 <!-- In the first use-case of `pbf()`, an `lm()` is ran for each sample. All `lm` objects are stored in a list which is passed to `pbf()`. -->
@@
 
-```{r, echo = TRUE, eval = FALSE}
+```{r, echo = TRUE, eval = TRUE, results = "hide"}
 # Add the additional predictor to the model, label the effect beta2
-model_nl <- paste(model_nl, "group =~ grp_1 + grp_2 + grp_3
-                             conservative ~ beta2 * group", sep = "\n")
+model_nl <- c(model_nl, "group =~ grp_1 + grp_2 + grp_3
+                         con ~ beta2 * group")
 
@@
 # Obtain BF for each sample; note that the Dutch sample has two hypotheses
-bf_nl <- bain(results_nl, hypothesis = "beta > 0;beta2 < 0", standardize = TRUE)
+bf_nl <- bain(results_nl, hypothesis = "beta > 0;
+                                        beta2 < 0", 
+              standardize = TRUE)
 bf_dk <- bain(results_dk, "beta > 0")
@@
 Note that, when analyzing a single parameter per sample, the standard error is sufficient to construct the asymptotic covariance matrix.
-Thus, this method is suitable for data that have been prepared for meta-analysis.
+Thus, this method can be applied to data that have been prepared for classic meta-analysis (effect sizes and their sampling variances).
 Importantly, unlike meta-analysis, the present method is suitable for conceptual replications.
@@
 
-Note that we intentionally manipulate the data to fit these types of analyses;
-for example, we compute mean scale scores and dichotomize a continuous variable to conduct a t-test.
+Note that we intentionally manipulate the data to illustrate these different analyses;
+for example, we compute mean scale scores and dichotomize the continuous conservatism scale to conduct a t-test.
 We do not advocate these practices for applied research.
@@
 # NL: Conduct t-test using Cohen's D
-NL$group <- cut(NL$conservative, breaks = 2, labels = c("liberal", "conservative"))
+NL$group <- cut(NL$conservative, breaks = 2,
+                labels = c("liberal", "conservative"))
 sample_sizes <- table(NL$group)
@@
 NL_est <- diff(tapply(NL$family, NL$group, mean)) / pooled_sd
-NL_var <- (sum(sample_sizes) / prod(sample_sizes)) + (NL_est^2 / (2*sum(sample_sizes)))
+NL_var <- (sum(sample_sizes) / prod(sample_sizes)) +
+  (NL_est^2 / (2*sum(sample_sizes)))
 
@@
 Then, we use `bain.default()` to evaluate the central hypothesis on each parameter estimate. 
-The pbf can then be calculated as in the previous use case.
+The `pbf()` function can be called on a list of the resulting bain objects.
 
-```{r}
+```{r eval = TRUE, echo = TRUE, results='hide'}
 # Use bain.default() to obtain BF for the central hypothesis 
@@
 
-<!--
 ```{r, eval = FALSE, echo = FALSE}
@@
 
-Then, each object is converted to a `bain` object with these estimates
-```{r}
+```{r eval = FALSE, echo = FALSE, results = "hide"}
 bain_sample1_ests <- bain(x = ests_sample1[['means']], 
@@
 bain_allsamples_ests = list(bain_sample1_ests, bain_sample2_ests, bain_sample3_ests)
+pbf(bain_allsamples_ests)
 ```
 
-And finally, on these objects, `pbf()` can be called.
-```{r}
-pbf(bain_allsamples_ests)
-```
--->
 <!-- It is interesting that these estimates are not identical to the other examples... Is that because we do not have ALL information available in this use-case, or is something off in the script? -->
 <!-- CJ: We should ask Herbert about this, but this example won't make it into this paper. If you want, you can send Herbert a reproducible example of the 'problem'!-->
+
+# Conclusion
+
+<!-- Researchers need to choose a method that best suits their research question, -->
+ <!-- the study highlights the importance of carefully considering the research question when selecting a method for meta-analysis, and the potential trade-offs that may exist between sensitivity and specificity when using different methods.
+-->
+
+In conclusion, this study evaluated the performance of the product Bayes factor as a method for evidence synthesis,
+and compared it against other commonly used evidence synthesis methods under different simulation conditions.
+Compared to the other methods,
+PBF had the highest overall accuracy.
+This was primarily due to its greater sensitivity.
+However, PBF had lower specificity than all other algorithms, suggesting a trade-off between sensitivity and specificity.
+The other algorithms showed ceiling effects in specificity, limiting their sensitivity.
+The performance of the PBF was most strongly affected by sample size,
+followed by the number of samples and reliability.
+We introduced a user-friendly implementation of the PBF in the `bain` R-package,
+and demonstrated its use with various analysis techniques in R,
+as well as with sufficient statistics that are already routinely coded for meta-analysis (i.e., effect sizes and their sampling variance).
+This means that researchers can now use the PBF to aggregate evidence in situations where classic meta-analytic methods are less suitable.
+for example, when one informative hypothesis has been tested in several replication studies,
+but these replication studies are quite heterogeneous because they sample from different populations and use different methods or analysis techniques.
+Especially when the number of replication studies is too small to adequately account for these sources of between-study heterogeneity,
+the PBF may be a useful method to aggregate evidence for the common informative hypothesis.
+Researchers should be aware that the PBF trades off increased sensitivity for decreased specificity,
+and that it addresses a different research question than other research synthesis methods.
+This highlights the importance of careful interpretation of the results,
+and consideration of the research question when selecting an aggregation method.
+In sum,
+our results suggest that PBF is a useful evidence synthesis method,
+which is now broadly accessible due to its inclusion in the `bain` R-package.
 

### 2023-04-25 manuscript.Rmd
Rebecca's edits

@@
 abstract: |
-  The product Bayes factor (PBF) can synthesize evidence in favor of an informative hypothesis across heterogeneous replication studies.
+  The product Bayes factor (PBF) can synthesize evidence for an informative hypothesis across heterogeneous replication studies.
 It is particularly useful when the number of studies is relatively low and conventional assumptions about between-studies heterogeneity are likely violated.
@@
 out <- readRDS("out.RData")
+resultaction = "markup"
 ```
@@
 and thus enable drawing overarching conclusions across multiple (replication) studies.
-In this paper, we introduce a new research synthesis method that uses Bayesian evidence synthesis to aggregate the evidence in favor of an informative hypothesis, quantified by the Bayes factor,
+This paper addresses Bayesian evidence synthesis,
+a research synthesis method that aggregates evidence for an informative hypothesis, quantified by the Bayes factor,
 across multiple studies.
 This method has the potential to provide a more comprehensive and accurate picture of the state of the literature, and to identify areas of consensus and disagreement among studies.
-We describe the method in detail and demonstrate its application through a tutorial example analysis.
+We describe the method in detail,
+benchmark its performance against other commonly used research synthesis methods,
+and demonstrate its application through a tutorial example analysis.
+To facilitate uptake of the method by applied researchers, we introduce an implementation of the method in the `bain` R-package.
+This implementation enables the use of Bayesian evidence synthesis for many commonly used statistical analyses in R.
 
@@
 <!-- even machine learning based methods cannot identify relevant moderators. -->
-It is still possible, however, to quantify the support these studs provide for the underlying informative hypothesis.
-To this end, Bayesian evidence synthesis (BES) aggregates the evidence in favor of an informative hypothesis $H_i$ across studies,
+It is still possible, however, to quantify the support these studies provide for the underlying informative hypothesis.
+To this end, Bayesian evidence synthesis (BES) aggregates the evidence for a theoretical relationship across studies,
 without imposing assumptions about heterogeneity [@kuiperCombiningStatisticalEvidence2013].
+Although this assumption is not necessary,
+for the sake of simplicity we assume that this theoretical relationship is evaluated via informative hypothesis $H_i$ in all studies.
 
 The amount of evidence for a hypothesis can be expressed as a Bayes factor, or BF.
-The BF can be interpreted as the ratio of evidence in favor of one hypothesis $H_1$ relative to another hypothesis $H_2$, so $BF_{12} = \frac{H_1}{H_2}$.
+The BF can be interpreted as the ratio of evidence for one hypothesis relative to another hypothesis.
+<!-- , so $BF_{12} = \frac{BF_{1u}}{BF_{2u}}$. -->
+<!-- Note that $BF_{12}$ is itself a ratio of two Bayes factors with a common denominator, which renders them comparable [see @guApproximatedAdjustedFractional2018]. -->
 Within the scope of this paper,
-all Bayes factors are the ratio of evidence in favor of an informative hypothesis $H_i$ relative to its complement $H_{!i}$, $BF_{c} = \frac{H_i}{H_!i}$. 
+all Bayes factors are the ratio of evidence for an informative hypothesis $H_i$ relative to its complement $H_{!i}$ [see @guApproximatedAdjustedFractional2018].
+<!-- This Bayes factor , $BF_{c} = \frac{H_i}{H_!i}$.  -->
 The subscript $_!$ represents the negation operator;
 in other words, $H_{!i}$ means "not $H_i$".
-Consequently, $BF_c$ represents the ratio of evidence in favor of $H_i$ divided by evidence against it.
-A value of $BF_c = 10$ means that the data provide ten times more support in favor of the hypothesis than against it.
+This Bayes factor, which we will refer to as $BF_c$, represents the ratio of evidence for $H_i$ divided by evidence against it.
+A value of $BF_c = 10$ means that the data provide ten times more support for the hypothesis than against it.
 
@@
 Given certain assumptions about between-studies heterogeneity, what is the average population effect size?
-BES, on the other hand, aggregates evidence in favor of an informative hypothesis across multiple studies.
+BES, on the other hand, aggregates evidence for an informative hypothesis across multiple studies.
 It thus answers the question: Do all these studies support the hypothesis of interest?
@@
 <!-- I do see that this would end up in multiple results (once using .1 and 0.2, once (say) 0 or 0.05 and .2; but also when using other cut-off values for PBF). -->
-The informative hypothesis was kept constant at $H_i: \rho > .1$.
+The informative hypothesis, set to be equal across studies,
+was $H_i: \rho > .1$.
 To examine the performance of the different evidence synthesis methods in a range of scenarios,
@@
 First was the presence or absence of a true population effect.
-A present effect was defined as $\rho = .2$;
+Given the informative hypothesis of $H_i: \rho > .1$,
+the presence of a true population effect was defined as $\rho = .2$ and
 a null effect was defined as $\rho = .1$.
@@
 power to reject a false null hypothesis of $\beta \in (.10, .30, .60, .80)$ power, respectively,
-assuming $\alpha = .05$ and a known effect size of $\rho = .1$.
+assuming $\alpha = .05$ and a known effect size of $\rho = .1$ [@cohenStatisticalPowerAnalysis1988].
 Third, we manipulated the number of independent samples (or: replication studies), $k \in (2, 3, 10)$.
-Fourth, the reliability of the two correlated variables was varied between $\alpha \in (0.6, 0.8, 1.0)$,
-where 0.6 is the lowest reliability conventionally considered to be acceptable, and 1 represents perfect reliability, as is
-assumed when analyzing correlations between observed items or scale scores.
+Fourth, the reliability of the two correlated variables was varied between $\alpha \in (0.6, 0.8, 1.0)$ to range from questionable to perfect reliability [@nunnallyPsychometricTheory2017].
+Questionable reliability is the lowest level considered to be acceptable in social scientific research, and perfect reliability is what is assumed when analyzing correlations between observed items or scale scores.
 For all unique combinations of these design factors,
@@
 The decision criterion was that the hypothesis was supported in the majority of samples.
-Thus, for example, if $H_a$ was supported in three out of five samples, our vote counting algorithm would find overall support for $H_a$ (and, by extension, $H_i$).
+Thus, for example, if $H_0$ was rejected in three out of five samples, our vote counting algorithm would find overall support for $H_a$ (and, by extension, $H_i$).
 
@@
 For each algorithm, inferential decisions made using the criteria described above were compared to the population status of the hypothesis (true or false).
-The resulting confusion matrix gives the number of decisions that were true positives (TP), true negatives (TN), false positives (FP) and false negatives (FN).
+The resulting confusion matrix gives the number of decisions that were true positives (TP), true negatives (TN), false positives (FP), and false negatives (FN).
 These quantities were summarized as sensitivity, $\frac{TP}{TP + FN}$,
@@
 ```
-First, we examined overall model performance across conditions.
+We examined overall model performance across conditions.
 PBF had a higher overall accuracy than other algorithms
-followed by IPD, then RMA and finally VC,
+followed by IPD, then RMA, and finally VC,
 see Table \@ref(tab:tabconf).
@@
 
-The figure indicates that, for PBF at higher levels of $k$,
+Figure \@ref(fig:figk) indicates that, for PBF at higher levels of $k$,
 lower sensitivity was traded for greater specificity.
@@
 For instance, when raw data is unavailable, IPD cannot be used,
-and when effect sizes are not available, only VC can be used.
+and when parameter estimates or effect sizes are not reported, only VC can be used.
 
@@
 If the informative hypothesis is supported, this does not necessarily mean that it is also true.
-Consider the hypothetical example that a test of the informative hypothesis that the earth is flat received support with $BF = 3.01$.
+Consider the hypothetical example that the informative hypothesis that the earth is flat was supported with $BF = 3.01$.
 Although the data support this hypothesis over its complement, the hypothesis is clearly wrong (the earth is spherical).
-If we would have tested another hypothesis, e.g., the earth is shaped like an American football, it would have received much more support, e.g. $BF = 1000$, even though it is also wrong.
+If we would have evaluated another hypothesis, e.g., the earth is shaped like an American football, it would have received much more support, e.g. $BF = 1000$, even though it is also wrong.
 A high Bayes factor thus does not mean that the hypothesis is true.
@@
 and does not provide information about the true state of affairs.
-
+A related limitation is that our simulation study used an arbitrary - albeit conventional - threshold for inference [@jeffreys1998theory].
+In applied research, it is more sensible to evaluate the weight of evidence, rather than resorting to a rule of thumb.
 <!-- In sum, this simulation study suggest that PBF may be a good choice when sensitivity is of utmost importance, but researchers should be mindful of the trade-off between specificity and sensitivity.
@@
 Van Leeuwen and colleagues conducted a theory-driven, preregistered study to address the research question whether political orientation and moral dispositions are associated.
-Suitable data were collected in three countries: the United states of America, Denmark and the Netherlands.
+Suitable data were collected in three countries: the United states of America, Denmark, and the Netherlands.
 Each sample contained multiple measures of political orientation and moral dispositions.
@@
 We will examine the informative hypothesis that self-reported importance of family morality is positively associated with a conservative socio-political orientation.
-First, we load the `bain` library and assign the data to three objects with convenient names:
-
-```{r eval = TRUE}
+We load the `bain` library and assign the data to three objects with convenient names:
+
+```{r eval = TRUE, echo = TRUE}
 library(bain)
@@
 <!-- Because the focus of this tutorial is on BES,  -->
-First, we briefly introduce the basic use of the `bain()` function, and how to interpret its output.
-We must estimate a model suitable for testing our informative hypothesis.
+We briefly introduce the basic use of the `bain()` function, and how to interpret its output.
+We must estimate a model suitable for evaluating our informative hypothesis.
 Because both scales consist of multiple items, we can use structural equation modeling (SEM) to perform latent variable regression [see @vanlissaTeacherCornerEvaluating2020]:
@@
 # The argument standardize = TRUE uses the standardized model parameters
-bf_nl <- bain(results_nl, hypothesis = "beta > 0", standardize = TRUE)
-```
-
-The informative hypothesis in this tutorial is $H_i: \beta > 0$, where $\beta$ (beta) is the standardized regression coefficient.
-This corresponds to a traditional one-sided null-hypothesis test.
-Note that this deviates from the original publication
-which used a minimal effect size of interest and specified $H_i: \beta > .1$.
+bf_nl <- bain(results_nl, hypothesis = "beta > .1", standardize = TRUE)
+```
+
+The informative hypothesis in this tutorial is $H_i: \beta > .1$, where $\beta$ (beta) is the standardized regression coefficient.
+Instead of a conventional null hypothesis, $H_0: \beta = 0$,
+the value of $.1$ was used as a minimal effect size of interest.
 The code below illustrates how to obtain a Bayes factor for this informative hypothesis,
@@
 if we had not done so, we could find the names of all model parameters by running `get_estimates(results_nl, standardize = TRUE)`.
-When evaluating the code below,
-note that the hypothesis is strongly supported when compared to its complement.
+The results indicate that the hypothesis is supported when compared to its complement.
 <!-- , $BF_c `report(bf_nl$fit$BF.c[1])`$. -->
@@
 
-```{r echo = TRUE, eval = TRUE, results='hide'}
+```{r echo = TRUE, eval = TRUE, results = resultaction}
 # Test that the effect labeled 'beta' is positive
-bf_nl <- bain(results_nl, hypothesis = "beta > 0", standardize = TRUE)
+bf_nl <- bain(results_nl, hypothesis = "beta > .1", standardize = TRUE)
 bf_nl
@@
 #specify Hypotheses to evaluate
-hypotheses <- "grp > 0"
+hypotheses <- "grp > .1"
 
@@
 
-```{r, eval = FALSE, echo = FALSE, results="hide"}
+```{r, eval = FALSE, echo = FALSE, results = resultaction}
 # Specify the models for DK and US
@@
 # Test the hypothesis that the effect size labeled 'beta' is positive
-bf_tot <- pbf(results, hypothesis = "beta > 0", standardize = TRUE)
-```
-
-As mentioned before, suitable data were collected to test the substantive hypothesis in three countries.
+bf_tot <- pbf(results, hypothesis = "beta > .1", standardize = TRUE)
+```
+
+As mentioned before, suitable data were collected to evaluate the substantive hypothesis in three countries.
 There are differences between countries that prevent analyzing these data as a multilevel model, however.
@@
 This is an appropriate situation to use the PBF to aggregate evidence across countries.
-First, we estimate a latent regression model for the remaining two countries, taking care to use the same label for the parameter of interest in all samples.
+Below, we estimate a latent regression model for the remaining two countries, taking care to use the same label for the parameter of interest in all samples.
 Then, we bind all three SEM-models in a list, and call PBF to evaluate the hypothesis of interest on all models and aggregate the evidence.
 As the BF in all three samples is positive, the resulting PBF is very large.
-We can thus conclude that the central hypothesis receives overwhelming evidence across samples.
+We can thus conclude that the central hypothesis receives overwhelming support across samples.
 
@@
 # Test the hypothesis that the effect size labeled 'beta' is positive
-pbf(results, hypothesis = "beta > 0", standardize = TRUE)
+pbf(results, hypothesis = "beta > .1", standardize = TRUE)
 ```
@@
 The example below illustrates this use case.
-As before, all analyses share one hypotheses in common ($H_i: \beta_{fam} > 0$),
-but the Dutch sample now contains a sample-specific hypothesis regarding the effect of group morality, namely that $\beta_{grp} < 0$.
+As before, all analyses share one hypotheses in common ($H_i: \beta_{fam} > .1$),
+but the Dutch sample now contains a sample-specific hypothesis regarding the effect of group morality, namely that $\beta_{grp} < .1$.
 The `pbf()` function is called on a list of `bain` objects.
@@
 
-```{r, echo = TRUE, eval = TRUE, results = "hide"}
+```{r, echo = TRUE, eval = TRUE, results = resultaction}
 # Add the additional predictor to the model, label the effect beta2
@@
 # Obtain BF for each sample; note that the Dutch sample has two hypotheses
-bf_nl <- bain(results_nl, hypothesis = "beta > 0;
-                                        beta2 < 0", 
+bf_nl <- bain(results_nl, hypothesis = "beta > .1;
+                                        beta2 < .1", 
               standardize = TRUE)
-bf_dk <- bain(results_dk, "beta > 0")
-bf_us <- bain(results_us, "beta > 0")
+bf_dk <- bain(results_dk, "beta > .1")
+bf_us <- bain(results_us, "beta > .1")
 
@@
 
-A third use-case occurs when the raw data from different samples is not available.
+A third use-case occurs when the raw data from different samples are not available.
 This may happen, for example, when aggregating findings from the published literature (similar to meta-analysis).
@@
 It does not require uniform effect size measures across studies.
-The example below illustrates how to aggregate evidence in favor of one hypotheses across three studies that each used different methods.
+The example below illustrates how to aggregate evidence for one hypotheses across three studies that each used different methods.
 <!-- CJ: What I'd like to do here is show how to aggregate a t-test, a regression coefficient and a correlation coefficient -->
 
-The present use case effectively tests the following hypothesis:
+The present use case evaluates the following hypothesis:
 *There is a positive association between family morality and political conservatism*.
-This conceptual hypothesis is tested differently in the three samples, resulting in three different types of statistics and distinct sample-specific hypotheses:
-
-1. A t-test was performed using the NL data; using Cohen's D gives $H_i^{NL}: \delta_{conservative > liberal} > 0$
-1. A bivariate regression coefficient for `grp` was calculated using the DK data, giving $H_i^{DK}: b_{fam} > 0$
-1. A correlation coefficient was calculated using the US data, giving $H_i^{US}: \rho_{fam,con} > 0$
+This conceptual hypothesis is evaluated differently in the three samples, resulting in three different types of statistics and distinct sample-specific hypotheses:
+
+1. A t-test was performed using the NL data; using Cohen's D gives $H_i^{NL}: \delta_{conservative > liberal} > 0$, where $\delta$ is the mean difference between groups.
+1. A bivariate regression coefficient was calculated using the DK data, giving $H_i^{DK}: \beta_{fam} > 0$
+1. A correlation coefficient was calculated using the US data, giving $H_i^{US}: \rho_{fam,con} > 0$, where $\rho$ is the correlation between family morality and conservatism.
 
@@
 First we obtain the relevant parameter estimates and their sampling variances,
-which allows us to test the specific hypotheses in `bain`:
+which allows us to evaluate the specific hypotheses in `bain`:
 
@@
 
-```{r eval = TRUE, echo = TRUE, results='hide'}
+```{r eval = TRUE, echo = TRUE, results = resultaction}
 # Use bain.default() to obtain BF for the central hypothesis 
@@
 
-The results suggest substantial evidence in favor of the hypothesis that there is a positive association between family morality and political conservatism.
+The results suggest substantial evidence for the hypothesis that there is a positive association between family morality and political conservatism.
 Although each study used a different method to assess this hypothesis,
-their evidence can be synthesized using the pbf.
 
@@
 
-```{r eval = FALSE, echo = FALSE, results = "hide"}
+```{r eval = FALSE, echo = FALSE, results = resultaction}
 bain_sample1_ests <- bain(x = ests_sample1[['means']], 
@@
 The performance of the PBF was most strongly affected by sample size,
-followed by the number of samples and reliability.
 We introduced a user-friendly implementation of the PBF in the `bain` R-package,
@@
 This means that researchers can now use the PBF to aggregate evidence in situations where classic meta-analytic methods are less suitable.
-for example, when one informative hypothesis has been tested in several replication studies,
+For example, when one informative hypothesis has been evaluated in several replication studies,
 but these replication studies are quite heterogeneous because they sample from different populations and use different methods or analysis techniques.

### 2023-04-25 manuscript.Rmd
prepare preprint

@@
   - name: "Caspar J. Van Lissa"
-    affiliation: "1,2"
+    affiliation: "1"
     corresponding: yes
-    address: "Padualaan 14, 3584CH Utrecht, The Netherlands"
-    email: "c.j.vanlissa@uu.nl"
+    address: "Professor Cobbenhagenlaan 125, 5037 DB Tilburg, The Netherlands"
+    email: "c.j.vanlissa@tilburguniversity.edu"
     role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
@@
 authornote: |
-  This is a preprint paper, generated from Git Commit # `r substr(gert::git_commit_id(),1,7)`. This work was funded by a NWO Veni Grant (NWO Grant Number VI.Veni.191G.090), awarded to the lead author.
+  This is a preprint paper, generated from Git Commit # `r substr(gert::git_commit_id(),1,8)`. This work was funded by a NWO Veni Grant (NWO Grant Number VI.Veni.191G.090), awarded to the lead author.
 
@@
 It is particularly useful when the number of studies is relatively low and conventional assumptions about between-studies heterogeneity are likely violated.
-The present paper introduces a user-friendly implementation of the PBF in the bain R-package, and demonstrates its use in several tutorial examples.
+The present paper introduces a user-friendly implementation of the PBF in the `bain` R-package.
   The method was validated in a simulation study that manipulated sample size, number of replication samples, and reliability.
-PBF had a higher overall accuracy when benchmarked against other evidence synthesis methods, including random-effects meta-analysis (RMA).
+Several tutorial examples demonstrate the use of the method in distinct use cases.
+Results of the simulation study show that PBF had a higher overall accuracy when benchmarked against other evidence synthesis methods, including random-effects meta-analysis (RMA).
 This was primarily due to PBF’s greater sensitivity in detecting a true effect.
@@
 PBF thus appears to be a promising method for meta-analysis of heterogeneous conceptual replication studies.
-Nonetheless, users should be aware of its lower specificity, and the fact that it takes a Bayesian approach to inference and addresses a qualitatively different research question than other evidence synthesis methods.
+Nonetheless, users should be aware of its lower specificity, and the fact that the Bayesian approach to inference addresses a qualitatively different research question than other evidence synthesis methods.
 
 keywords          : "bayes factor, evidence synthesis, bayesian, meta-analysis"
-wordcount         : "5356"
+wordcount         : "5039"
 
@@
 out <- readRDS("out.RData")
-resultaction = "markup"
+resultaction = "hide" # "markup"
 ```
@@
 Third, when there are a few systematic differences between studies, these can be accounted for using meta-regression.
-Finally, when there are many potential variables that cause systematic differences and it is not known beforehand which are relevant, exploratory techniques like random forest meta-analysis and penalized meta-regression can be used to identify relevant moderators [@vanlissaSelectRelevantModerators2021].
+Finally, when there are many potential variables that cause systematic differences and it is not known beforehand which are relevant, exploratory techniques like random forest meta-analysis and penalized meta-regression can be used to identify relevant moderators [@vanlissaSelectingRelevantModerators2023].
 However, accounting for moderators requires a relatively high number of observations per moderator,
 which may not be available.
-Each of these approaches thus makes different assumptions about the nature of heterogeneity [see "Models for meta-analysis" in @vanlissaSmallSampleMetaanalyses2020].
-
-A crucial shortcoming of existing research synthesis methods is that the aforementioned assumptions
+
+Each of the aforementioned approaches makes different assumptions about the nature of heterogeneity [see "Models for meta-analysis" in @vanlissaSmallSampleMetaanalyses2020].
+A crucial shortcoming of existing research synthesis methods is that these assumptions
 may not be tenable when meta-analyzing studies that investigate the same informative hypothesis,
@@
 
-This paper introduces the first implementation of BES in user-friendly free open source software (FOSS).
+This paper introduces the first implementation of BES in user-friendly free open source software.
 A function `pbf()` was contributed to the `bain` R-package for Bayesian informative hypothesis evaluation, version `0.2.9`.
 <!-- `r as.character(packageVersion("bain"))`. -->
-This paper presents a simulation study to validate the method and benchmark it against alternative evidence synthesis methods,
-and illustrates several use cases through reproducible examples.
+This paper presents a simulation study to validate the method and benchmark it against alternative evidence synthesis methods.
+It additionally illustrates several use cases through reproducible examples.
 
@@
 As a benchmark for comparison, we included several other algorithms that might feasibly be used by researchers who intend to examine whether a hypothesis is true across several independent samples.
-The first was *vote counting*: counting the number of significant effects.
+The first benchmark was *vote counting*: counting the number of significant effects.
 Although this method is still in use for aggregating conceptual replications,
@@
 
-The third algorithm was *random-effects meta-analysis* (RMA),
+The second benchmarking algorithm was *random-effects meta-analysis* (RMA),
 which is the current gold standard for evidence synthesis [@viechtbauerConductingMetaanalysesMetafor2010].
-For this algorithm, the null-hypothesis was rejected if the lower bound of a 90% confidence interval for the overall effect size excluded $H_i$.
-Note that this corresponds to a test at $\alpha = .05$,
+For this algorithm, the null-hypothesis was rejected if a 90% confidence interval excluded the hypothesized value under $H_i$.
+Note that a 90% confidence interval corresponds to a test at $\alpha = .05$,
 because all effects in the simulation are directional.
 
-The fourth algorithm was *individual participant data* (IPD) meta-analysis [@rileyMetaanalysisIndividualParticipant2010].
+The third benchmarking algorithm was *individual participant data* (IPD) meta-analysis [@rileyMetaanalysisIndividualParticipant2010].
 Like classic meta-analysis (RMA), IPD is a multilevel model, clustered by sample.
@@
 With this in mind, it is informative to benchmark it against both of these methods.
-IPD was also evaluated using the lower bound of a 90% confidence interval for the overall effect size.
+Just as for RMA, a 90% confidence interval was used for inference.
 <!-- Rebecca: -->
@@
 significance tests were uninformative.
-We thus focused on interpreting the effect size of the effects of design factors.
+We thus focused on interpreting the effect sizes of the design factors.
 The performance of PBF was most impacted by sample size $n$, followed by the number of groups $k$, and reliability.
@@
 
-Figure \@ref(fig:figk) indicates that, for PBF at higher levels of $k$,
-lower sensitivity was traded for greater specificity.
-The other algorithms did not show this pattern, as their specificity was at a ceiling.
-This difference in pattern of effects explains why number of samples had a moderate effect on the difference between algorithms (see Table \@ref(tab:tabeffect)).
+Figure \@ref(fig:figk) indicates that PBF showed increasing specificity at higher levels of $k$,
+while sensitivity was relatively unaffected.
+RMA and IPD showed a similar pattern,
+although their specificity was at a ceiling.
 Only VC showed decreasing sensitivity with an increasing number of groups;
 this is because the probability of obtaining false negatives increases with the number of groups [as also noted by @hedgesVotecountingMethodsResearch1980].
+This difference in pattern of effects explains why number of samples had a moderate effect on the difference between algorithms (see Table \@ref(tab:tabeffect)).
 
@@
 
-<!-- In the first use-case of `pbf()`, an `lm()` is ran for each sample. All `lm` objects are stored in a list which is passed to `pbf()`. -->
+<!-- In the first use case of `pbf()`, an `lm()` is ran for each sample. All `lm` objects are stored in a list which is passed to `pbf()`. -->
 <!-- `pbf()` supports lists of all objects that `bain()` also supports.
@@
 
-<!-- In the first use-case of `pbf()`, an `lm()` is ran for each country.  -->
+<!-- In the first use case of `pbf()`, an `lm()` is ran for each country.  -->
 <!-- All `lm` objects are stored in a list which is passed to `pbf()`. -->
@@
 
-A third use-case occurs when the raw data from different samples are not available.
+A third use case occurs when the raw data from different samples are not available.
 This may happen, for example, when aggregating findings from the published literature (similar to meta-analysis).
@@
 Although each study used a different method to assess this hypothesis,
+their evidence can be synthesized using `pbf()`.
 
@@
 
-<!-- It is interesting that these estimates are not identical to the other examples... Is that because we do not have ALL information available in this use-case, or is something off in the script? -->
+<!-- It is interesting that these estimates are not identical to the other examples... Is that because we do not have ALL information available in this use case, or is something off in the script? -->
 <!-- CJ: We should ask Herbert about this, but this example won't make it into this paper. If you want, you can send Herbert a reproducible example of the 'problem'!-->
@@
 The performance of the PBF was most strongly affected by sample size,
+followed by the number of samples and reliability.
 We introduced a user-friendly implementation of the PBF in the `bain` R-package,
@@
 \newpage
+# Highlights
+
+* Many research synthesis methods make strong assumptions about between-studies heterogeneity that are violated when studies are conceptually replicated.
+* The product Bayes factor (PBF) aggregates evidence for an informative hypothesis across conceptual replication studies without imposing assumptions about heterogeneity.
+* This paper introduces a user-friendly way to compute the PBF for a variety of widely used models via the `pbf()` function in the `bain` R-package.
+* A simulation study shows favorable performance for PBF relative to random effects meta-analysis, individual participant data meta-analysis, and vote counting.
+* Three tutorial examples illustrate distinct use cases of the method.
+
+# Data Availability Statement
+
+All analysis code is available in a version-controlled repository at <https://github.com/cjvanlissa/bayesynth>.
+
+\newpage
 

### 2024-04-09 simulation.Rmd
update 2024-04-09 14:52:13.376312

--- 2023-04-25 manuscript.Rmd
@@
 
-bibliography      : ["product_bayes.bib", "references.bib"]
+bibliography      : ["product_bayes.bib"]
 
@@
 ```
-
-Recent years have seen a crisis of confidence over the reliability of published results in psychology, and science more broadly [@brembs2018prestigious].
-Replication research has emerged as one potential way to address this crisis and derive knowledge that will stand the test of time [see @lavelleWhenCrisisBecomes2021]. 
-In step with this interest in replication research,
-research synthesis methods have become increasingly popular.
-These methods aggregate research findings,
-and thus enable drawing overarching conclusions across multiple (replication) studies.
-This paper addresses Bayesian evidence synthesis,
-a research synthesis method that aggregates evidence for an informative hypothesis, quantified by the Bayes factor,
-across multiple studies.
-This method has the potential to provide a more comprehensive and accurate picture of the state of the literature, and to identify areas of consensus and disagreement among studies.
-We describe the method in detail,
-benchmark its performance against other commonly used research synthesis methods,
-and demonstrate its application through a tutorial example analysis.
-To facilitate uptake of the method by applied researchers, we introduce an implementation of the method in the `bain` R-package.
-This implementation enables the use of Bayesian evidence synthesis for many commonly used statistical analyses in R.
-
-A key challenge in quantitative research synthesis is dealing with between-studies heterogeneity [@higginsReevaluationRandomeffectsMetaanalysis2009].
-When studies examine the same research question in different laboratories, use idiosyncratic methods, and sample from distinct populations, these between-study differences can introduce heterogeneity in findings.
-The most common quantitative research synthesis method is meta-analysis,
-in which results of different studies are aggregated to estimate an aggregate effect size [@borensteinIntroductionMetaAnalysis2009].
-In meta-analysis, heterogeneity can be accounted for in four ways [see @vanlissaSmallSampleMetaanalyses2020].
-First, if studies are exact replications,
-one may assume that no heterogeneity in the outcome exists and a fixed-effect meta-analysis can be conducted to estimate the common population effect. 
-Second, when heterogeneity between studies can be assumed to be random,
-random-effects meta-analysis can be used to estimate the mean of a distribution of population effects. 
-Third, when there are a few systematic differences between studies, these can be accounted for using meta-regression.
-Finally, when there are many potential variables that cause systematic differences and it is not known beforehand which are relevant, exploratory techniques like random forest meta-analysis and penalized meta-regression can be used to identify relevant moderators [@vanlissaSelectingRelevantModerators2023].
-However, accounting for moderators requires a relatively high number of observations per moderator,
-which may not be available.
-
-Each of the aforementioned approaches makes different assumptions about the nature of heterogeneity [see "Models for meta-analysis" in @vanlissaSmallSampleMetaanalyses2020].
-A crucial shortcoming of existing research synthesis methods is that these assumptions
-may not be tenable when meta-analyzing studies that investigate the same informative hypothesis,
-but are otherwise very heterogeneous.
-The situation may arise where each study is uniquely identified by a combination of linearly dependent moderators.
-In this case, it is no longer possible to synthesize *effect sizes* while accounting for heterogeneity using statistical methods.
-<!-- even machine learning based methods cannot identify relevant moderators. -->
-It is still possible, however, to quantify the support these studies provide for the underlying informative hypothesis.
-To this end, Bayesian evidence synthesis (BES) aggregates the evidence for a theoretical relationship across studies,
-without imposing assumptions about heterogeneity [@kuiperCombiningStatisticalEvidence2013].
-Although this assumption is not necessary,
-for the sake of simplicity we assume that this theoretical relationship is evaluated via informative hypothesis $H_i$ in all studies.
-
-The amount of evidence for a hypothesis can be expressed as a Bayes factor, or BF.
-The BF can be interpreted as the ratio of evidence for one hypothesis relative to another hypothesis.
-<!-- , so $BF_{12} = \frac{BF_{1u}}{BF_{2u}}$. -->
-<!-- Note that $BF_{12}$ is itself a ratio of two Bayes factors with a common denominator, which renders them comparable [see @guApproximatedAdjustedFractional2018]. -->
-Within the scope of this paper,
-all Bayes factors are the ratio of evidence for an informative hypothesis $H_i$ relative to its complement $H_{!i}$ [see @guApproximatedAdjustedFractional2018].
-<!-- This Bayes factor , $BF_{c} = \frac{H_i}{H_!i}$.  -->
-The subscript $_!$ represents the negation operator;
-in other words, $H_{!i}$ means "not $H_i$".
-This Bayes factor, which we will refer to as $BF_c$, represents the ratio of evidence for $H_i$ divided by evidence against it.
-A value of $BF_c = 10$ means that the data provide ten times more support for the hypothesis than against it.
-
-When multiple studies each provide evidence for $H_i$ in the form of complement Bayes factors,
-these Bayes factors can be synthesized across studies by taking their product [@kuiperCombiningStatisticalEvidence2013].
-The resulting product Bayes factor (PBF) summarizes the total evidence for the hypothesis.
-The only assumption of the PBF is that all study-specific hypotheses provide evidence about the same underlying theoretical relationship.
-Note that other approaches to BES exist;
-for instance, it is possible to use the posterior of one study as the prior for a replication study, and thus accumulate evidence across studies [see @heckReviewApplicationsBayes2022].
-Such applications are out of scope of the present paper,
-which addresses the PBF approach to BES.
-
-Although meta-analysis and BES are both research synthesis methods, they answer different research questions.
-Meta-analysis estimates the point estimate or distribution of a population effect size. 
-It pools estimates of this effect size across multiple studies to obtain an overall estimate of the effect size.
-It thus answers questions like:
-Given certain assumptions about between-studies heterogeneity, what is the average population effect size?
-BES, on the other hand, aggregates evidence for an informative hypothesis across multiple studies.
-It thus answers the question: Do all these studies support the hypothesis of interest?
-Both methods are appropriate for different research questions, and provide complementary information.
 
@@
 A second important caveat is that most of the algorithms did not reach
-a level of sensitivity that is considered acceptable from a perspective of statistical power [i.e., greater than .80, @cohen2013statistical].
+a level of sensitivity that is considered acceptable from a perspective of statistical power [i.e., greater than .80, @cohenStatisticalPowerAnalysis1988].
 PBF performed notably better than other algorithms,
@@
 <!-- OTHER PBF APPROACHES -->
-<!-- One possible issue with the PBF as defined in this research comes from an article of @rouder2011bayes.  -->
 <!-- The article states that a meta-analytic Bayes factor is not as straightforward as taking the product of $k$ individual Bayes factors. -->
@@
 <!-- This means that there is no assumption of a constant population effect and/or equal types of test statistics across studies. -->
-<!-- Furthermore, as the article of @rouder2011bayes states itself, pooling data is preferred if all data is independent and drawn from a common distribution.  -->
 <!-- The study idiosyncrasies encountered in evidence synthesis should wary the researcher in blindly making the assumption of a common distribution. -->
@@
 <!-- Again, this has to do with the small simulated difference between the true effect and the boundary value. -->
-
-# Tutorial
-
-This tutorial demonstrates how to synthesize evidence for an informative hypothesis across heterogeneous replications using the Product Bayes Factor (PBF). 
-We assume that users have installed the free open source statistical programming language R [@R-base].
-The R-package `bain` version `0.2.9` or later is required, which can be installed by running `install.packages("bain")` in the R console.
-The data used in this tutorial are included in the `bain` package,
-and have been simulated based on the data presented in [@vanleeuwen2022morality].
-A more detailed description of the datasets is found in [@vanleeuwen2022morality];
-additionally, the dataset documentation is accessed by running `?synthetic_us`, `?synthetic_dk` or `?synthetic_nl` in the R console.
-Van Leeuwen and colleagues conducted a theory-driven, preregistered study to address the research question whether political orientation and moral dispositions are associated.
-Suitable data were collected in three countries: the United states of America, Denmark, and the Netherlands.
-Each sample contained multiple measures of political orientation and moral dispositions.
-In the original publication, the PBF was used to aggregate evidence across scales and countries to obtain an overall measure of support for the central hypothesis.
-This tutorial follows the same rationale, but uses only one effect size per sample,
-and varies the way this effect size is computed to illustrate the more typical use case where the same informative hypothesis has been studied in different ways in multiple studies.
-We will examine the informative hypothesis that self-reported importance of family morality is positively associated with a conservative socio-political orientation.
-We load the `bain` library and assign the data to three objects with convenient names:
-
-```{r eval = TRUE, echo = TRUE}
-library(bain)
-NL <- synthetic_nl
-DK <- synthetic_dk
-US <- synthetic_us
-```
-
-<!-- skeleton
-1.) explain calculation of single BF and what it means, refer to previous tutorial on the bayes factor
-2.) introduce the case where there are multiple studies and we want to evaluate a single hypothesis
-3.) introduce pbf(), its uses, outputs and interpretation
--->
-
-### How to use bain
-
-<!-- CJ: Is dit niet teveel herhaling van wat we in de inleiding zeggen?? -->
-<!-- In conventional Null-hypothesis significance testing (NHST), as the name implies, data is analysed to generate evidence for or against the null-hypothesis.  -->
-<!-- This method evaluates a single hypothesis, the null-hypothesis, and its output is dichotomous in nature: whether a theoretical relationship is probable or improbable to exist given the data.  -->
-<!-- However, there are many more hypotheses that can theoretically be tested and researchers may not be interested in the null-hypothesis.
-They may also not only be interested in whether the hypothesis is probable, but also by how much.
--->
-<!-- In the Bayesian framework, the Bayes Factor (BF) is a test statistic that gives the amount of support for any possible desired hypothesis.  -->
-<!-- Because the focus of this tutorial is on BES,  -->
-We briefly introduce the basic use of the `bain()` function, and how to interpret its output.
-We must estimate a model suitable for evaluating our informative hypothesis.
-Because both scales consist of multiple items, we can use structural equation modeling (SEM) to perform latent variable regression [see @vanlissaTeacherCornerEvaluating2020]:
-<!--EB: For simplicities sake, I for now create simple average scores without taking into account reverse coding. This should likely be improved with factor scores -->
-<!-- CJ: Keep the code SUPER simple. No functions etc -->
-
-```{r echo = TRUE, eval = TRUE}
-# Load lavaan package for SEM
-library(lavaan)
-
-# Specify SEM-model for latent variable regression
-model_nl <- "
-fam =~ fam_1 + fam_2 + fam_3
-con =~ sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
-       sepa_eco_1 + sepa_eco_2 + sepa_eco_3 + sepa_eco_4 + sepa_eco_5
-con ~ beta * fam"
-
-# Estimate the model in lavaan
-results_nl <- sem(model = model_nl, data = NL)
-```
-```{r echo = FALSE, eval = FALSE}
-# Load relevant packages
-library(bain)
-library(lavaan)
-
-# Specify a SEM-model with two latent variables and latent regression
-model_nl <- "
-fam =~ fam_1 + fam_2 + fam_3
-con =~ sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
-                sepa_eco_1 + sepa_eco_2 + sepa_eco_3 + sepa_eco_4 + sepa_eco_5
-con ~ beta * fam"
-
-# Estimate the model in lavaan
-results_nl <- sem(model = model_nl, data = NL)
-
-# Test the hypothesis that the effect size labeled 'beta' is positive
-# The argument standardize = TRUE uses the standardized model parameters
-bf_nl <- bain(results_nl, hypothesis = "beta > .1", standardize = TRUE)
-```
-
-The informative hypothesis in this tutorial is $H_i: \beta > .1$, where $\beta$ (beta) is the standardized regression coefficient.
-Instead of a conventional null hypothesis, $H_0: \beta = 0$,
-the value of $.1$ was used as a minimal effect size of interest.
-The code below illustrates how to obtain a Bayes factor for this informative hypothesis,
-using the output of the SEM analysis above.
-We can refer to the parameter `beta` by name because we labeled it in the `lavaan` syntax;
-if we had not done so, we could find the names of all model parameters by running `get_estimates(results_nl, standardize = TRUE)`.
-The results indicate that the hypothesis is supported when compared to its complement.
-<!-- , $BF_c `report(bf_nl$fit$BF.c[1])`$. -->
-For a more in-depth tutorial on `bain()`, see @hoijtink2019tutorial, and for further guidance on the use of `bain()` for SEM, see @vanlissaTeacherCornerEvaluating2020.
-
-```{r echo = TRUE, eval = TRUE, results = resultaction}
-# Test that the effect labeled 'beta' is positive
-bf_nl <- bain(results_nl, hypothesis = "beta > .1", standardize = TRUE)
-bf_nl
-```
-
-<!-- Socio-political orientation and group loyalty have been aggregated to variables `soc` and `grp` respectively.  -->
-<!-- Both are continuous variables with higher scores indicating either conservatism or higher group loyalty.  -->
-<!-- In NHST we specify the null hypothesis that there is no association between `soc` and `grp`.  -->
-<!-- One method to test the hypothesis is to calculate a standardized regression coefficient. -->
-
-<!-- $H_0: \beta_{grp} = 0$. -->
-
-<!-- However, we may have prior expectations that higher group loyalty is somewhat related to  higher conservatism.  -->
-<!-- An informative hypothesis could be: -->
-
-<!-- $H_i: \beta_{grp} > 0$ -->
-
-<!-- SESAMESIM DATA: Let's say we are interested in whether the status of the site children grew up in (`site`) affects their score on a numbers test after watching Sesame Street for a year (`postnumb`).
-`postnumb` is a continuous variable which ranges from 0-63.
-`site` is a nominal variable indicating group membership with possible values 1 = disadvantaged inner city, 2 = advantaged suburban, 3 = advantaged rural, 4 = disadvantaged rural and 5 = disadvantaged Spanish speaking.
-In NHST we specify the null-hypothesis that the mean `postnumb` is equal for all groups in  site`:
-
-$H_0: \mu_{site1} =\mu_{site2} = \mu_{site3} = \mu_{site4} =\mu_{site5}$
-
-However, we may have prior expectations that the children growing up in an advantaged site will have a higher score than children growing up in a disadvantaged site. Our informative hypothesis is:
-
-$H_i: \mu_{site2},\mu_{site3} > \mu_{site1},\mu_{site4},\mu_{site5}$ -->
-
-<!-- Because it is very improbable that $\beta_{grp}$ is exactly equal to 0, more information can be won by testing $H_i$. -->
-<!-- We can test $H_i$ by running a linear regression on the data and passing the `lm` object with the specific hypotheses to `bain`.  -->
-<!-- For a single study, such as the study that investigated the Netherlands, the Bayes factor would be obtained as follows: -->
-
-```{r, echo = FALSE, eval = FALSE}
-# load bain package
-library(bain)
-
-# specify model, allow meaningful zero-point
-fit <- lm(soc ~ grp, data = NLa)
-
-#specify Hypotheses to evaluate
-hypotheses <- "grp > .1"
-
-fit_bain <- bain(fit, hypothesis = hypotheses)
-print(fit_bain)
-
-```
-<!-- SESAMESIM DATA: The output of `bain` seems to support neither hypothesis.
-The `fit` for both hypotheses is approximately zero which means that almost none of the data are in agreement with the restrictions the specified hypotheses pose.
-The `BF.c`, which is the Bayes Factor of the hypothesis versus its complement, is for both hypotheses close to zero.
-This implies that the complement of both hypotheses is $\frac{1}{BF.c}$ times more likely than the passed hypotheses.
-Finally, `PMPc`, the Posterior Model Probabilities of the specified hypotheses with $H_c$ included, is also close to zero for both specified hypothesis.
-This means that *out of* the specified hypotheses and their complement, the probability of $H_c$ being true is 99.9%.
--->
-
-### Aggregating evidence across studies
-
-```{r, eval = FALSE, echo = FALSE, results = resultaction}
-# Specify the models for DK and US
-model_dk <- "
-fam =~ fam_1 + fam_2 + fam_3
-con =~ sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
-sepa_eco_1 + sepa_eco_2 + sepa_eco_3 + sepa_eco_4 + sepa_eco_5
-con ~ beta * fam"
-model_us <- "
-fam =~ fam_1 + fam_2 + fam_3
-con =~
-secs_soc_1 + secs_soc_2 + secs_soc_3 + secs_soc_4 + secs_soc_5 + secs_soc_6 + secs_soc_7 +
-secs_eco_1 + secs_eco_2 + secs_eco_3 + secs_eco_4 + secs_eco_5
-con ~ beta * fam"
-
-# Estimate the model in lavaan
-results_dk <- sem(model = model_dk, data = DK)
-results_us <- sem(model = model_us, data = US)
-
-# Bind the models into a list
-results <- list(results_nl, results_dk, results_us)
-# Test the hypothesis that the effect size labeled 'beta' is positive
-bf_tot <- pbf(results, hypothesis = "beta > .1", standardize = TRUE)
-```
-
-As mentioned before, suitable data were collected to evaluate the substantive hypothesis in three countries.
-There are differences between countries that prevent analyzing these data as a multilevel model, however.
-For instance, conservatism was measured using different scales.
-This is an appropriate situation to use the PBF to aggregate evidence across countries.
-Below, we estimate a latent regression model for the remaining two countries, taking care to use the same label for the parameter of interest in all samples.
-Then, we bind all three SEM-models in a list, and call PBF to evaluate the hypothesis of interest on all models and aggregate the evidence.
-As the BF in all three samples is positive, the resulting PBF is very large.
-We can thus conclude that the central hypothesis receives overwhelming support across samples.
-
-```{r, eval = TRUE, echo = TRUE}
-# Specify the models for DK and US
-model_dk <- "
-fam =~ fam_1 + fam_2 + fam_3
-con =~
-sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
-sepa_eco_1 + sepa_eco_2 + sepa_eco_3 + sepa_eco_4 + sepa_eco_5
-con ~ beta * fam"
-model_us <- "
-fam =~ fam_1 + fam_2 + fam_3
-con =~
-secs_soc_1 + secs_soc_2 + secs_soc_3 + secs_soc_4 + secs_soc_5 +
-secs_soc_6 + secs_soc_7 +
-secs_eco_1 + secs_eco_2 + secs_eco_3 + secs_eco_4 + secs_eco_5
-con ~ beta * fam"
-
-# Estimate the model in lavaan
-results_dk <- sem(model = model_dk, data = DK)
-results_us <- sem(model = model_us, data = US)
-
-# Bind the models into a list
-results <- list(results_nl, results_dk, results_us)
-# Test the hypothesis that the effect size labeled 'beta' is positive
-pbf(results, hypothesis = "beta > .1", standardize = TRUE)
-```
-
-<!-- In the first use case of `pbf()`, an `lm()` is ran for each sample. All `lm` objects are stored in a list which is passed to `pbf()`. -->
-<!-- `pbf()` supports lists of all objects that `bain()` also supports.
-All supported types are specified in bain vignette *REF*.
-One exception is the use of named vectors, which will be covered later.
--->
-<!-- ```{r} -->
-<!-- lm_sample1 <- lm(postnumb~site-1, sample1) -->
-<!-- lm_sample2 <- lm(postnumb~site-1, sample2) -->
-<!-- lm_sample3 <- lm(postnumb~site-1, sample3) -->
-
-<!-- lm_allsamples <- list(lm_sample1, lm_sample2, lm_sample3) -->
-
-<!-- pbf(lm_allsamples, hypotheses) -->
-<!-- ``` -->
-
-<!-- The function returns a `data.frame` containing the passed hypotheses in the rows and in the columns the PBF and the individual Bayes Factor for each sample individually.
-The `PBF` of both hypotheses is close to zero, meaning that the combined evidence for both hypotheses over all studies is low.
-When looking at the individual Bayes Factors, we also see that the PBF is lower than all of the Bayes Factors individually.
-For sample 3 the specified hypotheses are even more likely than their complement.
-This example illustrates another strength of the PBF; being able to evaluate a hypothesis more extremely than the most extreme individual evaluation.
-If all samples do not support a hypothesis, than the number of times the hypothesis is not supported should be included in its evaluation.
-This leads to a more extreme combined evaluation than any individual evaluation.
-Likewise, if there is very strong support in only a small minority of the samples, while the majority of samples shows no support at all, than it is likely that the hypothesis is not true.
--->
-
-<!-- Sometimes there is access to the data of not only a single, but multiple studies that all contain the necessary information to evaluate our hypothesis $H_i$.  -->
-<!-- To evaluate $H_i$, it is possible to synthesize the support for $H_i$ over all studies using the pbf calculated with `pbf()`.  -->
-<!-- In this example, data to evaluate $H_i$ is available for three countries. -->
-
-<!-- In the first use case of `pbf()`, an `lm()` is ran for each country.  -->
-<!-- All `lm` objects are stored in a list which is passed to `pbf()`. -->
-<!-- `pbf()` supports lists of all objects that `bain()` also supports.  -->
-<!-- All supported types are specified in the `bain` vignette.  -->
-<!-- One exception is the use of named vectors, which will be covered later. -->
-<!-- ```{r} -->
-<!-- fit_US <- lm(soc ~ grp + her, data = USa) -->
-<!-- fit_DK <- lm(soc ~ grp + her, data = DKa) -->
-<!-- fit_NL <- lm(soc ~ grp + her, data = NLa) -->
-
-<!-- all_fits <- list(fit_US, fit_DK, fit_NL) -->
-
-<!-- pbf(all_fits, hypotheses) -->
-<!-- ``` -->
-
-### Using bain objects
-
-The `pbf()` function also accepts multiple `bain` objects.
-This makes it possible to, for example, evaluate different sets of hypotheses on different data sets before using the resulting `bain` objects to aggregate the evidence for all common hypotheses across datasets.
-The example below illustrates this use case.
-As before, all analyses share one hypotheses in common ($H_i: \beta_{fam} > .1$),
-but the Dutch sample now contains a sample-specific hypothesis regarding the effect of group morality, namely that $\beta_{grp} < .1$.
-The `pbf()` function is called on a list of `bain` objects.
-Note that, in this case, `pbf()` does not require an argument `hypothesis`, as the hypotheses are contained in the `bain` objects.
-
-```{r, echo = TRUE, eval = TRUE, results = resultaction}
-# Add the additional predictor to the model, label the effect beta2
-model_nl <- c(model_nl, "group =~ grp_1 + grp_2 + grp_3
-                         con ~ beta2 * group")
-
-# Estimate the model in lavaan
-results_nl <- sem(model = model_nl, data = NL)
-
-# Obtain BF for each sample; note that the Dutch sample has two hypotheses
-bf_nl <- bain(results_nl, hypothesis = "beta > .1;
-                                        beta2 < .1", 
-              standardize = TRUE)
-bf_dk <- bain(results_dk, "beta > .1")
-bf_us <- bain(results_us, "beta > .1")
-
-# Bind bain objects into a list
-bfs <- list(bf_nl, bf_dk, bf_us)
-
-# Call pbf on that list
-pbf(bfs)
-```
-
-As can be seen, the results are equivalent to the results in the previous example.
-The sample-specific hypothesis has been left out, and common hypotheses are retained and aggregated.
-If there are no common hypotheses across all objects, `pbf()` throws an error.
-
-```{r, echo = FALSE, eval = FALSE}
-bains_different_hyps <- list(
-  bain(fit_US, "grp = 0"),
-  bain(fit_DK, "grp > .1"),
-  bain(fit_NL, "grp < .1")
-)
-
-tryCatch({
-  pbf(bains_different_hyps)},
-  error = function(e){message(e)}
-      )
-
-```
-<!--The function returns an error stating that the samples have no hypothesis in common and thus the PBF cannot be calculated.-->
-
-### Using sufficient statistics
-
-A third use case occurs when the raw data from different samples are not available.
-This may happen, for example, when aggregating findings from the published literature (similar to meta-analysis).
-In this case, one can use the default interface of `bain`, as explained in [@hoijtink2019tutorial].
-This function requires four arguments: A named vector of parameter estimates, their asymptotic covariance matrix, the original sample size, and the number of within-group and between-group parameters.
-Note that, when analyzing a single parameter per sample, the standard error is sufficient to construct the asymptotic covariance matrix.
-Thus, this method can be applied to data that have been prepared for classic meta-analysis (effect sizes and their sampling variances).
-Importantly, unlike meta-analysis, the present method is suitable for conceptual replications.
-It does not require uniform effect size measures across studies.
-The example below illustrates how to aggregate evidence for one hypotheses across three studies that each used different methods.
-<!-- CJ: What I'd like to do here is show how to aggregate a t-test, a regression coefficient and a correlation coefficient -->
-
-The present use case evaluates the following hypothesis:
-*There is a positive association between family morality and political conservatism*.
-This conceptual hypothesis is evaluated differently in the three samples, resulting in three different types of statistics and distinct sample-specific hypotheses:
-
-1. A t-test was performed using the NL data; using Cohen's D gives $H_i^{NL}: \delta_{conservative > liberal} > 0$, where $\delta$ is the mean difference between groups.
-1. A bivariate regression coefficient was calculated using the DK data, giving $H_i^{DK}: \beta_{fam} > 0$
-1. A correlation coefficient was calculated using the US data, giving $H_i^{US}: \rho_{fam,con} > 0$, where $\rho$ is the correlation between family morality and conservatism.
-
-Note that we intentionally manipulate the data to illustrate these different analyses;
-for example, we compute mean scale scores and dichotomize the continuous conservatism scale to conduct a t-test.
-We do not advocate these practices for applied research.
-
-First we obtain the relevant parameter estimates and their sampling variances,
-which allows us to evaluate the specific hypotheses in `bain`:
-
-```{r eval = TRUE, echo = TRUE}
-# Create mean scale scores
-NL <- data.frame(
-  family = rowMeans(NL[c("fam_1", "fam_2", "fam_3")]),
-  conservative = rowMeans(NL[c("sepa_soc_1", "sepa_soc_2", "sepa_soc_3",
-                               "sepa_soc_4", "sepa_soc_5", "sepa_eco_1",
-                               "sepa_eco_2", "sepa_eco_3", "sepa_eco_4",
-                               "sepa_eco_5")]))
-DK <- data.frame(
-  family = rowMeans(DK[c("fam_1", "fam_2", "fam_3")]),
-  conservative = rowMeans(DK[c("sepa_soc_1", "sepa_soc_2", "sepa_soc_3",
-                               "sepa_soc_4", "sepa_soc_5", "sepa_eco_1",
-                               "sepa_eco_2", "sepa_eco_3", "sepa_eco_4",
-                               "sepa_eco_5")]))
-
-US <- data.frame(
-  family = rowMeans(US[c("fam_1", "fam_2", "fam_3")]),
-  conservative = rowMeans(US[c("secs_soc_1", "secs_soc_2", "secs_soc_3",
-                               "secs_soc_4", "secs_soc_5", "secs_soc_6",
-                               "secs_soc_7", "secs_eco_1", "secs_eco_2",
-                               "secs_eco_3", "secs_eco_4", "secs_eco_5")]))
-
-# NL: Conduct t-test using Cohen's D
-NL$group <- cut(NL$conservative, breaks = 2,
-                labels = c("liberal", "conservative"))
-sample_sizes <- table(NL$group)
-sds <- tapply(NL$family, NL$group, sd)
-pooled_sd <- sqrt(sum((sample_sizes - 1) * sds) / (sum(sample_sizes) - 2))
-NL_est <- diff(tapply(NL$family, NL$group, mean)) / pooled_sd
-NL_var <- (sum(sample_sizes) / prod(sample_sizes)) +
-  (NL_est^2 / (2*sum(sample_sizes)))
-
-# DK: Conduct bivariate regression
-DK_fit <- lm(conservative ~ family, data = DK)
-DK_est <- coef(DK_fit)["family"]
-DK_var <- vcov(DK_fit)["family", "family"]
-
-# US: Correlation coefficient
-US_est <- cor(US)[1, 2]
-US_var <- (1 - US_est^2)^2 / (nrow(US) - 1)
-
-# Name the estimates so hypotheses will be the same
-names(NL_est) <- names(DK_est) <- names(US_est) <- "parameter"
-```
-
-Then, we use `bain.default()` to evaluate the central hypothesis on each parameter estimate. 
-The `pbf()` function can be called on a list of the resulting bain objects.
-
-```{r eval = TRUE, echo = TRUE, results = resultaction}
-# Use bain.default() to obtain BF for the central hypothesis 
-NL_bain <- bain(x = NL_est, 
-                Sigma = matrix(NL_var, 1, 1),
-                n = nrow(NL),
-                hypothesis = "parameter > 0",
-                joint_parameters = 1)
-DK_bain <- bain(x = DK_est,
-                Sigma = matrix(DK_var, 1, 1),
-                n = nrow(DK),
-                hypothesis = "parameter > 0",
-                joint_parameters = 1)
-US_bain <- bain(x = US_est,
-                Sigma = matrix(US_var, 1, 1),
-                n = nrow(US),
-                hypothesis = "parameter > 0",
-                joint_parameters = 1)
-
-# Aggregate evidence using pbf()
-pbf(list(US_bain, DK_bain, NL_bain))
-```
-
-The results suggest substantial evidence for the hypothesis that there is a positive association between family morality and political conservatism.
-Although each study used a different method to assess this hypothesis,
-their evidence can be synthesized using `pbf()`.
-
-```{r, eval = FALSE, echo = FALSE}
-get_estimates <- function(sample){
-  number_sites <- length(levels(sesamesim$site))
-
-  sample_size <- nrow(sample)
-
-  site_estimates <- sapply(1:number_sites, function(i){
-    rbind(
-      mean(sample[sample['site'] == i, "postnumb"]),
-      sd(sample[sample['site'] == i, "postnumb"])
-      )
-    })
-
-  est_names <- sprintf("site%d", 1:number_sites)
-  site_means <- site_estimates[1,]
-  names(site_means) <- est_names
-
-  covmat <- diag((site_estimates[2,] / sqrt(sample_size))^2)
-  dimnames(covmat) <- list(est_names, est_names)
-  return(list(means = site_means, covmat = covmat, sample_size = sample_size))
-}
-
-ests_sample1 <- get_estimates(sample1)
-ests_sample2 <- get_estimates(sample2)
-ests_sample3 <- get_estimates(sample3)
-
-```
-
-```{r eval = FALSE, echo = FALSE, results = resultaction}
-bain_sample1_ests <- bain(x = ests_sample1[['means']], 
-                          Sigma = ests_sample1[['covmat']], 
-                          n = ests_sample1[['sample_size']], 
-                          hypothesis = hypotheses)
-
-bain_sample2_ests <- bain(x = ests_sample2[['means']], 
-                          Sigma = ests_sample2[['covmat']], 
-                          n = ests_sample2[['sample_size']], 
-                          hypothesis = hypotheses)
-
-bain_sample3_ests <- bain(x = ests_sample3[['means']], 
-                          Sigma = ests_sample3[['covmat']], 
-                          n = ests_sample3[['sample_size']], 
-                          hypothesis = hypotheses)
-
-bain_allsamples_ests = list(bain_sample1_ests, bain_sample2_ests, bain_sample3_ests)
-pbf(bain_allsamples_ests)
-```
-
-<!-- It is interesting that these estimates are not identical to the other examples... Is that because we do not have ALL information available in this use case, or is something off in the script? -->
-<!-- CJ: We should ask Herbert about this, but this example won't make it into this paper. If you want, you can send Herbert a reproducible example of the 'problem'!-->
-
-# Conclusion
-
-<!-- Researchers need to choose a method that best suits their research question, -->
- <!-- the study highlights the importance of carefully considering the research question when selecting a method for meta-analysis, and the potential trade-offs that may exist between sensitivity and specificity when using different methods.
--->
-
-In conclusion, this study evaluated the performance of the product Bayes factor as a method for evidence synthesis,
-and compared it against other commonly used evidence synthesis methods under different simulation conditions.
-Compared to the other methods,
-PBF had the highest overall accuracy.
-This was primarily due to its greater sensitivity.
-However, PBF had lower specificity than all other algorithms, suggesting a trade-off between sensitivity and specificity.
-The other algorithms showed ceiling effects in specificity, limiting their sensitivity.
-The performance of the PBF was most strongly affected by sample size,
-followed by the number of samples and reliability.
-We introduced a user-friendly implementation of the PBF in the `bain` R-package,
-and demonstrated its use with various analysis techniques in R,
-as well as with sufficient statistics that are already routinely coded for meta-analysis (i.e., effect sizes and their sampling variance).
-This means that researchers can now use the PBF to aggregate evidence in situations where classic meta-analytic methods are less suitable.
-For example, when one informative hypothesis has been evaluated in several replication studies,
-but these replication studies are quite heterogeneous because they sample from different populations and use different methods or analysis techniques.
-Especially when the number of replication studies is too small to adequately account for these sources of between-study heterogeneity,
-the PBF may be a useful method to aggregate evidence for the common informative hypothesis.
-Researchers should be aware that the PBF trades off increased sensitivity for decreased specificity,
-and that it addresses a different research question than other research synthesis methods.
-This highlights the importance of careful interpretation of the results,
-and consideration of the research question when selecting an aggregation method.
-In sum,
-our results suggest that PBF is a useful evidence synthesis method,
-which is now broadly accessible due to its inclusion in the `bain` R-package.
-
-\newpage
-# Highlights
-
-* Many research synthesis methods make strong assumptions about between-studies heterogeneity that are violated when studies are conceptually replicated.
-* The product Bayes factor (PBF) aggregates evidence for an informative hypothesis across conceptual replication studies without imposing assumptions about heterogeneity.
-* This paper introduces a user-friendly way to compute the PBF for a variety of widely used models via the `pbf()` function in the `bain` R-package.
-* A simulation study shows favorable performance for PBF relative to random effects meta-analysis, individual participant data meta-analysis, and vote counting.
-* Three tutorial examples illustrate distinct use cases of the method.
-
-# Data Availability Statement
-
-All analysis code is available in a version-controlled repository at <https://github.com/cjvanlissa/bayesynth>.
 

### 2024-04-09 tutorial.Rmd
update 2024-04-09 14:52:13.376312

---
title             : "A Tutorial on Aggregating Evidence from Conceptual Replication Studies using the Product Bayes Factor"
shorttitle        : "PRODUCT BAYES FACTOR"

author:
  - name: "Caspar J. Van Lissa"
    affiliation: "1"
    corresponding: yes
    address: "Professor Cobbenhagenlaan 125, 5037 DB Tilburg, The Netherlands"
    email: "c.j.vanlissa@tilburguniversity.edu"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Formal Analysis
      - Funding acquisition
      - Methodology
      - Project administration
      - Software
      - Supervision
      - Writing – original draft
      - Writing – review & editing
  - name: "Eli-Boaz Clapper"
    affiliation: "2"
    role:
      - Formal Analysis
      - Writing – original draft
      - Writing – review & editing
  - name: "Rebecca Kuiper"
    affiliation: "2"
    role:
      - Conceptualization
      - Writing – review & editing
affiliation:
  - id            : "1"
    institution   : "Tilburg University, dept. Methodology & Statistics"
  - id            : "2"
    institution   : "Utrecht University, dept. Methodology & Statistics"

authornote: |
  This is a preprint paper, generated from Git Commit # `r substr(gert::git_commit_id(),1,8)`. This work was funded by a NWO Veni Grant (NWO Grant Number VI.Veni.191G.090), awarded to the lead author.

abstract: |
  The product Bayes factor (PBF) synthesizes evidence for an informative hypothesis across heterogeneous replication studies. It can be used when fixed- or random effects meta-analysis fall short. For example, when effect sizes are incomparable and cannot be pooled, or when studies diverge significantly in the populations, study designs, and measures used. PBF shines as a solution for small sample meta-analyses, where the number of between-study differences is often large relative to the number of studies, precluding the use of meta-regression to account for these differences. Users should be mindful of the fact that the PBF answers a qualitatively different research question than other evidence synthesis methods. For example, whereas fixed-effect meta-analysis estimates the size of a population effect, the PBF quantifies to what extent an informative hypothesis is supported in all included studies. This tutorial paper showcases the user-friendly PBF functionality within the `bain` R-package. This new implementation of an existing method was validated using an extensive simulation study, available in an Online Supplement. The simulation found that PBF had a high overall accuracy, due to greater sensitivity and lower specificity, compared to random-effects meta-analysis, individual participant data meta-analysis, and vote counting. This tutorial demonstrates applications of the method in several examples based on published research. The examples use datasets included in `bain`, so readers can reproduce the examples, or apply the code to their own data. The PBF is a promising method for synthesizing evidence for informative hypotheses across conceptual replications that are not suitable for conventional meta-analysis.

keywords          : "bayes factor, evidence synthesis, bayesian, meta-analysis"
wordcount         : "5039"

bibliography      : ["product_bayes.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
knit: worcs::cite_essential
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
run_everything <- TRUE
library("papaja")
library(tidySEM)
library(kableExtra)
library(bain)
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, warning = FALSE)
library(targets)
tar_load_everything()
```

# Introduction

Recent years have seen a crisis of confidence over the reliability of published results in psychology, and science more broadly [@brembsPrestigiousScienceJournals2018].
This crisis has led to an increase in replication research as a way to derive knowledge that will stand the test of time [see @lavelleWhenCrisisBecomes2021]. 
In step with this interest in replication research,
research synthesis methods have become increasingly popular.
Such methods aggregate research findings,
and thus enable drawing overarching conclusions across multiple replication studies.
Meta-analysis is the most popular quantitative research synthesis method.
It estimates an overall population effect size based on several close replication studies.
One limitation is that its assumptions about the nature of between-studies heterogeneity may be untenable in some practical applications.
The Product Bayes Factor (PBF) is a valid alternative in such cases [@kuiperCombiningStatisticalEvidence2013].
Instead of estimating a pooled effect size across studies,
the PBF computes a Bayes factor that quantifies the amount of support the studies provide for a common informative hypothesis.
This tutorial introduces a novel user-friendly implementation of this existing method in the `bain` R-package.
Its performance was recently validated using a simulation study [@klugkistBayesianEvidenceSynthesis2023],
and the present study includes an Online Supplement with an additional simulation study to benchmark the present implementation against other commonly used research synthesis methods.
We demonstrate the use of the PBF in a series of reproducible Tutorial examples, based on re-analyis of data from published studies that previously used the PBF [@kuiperCombiningStatisticalEvidence2013; @vanleeuwenMoralityCooperationPolitics2024].

Differences between studies present a fundamental challenge to research synthesis methods [@higginsReevaluationRandomeffectsMetaanalysis2009].
Even when studies test the same hypothesis,
they often do so in different laboratories,
with idiosyncratic methods, and sample from distinct populations.
These between-study differences can introduce heterogeneity in the effect sizes found.
Meta-analysis is the most common quantitative research synthesis method.
It aggregates results of different studies by estimating an overall effect size [@borensteinIntroductionMetaAnalysis2009].
In meta-analysis, heterogeneity can be accounted for in four ways [see @vanlissaSmallSampleMetaanalyses2020].
First, if studies are exact replications, researchers might assume that they share a single true effect and estimate its value using fixed-effect meta-analysis.
Second, if heterogeneity between studies can be assumed to be normally distributed, random-effects meta-analysis can be used to estimate the mean and variance of this distribution.
This assumption holds, for example, when the effect size is influenced in a minor way by many uncorrelated factors.
Third, when there are a few moderators that cause larger systematic differences between studies,
their influence can be accounted for using meta-regression.
Note that it must be known beforehand which moderators are relevant,
and there must be sufficient studies relative to the number of moderators.
Fourth, when there are many moderators that could cause systematic differences, but it is not known beforehand which of them are relevant,
exploratory techniques like random forest meta-analysis and penalized meta-regression can be used to identify relevant moderators [@vanlissaSelectingRelevantModerators2023a].
Each of these approaches makes different assumptions about the nature of heterogeneity [see "Models for meta-analysis" in @vanlissaSmallSampleMetaanalyses2020].

A significant challenge arises in situations where none of the aforementioned assumptions are tenable.
This occurs when studies all assess the same informative hypothesis,
but differ in fundamental ways that preclude their effect sizes from being aggregated.
Consider, for example, the situation that arises when the number of moderators is large relative to the number of studies.
This issue is exemplified in the meta-analytic dataset from @kuiperCombiningStatisticalEvidence2013 in Table \@ref(tab:tabkuiper).
Each of the four studies is unique in its combination of moderator values,
making it impossible to control for their influence using meta-regression [@kuiperCombiningStatisticalEvidence2013].
The problem of multicollinear moderators is immediately evident in small datasets like this, but it also occurs in larger datasets, and may be less conspicuous [e.g., @kweeAnxiolyticEffectsEndocannabinoid2023].
Another example of fundamentally incomparable studies is the following:
Consider four studies which have sampled from the same populations, used the same study designs, the same measurement instruments, and the same statistical models.
At first glance, these studies might seem ideally suited for (fixed-effect) meta-analysis.
The situation changes, however, if each study controlled for a different set of covariates.
This would render the estimands of the effect size of interest fundamentally incomparable [@pearlSevenToolsCausal2019].
Another example occurs when transforming different effect sizes to a common metric (e.g., odds ratio to correlation coefficient).
While this is a common practice, researchers may be unaware that conversions often involve strong assumptions and may result in estimates with unknown distributional properties [@vanassenEndJustifiesAll2023].
These examples illustrate the general problem that it may not always be possible to estimate a pooled effect size across studies.
In such cases, the PBF can still be used to determine whether all studies support a common informative hypothesis.

```{r tabkuiper, eval = TRUE, echo = FALSE}
library(bain)
tab_kuiper <- bain::kuiper2013
tab_kuiper$Type <- c("Survey", "Experiment", "Experiment", "Experiment")
tab_kuiper$Sample <- c("Managers", "Managers", "Students", "Students")
tab_kuiper$Analysis <- c("Linear regression", "Linear regression", "Probit regression", "Multilevel model")
papaja::apa_table(tab_kuiper, digits = 2, caption = "Data from Kuiper et al., 2013.")
```

```{r tabmulticol, eval = TRUE, echo = FALSE, results = 'asis'}
library(metafor)
dat <- escalc(measure="RD", n1i=n1i, n2i=n2i, ai=ai, ci=ci, data=dat.hine1989)
dat <- dat[sample(nrow(dat), 4), ]
tab <- data.frame(Study = 1:4,
                  Sample = c("Europe", "Europe", "USA", "China"),
                  Design = c("Experiment", "Survey", "Experiment", "Survey"),
                  Outcome = c("Reaction time", "Self-report", "Self-report", "Reaction time"),
                  yi = dat$yi,
                  vi = dat$vi)
papaja::apa_table(tab, caption = "Example of incomparabe studies", digits = 2)
```

### What are Informative Hypotheses?

Informative hypotheses relate the parameters of interest (e.g., $\beta$) to each other and/or to numerical constants by means of (in)equality constraints. The conventional null hypothesis $H_0$ can be thought of as a special case of an informative hypothesis, which specifies that the parameter is equal to the value zero, $H_i: \beta = 0$. But informative hypotheses can be much more complex, and ideally, should closely represent the theoretical expectation. If one is confident that a new drug  one might hypothesize that a parameter is larger than a smallest effect size of interest, $H_i: \beta > 0.1$ [see @lakensEquivalenceTestsPractical2017]. If a dose-response effect is expected for a certain drug, one might hypothesize that three treatments of increasing intensity will have effects of increasing size, and that all these treatments will exceed the smallest effect size of interest, $H_i: (\beta_1 < \beta_2 < \beta_3) > 0.1$.

### Bayesian Informative Hypothesis Tests

The amount of evidence for an informative hypothesis $H_i$ relative to another hypothesis can be expressed as a Bayes factor, or BF.
Bayes factors for one hypothesis are thus always defined in relation to another hypothesis.
In this tutorial paper, all informative hypotheses $H_i$ are compared to their complement $H_{!i}$ [see @guApproximatedAdjustedFractional2018].
This complement $H_{!i}$ can be interpreted as "not $H_i$".
This Bayes factor, which we will refer to as $BF_c$, represents the ratio of evidence for $H_i$ relative to the evidence against it.
A value of $BF_c = 10 = \frac{10}{1}$ means that the data provide ten times more support for the hypothesis than against it;
a value of $BF_c = 0.1= \frac{1}{10}$ means that the data provide ten times less support for the hypothesis than against it.

Testing hypotheses using the Bayes factor deviates in important ways from conventional null-hypothesis significance testing.
It involves thinking about probability in a more subjective way, which allows one to express confidence in the truth of a hypothesis.
This deviates from frequentist thinking, where probability is defined in objective terms, as a function of the frequency with which an event would occur if a procedure would be repeated many times.
As a result of these differences in interpretation, Bayes factors provide a continuous measure of evidence for the informative hypothesis, rather than a binary decision to (not) reject a typically uninformative null hypothesis.
Bayes factors can also be used to evaluate relative support for competing informative hypotheses.
Moreover, they are robust to multiple- and sequential testing.
The Bayes factor approach also has downsides; most notably, the absence of hard-and-fast decision criteria [although some have suggested cut-offs of $BF > 3$ or $BF > 10$ as conclusive evidence, @rouderBayesianTestsAccepting2009].
Importantly, the Bayes factor is not inherently better or worse than null-hypothesis significance testing.
It is just an alternative inferential procedure, which - under some conditions - performs similarly to null-hypothesis significance tests [@mulderPriorAdjustedDefault2014].

### How are Bayes Factors Calculated?

This tutorial uses the *Approximated adjusted fractional Bayes factor* (AAFBF) first introduced by @mulderPriorAdjustedDefault2014.
Conceptually, AAFBFs are computed by integrating so-called posterior and prior distributions with respect to the parameter constraints specified in the hypothesis, and taking the ratio of these quantities.
The term *approximated* in AAFBF refers to the fact that the prior and posterior are both approximated by normal distributions,
which makes it possible to compute these Bayes factors from sufficient statistics.
The (multivariate) normal posterior is centered around the maximum likelihood estimate of the parameter(s),
and its its covariance consists of the asymptotic (co)variance matrix of the parameters.
For a single parameter, this simplifies to the parameter estimate and its sampling variance (the squared standard error).
The term *adjusted* in AAFBF refers to the fact that the center of the prior distribution is adjusted based on the set of informative hypotheses,
and term *fractional* refers to the fact that the the prior covariance matrix is based on a fraction of the information contained in the posterior covariance matrix.
Computation is simplified by sampling random values from these normal distributions instead of integrating over them analytically.
This introduces slight random fluctuations in the results.
To make the analyses reproducible, it is therefore important to set a "random seed",
which ensures that the same pseudo-random numbers will be generated upon repeated evaluation of the code.
This is done throughout the tutorial examples.
For a more detailed technical description of the AAFBF, readers can turn to @guApproximatedAdjustedFractional2018; for mathematical derivations, see @mulderPriorAdjustedDefault2014.

### The Product Bayes Factor

The PBF is a Bayesian evidence synthesis (BES) technique that aggregates the evidence for a theoretical relationship across studies,
without imposing assumptions about heterogeneity.
As explained in @kuiperCombiningStatisticalEvidence2013, aggregating evidence across studies comes down to taking the product of Bayes factors from these individual studies.
Note that Bayes factors are always a ratio of evidence in favor of one hypothesis over another hypothesis.
The PBF assumes that these hypotheses are a priori equally likely.

When multiple studies each provide evidence for $H_i$ in the form of complement Bayes factors,
these Bayes factors can be synthesized across studies by taking their product [@kuiperCombiningStatisticalEvidence2013].
The resulting product Bayes factor (PBF) summarizes the total evidence for the hypothesis.
The only assumption of the PBF is that all study-specific hypotheses provide evidence about the same underlying theoretical relationship.
Note that other approaches to BES exist;
for instance, it is possible to use the posterior of one study as the prior for a replication study, and thus accumulate evidence across studies [see @heckReviewApplicationsBayes2022].
Such applications are out of scope of the present paper,
which addresses the PBF approach to BES.

Although meta-analysis and PBF are both research synthesis methods, they answer different research questions.
Meta-analysis estimates the point estimate or distribution of a population effect size. 
It pools estimates of this effect size across multiple studies to obtain an overall estimate of the effect size.
If all assumptions are met, this is a consistent estimate of the population effect size.
It thus answers questions like:
Given certain assumptions about between-studies heterogeneity, what is the average population effect size?
The PBF, on the other hand, aggregates evidence for an informative hypothesis across multiple studies.
It thus answers the question: Do all studies support the hypothesis of interest?
The PBF is emphatically *not* a consistent estimator of the amount of support the hypothesis would receive based on individual participant data.
Both methods thus answer different research questions,
and provide complementary information.

### Are Studies' Sample Sizes Taken into Account?

Conventional meta-analysis can be conceptualized as a weighted average of study effect sizes [@borensteinIntroductionMetaAnalysis2009].
Each study's sample size indirectly affects its weight, via its influence on the sampling variance (the squared standard error).
The approximate fractional Bayes factors also accounts for sample size to some extent,
as the posterior (co)variance used to calculate the Bayes factors is the sampling variance of the estimate,
just as in meta-analysis.
The prior distribution is also based, to some extent, on the sample size [@hoijtinkTutorialTestingHypotheses2019].
The resulting Bayes factor for a true hypothesis increases monotonically with increasing sample size.
Thus, the overall support for a true hypothesis will be greater if it is based on larger, rather than smaller, samples.

## Simulation Study

This tutorial paper introduces a new implementation of the existing PBF method.
A simulation study validating this new implementation is available in an Online Supplement, and the reproducible code is available at [MASKED FOR REVIEW].
We simulated a PBF analysis of correlation coefficients and manipulated the presence or absence of a true population effect, the sample size per study, the number of replication studies, and the the reliability of the correlated variables. We compared the performance of the PBF against vote counting [@hedgesVotecountingMethodsResearch1980], random-effects meta-analysis [@viechtbauerConductingMetaanalysesMetafor2010], and individual participant data meta-analysis [@rileyMetaanalysisIndividualParticipant2010].
PBF had the highest overall accuracy, primarily due to its greater sensitivity to detecting a true effect.
However, PBF had lower specificity than all other algorithms, suggesting a trade-off between sensitivity and specificity.
The other algorithms showed ceiling effects in specificity, limiting their sensitivity.
The performance of the PBF was most strongly affected by sample size,
followed by the number of samples and reliability.

## This Tutorial Paper

This paper introduces the first implementation of the PBF in user-friendly free open source software.
This tutorial requires the `bain` R-package for Bayesian informative hypothesis evaluation, version `0.2.11`, which contains the `pbf()` function and all tutorial data.
The Online Supplement presents a simulation study that validates our implementation of the PBF and benchmark it against alternative evidence synthesis methods, including random-effects restricted maximum likelihood meta-analysis, individual participant data meta-analysis, and vote counting.
This paper focuses on applications of the PBF, illustrated by several reproducible examples.
This tutorial was, itself, made reproducible using the Workflow for Open Reproducible Code in Science (WORCS).
The code archive is available at <https://github.com/cjvanlissa/bayesynth>.

# Tutorial

This tutorial demonstrates how to synthesize evidence for an informative hypothesis across heterogeneous replications using the Product Bayes Factor (PBF). 
We assume that users have installed the free open source statistical programming language R [@rcoreteamLanguageEnvironmentStatistical2022].
The R-package `bain` can be installed by running `install.packages("bain")` in the R console.
The datasets used in this tutorial are all included in the `bain` package.
The `kuiper2013` dataset is based on @kuiperCombiningStatisticalEvidence2013 and the original publications meta-analyzed therein.
Its documentation is accessed by running `?kuiper2013` in the R console.
The other datasets were simulated based on data presented in [@vanleeuwenMoralityCooperationPolitics2024].
Their documentation is accessed by running `?synthetic_us`, `?synthetic_dk` or `?synthetic_nl`.
While we introduce the basic functionality of the `bain` package,
we direct the interested reader to @hoijtinkTutorialTestingHypotheses2019 for a general introduction to `bain`,
and to @vanlissaTeacherCornerEvaluating2020 for an in-depth tutorial on informative hypothesis tests for Structural Equation Models using `bain`.

## Tutorial 1: When Meta-Analysis Falls Short

The PBF was first introduced by @kuiperCombiningStatisticalEvidence2013
for cases where random-effects meta-analysis would otherwise be used,
but is deemed inappropriate because its assumptions are likely to be violated.
This Tutorial illustrates the use of the PBF in such cases,
and compares it to meta-analysis to show that one can straightforwardly perform a PBF analysis with a dataset prepared for meta-analysis.
As explained in the introduction, if the hypothesis of interest pertains to only one parameter, then the Bayes factor can be computed using the estimate and its standard error [@hoijtinkTutorialTestingHypotheses2019].
It is also possible to formulate more complex hypotheses that involve multiple parameters; for example, in structural equation models [see @vanlissaTeacherCornerEvaluating2020].
This requires access to the parameter covariance matrix.
The present example focuses on sufficient statistics to illustrate how one might complement a conventional meta-analysis with a PBF analysis and report both in the same paper.
This way, readers can determine whether they trust the assumption of normally distributed heterogeneity and interpret the random-effects estimate,
or doubt the assumption and interpret the PBF instead.

The data for this example is shown in Table \@ref(tab:tabkuiper).
Run `?kuiper2013` to view its documentation.
Kuiper and colleagues [-@kuiperCombiningStatisticalEvidence2013] set out to aggregate evidence for the effect of prior interactions between partners on trust in (economic) exchange relations across four heterogeneous replication studies.
All studies investigated the informative hypothesis that past (experience) with a seller has a positive effect on trust.
Batenburg et al. (2003) analyzed survey data using linear regression with covariates; Buskens and Raub (2002) analyzed experimental data using linear regression; Buskens and Weesie (2000) used an experimental design with a binary outcome, analyzed using probit regression; and Buskens, Raub, and Van der Veer (2010) used a longitudinal experimental design, analyzing the data with a three-level logistic regression.
These studies each provide a regression coefficient assessing the effect of past experience on trust, and its estimated sampling variance (squared standard error).
However, because the studies differ in terms of design (survey vs. experiment),
operationalization of variables, measurement level of variables, and statistical model used,
these regression coefficients are not directly comparable and should not be pooled using meta-analysis.
The authors developed the PBF to determine whether all studies support the informative hypothesis.

Although there is one clear informative hypothesis, we follow the original study and estimate the evidence for three competing hypotheses: $H_1: \beta = 0$ (the null hypothesis that the effect of past on trust is zero), $H_2: \beta > 0$ (a directional hypothesis that the effect of past on trust is positive), and $H_3: \beta < 0$ (a directional hypothesis that the effect of past on trust is negative).
Note that Kuiper and colleagues computed the Bayes factors and posterior model probabilities by hand, using custom priors for the first study, and using the posterior of the first study as prior for subsequent studies.
The present tutorial instead uses approximate fractional Bayes factors computed using `bain`, with default settings. 
The prior for these Bayes factors is derived from the data.
Consequently, the numerical results differ from Kuiper and colleagues, although the conclusions remain the same.

We first conduct a random-effects meta-analysis using the function `rma()` from the `metafor` package.
Then, we perform a PBF analysis using the `pbf()` function in the `bain` package.
This allows us to compare the interface of both functions and their results.
First, we will load the required packages.
This also makes the `kuiper2013` available.
Printing it to the console should give the same result as Table \@ref(tab:tabkuiper).

```{r echo = TRUE, eval = TRUE, results='hide'}
library(metafor)
library(bain)
kuiper2013
```

To perform a random-effects meta-analysis, run the following code:

```{r echo = TRUE, eval = FALSE}
rma(yi = kuiper2013$beta, vi = kuiper2013$vi)
```

```{r echo = FALSE, eval = TRUE}
out <- capture.output(print(res_rma, digits = 2))
out <- out[c(2:4, 8:15)]
out <- out[!out == ""]
cat(out, sep = "\n")
```

Note that the pooled effect size is $\beta `r report(res_rma[["b"]][1,1])`$.
Considering our hypothesis is one-sided, we can divide the p-value by two, and report $p `r report(res_rma[["pval"]]/2)`$.
Thus, there is a significant positive effect of prior interactions on trust.
However, this analysis assumes a normal distribution of population effect sizes.
We doubt this assumption, because the studies are all qualitatively different.
While empirical evidence for this violation of assumptions is given by a significant heterogeneity test,
we do not need to make a purely data-driven decision [@wichertsDegreesFreedomPlanning2016].
We can support the assumption that population effect sizes are *not* normally distributed on theoretical grounds.

We now perform the PBF analysis, using the `pbf()` method for numeric input (see `?pbf`).
This interface is very similar to `rma()`, and is specifically designed for applications where PBF is applied to meta-analytic datasets.
Because PBF relies on Monte Carlo estimation (i.e., randomly sampling values),
however, it is advisable to set a seed to make the analysis reproducible.
Throughout this tutorial we use the value `set.seed(1)`,
but in your analyses, make sure to select a different unique value.
Run the following code:

```{r echo = TRUE, eval = FALSE}
set.seed(1)
pbf(yi = kuiper2013$beta,
    vi = kuiper2013$vi,
    ni = kuiper2013$n,
    hypothesis = "y = 0; y > 0; y < 0")
```
```{r echo = FALSE, eval = TRUE}
print_pbf <- function(x){
  out <- x
  out[,] <- sapply(x, formatC, digits = 2, format = "f")
  if(any(nchar(unlist(out)) > 8)){
    for(i in 1:nrow(out)){
      for(j in 1:ncol(out)){
        if(nchar(out[i,j]) > 8){
          out[i,j] <- formatC(x[i,j], format = "e", digits = 2)
        }
      }
    }
  }
  print(out)
}
print_pbf(res_pbf_t1)
```

Note that the `yi` and `vi` arguments are the same as those of `rma()`.
Additional argument `ni` is used to construct the prior for the approximate Bayes factors [@hoijtinkTutorialTestingHypotheses2019].
Importantly, the `hypothesis` argument determines which informative hypotheses are tested.
Its default value of `y = 0` is similar to a null hypothesis test.
Here, we override this default value to test our three informative hypotheses.
The resulting output shows Bayes factors for each of the three hypotheses (PBF column), as well as the study-specific evidence for each hypothesis (remaining columns).
Note that the PBF for hypotheses $H_1$ and $H_3$ are approximately zero; there is essentially no support in the data that the effect of prior interaction on trust is equal to, or smaller than, zero.
Support for $H_2$ is overwhelming, however.
Thus, we can conclude that all four included studies support the informative hypothesis that there is a positive effect of prior interaction on trust.

## Tutorial 2: Computing a Bayes Factor

The preceding example used a simplified interface to the PBF, designed to be similar to `rma()` to facilitate analyzing data initially prepared for meta-analysis.
The simplified interface internally performs several intermediate steps.
These next tutorials go through those steps one by one, to help users understand the calculations involved.
As these calculations require individual participant data,
we cannot use the sufficient statistics from Tutorial 1.
For these tutorials we use the synthetic data based on
Van Leeuwen and colleagues [@vanleeuwenMoralityCooperationPolitics2024].
They conducted a theory-driven, preregistered study of the informative hypothesis that higher self-reported moral dispositions would be associated with a more conservative socio-political orientation.
Data were collected in three countries: the United states of America, Denmark, and the Netherlands.
Each sample contained multiple measures of political orientation and moral dispositions.
In the original publication, the PBF was used to aggregate evidence across scales and countries to obtain an overall measure of support for the informative hypothesis.
This tutorial follows the same rationale, but uses only one effect size per sample.
We intentionally vary the way this effect size is computed to illustrate the use of the PBF in cases when the same informative hypothesis has been studied in different ways in multiple studies.
Specifically, we will examine the informative hypothesis that self-reported importance of family morality is positively associated with a conservative socio-political orientation.

We must estimate a model suitable for evaluating this informative hypothesis.
Because both scales consist of multiple items, we can use structural equation modeling (SEM) to perform latent variable regression [see @vanlissaTeacherCornerEvaluating2020]:

```{r echo = TRUE, eval = FALSE}
# Load lavaan package for SEM
library(lavaan)

# Specify SEM-model for latent variable regression
model_nl <- "
fam =~ fam_1 + fam_2 + fam_3
con =~ sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
       sepa_eco_1 + sepa_eco_2 + sepa_eco_3 + sepa_eco_4 + sepa_eco_5
con ~ beta * fam"

# Estimate the model in lavaan
results_nl <- sem(model = model_nl, data = synthetic_nl)
```

Whereas the previous Tutorial used a test value of zero, similar to conventional frequentist null hypothesis testing, this tutorial uses a smallest effect size of interest [@lakensEquivalenceTestsPractical2017].
Thus, our informative hypothesis is $H_i: \beta > .1$, where $\beta$ (beta) is the standardized (partial) regression coefficient, and we hypothesize that its value will be at least $.1$.

The code below illustrates how to use the `bain()` function to obtain a Bayes factor for this informative hypothesis.
In the `hypothesis` argument, we can refer to the parameter `beta` by name because we labeled it in the `lavaan` syntax above.
If we had not done labeled the parameter, we could inspect all parameter names and their values by running `get_estimates(results_nl, standardize = TRUE)`.

```{r echo = TRUE, eval = FALSE}
# Test that the effect labeled 'beta' is positive
set.seed(1)
bf_nl <- bain(results_nl,
              hypothesis = "beta > .1",
              standardize = TRUE)
bf_nl
```

```{r echo = FALSE, eval = TRUE}
print_bain <- function(x, stats = c("Fit", "Com", "BF.u", "BF.c","PMPa", "PMPb"),
                       digits = 3,
                       na.print = "", ...){

  fits <- as.matrix(x$fit)

  if (dim(fits)[2] == 12){stats[7] <- "PMPc"}

  dat <- fits[, stats]
  miss_val <- is.na(dat)
  dat <- formatC(dat, digits = digits, format = "f")
  dat[miss_val] <- ""
  model_type <- class(x$model)[1]
  if(model_type == "lm"){
    model_type <- paste0(model_type, " (", attr(x, "which_model"), ")")
  }
  cat("Bayesian informative hypothesis testing for an object of class ", model_type, ":\n\n", sep = "")

  prmatrix(dat,
           quote = FALSE,
           na.print = na.print)

  if (dim(fits)[2] == 12){
  cat("\nHypotheses:\n ", paste(rownames(dat)[c(-nrow(dat),-(nrow(dat)-1))], ": ", x$hypotheses, sep = "", collapse = "\n  "))}
  if (dim(fits)[2] == 11){
    cat("\nHypotheses:\n ", paste(rownames(dat)[-nrow(dat)], ": ", x$hypotheses, sep = "", collapse = "\n  "))}

}
print_bain(res_t2, digits = 2)
```

The results indicate that the informative hypothesis receives about 23 times as much support from the data relative to its complement, which is convincing evidence.

## Tutorial 3: Aggregating Bayes Factors

As mentioned before, suitable data were collected to evaluate the informative hypothesis in three countries.
There are differences between countries that prevent analyzing these data as a multilevel model, however.
For instance, conservatism was measured using different scales.
This is an appropriate situation to use the PBF to aggregate evidence across countries.
Below, we estimate a latent regression model for the remaining two countries, taking care to use the same label for the parameter of interest in all samples.
Then, we bind all three SEM-models in a list, and call PBF to evaluate the hypothesis of interest on all models and aggregate the evidence.
As the BF in all three samples is positive, the resulting PBF is very large.
We can thus conclude that the central hypothesis receives overwhelming support across samples.

```{r, eval = FALSE, echo = TRUE}
# Specify the models for synthetic_dk and synthetic_us
model_dk <- "
fam =~ fam_1 + fam_2 + fam_3
con =~
sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
sepa_eco_1 + sepa_eco_2 + sepa_eco_3 + sepa_eco_4 + sepa_eco_5
con ~ beta * fam"
model_us <- "
fam =~ fam_1 + fam_2 + fam_3
con =~
secs_soc_1 + secs_soc_2 + secs_soc_3 + secs_soc_4 + secs_soc_5 +
secs_soc_6 + secs_soc_7 +
secs_eco_1 + secs_eco_2 + secs_eco_3 + secs_eco_4 + secs_eco_5
con ~ beta * fam"

# Estimate the model in lavaan
results_dk <- sem(model = model_dk, data = synthetic_dk)
results_us <- sem(model = model_us, data = synthetic_us)

# Bind the models into a list
results <- list(results_nl, results_dk, results_us)
# Test the hypothesis that the effect size labeled 'beta' is positive
set.seed(1)
pbf(results, hypothesis = "beta > .1", standardize = TRUE)
```
```{r, eval = TRUE, echo = FALSE}
print_pbf(res_pbf_t3)
```

## Tutorial 4: Using `bain` Objects

The `pbf()` function also accepts multiple `bain` objects.
This makes it possible to, for example, evaluate different sets of hypotheses on different data sets before using the resulting `bain` objects to aggregate the evidence for all common hypotheses across datasets.
The example below illustrates this use case.
As before, all analyses share one hypotheses in common ($H_i: \beta_{fam} > .1$),
but the Dutch sample now contains a sample-specific hypothesis regarding the effect of group morality, namely that $\beta_{grp} < .1$.
Note that controlling for the effect of group morality affects the effect of family morality on conservatism, which is now a partial effect.
This also affects the Bayes factor for this sample.
The `pbf()` function is called on a list of `bain` objects.
Note that, in this case, `pbf()` does not require an argument `hypothesis`, as the hypotheses are contained in the individual `bain` objects.

```{r, echo = TRUE, eval = FALSE}
# Add the additional predictor to the model, label the effect beta2
model_nl2 <- c(model_nl, "group =~ grp_1 + grp_2 + grp_3
                         con ~ beta2 * group")

# Estimate the model in lavaan
results_nl2 <- sem(model = model_nl2, data = synthetic_nl)

# Obtain BF for each sample
# Note that the Dutch sample has two hypotheses:
set.seed(1)
bf_nl2 <- bain(results_nl2,
               hypothesis = "beta > .1; beta2 < .1", 
               standardize = TRUE)
bf_dk <- bain(results_dk, hypothesis = "beta > .1", standardize = TRUE)
bf_us <- bain(results_us, hypothesis = "beta > .1", standardize = TRUE)
```

```{r, echo = TRUE, eval = FALSE}
# Bind bain objects into a list
bfs <- list(bf_nl2, bf_dk, bf_us)

# Call pbf on that list
pbf(bfs)
```

```{r, echo = FALSE, eval = TRUE}
print_pbf(res_pbf4)
```

In the output of this analysis, hypotheses common to all `bain` objects are retained and aggregated, but the sample-specific hypothesis of the first object is omitted.
If there are no common hypotheses across all objects, `pbf()` throws an error.
As can be seen, the results are somewhat different due to the additional control variable - but the conclusions are equivalent to the previous tutorial.
Note the location of the `set.seed()` command: it is called before the initial `bain()` call,
because the `bain()` function relies on random sampling - but `pbf()` does not.
Compare this to Tutorial 3, where we called `set.seed()` before `pbf()`.
In that case, `pbf()` called the `bain()` function internally.

## Tutorial 5: Using Sufficient Statistics

Bringing our examples full circle, we once again illustrate how to compute the PBF from sufficient statistics - but in this case, we will calculate them ourselves instead of extracting them from published papers as in Tutorial 1.
In this Tutorial, we use the default interface of `bain`, as explained in [@hoijtinkTutorialTestingHypotheses2019].
This function requires four arguments: A named vector of parameter estimates, their asymptotic covariance matrix, the original sample size, and the number of within-group and between-group parameters.
Note that, when analyzing a single parameter per sample, the standard error is sufficient to construct the asymptotic covariance matrix.

The present use case evaluates the following hypothesis:
*There is a positive association between family morality and political conservatism*.
This conceptual hypothesis is evaluated differently in the three samples, resulting in three different types of statistics and distinct sample-specific hypotheses:

1. A t-test was performed using the synthetic_nl data; using Cohen's D gives $H_i^{synthetic_nl}: \delta_{conservative > liberal} > 0$, where $\delta$ is the mean difference between groups.
1. A bivariate regression coefficient was calculated using the synthetic_dk data, giving $H_i^{synthetic_dk}: \beta_{fam} > 0$
1. A correlation coefficient was calculated using the synthetic_us data, giving $H_i^{synthetic_us}: \rho_{fam,con} > 0$, where $\rho$ is the correlation between family morality and conservatism.

Note that we intentionally manipulate the data to illustrate these different analyses;
for example, we compute mean scale scores and dichotomize the continuous conservatism scale to conduct a t-test.
We do not advocate these practices for applied research.

First we obtain the relevant parameter estimates and their sampling variances,
which allows us to evaluate the specific hypotheses in `bain`:

```{r eval = FALSE, echo = TRUE}
# Create mean scale scores
synthetic_nl <- data.frame(
  family = rowMeans(synthetic_nl[c("fam_1", "fam_2", "fam_3")]),
  conservative = rowMeans(synthetic_nl[c("sepa_soc_1", "sepa_soc_2", "sepa_soc_3",
                               "sepa_soc_4", "sepa_soc_5", "sepa_eco_1",
                               "sepa_eco_2", "sepa_eco_3", "sepa_eco_4",
                               "sepa_eco_5")]))
synthetic_dk <- data.frame(
  family = rowMeans(synthetic_dk[c("fam_1", "fam_2", "fam_3")]),
  conservative = rowMeans(synthetic_dk[c("sepa_soc_1", "sepa_soc_2", "sepa_soc_3",
                               "sepa_soc_4", "sepa_soc_5", "sepa_eco_1",
                               "sepa_eco_2", "sepa_eco_3", "sepa_eco_4",
                               "sepa_eco_5")]))

synthetic_us <- data.frame(
  family = rowMeans(synthetic_us[c("fam_1", "fam_2", "fam_3")]),
  conservative = rowMeans(synthetic_us[c("secs_soc_1", "secs_soc_2", "secs_soc_3",
                               "secs_soc_4", "secs_soc_5", "secs_soc_6",
                               "secs_soc_7", "secs_eco_1", "secs_eco_2",
                               "secs_eco_3", "secs_eco_4", "secs_eco_5")]))

# synthetic_nl: Conduct t-test using Cohen's D
synthetic_nl$group <- cut(synthetic_nl$conservative, breaks = 2,
                labels = c("liberal", "conservative"))
sample_sizes <- table(synthetic_nl$group)
sds <- tapply(synthetic_nl$family, synthetic_nl$group, sd)
pooled_sd <- sqrt(sum((sample_sizes - 1) * sds) / (sum(sample_sizes) - 2))
NL_est <- diff(tapply(synthetic_nl$family, synthetic_nl$group, mean)) / pooled_sd
NL_var <- (sum(sample_sizes) / prod(sample_sizes)) +
  (NL_est^2 / (2*sum(sample_sizes)))

# synthetic_dk: Conduct bivariate regression
DK_fit <- lm(conservative ~ family, data = synthetic_dk)
DK_est <- coef(DK_fit)["family"]
DK_var <- vcov(DK_fit)["family", "family"]

# synthetic_us: Correlation coefficient
US_est <- cor(synthetic_us)[1, 2]
US_var <- (1 - US_est^2)^2 / (nrow(synthetic_us) - 1)

# Name the estimates so hypotheses will be the same
names(NL_est) <- names(DK_est) <- names(US_est) <- "parameter"
```

Then, we use `bain.default()` to evaluate the informative hypothesis on each parameter estimate. 
The `pbf()` function can be called on a list of the resulting bain objects.

```{r eval = FALSE, echo = TRUE}
# Use bain.default() to obtain BF for the central hypothesis 
NL_bain <- bain(x = NL_est, 
                Sigma = matrix(NL_var, 1, 1),
                n = nrow(synthetic_nl),
                hypothesis = "parameter > 0",
                joint_parameters = 1)
DK_bain <- bain(x = DK_est,
                Sigma = matrix(DK_var, 1, 1),
                n = nrow(synthetic_dk),
                hypothesis = "parameter > 0",
                joint_parameters = 1)
US_bain <- bain(x = US_est,
                Sigma = matrix(US_var, 1, 1),
                n = nrow(synthetic_us),
                hypothesis = "parameter > 0",
                joint_parameters = 1)

# Aggregate evidence using pbf()
pbf(list(US_bain, DK_bain, NL_bain))
```

```{r echo = FALSE, eval = TRUE}
print_pbf(res_pbf_t5)
```

The results suggest substantial evidence for the hypothesis that there is a positive association between family morality and political conservatism.
Although each study used a different method to assess this hypothesis,
their evidence can be synthesized using `pbf()`.

# Conclusion

<!-- Researchers need to choose a method that best suits their research question, -->
 <!-- the study highlights the importance of carefully considering the research question when selecting a method for meta-analysis, and the potential trade-offs that may exist between sensitivity and specificity when using different methods. -->

The Product Bayes Factor (PBF) is a suitable solution when aggregating evidence across conceptual replication studies that are too heterogeneous to meet the assumptions of conventional meta-analysis.
In contrast to meta-analysis, the PBF does not estimate the value of an overall effect size and its heterogeneity.
Instead, the PBF quantifies support for a common informative hypothesis across studies.
While this may be perceived as a loss of information, the PBF is used in situations where the assumptions of conventional meta-analysis are violated, and consequently, such estimates would not be informative.

One important distinction between meta-analysis and PBF is that the former provides ever-more precise estimates of the overall effect size as more studies are added.
There is no straightforward parallel for the PBF.
If all aggregated studies support the informative hypothesis, the PBF will increase as more studies are added.
However, if the studies offer mixed evidence of the informative hypothesis, the PBF can be inconclusive, even for a large number of studies.
This is because the PBF essentially answers the question:
do these studies all support the informative hypothesis?
If the answer is no, the results will reflect that.

We introduced a user-friendly implementation of the PBF in the `bain` R-package,
and demonstrated its use with various analysis techniques in R,
as well as with sufficient statistics that are already routinely coded for meta-analysis (i.e., effect sizes and their sampling variance).
This means that researchers can now use the PBF to aggregate evidence in situations where classic meta-analytic methods are less suitable.
For example, when one informative hypothesis has been evaluated in several replication studies,
but these replication studies are quite heterogeneous because they sample from different populations and use different methods or analysis techniques.
Especially when the number of replication studies is too small to adequately account for these sources of between-study heterogeneity,
the PBF may be a useful method to aggregate evidence for the common informative hypothesis.
Researchers should be aware that the PBF trades off increased sensitivity for decreased specificity,
and that it addresses a different research question than other research synthesis methods.
This highlights the importance of careful interpretation of the results,
and consideration of the research question when selecting an aggregation method.
In sum,
our results suggest that PBF is a useful evidence synthesis method,
which is now broadly accessible due to its inclusion in the `bain` R-package.

\newpage
# Highlights

* Many research synthesis methods make strong assumptions about between-studies heterogeneity that are violated when studies are conceptually replicated.
* The product Bayes factor (PBF) aggregates evidence for an informative hypothesis across conceptual replication studies without imposing assumptions about heterogeneity.
* This paper introduces a user-friendly way to compute the PBF for a variety of widely used models via the `pbf()` function in the `bain` R-package.
* A simulation study shows favorable performance for PBF relative to random effects meta-analysis, individual participant data meta-analysis, and vote counting.
* Three tutorial examples illustrate distinct use cases of the method.

# Data Availability Statement

All analysis code is available in a version-controlled repository at <https://github.com/cjvanlissa/bayesynth>.

# Conflict of Interest Statement

The authors declare no conflict of interest.

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup

### 2024-04-09 _targets.R

@@ -1,0 +1,206 @@
+# Created by use_targets().
+# Follow the comments below to fill in this target script.
+# Then follow the manual to check and run the pipeline:
+#   https://books.ropensci.org/targets/walkthrough.html#inspect-the-pipeline
+
+# Load packages required to define the pipeline:
+library(targets)
+
+# Set target options:
+tar_option_set(packages = c("tidySEM", "kableExtra", "bain", "lavaan", "metafor"))
+
+list(
+  tar_target(
+    name = res_rma,
+    command = rma(yi = kuiper2013$beta,
+                  vi = kuiper2013$vi)
+  )
+  , tar_target(
+    name = res_pbf_t1,
+    command = {
+      set.seed(1)
+      pbf(yi = kuiper2013$beta, vi = kuiper2013$vi, ni = kuiper2013$n,
+          hypothesis = "y = 0; y > 0; y < 0")}
+  )
+  
+  , tar_target(
+    name = res_t2,
+    command = {
+      
+      # Load lavaan package for SEM
+      library(lavaan)
+      
+      # Specify SEM-model for latent variable regression
+      model_nl <- "
+fam =~ fam_1 + fam_2 + fam_3
+con =~ sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
+       sepa_eco_1 + sepa_eco_2 + sepa_eco_3 + sepa_eco_4 + sepa_eco_5
+con ~ beta * fam"
+      
+      # Estimate the model in lavaan
+      results_nl <- sem(model = model_nl, data = synthetic_nl)
+      # Test that the effect labeled 'beta' is positive
+      set.seed(1)
+      bain(results_nl,
+           hypothesis = "beta > .1",
+           standardize = TRUE)
+      
+      
+    }
+  )
+  
+  , tar_target(
+    name = res_pbf_t3,
+    command = {
+      model_nl <- "
+fam =~ fam_1 + fam_2 + fam_3
+con =~ sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
+       sepa_eco_1 + sepa_eco_2 + sepa_eco_3 + sepa_eco_4 + sepa_eco_5
+con ~ beta * fam"
+      
+      # Estimate the model in lavaan
+      results_nl <- sem(model = model_nl, data = synthetic_nl)
[…]

### 2024-09-11 tutorial.Rmd
fix reproducibility 1

@@
 library(targets)
-tar_load_everything()
 ```
@@
 ```{r echo = FALSE, eval = TRUE}
+tar_load("res_rma")
 out <- capture.output(print(res_rma, digits = 2))
@@
 }
+tar_load("res_pbf_t1")
 print_pbf(res_pbf_t1)
@@
 }
+tar_load("res_t2")
 print_bain(res_t2, digits = 2)
@@
 ```{r, eval = TRUE, echo = FALSE}
+tar_load("res_pbf_t3")
 print_pbf(res_pbf_t3)
@@
 ```{r, echo = FALSE, eval = TRUE}
+tar_load("res_pbf4")
 print_pbf(res_pbf4)
@@
 ```{r echo = FALSE, eval = TRUE}
+tar_load("res_pbf_t5")
 print_pbf(res_pbf_t5)

### 2024-09-11 _targets.R

@@ -202,5 +202,5 @@ con ~ beta * fam"
                pbf(list(US_bain, DK_bain, NL_bain))
              })
 
-, tarchetypes::tar_render(manuscript, "tutorial.Rmd", cue = tar_cue(mode = "always"))
+# , tarchetypes::tar_render(manuscript, "tutorial.Rmd", cue = tar_cue(mode = "always"))
 )

### 2024-09-11 _targets.R
update 2024-09-11 11:36:43.747856

@@ -201,6 +201,6 @@ con ~ beta * fam"
                # Aggregate evidence using pbf()
                pbf(list(US_bain, DK_bain, NL_bain))
              })
-
+# 
 # , tarchetypes::tar_render(manuscript, "tutorial.Rmd", cue = tar_cue(mode = "always"))
 )
